### 2019-10-19 Common/chapters/technical_info.tex
Julien and Johan: add technical part about NIfTI and DICOM

% !TEX root = ../main.tex

\chapter{Technical information}
\label{ch:technical_info}

\section{NIfTI}

\subsubsection{Origin of the format}
The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format. The main problem of the latter was “the lack of adequate information about orientation in space, such that the stored data could not be unambiguously interpreted” .  For instance, there was a real confusion to determine the left and right sides of brain images.  Hence, the NIfTI file format was defined to obviate this major issue.

\subsubsection{Organization of data}
Unlike the ANALYZE format that used two files to store the meta-data and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility. This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data. The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.

\subsubsection{Overview of the header structure}
With the goal of preserving the compatibility between the “analyze” and the NIfTI format, both headers have the same size of 348 bytes. “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”. Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}

\section{DICOM}

\subsection{Origin}

The acronym DICOM stands for Digital Imaging and Communications in Medicine. Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks. For that reason, the ACR (American College of Radiology) and the NEMA (National Electrical Manufacturers Association) created a committee to build a standard. After two iterations with other names, DICOM was created in 1993. It standardized the representation of medical images and their transmission since it provided a network protocol built on top of TCP/IP.

\subsection{Data format}

DICOM files can be viewed as containers of attributes, also called tags. The values of the pixels themselves are stored under the "Pixel Data" tag. Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.

Other useful information such as the patient name and ID is directly stored within the DICOM files. This approach aims at linking each image to a specific person and event in order not to mix them up. Each DICOM file can be seen as part of a bigger dataset.

\subsection{Processing images}

When manipulating DICOM files, multiple details must be taken into account.

\subsubsection{Order}
First of all, the name of the files within datasets is a 6 digit number, from 000000 to the number of images minus one. However, this order doesn’t match the real order of the images. In fact, the correct order is given by the "Instance Number" tags contained in the various files. Therefore, converted images must be sorted by instance number.

\subsubsection{Data manipulation}
Then, CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other. DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used. Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations. 
Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.

If the tags "Window Width" and "Window Center" (one always come with the other) are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
\begin{itemize}
	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
	\item Samples per pixel: the number of channels for each pixel

\end{itemize}

Examples: 
\begin{itemize}
\item 1 bit, 1 sample: black and white
\item 8 bits, 1 sample: grayscale
\item 8 bits, 3 samples: RGB
\item 16 bits, 1 sample: grayscale

\end{itemize}

Otherwise, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one. To do this, two step are required:

\begin{enumerate}
	\item \textbf{Apply the Hounsfield correction}\newline
	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air. The correction is calculated thanks to the following formula:
\begin{equation}
	HU = m * P + b
\end{equation}
	where~$m$ is the rescale slope,~$P$ the pixel value, ~$b$ the rescale intercept.

	\item \textbf{Apply a linear interpolation}\newline
	The result of the first operation then goes through a linear interpolation to get the right pixel values, based on the following conditions:

\begin{equation}
\textrm{if } (P \leq c - 0.5 - \frac{w-1}{2}) \textrm{, then }y = y_{min}
\end{equation}

\begin{equation}
\textrm{else if } (P > c - 0.5 + \frac{w-1}{2}) \textrm{, then }y = y_{max}
\end{equation}

\begin{equation}
\textrm{else } y = (\frac{P - (c - 0.5)}{w-1}) + 0.5) * (y_{max} - y_{min}) + y_{min}
\end{equation}

where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range, ~$y_{max}$ the maximal value of the output range.

\end{enumerate}

\subsubsection{Export}
Medical data can be represented over 16 bits, while exporting images to PNG require 8-bit images. To transpose 16 bits to 8, the following procedure can be applied:

\begin{enumerate}
	\item Find maximum value of the current image $Pixel_{min}$
	\item Find minimum value of the current image $Pixel_{max}$
	\item Pixel normalization: $Pixel = \frac{(Pixel - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
	\item Modify object type to 8 bits
	\item Export to PNG
\end{enumerate}

### 2019-10-21 Common/chapters/technical_info.tex
Julien and Johan: more details about DICOM and paragraph about RAW and MHD

@@
 \label{ch:technical_info}
-
-\section{NIfTI}
-
-\subsubsection{Origin of the format}
-The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
-The main problem of the latter was “the lack of adequate information about orientation in space, such that the stored data could not be unambiguously interpreted” .
-For instance, there was a real confusion to determine the left and right sides of brain images.
-Hence, the NIfTI file format was defined to obviate this major issue.
-
-\subsubsection{Organization of data}
-Unlike the ANALYZE format that used two files to store the meta-data and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
-This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
-The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.
-
-\subsubsection{Overview of the header structure}
-With the goal of preserving the compatibility between the “analyze” and the NIfTI format, both headers have the same size of 348 bytes.
-“Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
-Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}
 
@@
 \subsubsection{Data manipulation}
-Then, CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
+CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
 DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used.
@@
 
-If the tags "Window Width" and "Window Center" (one always come with the other) are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
+The procedure depends on the tags "Window Width" and "Window Center" (one always come with the other).
+These are used to represent a range of values corresponding to the pixel values in the data.
+For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.
+
+If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
 \begin{itemize}
@@
 
-Examples: 
+\noindent Examples: 
 \begin{itemize}
@@
 
-Otherwise, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one. To do this, two step are required:
+If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one. To do this, two step are required:
 
@@
 	\item \textbf{Apply the Hounsfield correction}\newline
-	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air. The correction is calculated thanks to the following formula:
+	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air.
+Provided that the rescale slope and the rescale intercept are included in the DICOM header, the correction is applied thanks to the following formula:
 \begin{equation}
@@
 
-	\item \textbf{Apply a linear interpolation}\newline
-	The result of the first operation then goes through a linear interpolation to get the right pixel values, based on the following conditions:
+	\item \textbf{Apply a linear transformation}\newline
+	The result of the first operation then goes through a linear transformation based on the following conditions:
 
-\begin{equation}
+\begin{equation}\label{eq:dicom1}
 \textrm{if } (P \leq c - 0.5 - \frac{w-1}{2}) \textrm{, then }y = y_{min}
@@
 
-\begin{equation}
+\begin{equation}\label{eq:dicom2}
 \textrm{else if } (P > c - 0.5 + \frac{w-1}{2}) \textrm{, then }y = y_{max}
@@
 
-\begin{equation}
+\begin{equation}\label{eq:dicom3}
 \textrm{else } y = (\frac{P - (c - 0.5)}{w-1}) + 0.5) * (y_{max} - y_{min}) + y_{min}
@@
 
-where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range, ~$y_{max}$ the maximal value of the output range.
+where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255).
+Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.
 
@@
 
-\subsubsection{Export}
-Medical data can be represented over 16 bits, while exporting images to PNG require 8-bit images. To transpose 16 bits to 8, the following procedure can be applied:
+\section{NIfTI}
+
+\subsubsection{Origin}
+The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
+The main problem of the latter was that it was lacking information about orientation in space.
+Therefore, the interpretation of stored data could be problematic and inconsistent.
+For instance, there was a real confusion to determine the left and right sides of brain images.
+Hence, the NIfTI file format was defined to overcome this major issue.
+
+\subsubsection{Data format}
+Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
+This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
+The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.
+
+\subsubsection{Overview of the header structure}
+With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+“Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
+Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}
+
+\section{RAW and MHD}
+
+Some datasets use a combination of RAW and MHD files.
+The latter contain metainformation about their corresponding RAW file(s) which contain the data.
+In most cases, each MHD file points to a unique RAW file whose name is the same as the MHD file name.
+A single RAW file can be used to represent three-dimensional data, i.e.
+the combination of multiple two-dimensional images.
+Libraries such as \mbox{SimpleITK} in Python allow to manipulate RAW images in an easy way.
+
+\section{Conversion}
+
+Medical data is often represented over 16 bits. However, exporting images to PNG require 8-bit images. To transpose a 16-bit image to an 8-bit one, the following procedure was applied:
 

### 2019-11-30 Common/chapters/data_processing.tex
Julien and Johan: basics of data processing

--- 2019-10-21 Common/chapters/technical_info.tex
@@
 
-\chapter{Technical information}
-\label{ch:technical_info}
+\chapter{Data processing}
+\label{ch:data_processing}
 
-\section{DICOM}
+\section{DICOM file format}
 
@@
 \subsubsection{Order}
-First of all, the name of the files within datasets is a 6 digit number, from 000000 to the number of images minus one.
+First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
 However, this order doesn’t match the real order of the images.
@@
 \subsubsection{Data manipulation}
+\label{sec:data_manipulation}
 CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
@@
 
-If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one. To do this, two step are required:
+\noindent If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one.
+To do this, two steps are required:
 
@@
 
-\section{NIfTI}
+\section{NIfTI file format}
 
@@
 
-\section{RAW and MHD}
+\section{RAW and MHD file formats}
 
@@
 
-Medical data is often represented over 16 bits. However, exporting images to PNG require 8-bit images. To transpose a 16-bit image to an 8-bit one, the following procedure was applied:
+Medical data is often represented over 16 bits.
+However, exporting images to PNG require 8-bit images.
+To transpose a 16-bit image to an 8-bit one, the procedure described by \ref{alg:16_to_8_bits_conversion} was applied.
 
-\begin{enumerate}
-	\item Find maximum value of the current image $Pixel_{min}$
-	\item Find minimum value of the current image $Pixel_{max}$
-	\item Pixel normalization: $Pixel = \frac{(Pixel - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
-	\item Modify object type to 8 bits
-	\item Export to PNG
-\end{enumerate}
+\begin{algorithm}
+    \caption{16 to 8 bits conversion}
+    \label{alg:16_to_8_bits_conversion}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{16\_to\_8\_bits\_conversion}{$pixel\_array$}
+        		\State{$Pixel_{min} \gets$ minimal pixel value in $pixel\_array$}
+        		\State{$Pixel_{max} \gets$ maximal pixel value in $pixel\_array$}
+        		\For{$pixel\_value$ in $pixel\_array$}
+        			\State $pixel\_value \gets \frac{(pixel\_value - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
+        		\EndFor
+        		\State Modify object type to 8 bits
+        		\State Export $pixel\_array$ to PNG
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{PROSTATEx: From DICOM to NumPy arrays}
+
+The PROSTATEx dataset comes with two CSV files for the training set.
+The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance.
+Multiple findings can be associated with the same patient (ProxID).
+The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding.
+Important labels are "ProxID" (patient ID), "fid" (finding ID, from 1 to $\infty$), "ClinSig" (clinical significance, TRUE or FALSE), "DCMSerNumber" (digit before the dash in the folder name containing DICOM files) and "ijk" (position of the lesion: slice number $k$ at coordinates $(i,j)$, $i,j,k \in [0,\infty]$).
+Both CSV files are complementary to each other.
+Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays.
+
+\begin{algorithm}
+    \caption{PROSTATEx preprocessing}
+    \label{alg:PROSTATEx_preprocessing}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{main}{$dataset\_folder, findings\_CSV, slices\_CSV, output\_folder$}
+        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
+        		\State $findings \gets read\_CSV(findings\_CSV)$ \Comment{ProstateX-Findings-Train.csv}
+        		\State $slices \gets read\_CSV(slices\_CSV)$ \Comment{ProstateX-Images-Train.csv}
+            \State $meta \gets merge(findings, slices)$\Comment{Both CSV files are complementary to each other.}\\
+            \For{$row$ in $meta$}
+            		\State $patient\_id \gets row["ProxID"]$
+            		\State $finding\_id \gets row["fid"]$
+            		\State $mri\_type\_number \gets row["DCMSerNumber"]$
+            		\State $clinical\_significance \gets row["ClinSig"]$
+            		\State $img\_i, img\_j, img\_k \gets row["ijk"]$
+            		\State $slice\_number \gets img\_k + 1$ \Comment{CSV indexing in $[0,\infty]$, DICOM in $[1,\infty]$}\\
+                \For{$visit\ in\ patient\_id$'s folder}
+                		\For{$mri\_type$ in $visit$}
+                			\If {$mri\_type$ starts with "$mri\_type\_number$-"}
+                				\For{$dicom\_file$ in $mri\_type$}
+                					\If {$slice\_number == dicom\_file.InstanceNumber$}
+                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:data_manipulation}}
+                						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
+                					\EndIf
+                				\EndFor
+                			\EndIf
+            			\EndFor
+            		\EndFor
+            \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Lung CT Challenge - From DICOM to NumPy arrays}
+
+Lung CT Challenge is composed of two different subdatasets: one is considered as a calibration set (10 patients) and the other as a test set (60 patients).
+Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a training set.
+
+Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images.
+In order to facilitate label managing, two CSV files were manually created: \textit{TestSet.csv} and \textit{CalibrationSet.csv}.
+Contrary to PROSTATEx, more than two labels were used for this dataset.
+Both "malignant" and "Primary lung cancer" were considered as positive, whereas "benign" and "Benign nodule" as negative.
+A third label called "Suspicious malignant nodule" appeared two times.
+Since the diagnosis was not clearly defined for those images, they were not treated and included in the training data in order to avoid any noise.
+Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various processing steps.
+
+\begin{algorithm}
+    \caption{Lung CT Challenge preprocessing}
+    \label{alg:LungCTChallenge_preprocessing}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{main}{$dataset\_folder, train\_CSV, test\_CSV, output\_folder$}
+        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
+        		\State $csv\_training \gets read\_CSV(train\_CSV)$ \Comment{CalibrationSet.csv}
+        		\State $csv\_test \gets read\_CSV(test\_CSV)$ \Comment{TestSet.csv}
+            \State $csv\_concatenated \gets concat(csv\_training, csv\_test)$\Comment{Both CSV files contain similar information about different patients.}\\
+            \For{$row$ in $csv\_concatenated$}
+            		\State $patient\_id \gets row["Scan Number"]$
+            		\State $slice\_number \gets row["Nodule Center Image"]$ \Comment{Value in $[1,\infty]$}
+            		\State $finding\_id \gets row["Nodule Number"]$
+            		\State $clinical\_significance \gets row["Diagnosis"]$\\
+            		\If{$clinical\_significance$ == $"malignant"$ or $"Primary\ lung\ cancer"$}
+            			\State{$clinical\_significance \gets True$}
+            		\ElsIf{$clinical\_significance$ == $"benign"$ or $"Benign\ nodule"$}
+            			\State{$clinical\_significance \gets False$}
+            		\ElsIf {$"clinical\_significance$ == $"Suspicious\ malignant\ nodule"$}
+            			\State{Continue}
+            		\EndIf\\
+                \For{$visit\ in\ patient\_id$'s folder}
+                		\For{$mri\_type$ in $visit$}
+                			\For{$dicom\_file$ in $mri\_type$}
+                				\If {$slice\_number == dicom\_file.InstanceNumber$}
+                					\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:data_manipulation}}
+                					\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
+                				\EndIf
+                			\EndFor
+            			\EndFor
+            		\EndFor
+            \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{NumPy arrays to PNG files}
+Exporting to NumPy arrays instead of image files directly has multiple advantages.
+First of all, converting medical files to NumPy arrays or PNG files takes more time than converting NumPy arrays to image files.
+Reason for that is that the former requires a lot of operation as well as pixel normalization, whereas the latter is a mere conversion of one format to the other.
+This makes data more reusable.
+Then, it facilitates the debugging process by separating the conversion from medical files to PNG files into two different steps.
+Finally, the same processing script can be used to convert NumPy arrays from any dataset to PNG images, which eases operations such as generating different training-validation splits, augmenting data, etc.
+
+\begin{algorithm}
+    \caption{Create dataset - NumPy arrays to PNG conversion, organizing data into training and validation data}
+    \label{alg:create_dataset}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{create\_pngs}{$dataset\_folder, output\_folder, split$}
+        		\State Create output directories: $"output\_folder/[Train|Val]/[0|1]"$\\
+        		\State $true\_nparrays\_dict \gets \{"patient\_id" : [file\_name\_1, ...]\}$
+        		\State $number\_of\_patients\_true \gets len(true\_nparrays\_dict)$
+        		\State $number\_of\_training\_patients\_true \gets \lfloor number\_of\_patients\_true * split \rfloor$
+        		\State $false\_nparrays\_dict \gets \{"patient\_id" : [file\_name\_1, ...]\}$\\
+        		\State $index \gets 0$
+        		\For{$patient\_id, file\_names$ in $true\_nparrays\_dict$}
+        			\If{$index < number\_of\_training\_patients\_true$} \Comment{Training set, True}
+        				\For{$file\_name$ in $file\_names$}
+        					\State $image\_array \gets load(file\_name)$
+        					\State $image \gets convertArrayToGrayscaleImage(image\_array)$
+        					\State Save image to $"output\_folder/Train/1"$
+        				\EndFor
+        			\Else \Comment{Validation set, True}
+        				\For{$file\_name$ in $file\_names$}
+        					\State $image\_array \gets load(file\_name)$
+        					\State $image \gets convertArrayToGrayscaleImage(image\_array)$
+        					\State Save image to $"output\_folder/Val/1"$
+        				\EndFor
+        			\EndIf
+        			\State{$index \gets index + 1$}
+        		\EndFor\\
+        		\State $number\_of\_training\_patients\_false \gets \lfloor number\_of\_patients\_false * split \rfloor$
+        		\State $false\_nparrays\_dict \gets \{"patient\_id" : [file\_name\_1, ...]\}$
+        		\State $number\_of\_patients\_false \gets len(false\_nparrays\_dict)$
+        		\State $false\_nparrays\_dict \gets \{"patient\_id" : [file\_name\_1, ...]\}$\\
+        		\State $index \gets 0$
+        		\For{$patient\_id, file\_names$ in $false\_nparrays\_dict$}
+        			\If{$index < number\_of\_training\_patients\_false$} \Comment{Training set, False}
+        				\For{$file\_name$ in $file\_names$}
+        					\State $image\_array \gets load(file\_name)$
+        					\State $image \gets convertArrayToGrayscaleImage(image\_array)$
+        					\State Save image to $"output\_folder/Train/0"$
+        				\EndFor
+        			\Else \Comment{Validation set, False}
+        				\For{$file\_name$ in $file\_names$}
+        					\State $image\_array \gets load(file\_name)$
+        					\State $image \gets convertArrayToGrayscaleImage(image\_array)$
+        					\State Save image to $"output\_folder/Val/0"$
+        				\EndFor
+        			\EndIf
+        			\State{$index \gets index + 1$}
+        		\EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+Algorithm \ref{alg:create_dataset} describes how data was split and organized in order for PyTorch to make use of it.
+To do so, they were split by class (True $\equiv$ 1, False $\equiv$ 0) and role (training or validation).
+Furthermore, data are split by patients and not by slices.
+In fact, multiple slices are usually assigned to each patient.
+Instead of considering each slice separately, which allows to divide a single patient's slices into the training and validation folders, they were treated as a whole.
+To sum up, an 80-20 split between the training and validation data (split argument set to $0.8$) will use 80\% of the patients as training data and 20\% of the patients as validation data, regardless of the number of slices.

### 2019-11-30 Common/chapters/data_processing.tex
Julien and Johan: create basic structure and chapters

@@
 However, exporting images to PNG require 8-bit images.
-To transpose a 16-bit image to an 8-bit one, the procedure described by \ref{alg:16_to_8_bits_conversion} was applied.
+To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
 
@@
 
-\section{PROSTATEx: From DICOM to NumPy arrays}
+\section{Processing flow}
+
+Schema Datasets -> xxx\_preprocessing.py -> create\_dataset.py -> Train/Val
+
+\subsection{PROSTATEx: From DICOM to NumPy arrays}
 
@@
 
-\section{Lung CT Challenge - From DICOM to NumPy arrays}
+\subsection{Lung CT Challenge - From DICOM to NumPy arrays}
 
@@
             \For{$row$ in $csv\_concatenated$}
-            		\State $patient\_id \gets row["Scan Number"]$
-            		\State $slice\_number \gets row["Nodule Center Image"]$ \Comment{Value in $[1,\infty]$}
-            		\State $finding\_id \gets row["Nodule Number"]$
+            		\State $patient\_id \gets row["Scan\ Number"]$
+            		\State $slice\_number \gets row["Nodule\ Center\ Image"]$ \Comment{Value in $[1,\infty]$}
+            		\State $finding\_id \gets row["Nodule\ Number"]$
             		\State $clinical\_significance \gets row["Diagnosis"]$\\
@@
 
-\section{NumPy arrays to PNG files}
+\subsection{NumPy arrays to PNG files}
 Exporting to NumPy arrays instead of image files directly has multiple advantages.

### 2019-11-30 Common/chapters/deep_learning.tex
Julien and Johan: create basic structure and chapters

% !TEX root = ../main.tex

\chapter{Deep learning}
\label{ch:deep_learning}

Deep learning.

\section{Introduction to deep learning}

\subsection{What is a neural network?}

\subsection{Historical background}

\section{Neural networks basics}

\subsection{Perceptrons}

\subsection{Multilayer perceptron}

\subsection{Forward propagation}

\subsection{Backpropagation}

\subsubsection{Loss function}

\subsubsection{Weight update}

\section{Training a neural network}

\subsection{Data}
Training set, validation set, test set, supervised learning, unsupervised learning,

\subsection{Weight initialization}

\subsection{Hyperparameters}
lr, batch size, epochs, optimizer, learning rate decay, ...

\section{Deep learning models}
xxx

\subsubsection{Convolutional Neural Networks}

\subsubsection{Recurrent Neural Networks}

\section{Transfer learning}

\section{Deep learning frameworks}

### 2019-11-30 Common/chapters/literature_review.tex
Julien and Johan: create basic structure and chapters

% !TEX root = ../main.tex

\chapter{Literature review}
\label{ch:literature_review}

Literature review.

### 2019-11-30 Common/chapters/medical.tex
Julien and Johan: create basic structure and chapters

% !TEX root = ../main.tex

\chapter{Medical information}
\label{ch:medical}

\section{Cancer}

\subsection{Origin}

\subsection{Chronological evolution}

\section{Medical imaging}

\subsection{Types}

\subsubsection{MRI}

\subsubsection{CT}

\subsubsection{Mammography}

### 2019-11-30 Common/chapters/deep_learning.tex
Johan and Julien: Beginning of deep learning chapter

@@
 
-Deep learning.
+\section{Introduction to deep learning}
+Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence.
+Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them.
+In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way accordingly.
+There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc.
+What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}), that are combined to form deep neural networks.\\
+Neural networks are inspired from the biological neural networks of the brain.
+These systems try to learn how to solve a problem based on the data they receive as input.
+Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, personal assistants, art creation, robotics, etc.
+The presence of deep learning techniques in all mentioned use cases clearly attests that this technology is a real breakthrough.
+Furthermore, since this field has recently gained interest   (see section \ref{historical_background}), multiple researches are still ongoing, which suggests that many exciting new applications will certainly be discovered.
 
-\section{Introduction to deep learning}
+\subsection{Historical background}
+\label{historical_background}
+As described in figure \ref{history} from article \cite{14}, the theoretical foundations of deep learning appeared long before the invention of the computers.
+From the first attempts to understand the human brain until today, a long way has been made to establish the basic components of modern neural networks.
+One could ask: why did deep learning take off recently if the theory has been around for a long time?\\
+First part of the answer is about the computational power.
+Indeed, deep learning algorithms need a lot of data to work properly and this requires powerful CPU/GPU that either didn't exist or were only available to a minority of people.
+\\
+One other main reason concerns the lack of data.
+Since deep learning algorithms "learn" from data, if not data is available, no learning is possible.
+Consequently, the era of Big Data has enhanced the deep learning possibilities.\\
+Finally, before the year 2012, the abilities of neural networks had not been publicly proven.
+This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluate their algorithms on several visual recognition tasks).
+In fact, the deep convolutional neural network called "AlexNet" achieved 16\% of classification error rate, whereas the previous best scores were around 25\%.
+This victory marked the beginning of a new craze for these types of algorithms.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/history.png}
+\caption{Milestones }
+\label{history}
+\end{figure}
 
 \subsection{What is a neural network?}
+\label{what_is_a_neural_network}
 
-\subsection{Historical background}
+From a descriptive point of view, neural networks can simply be seen as a non linear applications that associate an input to an output with respect to certain parameters.
+The input can be images, sounds or whatever features that is numerical.
+The output of a neural network depends on the problem it tries to solve.
+In computer vision, the most common types of outputs  are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
+The figure \ref{neural_network} globally illustrates the concept of a neural network: in a fully connected neural network,  all inputs are given to all neurons of the first layer.
+Each neuron sums the incoming inputs, passes the result through a non linear function and sends it to all neuron of the next layer along an edge.
+Each edge between two neurons has a certain weight (a numerical value) that weights the importance of the output of the previous neuron.\\
+From a mathematical point of view, a neural network can be defined as a non linear function $f$ that associates to an input $x$ an output $y$ with respect to parameters $\theta$.
+\begin{equation}
+y = f(x,\theta)
+\end{equation}
+The parameters $\theta$ are estimated from the training samples.
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/neural_network.png}
+\caption{Neural Network example}
+\label{neural_network}
+\end{figure}
 
@@
 \subsection{Perceptrons}
+\label{perceptron}
 

### 2019-12-01 Common/chapters/deep_learning.tex
Julien and Johan: make neural network with tikz

@@
 The parameters $\theta$ are estimated from the training samples.
+%\begin{figure}[!h]
+%\centering
+%\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/neural_network.png}
+%\caption{Neural Network example}
+%\label{neural_network}
+%\end{figure}
+
 \begin{figure}[!h]
 \centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/neural_network.png}
+	\begin{tikzpicture}[x=1.5cm, y=1.5cm, >=stealth]
+
+	\foreach \m/\l [count=\y] in {1,2,3,4}
+	  \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};
+
+	\foreach \m [count=\y] in {1,2}
+	  \node [every neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2-\y*1.25) {};
+
+	\foreach \m [count=\y] in {1}
+	  \node [every neuron/.try, neuron \m/.try ] (output-\m) at (4,0.15) {};
+
+	% input layer
+	\foreach \l [count=\i] in {1,2,3,4}
+	  \draw [<-] (input-\i) -- ++(-1,0)
+	    node [above, midway] {$I_\l$};
+
+	% hidden layer
+	\foreach \l [count=\i] in {1,2}
+	  \node [above] at (hidden-\i.north) {$H_\l$};
+
+	% output neurons
+	\foreach \l [count=\i] in {1}
+	  \draw [->] (output-\i) -- ++(1,0)
+	    node [above, midway] {$O_\l$};
+
+	% input -> hidden
+	\foreach \i in {1,...,4}
+	  \foreach \j in {1,...,2}
+	    \draw [->] (input-\i) -- (hidden-\j);
+
+	% hidden -> output
+	\foreach \i in {1,...,2}
+	  \foreach \j in {1}
+	    \draw [->] (hidden-\i) -- (output-\j);
+
+	% labels above layers
+	\foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
+	  \node [align=center, above] at (\x*2,2) {\l \\ layer};
+
+	\end{tikzpicture}
 \caption{Neural Network example}

### 2019-12-01 Common/chapters/data_processing.tex
Julien and Johan: add data augmentation and visualization

@@
 \subsubsection{Data manipulation}
-\label{sec:data_manipulation}
+\label{sec:dicom_data_manipulation}
 CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
@@
                 					\If {$slice\_number == dicom\_file.InstanceNumber$}
-                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:data_manipulation}}
+                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
                 						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
@@
                 				\If {$slice\_number == dicom\_file.InstanceNumber$}
-                					\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:data_manipulation}}
+                					\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
                 					\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
@@
 To sum up, an 80-20 split between the training and validation data (split argument set to $0.8$) will use 80\% of the patients as training data and 20\% of the patients as validation data, regardless of the number of slices.
+
+\subsection{Data augmentation}
+
+The amount of publicly available data is relatively limited.
+However, deep learning models require a lot of training data to be able to learn and generalize well.
+Therefore, augmenting data allows to create more training examples from the ones available.
+To achieve this, multiple techniques such as rotation, flipping, cropping and shifting can be used.
+Algorithm \ref{alg:augment} describes how data was augmented.
+Given an MR image, a large patch centered on the lesion was cropped first, which makes the rotation process easier.
+In fact, rotating the image without cropping it first also moves the region of interest.
+In this case, finding the exact coordinates becomes more complicated.
+Therefore, a first large patch was created and then rotated.
+Afterwards, a smaller patch around the region of interest was extracted because of the potential padding induced by the rotation.
+This guarantees the addition of the minimal amount of padding pixels around the image to plug the holes, which decreases the probability of adding artificial information to the image.
+
+\begin{algorithm}
+    \caption{Create augmented dataset - Augmentation method}
+    \label{alg:augment}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{augment}{$image, img\_i, img\_j, available\_combinations\_list$}
+        		%\Comment{image: Pillow image, img\_[i,j]: position of the lesion, available\_combinations\_list: tuples (degree, prob\_flipping)}\\
+        		\State{$index \gets$ random index in $[0, $len(available\_combinations\_list)$ - 1]$}
+        		\State{$degree, prob\_flipping = available\_combinations\_list[index]$}
+        		\State{Delete element $available\_combinations\_list[index]$} \Comment{Avoid duplication}\\
+        		\State{$temp\_image \gets$ $image$.crop(64x64, center=$(img\_i, img\_j)$)} \Comment{$1^{st}$ crop}
+        		\State{$temp\_image \gets\ temp\_image$.rotate(degree)} \Comment{Rotate $1^{st}$ crop}\\
+        		\State{$width, height \gets$ $temp\_image$.size}
+        		\State{$x\_center \gets \lfloor \frac{width}{2} \rfloor, y\_center \gets \lfloor \frac{height}{2} \rfloor$}\Comment{Find center pixel of $1^{st}$ crop}
+        		\State{$temp\_image \gets$ $image$.crop(32x32, center=$(x\_center, y\_center)$)}\Comment{$2^{nd}$ crop}\\
+        		\If{$prob\_flipping > 0.5$}
+        			\State{$temp\_image \gets temp\_image.horizontal\_flip()$}
+        		\EndIf\\
+        		\Return $temp\_image$
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\subsection{Data visualization and verification}
+
+Processing data manually increases the probability of making mistakes.
+For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developped.
+Their goal was to compare our visual representation of an image to the one obtained in professional pieces of software.
+Also, PNG conversion was meticulously tested by comparing the pixel values from the original NumPy arrays with the PNG pixel values.
+
+\subsubsection{DICOM}
+
+Visualizing DICOM files is pretty straightforward since each file represent a single two-dimensional image.
+However, the way this standard works require a lot of pixel transformation and normalization to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may cause errors.
+Our tools allows to display a single DICOM file as well as a sequence of files if the function is feeded with a directory.
+Users can then scroll through the Z-axis, displaying the next or previous slice.
+
+\textbf{SCREENSHOT}
+
+\subsubsection{NIfTI}
+
+As NIfTI files can be three or four-dimensional (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the corresponding slice with respect to another timestamp.
+When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
+Furthermore, the three-dimensionality implies that the volume is viewable under three different perspectives.
+For example, a three-dimensional brain volume can display it from the top of the head to the bottom, from one ear to the other and from the back of the head to the person's face.
+Therefore, the user can choose a specific axis to navigate through.
+If no axis is chosen, all three perspectives are shown one after the other.
+
+\textbf{SCREENSHOT}
+
+\subsubsection{RAW}
+
+Medical RAW files are three-dimensional.
+Scrolling with the mouse goes through slices over a specific axis.
+Like NIfTI, the user can profit from the three-dimensionality by navigating through slices under three different perspectives.
+
+\textbf{SCREENSHOT}

### 2019-12-02 Common/chapters/deep_learning.tex
Johan and Julien:  Deep learning chapter continuation

@@
 These systems try to learn how to solve a problem based on the data they receive as input.
-Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, personal assistants, art creation, robotics, etc.
+Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc.
 The presence of deep learning techniques in all mentioned use cases clearly attests that this technology is a real breakthrough.
@@
 Indeed, deep learning algorithms need a lot of data to work properly and this requires powerful CPU/GPU that either didn't exist or were only available to a minority of people.
-\\
 One other main reason concerns the lack of data.
 Since deep learning algorithms "learn" from data, if not data is available, no learning is possible.
-Consequently, the era of Big Data has enhanced the deep learning possibilities.\\
+Consequently, the era of Big Data has enhanced the deep learning possibilities.
+These two points are summarised: "The increase in model size overtime, due to the availability of faster CPUs, the advent of general purpose GPUs, fasternetwork, connectivity and better software infrastructure for distributed computing, is one of the most important trends in the history of deep learning.
+This trend is generally expected to continue well into the future"\cite{15}.\\
 Finally, before the year 2012, the abilities of neural networks had not been publicly proven.
@@
 \label{perceptron}
+Perceptrons are the main components of neural networks.
+They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts"\cite{13}.
+Today, they are called "artificial neurons" or simply "neurons".\\
+A perceptron $j$ is a function $f$ of the input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non linear activation function $\sigma$:
+\begin{equation}
+y_{j} = f_{j}(x) = \sigma((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
+\end{equation}
+
+Schematically, a perceptron can be represented as shown in figure \ref{perceptron_model}.
+Each input is multiplied with its corresponding weight and the sum of this result is then passed through a non linear function, called an "activation function".
+This activation function acts like a threshold that decides if the sum can be given to the next perceptron or not.
+There exist multiple activation functions (see \ref{activation_functions}).
+It is extremely important to use a non linear functions instead of a linear function.
+Indeed, in a neural network where perceptrons are assembled together, the output of a perceptron is given as input to the others.
+Consequently, if linear functions are used, a linear output would be given as input to another linear function.
+Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create big neural networks.
+
+\tikzset{basic/.style={draw,fill=white!20,text width=1em,text badly centered}}
+\tikzset{input/.style={basic,circle}}
+\tikzset{weights/.style={basic,rectangle}}
+\tikzset{functions/.style={basic,circle,fill=white!10}}
+
+\begin{figure}[!h]
+\centering
+	\begin{tikzpicture}
+
+	\node[functions] (center) {};
+        \node[below of=center,font=\scriptsize,text width=4em] {Activation function};
+        \draw[thick] (0.5em,0.5em) -- (0,0.5em) -- (0,-0.5em) -- (-0.5em,-0.5em);
+        \draw (0em,0.75em) -- (0em,-0.75em);
+        \draw (0.75em,0em) -- (-0.75em,0em);
+        \node[right of=center] (right) {};
+            \path[draw,->] (center) -- (right);
+        \node[functions,left=5em] (left) {$\sum$};
+            \path[draw,->] (left) -- (center);
+        \node[weights,left=15em] (2) {$w_2$} -- (2) node[input,left of=2] (l2) {$x_2$};
+            \path[draw,->] (l2) -- (2);
+            \path[draw,->] (2) -- (left);
+        \node[below of=2] (dots) {$\vdots$} -- (dots) node[left of=dots] (ldots) {$\vdots$};
+        \node[weights,below of=dots] (n) {$w_n$} -- (n) node[input,left of=n] (ln) {$x_n$};
+            \path[draw,->] (ln) -- (n);
+            \path[draw,->] (n) -- (left);
+        \node[weights,above of=2] (1) {$w_1$} -- (1) node[input,left of=1] (l1) {$x_1$};
+            \path[draw,->] (l1) -- (1);
+            \path[draw,->] (1) -- (left);
+        \node[weights,above of=1] (0) {$w_0$} -- (0) node[input,left of=0] (l0) {$1$};
+            \path[draw,->] (l0) -- (0);
+            \path[draw,->] (0) -- (left);
+        \node[below of=ln,font=\scriptsize] {inputs};
+        \node[below of=n,font=\scriptsize] {weights};
+
+	\end{tikzpicture}
+\caption{The perceptron model}
+\label{perceptron_model}
+\end{figure}
 
@@
 
+\subsection{Notation}
+In order to respect a consistent mathematical notation to describe neural networks, this work will use the notation of Andrew Y.
+Ng, one of the pioneers of deep learning, that can be found in \cite{16}.\\
+
+\noindent \textbf{General comment}
+\begin{itemize}
+\item Superscript (i) will denote the $i^{th}$ training example.
+\item Superscript [l] will denote the $l^{th}$ layer of the neural network.
+\end{itemize}
+
+\noindent \textbf{Sizes}
+\begin{itemize}
+\item $m$: number of examples in the dataset.
+\item $n_{x}$: input size.
+\item $n_{y}$: output size (or number of classes).
+\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer. 
+\item $L$: number of layers in the network.
+\end{itemize}
+
+\noindent \textbf{Neural networks components}
+\begin{itemize}
+\item $X\in \R$ is the input matrix matrix of a neural network.
+\item $x^{(i)} \in \R^{n_{x}}$ is the $i^{th}$ example (sample) represented as a column vector.
+\item $Y \in \R^{n_{y} \times m}$ is the label matrix.
+\item $y^{(i) \in \R^{n_{y}}}$ is the output label for the $i^{th}$ example.
+\item $W^{[l]} \in \R$ \textsuperscript{\# of neurons in the next layer = j  x \# of neurons in the previous layer = k} is the weight matrix at layer $[l]$.
+Unformally, a weight matrix has this form:\\\\
+
+\item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
+\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the network.
+\item $g^{[l]}(x)$ is the $l^{th}$ activation function.  
+\item $z^{[l]} = W_{x}x^{(i)} + b^{[l]}$ denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function.\\
+\end{itemize}
+
+\noindent \textbf{Forward propagation equations}
+\begin{itemize}
+\item $a = g^{[l]}(W_{x}x^{(i)} + b^{[l]}) = g^{[l]}(z^{[l]})$ where $g^{[l]}$ denotes the $l^{th}$ layer activation function.
+\item $a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) $ is the general activation formula at $l^{th}$ layer, 
+\end{itemize}
+
+\subsection{Activation functions}
+\label{activation_functions}
+Activation functions are used to bring non-linearity in neural networks. There exist
+
 \subsection{Forward propagation}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/forward_propagation.png}
+\caption{Forward propagation }
+\label{forward_propagation}
+\end{figure}
 

### 2019-12-02 Common/chapters/data_processing.tex
Julien and Johan: add screenshots for visualization tool and remake forward propagation screenshot

@@
 
-\textbf{SCREENSHOT}
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_lung_dcm.png}
+\caption{Visualization of a folder of DICOM files}
+\label{fig:visualize_lung_dcm}
+\end{figure}
 
@@
 
-\textbf{SCREENSHOT}
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_brain_nii.png}
+\caption{Four-dimensional NIfTI visualization}
+\label{fig:visualize_brain_nii}
+\end{figure}
 
@@
 
-\textbf{SCREENSHOT}
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_liver_raw.png}
+\caption{Three-dimensional RAW visualization}
+\label{fig:visualize_liver_raw}
+\end{figure}

### 2019-12-02 Common/chapters/medical.tex
Julien and Johan: Beginning of medical part

@@
 
-\subsection{Origin}
+\subsection{Basics}
+An accumulation of cells forming a mass is called a tumor.
+These tumors are detectable thanks to medical imaging (see section \ref{sec:medical_imaging}) and other symptoms.
+However, not every tumor is as dangerous as the other, as it can be benign (does not contain cancerous cells) or malignant (contains cancerous cells).
 
-\subsection{Chronological evolution}
+The term cancer refers to different phenomena which involve mutation, abnormal multiplication and spreading of cells.
+As stated by Hanahan et al.
+in "The Hallmarks of Cancer" \cite{19} and "The Hallmarks of Cancer: The Next Generation" \cite{20}, every malignant tumor acquires six different capabilities during its evolution:
+\begin{itemize}
+	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells don't wait for the body's approval to grow and proliferate, contrary to normal cells.
+They become the only responsible for their multiplication.
+	\item \textbf{"Evading growth suppressors"}\\
+The body sends signals to contain cell growth within a tissue. Cancerous cells are insensitive to these. 
+	\item \textbf{"Activating invasion and metastasis"}\\
+Metastases are cells whose role is to propagate to other parts of the body in order to colonize and create new tumors. 
+	\item \textbf{"Enabling replicative immortality"}\\
+Healthy cells replication is limited to a certain amount, which is not the case for cancerous cells. 
+	\item \textbf{"Inducing angiogenesis"}
+\\Angiogenesis is the process of creating new blood vessels. Tumors have an influence on angiogenesis around them, since they need vascularization to continue growing. 
+	\item \textbf{"Resisting cell death"}
+\\Apoptosis is the programmed death of cells, which is part of the continuous regeneration of every cell within a body. Cancerous cells survive this programmed death.
+
+\end{itemize}
+
+\subsection{Seriousness}
+
+Most cancers can be staged thanks to the TNM system.
+The T corresponds to the tumor size and its location; the N corresponds to whether or not the tumor has spread to draining lymph nodes; the M corresponds to the presence or absence of metastases in other parts of the body \cite{21}.
+These pieces of information are used to classify cancer between four (I to IV) or sometimes five (0 to IV) different stages, reflecting the progression and the seriousness of the illness \cite{22}.
+The earlier they are detected, the higher the chances of recovering are.
+
+%\subsection{Chronological evolution}
 
 \section{Medical imaging}
+\label{sec:medical_imaging}
 

### 2019-12-02 Common/chapters/deep_learning.tex
Johan and Julien: deep learning chapter continued

@@
 In computer vision, the most common types of outputs  are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
-The figure \ref{neural_network} globally illustrates the concept of a neural network: in a fully connected neural network,  all inputs are given to all neurons of the first layer.
-Each neuron sums the incoming inputs, passes the result through a non linear function and sends it to all neuron of the next layer along an edge.
-Each edge between two neurons has a certain weight (a numerical value) that weights the importance of the output of the previous neuron.\\
 From a mathematical point of view, a neural network can be defined as a non linear function $f$ that associates to an input $x$ an output $y$ with respect to parameters $\theta$.
@@
 %\end{figure}
+
+\section{Neural networks basics}
+
+\subsection{Notation}
+In order to respect a consistent mathematical notation to describe neural networks, this work will use the notation of Andrew Y.
+Ng, one of the pioneers of deep learning, that can be found in \cite{16}.\\
+
+\noindent \textbf{General comment}
+\begin{itemize}
+\item Superscript (i) will denote the $i^{th}$ training example.
+\item Superscript [l] will denote the $l^{th}$ layer of the neural network.
+\end{itemize}
+
+\noindent \textbf{Sizes}
+\begin{itemize}
+\item $m$: number of examples in the dataset.
+\item $n_{x}$: input size.
+\item $n_{y}$: output size (or number of classes).
+\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer. 
+\item $L$: number of layers in the network.
+\end{itemize}
+
+\noindent \textbf{Neural networks components}
+\begin{itemize}
+\item $X\in \R$ is the input matrix matrix of a neural network.
+\item $x^{(i)} \in \R^{n_{x}}$ is the $i^{th}$ example (sample) represented as a column vector.
+\item $Y \in \R^{n_{y} \times m}$ is the label matrix.
+\item $y^{(i) \in \R^{n_{y}}}$ is the output label for the $i^{th}$ example.
+\item $W^{[l]} \in \R$ \textsuperscript{\# of neurons in the next layer = j  x \# of neurons in the previous layer = k} is the weight matrix at layer $[l]$.
+\item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
+\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the network.
+\item $g^{[l]}(x)$ is the $l^{th}$ activation function.  
+\item $z^{[l]} = W_{x}x^{(i)} + b^{[l]}$ denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function.\\
+\end{itemize}
+
+\noindent \textbf{Forward propagation equations}
+\begin{itemize}
+\item $a = g^{[l]}(W_{x}x^{(i)} + b^{[l]}) = g^{[l]}(z^{[l]})$ where $g^{[l]}$ denotes the $l^{th}$ layer activation function.
+\item $a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) $ is the general activation formula at $l^{th}$ layer.
+\item $J(x, W, b, y) or J(\hat{y}, y)$ denote the cost function.
+\end{itemize}
+
+\subsection{Perceptrons}
+\label{perceptron}
+Perceptrons are the main components of neural networks.
+They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts"\cite{13}.
+Today, they are called "artificial neurons" or simply "neurons".\\
+A perceptron $j$ is a function $f$ of the input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non linear activation function $g$:
+\begin{equation}
+y_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
+\end{equation}
+
+Schematically, a perceptron can be represented as shown in figure \ref{perceptron_model}.
+Each input is multiplied with its corresponding weight and the sum of this result is then passed through a non linear function, called an "activation function".
+This activation function acts like a threshold that decides the proportion of the result that has to be given further.
+There exist multiple activation functions (see \ref{activation_functions}).
+It is extremely important to use a non linear functions instead of a linear function.
+Indeed, in a neural network where perceptrons are assembled together, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
+Consequently, if only linear functions were used, linear outputs would be given as inputs to other linear functions.
+Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create neural networks of multiple layers.
+
+\tikzset{basic/.style={draw,fill=white!20,text width=1em,text badly centered}}
+\tikzset{input/.style={basic,circle}}
+\tikzset{weights/.style={basic,rectangle}}
+\tikzset{functions/.style={basic,circle,fill=white!10}}
+
+\begin{figure}[!h]
+\centering
+	\begin{tikzpicture}
+
+	\node[functions] (center) {};
+        \node[below of=center,font=\scriptsize,text width=4em] {Activation function};
+        \draw[thick] (0.5em,0.5em) -- (0,0.5em) -- (0,-0.5em) -- (-0.5em,-0.5em);
+        \draw (0em,0.75em) -- (0em,-0.75em);
+        \draw (0.75em,0em) -- (-0.75em,0em);
+        \node[right of=center] (right) {};
+            \path[draw,->] (center) -- (right);
+        \node[functions,left=5em] (left) {$\sum$};
+            \path[draw,->] (left) -- (center);
+        \node[weights,left=15em] (2) {$w_2$} -- (2) node[input,left of=2] (l2) {$x_2$};
+            \path[draw,->] (l2) -- (2);
+            \path[draw,->] (2) -- (left);
+        \node[below of=2] (dots) {$\vdots$} -- (dots) node[left of=dots] (ldots) {$\vdots$};
+        \node[weights,below of=dots] (n) {$w_n$} -- (n) node[input,left of=n] (ln) {$x_n$};
+            \path[draw,->] (ln) -- (n);
+            \path[draw,->] (n) -- (left);
+        \node[weights,above of=2] (1) {$w_1$} -- (1) node[input,left of=1] (l1) {$x_1$};
+            \path[draw,->] (l1) -- (1);
+            \path[draw,->] (1) -- (left);
+        \node[weights,above of=1] (0) {$w_0$} -- (0) node[input,left of=0] (l0) {$1$};
+            \path[draw,->] (l0) -- (0);
+            \path[draw,->] (0) -- (left);
+        \node[below of=ln,font=\scriptsize] {inputs};
+        \node[below of=n,font=\scriptsize] {weights};
+
+	\end{tikzpicture}
+\caption{The perceptron model}
+\label{perceptron_model}
+\end{figure}
+
+\subsection{Activation functions}
+\label{activation_functions}
+Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to be pass the sum through an activation function that will decide the proportion of the result to send further.
+Activation functions have to be non-linear in order to approximate extremely complex functions.
+In fact, neural networks are considered as universal approximators.
+Article \cite{17} claims that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense."According to \cite{18} the most commonly used activation functions are:\\\\
+ \noindent \textbf{Sigmoid function}\\
+ The sigmoid function is defined as:
+ \begin{equation}
+  g(x) = \frac{1}{1+e^{-x}}
+ \end{equation}
+ Its derivative is:
+ \begin{equation}
+ g'(x) = g(x)(1-g(x))
+ \end{equation}
+
+\begin{figure}[h!]
+  \begin{center}
+    \begin{tikzpicture}
+      \begin{axis}[
+      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
+          xlabel={$x$}, 
+          ylabel={$y$},
+          xlabel=X Axis, 
+          ylabel=Y Axis,
+          samples=41, 
+          grid, 
+          thick,
+          domain=-10:10,
+		legend style={at={(0,1)},anchor=north west}
+        ]
+        \addplot+[no marks] {1/(1+exp(-x))};
+        \addlegendentry{$g(x)$}
+        \addplot+[no marks] {(1/(1+exp(-x))) * (1-(1/(1+exp(-x))))};
+        \addlegendentry{$g(x)(1-g(x))$}
+      \end{axis}
+    \end{tikzpicture}
+    \caption{The sigmoid function and its derivative}
+  \end{center}
+\end{figure}
+
+ \noindent \textbf{Tanh function}\\
+ \begin{equation}
+ g(x) = tanh(x) = \frac{1-e^{-2x}}{1+e^{-2x}}
+ \end{equation}
+
+ Its derivative is:
+ \begin{equation}
+ g'(x) = tanh'(x) = 1-tanh^{2}(x)
+ \end{equation}
+
+  \begin{figure}[h!]
+  \begin{center}
+    \begin{tikzpicture}
+      \begin{axis}[
+      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
+          xlabel={$x$}, 
+          ylabel={$y$},
+          xlabel=X Axis, 
+          ylabel=Y Axis,
+          samples=41, 
+          grid, 
+          thick,
+          domain=-10:10,
+		legend style={at={(0,1)},anchor=north west}
+        ]
+        \addplot+[no marks] {(1-exp(-2*x))/(1+exp(-2*x))};
+        \addlegendentry{$tanh(x)$}
+        \addplot+[no marks] {1-((1-exp(-2*x))/(1+exp(-2*x)))^2};
+        \addlegendentry{$1-tanh^{2}(x)$}
+      \end{axis}
+    \end{tikzpicture}
+    \caption{The tanh function and its derivative}
+  \end{center}
+\end{figure}
+
+ \noindent \textbf{ReLU function}\\
+ \begin{equation}
+ g(x) = ReLU(x) = 
+ \begin{cases}
+    1 & \text{if } x\geq 0\\
+    0              & \text{otherwise}
+\end{cases}
+ \end{equation}
+
+ \begin{figure}[h!]
+  \begin{center}
+    \begin{tikzpicture}[
+    declare function={
+    func(\x)= (\x<=0) * (0);
+  }
+  ]
+      \begin{axis}[
+      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
+          xlabel={$x$}, 
+          ylabel={$y$},
+          xlabel=X Axis, 
+          ylabel=Y Axis,
+          samples=41, 
+          grid, 
+          thick,
+          domain=-10:10,
+		legend style={at={(0,1)},anchor=north west}
+        ]
+        \addplot+[no marks] {(x>=0)*x};
+        \addlegendentry{$ReLU(x)$}
+        \addplot[domain=-20:0,red] {0};
+		\addplot[domain=0:20, red] {1};
+        \addlegendentry{ReLU'(x)}
+      \end{axis}
+    \end{tikzpicture}
+    \caption{The ReLU function and its derivative(not defined when x=0)}
+  \end{center}
+\end{figure}
+
+ \noindent \textbf{ELU function}\\
+ \begin{equation}
+ g(x) = 
+ \begin{cases}
+    x & \text{if } x\geq 0\\
+    e^{x}-1 & \text{otherwise }
+\end{cases}
+ \end{equation}
+
+ Its derivative is:
+ \begin{equation}
+g'(x) = 
+ \begin{cases}
+    1 & \text{if } x\geq 0\\
+    e^{x} & \text{otherwise }
+\end{cases}
+ \end{equation}
+
+  \begin{figure}[h!]
+  \begin{center}
+    \begin{tikzpicture}[
+      \begin{axis}[
+      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
+          xlabel={$x$}, 
+          ylabel={$y$},
+          xlabel=X Axis, 
+          ylabel=Y Axis,
+          samples=41, 
+          grid, 
+          thick,
+          domain=-10:10,
+		legend style={at={(0,1)},anchor=north west}
+        ]
+        \addplot+[no marks] {func(x)};
+        \addlegendentry{$ELU(x)$}
+        \addplot+[no marks] {6};
+        \addlegendentry{$ELU'(x)$}
+      \end{axis}
+    \end{tikzpicture}
+    \caption{The ELU function and its derivative}
+  \end{center}
+\end{figure}
+
+\subsection{Multilayer perceptrons}
+\label{multilayer_perceptron}
+The figure \ref{multilayer_perceptrons_figure} globally illustrates the concept of a multilayer perceptrons: all inputs are given to all neurons of the first layer.
+Each neuron sums the incoming inputs, passes the result through a non linear function and sends it to all neuron of the next layer along an edge.
+Each edge between two neurons (i.e perceptrons) has a certain weight (a numerical value) that weights the importance of the output of the previous neuron.\\
 
@@
 \caption{Neural Network example}
-\label{neural_network}
+\label{multilayer_perceptrons_figure}
 \end{figure}
 
-\section{Neural networks basics}
-
-\subsection{Perceptrons}
-\label{perceptron}
-Perceptrons are the main components of neural networks.
-They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts"\cite{13}.
-Today, they are called "artificial neurons" or simply "neurons".\\
-A perceptron $j$ is a function $f$ of the input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non linear activation function $\sigma$:
-\begin{equation}
-y_{j} = f_{j}(x) = \sigma((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
-\end{equation}
-
-Schematically, a perceptron can be represented as shown in figure \ref{perceptron_model}.
-Each input is multiplied with its corresponding weight and the sum of this result is then passed through a non linear function, called an "activation function".
-This activation function acts like a threshold that decides if the sum can be given to the next perceptron or not.
-There exist multiple activation functions (see \ref{activation_functions}).
-It is extremely important to use a non linear functions instead of a linear function.
-Indeed, in a neural network where perceptrons are assembled together, the output of a perceptron is given as input to the others.
-Consequently, if linear functions are used, a linear output would be given as input to another linear function.
-Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create big neural networks.
-
-\tikzset{basic/.style={draw,fill=white!20,text width=1em,text badly centered}}
-\tikzset{input/.style={basic,circle}}
-\tikzset{weights/.style={basic,rectangle}}
-\tikzset{functions/.style={basic,circle,fill=white!10}}
-
-\begin{figure}[!h]
-\centering
-	\begin{tikzpicture}
-
-	\node[functions] (center) {};
-        \node[below of=center,font=\scriptsize,text width=4em] {Activation function};
-        \draw[thick] (0.5em,0.5em) -- (0,0.5em) -- (0,-0.5em) -- (-0.5em,-0.5em);
-        \draw (0em,0.75em) -- (0em,-0.75em);
-        \draw (0.75em,0em) -- (-0.75em,0em);
-        \node[right of=center] (right) {};
-            \path[draw,->] (center) -- (right);
-        \node[functions,left=5em] (left) {$\sum$};
-            \path[draw,->] (left) -- (center);
-        \node[weights,left=15em] (2) {$w_2$} -- (2) node[input,left of=2] (l2) {$x_2$};
-            \path[draw,->] (l2) -- (2);
-            \path[draw,->] (2) -- (left);
-        \node[below of=2] (dots) {$\vdots$} -- (dots) node[left of=dots] (ldots) {$\vdots$};
-        \node[weights,below of=dots] (n) {$w_n$} -- (n) node[input,left of=n] (ln) {$x_n$};
-            \path[draw,->] (ln) -- (n);
-            \path[draw,->] (n) -- (left);
-        \node[weights,above of=2] (1) {$w_1$} -- (1) node[input,left of=1] (l1) {$x_1$};
-            \path[draw,->] (l1) -- (1);
-            \path[draw,->] (1) -- (left);
-        \node[weights,above of=1] (0) {$w_0$} -- (0) node[input,left of=0] (l0) {$1$};
-            \path[draw,->] (l0) -- (0);
-            \path[draw,->] (0) -- (left);
-        \node[below of=ln,font=\scriptsize] {inputs};
-        \node[below of=n,font=\scriptsize] {weights};
-
-	\end{tikzpicture}
-\caption{The perceptron model}
-\label{perceptron_model}
-\end{figure}
-
-\subsection{Multilayer perceptron}
-
-\subsection{Notation}
-In order to respect a consistent mathematical notation to describe neural networks, this work will use the notation of Andrew Y.
-Ng, one of the pioneers of deep learning, that can be found in \cite{16}.\\
-
-\noindent \textbf{General comment}
-\begin{itemize}
-\item Superscript (i) will denote the $i^{th}$ training example.
-\item Superscript [l] will denote the $l^{th}$ layer of the neural network.
-\end{itemize}
-
-\noindent \textbf{Sizes}
-\begin{itemize}
-\item $m$: number of examples in the dataset.
-\item $n_{x}$: input size.
-\item $n_{y}$: output size (or number of classes).
-\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer. 
-\item $L$: number of layers in the network.
-\end{itemize}
-
-\noindent \textbf{Neural networks components}
-\begin{itemize}
-\item $X\in \R$ is the input matrix matrix of a neural network.
-\item $x^{(i)} \in \R^{n_{x}}$ is the $i^{th}$ example (sample) represented as a column vector.
-\item $Y \in \R^{n_{y} \times m}$ is the label matrix.
-\item $y^{(i) \in \R^{n_{y}}}$ is the output label for the $i^{th}$ example.
-\item $W^{[l]} \in \R$ \textsuperscript{\# of neurons in the next layer = j  x \# of neurons in the previous layer = k} is the weight matrix at layer $[l]$.
-Unformally, a weight matrix has this form:\\\\
-
-\item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
-\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the network.
-\item $g^{[l]}(x)$ is the $l^{th}$ activation function.  
-\item $z^{[l]} = W_{x}x^{(i)} + b^{[l]}$ denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function.\\
-\end{itemize}
-
-\noindent \textbf{Forward propagation equations}
-\begin{itemize}
-\item $a = g^{[l]}(W_{x}x^{(i)} + b^{[l]}) = g^{[l]}(z^{[l]})$ where $g^{[l]}$ denotes the $l^{th}$ layer activation function.
-\item $a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) $ is the general activation formula at $l^{th}$ layer, 
-\end{itemize}
-
-\subsection{Activation functions}
-\label{activation_functions}
-Activation functions are used to bring non-linearity in neural networks. There exist
+A multilayer perceptron is an artifical neural network type.
+Article \cite{22} defines the multilayer perceptrons in the following way: "Multilayer perceptron are feedforward networks with one or more layers of units between the
+input and output layers. The output units represent a hyperplane in the space
+of the input patterns."\\
+A multilayer perceptrons is composed of multiple perceptrons that are arranged into multiple layers. When the input layer receives the signal, transmits it to the hidden layer
 

### 2019-12-02 Common/chapters/deep_learning.tex
Auto stash before merge of "master" and "origin/master"
Johan and Julien: correction of little mistakes

@@
     \begin{tikzpicture}[
+    declare function={
+    func(\x)= (\x>=0) * (x) +and (\x<0) * 4;
+  }
+    ]
       \begin{axis}[
@@
         ]
-        \addplot+[no marks] {func(x)};
+        \addplot+[no marks] {5};
         \addlegendentry{$ELU(x)$}
@@
 A multilayer perceptron is an artifical neural network type.
-Article \cite{22} defines the multilayer perceptrons in the following way: "Multilayer perceptron are feedforward networks with one or more layers of units between the
+Article \cite{23} defines the multilayer perceptrons in the following way: "Multilayer perceptron are feedforward networks with one or more layers of units between the
 input and output layers. The output units represent a hyperplane in the space

### 2019-12-05 Common/chapters/deep_learning.tex
Johan and Julien: Activation function subsection finished
Johan and Julien: multilayer perceptrons chapter finished + beginning of chapter "training a neural network"

@@
 
+\subsection{Supervised learning and Unsupervised learning}
+In machine learning, there mainly exists two different kinds of task.
+The first one is "supervised learning".
+It includes all learning algorithms, whose training samples contains labels in order to find the optimal mapping between the input and the output.
+The second one is "unsupervised learning".
+In contrast to supervised algorithms, the latter use non labeled data and its main goal is to infer the natural structure  present in the data.
+Since the model presented in this work belongs to the "supervised learning" category, all notions explained below refer to this kind of algorithms.
+
 \section{Neural networks basics}
@@
 \begin{equation}
+\label{perceptron_equation}
 y_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
@@
 \label{activation_functions}
-Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to be pass the sum through an activation function that will decide the proportion of the result to send further.
+Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result to send further.
 Activation functions have to be non-linear in order to approximate extremely complex functions.
@@
  \end{equation}
-
  Its derivative is:
@@
 		\addplot[domain=0:20, red] {1};
-        \addlegendentry{ReLU'(x)}
+        \addlegendentry{$ReLU'(x)$}
       \end{axis}
@@
         ]
-        \addplot+[no marks] {5};
+        \addplot[domain=0:20, blue] {x};
         \addlegendentry{$ELU(x)$}
-        \addplot+[no marks] {6};
+        \addplot[domain=0:20,red] {1};
         \addlegendentry{$ELU'(x)$}
+
+        \addplot[domain=-20:0,red] {exp(x)};
+        \addplot[domain=-20:0,blue] {exp(x)-1};
+
       \end{axis}
@@
 \label{multilayer_perceptron}
-The figure \ref{multilayer_perceptrons_figure} globally illustrates the concept of a multilayer perceptrons: all inputs are given to all neurons of the first layer.
-Each neuron sums the incoming inputs, passes the result through a non linear function and sends it to all neuron of the next layer along an edge.
-Each edge between two neurons (i.e perceptrons) has a certain weight (a numerical value) that weights the importance of the output of the previous neuron.\\
+A multilayer perceptron is an artifical neural network type. Article \cite{23} defines the multilayer perceptrons as "feedforward networks with one or more layers of units between the
+input and output layers. The output units represent a hyperplane in the space
+of the input patterns."\\
+A multilayer perceptrons is composed of $L$ layers each of them having $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
+\begin{itemize}
+\item The input layer: it is the neural network entry point for the data.
+Generally, the data is provided to the network in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $n_{y} \times batch\_size$.
+The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is their dimension.
+Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector.
+Finally, $y^{(i)}$ is the output label for the $i^{th}$ example.
+For instance, under the assumption that the number of samples is 100 and that the batch size is 32, 4 batches in total will be given as input to the network.
+\item The hidden layer(s): hidden layers stand for all layers that are between the input layer and the output layer.
+For each of them are associated weights and biases (W, b), respectively denoted by $W^{[l]} \in \R $ and $b^{[l]}$, where $W_{ij}^{l}$ correspond to the weights associated with the connection between perceptron $j$ in layer $l$ and perceptron $i$ in layer $l+1$.
+By analogy, $b_{i}^{[l]}$ is the bias associated with perceptron $i$ in layer $l+1$.
+The weights and the biases of the network are the parameters that have to be optimized in order to have the best mapping between the inputs and the output of the network (see section \ref{training_a_neural_network} ).
+\item The output layer: it is the last layer of the neural network.
+Its role is essential since it produces the prediction of the network for a given input.
+The prediction of a neural network is given by $\hat{y} \in R^{n_{y}}$ with $n_{y}$ beeing the number of different labels.
+In a classification task, whose goal is to assign to each input a specific class, the $\hat{y}$ could be the probability of belonging to each class.
+In that case, the \textit{softmax} activation function would be used.
+\end{itemize}
+The figure \ref{multilayer_perceptrons_figure} globally illustrates the concept of a multilayer perceptrons.
+In this example, the total number of layers $L$ is equal to 3, the input size $n_{x}$ is equal to 4 and the number of units in in each layer is $n_{h}^{1}= 4$, $n_{h}^{2}=2$, $n_{h}^{3}=1$.
+The network has the weights  $W^{(1)} \in \R^{2x3}$, $W^{(2)} \in \R^{1x2}$  and the biases  $b^{[1]}=2$, $b^{[2]}=1$.
 
@@
 	\end{tikzpicture}
-\caption{Neural Network example}
+\caption{Multilayer perceptrons}
 \label{multilayer_perceptrons_figure}
@@
 
-A multilayer perceptron is an artifical neural network type.
-Article \cite{23} defines the multilayer perceptrons in the following way: "Multilayer perceptron are feedforward networks with one or more layers of units between the
-input and output layers. The output units represent a hyperplane in the space
-of the input patterns."\\
-A multilayer perceptrons is composed of multiple perceptrons that are arranged into multiple layers. When the input layer receives the signal, transmits it to the hidden layer
+\section{Training a neural network}
+\label{training_a_neural_network}
+Training a neural network is composed of multiple steps.
+The first one is the "forward propagation step".
+This step consists in giving to the untrained neural network examples that need to be classified (or segmented, depending on the task) and to transmit them through all layers of the network.
+Passing all batches as input the network is called an "epoch".
+At the end of each batch the loss is computed using a "loss function".
+The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth.
+Then, the weights and biases of the networks are updated during a process called "back propagation" in order find the global minima of the loss function.
+This process is repeated a for a defined number of epochs.
 
@@
 
+\subsubsection{Loss computation}
+
 \subsection{Backpropagation}
 
-\subsubsection{Loss function}
-
 \subsubsection{Weight update}
-
-\section{Training a neural network}
 
@@
 \subsection{Hyperparameters}
+
 lr, batch size, epochs, optimizer, learning rate decay, ...
+
+From: https://stats.stackexchange.com/questions/153531/what-is-batch-size-in-neural-network
+Batch size: 
+Advantages of using a batch size < number of all samples:
+It requires less memory.
+Since you train the network using fewer samples, the overall training procedure requires less memory.
+That's especially important if you are not able to fit the whole dataset in your machine's memory.
+
+Typically networks train faster with mini-batches.
+That's because we update the weights after each propagation.
+In our example we've propagated 11 batches (10 of them had 100 samples and 1 had 50 samples) and after each of them we've updated our network's parameters.
+If we used all samples during propagation we would make only 1 update for the network's parameter
+
+The smaller the batch the less accurate the estimate of the gradient will be.
+In the figure below, you can see that the direction of the mini-batch gradient (green color) fluctuates much more in comparison to the direction of the full batch gradient (blue color).
 

### 2019-12-08 Common/chapters/medical.tex
Julien and Johan: describe medical imaging

@@
 
-\subsection{Types}
+Multiple types of medical imaging exist. The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans
+%, PET (Positron Emission Tomography) scans
+and mammographies.
 
-\subsubsection{MRI}
+MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the outputed images as a volume.
+Different settings, usually called sequences, make the look of the output vary, as shown on figure \ref{fig:PROSTATEx-t2-adc-dwi}.
+Unlike MRI, CT is based on X-rays instead of magnetic fields, but still provides a three-dimensional representation of a body part. Figure \ref{fig:visualize_lung_dcm} shows a lung CT scan.
 
-\subsubsection{CT}
-
-\subsubsection{Mammography}
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/PROSTATEx-t2-adc-dwi.png}
+\caption{MRI - PROSTATEx - From left to right: T2-weighted, ADC and DWI}
+\label{fig:PROSTATEx-t2-adc-dwi}
+\end{figure}

### 2019-12-09 Common/chapters/deep_learning.tex
Johan and Julien: deep learning chapter continued until backpropagation (need to be terminated)

@@
 \label{historical_background}
-As described in figure \ref{history} from article \cite{14}, the theoretical foundations of deep learning appeared long before the invention of the computers.
+As described in figure \ref{history} from article \cite{14}, the theoretical foundations of deep learning appeared long before the invention of computers.
 From the first attempts to understand the human brain until today, a long way has been made to establish the basic components of modern neural networks.
@@
 \label{perceptron_equation}
-y_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
+a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
 \end{equation}
@@
 It is extremely important to use a non linear functions instead of a linear function.
-Indeed, in a neural network where perceptrons are assembled together, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
-Consequently, if only linear functions were used, linear outputs would be given as inputs to other linear functions.
+Indeed, in a neural network the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
+Consequently, if only linear functions are used, linear outputs would be given as inputs to other linear functions.
 Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create neural networks of multiple layers.
@@
 \item The input layer: it is the neural network entry point for the data.
-Generally, the data is provided to the network in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $n_{y} \times batch\_size$.
+Generally, the data is provided to the network in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
 The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is their dimension.
 Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector.
+The total number of training samples is given by $m$.
 Finally, $y^{(i)}$ is the output label for the $i^{th}$ example.
@@
 The weights and the biases of the network are the parameters that have to be optimized in order to have the best mapping between the inputs and the output of the network (see section \ref{training_a_neural_network} ).
+Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see section \ref{weight_optimization}).
 \item The output layer: it is the last layer of the neural network.
@@
 \end{itemize}
+The advantage of organizing the weights, biases and inputs in matrices is due to the ability of modern CPU/GPU to quickly perform linear algebra computations.
+This way of structuring the network component is called "vectorization" and avoids at most using loops in the code, which could considerably slow down the computations.
 The figure \ref{multilayer_perceptrons_figure} globally illustrates the concept of a multilayer perceptrons.
@@
 This step consists in giving to the untrained neural network examples that need to be classified (or segmented, depending on the task) and to transmit them through all layers of the network.
-Passing all batches as input the network is called an "epoch".
+Passing all batches as input to a network is called an "epoch".
 At the end of each batch the loss is computed using a "loss function".
@@
 Then, the weights and biases of the networks are updated during a process called "back propagation" in order find the global minima of the loss function.
-This process is repeated a for a defined number of epochs.
+This process is repeated for a defined number of epochs.
 
 \subsection{Forward propagation}
+The forward propagation is used to transmit the input through the entire neural network. Mathematically, the forward propagation for a given layer $l$ is given by two equations:
+\begin{equation}
+z^{[l]} = W_{x}^{[l]}x^{(i)} + b^{[l]}
+\end{equation}
+This first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$.
+The second equation describes the effect of the activation function:
+\begin{equation}
+a^{[l+1]} = g^{[l]}(z^{[l]})
+\end{equation}
+Since the output of the activation function is then given as input to the neuron of the next layer, the forward propagation step can be generally defined as:
+\begin{equation}
+a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) 
+\end{equation}
+The figure \ref{forward_propagation} illustrates the computation of the forward propagation for the $l^{th}$ layer.
+The matrix of weight $W_{jk}^{l}$ correspond to the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$.
+The latter is multiplied by the output of previous layer $a_{j}^{[l-1]}$, the biased $b^{[l]}$ is added and the entire result is given as input to the activation function.
 
@@
 \subsubsection{Loss computation}
+"The loss function evaluates the error performed by the neural network when it tries
+to estimate the data to be predicted"\cite{18}.
+It is therefore useful to measure the penalty for a single input.
+On the contrary, when the goal is to have a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used.
+The latter is given by the equation:
+\begin{equation}
+J(\hat{y}, y)
+\end{equation}
+where $\hat{y}$ is the prediction of the neural network and $y$ the real label.\\
+There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
+\begin{equation}
+J(\hat{y}, y) = \frac{1}{m}[\sum_{i=1}^{m} (\hat{y}^{(i)} - y^{(i)})^{2}]
+\end{equation}
+For classification problems, the cross entropy function is regularly used.
+We distinguish the binary classification where the number of classes $n_{y}$ = 2 and the multiclass classification where $n_{y}$ > 2.
+In binary classification, the cross entropy is:
+\begin{equation}
+J(\hat{y}, y) = -\frac{1}{m}\sum_{i=1}^{m} [y_{i}*log(\hat{y}_{i}) + (1-y_{i})*log(1-\hat{y_i})]
+\end{equation}
+In multiclass classification, the categorical crossentropy is:
+\begin{equation}
+J(\hat{y}, y) = - \sum_{i=1}^{n_{y}} \sum_{j=1}^{m} (y_{ij}*log(\hat{y}_{ij}))
+\end{equation}
 
+Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weight in order to approach the minimum of the function.
+Therefore, deep learning problems can be considered as optimization problems.
+Solutions to these problems can be found using the gradient descent during the backpropagation.
 \subsection{Backpropagation}
+The backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
+Globally, "the intuition behind the backpropagation algorithm is as follows.
+Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network.
+Then, for each node $i$ in layer $l$, we would like to compute an "error term" $\partial^{(l)}_{i}$ that measures how much that node was "responsible" for any errors in our output.
+For an output node, we can directly measure the difference between the network’s activation and the true target value, and use that to define $\partial^{(n_{l})}_{i}$(where layer $n_{l}$ is the output layer).
+How about hidden units?
+For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input"\cite{24}.
 
@@
 \subsection{Weight initialization}
+\label{weight_optimization}
 

### 2019-12-11 Common/chapters/deep_learning.tex
Johan and Julien: deep learning chapter continued

@@
 Indeed, in a neural network the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
-Consequently, if only linear functions are used, linear outputs would be given as inputs to other linear functions.
+Consequently, if only linear functions are used, linear outputs are given as inputs to other linear functions.
 Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create neural networks of multiple layers.
@@
 \end{equation}
-Since the output of the activation function is then given as input to the neuron of the next layer, the forward propagation step can be generally defined as:
+Since the output of the activation function is then given as input to all neurons of the next layer, the forward propagation step can be generally defined as:
 \begin{equation}
@@
 Therefore, deep learning problems can be considered as optimization problems.
-Solutions to these problems can be found using the gradient descent during the backpropagation.
+Solutions to these problems can be found using the gradient descent algorithm during backpropagation.
 \subsection{Backpropagation}
-The backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
+Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
 Globally, "the intuition behind the backpropagation algorithm is as follows.
@@
 How about hidden units?
-For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input"\cite{24}.
+For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input"\cite{24}.\\
+In other words, after each forward pass through the entire network, backpropagation performs a backward pass with the goal of minimizing the cost function by adjusting model's parameters.
+The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network.
+The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ in point $x$ is given by:
+\begin{equation}
+\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{1}}]
+\end{equation}
+The gradient shows how much all parameters that constitute x need to change to minimize the function.
+In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$ The computation of all these gradients relies on the "chain rule".
+For weights parameters, the chain rule is:
+\begin{equation}
+\frac{\partial J}{\partial w_{jk}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial w_{jk}^{l}}
+\end{equation}
+Similarly, the chain rules has to be applied to biases since they are also parameters of the cost function:
+\begin{equation}
+\frac{\partial J}{\partial b_{j}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial b_{j}^{l}}
+\end{equation}
+Once the gradients of each parameters of the neural network are computed, they have to be updated. The weights update is described by the following equation:
+\begin{equation}
+W^{[l]} = W^{[l]} - \alpha * \frac{\partial J}{\partial W^{[l]}}
+\end{equation}
+The biases update corresponds to:
+\begin{equation}
+b^{[l]} = b^{[l]} - \alpha * \frac{\partial J}{\partial b^{[l]}}
+\end{equation}
+The hyperparameter $\alpha$ of these two equations is named "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
 
-\subsubsection{Weight update}
+\subsection{Metrics}
+In classification tasks, four separate output labels can occur:
+\begin{itemize}
+\item True Positive (TP):  an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{correct}.
+\item True Negative (TN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{correct}.
+\item False Positive (FP): an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{incorrect}.
+\item False Negative (FN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{incorrect}.
+\end{itemize}
+From these four categories, multiple metrics with their own specificities can be established \cite{25}:
+\begin{itemize}
+\item Accuracy: it is defined as the ratio of the correctly labeled subjects to the whole pool of subjects.
+\begin{equation}
+Accuracy = \frac{(TP+TN)}{TP+FP+FN+TN}
+\end{equation}
+Accuracy is a great measure in the case of symmetric data (i.e the number of FN $\approx$ FP and their cost is similar).
+When this condition is not fulfilled, accuracy can lead to bad models.
+For instance, let's define a binary classification model that always outputs class 0.
+If the data is composed of 99 samples from class 0 and 1 from class 1, the accuracy is equal to 99\% but the model is not smart.
+Consequently, this metric has to be used in addition to other metrics.
+
+\item Precision: it is the ratio of the correctly positive labeled subjects by the model to all positive labeled subjects.
+\begin{equation}
+Precision = \frac{TP}{(TP + FP)}
+\end{equation}
+This metric is recommended when the confidence of the true positives predicted by the model is important.
+For instance, this happens in spam blockers where it is preferable to have a spam in mailbox rather than a regular mail in the spam box.
+
+\item Recall (sensitivity):
+\begin{equation}
+Recall = \frac{TP}{(TP + FN)}
+\end{equation}
+This metric is recommended when the occurence of false negatives is intolerable and false positives are preferred.
+This makes perfect sense for disease detection models: labeling an healthy person as unhealthy is better than labeling an unhealthy person as healthy.
+
+\item F1-score:
+\begin{equation}
+F1-score = \frac{2* recall * precision}{(recall + precision)}
+\end{equation}
+F1-score considers both precision and recall and is the highest if these two metrics are balanced.
+This metric is perfectly suitable when the cost of false positives and false negatives is not the same.
+
+\item Specificity:
+\begin{equation}
+Specificity = \frac{TN}{(TN + FP)}
+\end{equation}
+This metric is recommended when the occurence of false positives is intolerable whereas true negatives are desired.
+For instance drug tests can not indicate false positives but they have to cover all true negatives.
+\item AUC-ROC:
+"The ROC curve is plotted with recall against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
+AUC - ROC curve is a performance measurement for classification problem at various thresholds settings.
+ROC is a probability curve and AUC represents degree or measure of separability.
+It tells how much model is capable of distinguishing between classes.
+Higher the AUC, better the model is at predicting 0s as 0s and 1s as 1s\cite{26}."
+
+\end{itemize}
 
 \subsection{Data}
+In deep learning, data is essential.
+As seen previously, neural networks learn features from it.
+Therefore, data has to be handled carefully and in the right way.
+Usually, it is split into three different sets:
+\begin{itemize}
+\item The training set: this is the dataset that is used to train the neural network (the weights and biases).
+\item The validation set: this the dataset that is used to evaluate a trained model.
+Usually, the evaluation on the validation set is made every $x$ (a fixed number) epochs.
+The validation set has to come from the same distribution as the training set but should contain exclusively unseen data.
+This last point is crucial since the validation set shows how well the neural network can generalize on unknown data.
+The validation set can also be used as indicator to decide when the training should be stopped in order to prevent overfitting.
+This technique is called "early stopping".
+\item The test set: this last dataset is used to establish the final evaluation of the model.
+
+\end{itemize}
 Training set, validation set, test set, supervised learning, unsupervised learning,

### 2019-12-12 Common/chapters/deep_learning.tex
Johan and Julien: Deep learning chapter continued

@@
 The weights and the biases of the network are the parameters that have to be optimized in order to have the best mapping between the inputs and the output of the network (see section \ref{training_a_neural_network} ).
-Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see section \ref{weight_optimization}).
+Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see section \ref{weight_initialization}).
 \item The output layer: it is the last layer of the neural network.
@@
 The first one is the "forward propagation step".
-This step consists in giving to the untrained neural network examples that need to be classified (or segmented, depending on the task) and to transmit them through all layers of the network.
+It consists in giving to the untrained neural network examples that need to be classified (or segmented, depending on the task) and to transmit them through all layers of the network.
 Passing all batches as input to a network is called an "epoch".
@@
 The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth.
-Then, the weights and biases of the networks are updated during a process called "back propagation" in order find the global minima of the loss function.
+Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minima of the loss function.
 This process is repeated for a defined number of epochs.
@@
 \begin{equation}
-a^{[l+1]} = g^{[l]}(z^{[l]})
+a^{[l]} = g^{[l]}(z^{[l]})
 \end{equation}
@@
 On the contrary, when the goal is to have a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used.
-The latter is given by the equation:
-\begin{equation}
-J(\hat{y}, y)
-\end{equation}
-where $\hat{y}$ is the prediction of the neural network and $y$ the real label.\\
-There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
+The latter is given by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.
+There exist multiple cost functions.\\
+For a regression problem, a commonly used loss function is the mean square error:
 \begin{equation}
@@
 Usually, the evaluation on the validation set is made every $x$ (a fixed number) epochs.
-The validation set has to come from the same distribution as the training set but should contain exclusively unseen data.
+The validation set needs to come from the same distribution as the training set but should contain exclusively unseen data.
 This last point is crucial since the validation set shows how well the neural network can generalize on unknown data.
-The validation set can also be used as indicator to decide when the training should be stopped in order to prevent overfitting.
+The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" (overfitting describes the behaviour of a model that too closely fits to the training set and don't generalize well).
+Indeed, if the validation loss continually increases for a certain number of epochs, going on with the training will increase the overfitting.
 This technique is called "early stopping".
-\item The test set: this last dataset is used to establish the final evaluation of the model.
+\item The test set: this last dataset is used to establish the final evaluation of the model and also contains unseen data. 
+\end{itemize}
+Regarding the way these three sets are split, it mostly depends on the number of samples available.
+If the latter is big, the data is split into training and testing using the ratio 80/20.
+Then the remaining training samples are also split into training and validation using the ratio 80/20.
+On the contrary, if there are few data available, K-folds cross-validation is a good practice.
+This technique consists in splitting the entire dataset into k-folds.
+One fold is picked as test set and the others are considered as training sets.
+The model is trained on training folds and tested on the test set.
+Then, another test set is picked and the same process is repeated until all possible tests set are picked.
+At the end of the process, all results of all test sets are averaged, which provides a good estimation of the model performances.
+This technique is summarized on figure \ref{cross_validation}.
 
-\end{itemize}
-Training set, validation set, test set, supervised learning, unsupervised learning,
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/cross_validation.png}
+\caption{5-folds cross-validation }
+\label{cross_validation}
+\end{figure}
 
 \subsection{Weight initialization}
-\label{weight_optimization}
+\label{weight_initialization}
+Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
+The initialization of the neural network weights is crucial since it will determine how quickly the network converges.
+The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
+If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all"\cite{27}.\\
+The simplest and least efficient technique to initialize neural network weights is to randomly generate them.
+The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions.
+Consequently, the slope of the gradient changes slowly and the learning takes a lot of time.\\
+To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
+\begin{equation}
+\frac{\sqrt{6}}{\sqrt{n_{h}^{[l]}+n_{h}^{[l+1]}}}
+\end{equation}
+where $n_{h}^{[l]}$ is the number of incoming network connection to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
+For activation functions that are not symmetric about zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative.
+It consists in multiplying the random initialized matrix weight by:
+\begin{equation}
+\frac{\sqrt{2}}{\sqrt{n_{h}^{[l]}}}
+\end{equation} 
+where $n_{h}^{[l]}$ is the number of incoming connections coming into a given layer from the previous layer's output.
 
 \subsection{Hyperparameters}
+Hyperparameters denotes parameters that cannot be directly learned from the data. In deep learning,
 

### 2019-12-17 Common/chapters/deep_learning.tex
Johan and Julien: Deep learning chapter continued

@@
 \end{equation}
-Similarly, the chain rules has to be applied to biases since they are also parameters of the cost function:
+Similarly, the chain rules has to be applied to the biases:
 \begin{equation}
@@
 For instance, let's define a binary classification model that always outputs class 0.
-If the data is composed of 99 samples from class 0 and 1 from class 1, the accuracy is equal to 99\% but the model is not smart.
+If the data is composed of 99 samples from class 0 and 1 sample from class 1, the accuracy is equal to 99\% but the model is not smart.
 Consequently, this metric has to be used in addition to other metrics.
@@
 
-\subsection{Hyperparameters}
-Hyperparameters denotes parameters that cannot be directly learned from the data. In deep learning,
+\subsection{Hyperparameters tuning}
+Hyperparameters denotes parameters that cannot be directly learned from the data.
+That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
+So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes at most the cost function keeping a good level of generalization).\\
+Regarding the learning rate $\alpha$, its value has to be neither too large or too small.
+A too large learning rate is recognizable analyzing the training loss curve: if "the loss is exploding or fluctuating" or if "it has stopped improving and it is wandering around a suboptimal local optima"\cite{28} the learning rate is too high and should be decreased.
+On the contrary, if "the learning is very slow and the loss is decreasing consistently" or if "the model is overfitting"\cite{28}, it is a clear sign that it should be increased.
+The learning rate can take a wide range of values.
+Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored"\cite{28}.
+There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it at each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate schedulers).\\
+Batch size is another important hyperparameter to tune.
+Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples"\cite{29}.
+Furthermore, "networks train faster because the weights update is done after each propagation"\cite{29}.
+Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be"\cite{29}.
+Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.\\
+Finally, the number of epochs during which the network is trained has to be carefully chosen.
+In fact, from a certain point in the training, neural network don't learn anymore useful features in the data and start overfitting.
+This point corresponds to the moment where, the validation loss does not decrease anymore and starts to continually increase.
+It is exactly at this moment that the training should be stopped.
+To achieve this goal, it is possible either to directly choose the right number of epochs or to use "early stopping", which stops the training as soon as the validation loss does not decrease for $x$ epochs.
 
-lr, batch size, epochs, optimizer, learning rate decay, ...
-
-From: https://stats.stackexchange.com/questions/153531/what-is-batch-size-in-neural-network
-Batch size: 
-Advantages of using a batch size < number of all samples:
-It requires less memory.
-Since you train the network using fewer samples, the overall training procedure requires less memory.
-That's especially important if you are not able to fit the whole dataset in your machine's memory.
-
-Typically networks train faster with mini-batches.
-That's because we update the weights after each propagation.
-In our example we've propagated 11 batches (10 of them had 100 samples and 1 had 50 samples) and after each of them we've updated our network's parameters.
-If we used all samples during propagation we would make only 1 update for the network's parameter
-
-The smaller the batch the less accurate the estimate of the gradient will be.
-In the figure below, you can see that the direction of the mini-batch gradient (green color) fluctuates much more in comparison to the direction of the full batch gradient (blue color).
-
-\section{Deep learning models}
-xxx
-
-\subsubsection{Convolutional Neural Networks}
-
-\subsubsection{Recurrent Neural Networks}
+\section{Convolutional Neural Networks}
 
@@
 
-\section{Deep learning frameworks}
+"Transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task.\cite{30}".
+The model dedicated to the second task uses all or parts of the first model (i.e.
+keeps the same weights and architecture or a part of them) and is then retrained on data available for this task.
+The first model can either be implemented from scratch if enough data is available or can simply be downloaded from institutions that release large and trained models for similar tasks.\\
+It mainly provides three benefits \cite{30}:
+\begin{itemize}
+\item Higher start: the initial skill (before refining the model) on the source model is higher than it otherwise would be.
+\item Higher slope: the rate of improvement of skill during training of the source model is steeper than it otherwise would be.
+\item Higher asymptote: the converged skill of the trained model is better than it otherwise would be.
+\end{itemize}
+Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated"\cite{30}.

### 2019-12-17 Common/chapters/deep_learning.tex
Julien and Johan: correct typos and rephrase some bits

@@
 Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them.
-In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way accordingly.
+In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way.
 There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc.
-What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}), that are combined to form deep neural networks.\\
+What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
 Neural networks are inspired from the biological neural networks of the brain.
@@
 Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc.
-The presence of deep learning techniques in all mentioned use cases clearly attests that this technology is a real breakthrough.
-Furthermore, since this field has recently gained interest   (see section \ref{historical_background}), multiple researches are still ongoing, which suggests that many exciting new applications will certainly be discovered.
+The presence of deep learning techniques in these use cases clearly attests that this technology is a breakthrough.
+Furthermore, since this field has recently gained interest (see section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.
 
@@
 \label{historical_background}
-As described in figure \ref{history} from article \cite{14}, the theoretical foundations of deep learning appeared long before the invention of computers.
-From the first attempts to understand the human brain until today, a long way has been made to establish the basic components of modern neural networks.
-One could ask: why did deep learning take off recently if the theory has been around for a long time?\\
-First part of the answer is about the computational power.
-Indeed, deep learning algorithms need a lot of data to work properly and this requires powerful CPU/GPU that either didn't exist or were only available to a minority of people.
+As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
+From the first attempts to understand the human brain until today, huge progres was made to be establish the basic components of modern neural networks.
+One could ask why deep learning took off recently if the theory was around for a long time.\\
+As stated by Goodfellow et al.
+\cite{15}, the first part of the answer is computing power.
+In fact, deep learning algorithms need a lot of data to work properly, which requires powerful CPUs/GPUs that either didn't exist or were only within few people's reach.
 One other main reason concerns the lack of data.
-Since deep learning algorithms "learn" from data, if not data is available, no learning is possible.
-Consequently, the era of Big Data has enhanced the deep learning possibilities.
-These two points are summarised: "The increase in model size overtime, due to the availability of faster CPUs, the advent of general purpose GPUs, fasternetwork, connectivity and better software infrastructure for distributed computing, is one of the most important trends in the history of deep learning.
-This trend is generally expected to continue well into the future"\cite{15}.\\
-Finally, before the year 2012, the abilities of neural networks had not been publicly proven.
+Since deep learning algorithms "learn" from data, learning is impossible if good-quality data are not available.
+The era of Big Data enhanced deep learning possibilities.
+%%%% Je mettrais soit l'un soit l'autre. Le paragraphe du dessus c'est le paragraphe du dessous avec des autres mots :)
+% These two points are summarised: "The increase in model size overtime, due to the availability of faster CPUs, the advent of general purpose GPUs, fasternetwork, connectivity and better software infrastructure for distributed computing, is one of the most important trends in the history of deep learning.
+This trend is generally expected to continue well into the future".\\
+Finally, before the year 2012, the abilities of neural networks were still to be proven.
 This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluate their algorithms on several visual recognition tasks).
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/history.png}
-\caption{Milestones }
+\caption{Milestones - Wang et al. \cite{14}}
 \label{history}
@@
 
-From a descriptive point of view, neural networks can simply be seen as a non linear applications that associate an input to an output with respect to certain parameters.
-The input can be images, sounds or whatever features that is numerical.
+From a descriptive point of view, neural networks can simply be seen as a non-linear applications that associate an input to an output with respect to certain parameters.
+The input can be an image, a sound or any input that can be converted into numerical features.
 The output of a neural network depends on the problem it tries to solve.
 In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
-From a mathematical point of view, a neural network can be defined as a non linear function $f$ that associates to an input $x$ an output $y$ with respect to parameters $\theta$.
+From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates to an input $x$ an output $y$ with respect to parameters $\theta$.
 \begin{equation}
@@
 
-\subsection{Supervised learning and Unsupervised learning}
-In machine learning, there mainly exists two different kinds of task.
+\subsection{Supervised and unsupervised learning}
+In machine learning, two different kinds of tasks exist.
 The first one is "supervised learning".
-It includes all learning algorithms, whose training samples contains labels in order to find the optimal mapping between the input and the output.
+It includes learning algorithms whose training samples are associtted to their labels in order to find the optimal mapping between the input and the output.
 The second one is "unsupervised learning".
-In contrast to supervised algorithms, the latter use non labeled data and its main goal is to infer the natural structure  present in the data.
-Since the model presented in this work belongs to the "supervised learning" category, all notions explained below refer to this kind of algorithms.
+In contrast to supervised algorithms, the latter rely on unlabeled data.
+Its main goal is to infer the natural structure  present in the data.
+Since the models presented in this work belong to the "supervised learning" category, notions explained below refer to this kind of algorithms.
 
@@
 \subsection{Notation}
-In order to respect a consistent mathematical notation to describe neural networks, this work will use the notation of Andrew Y.
-Ng, one of the pioneers of deep learning, that can be found in \cite{16}.\\
+In order to keep the mathematical description of neural network consistent, this work will use Andrew Y. Ng's notation \cite{16}, who was a pioneer in deep learning.\\
 
@@
 \begin{itemize}
-\item Superscript (i) will denote the $i^{th}$ training example.
-\item Superscript [l] will denote the $l^{th}$ layer of the neural network.
+\item Superscript (i) denotes the $i^{th}$ training example.
+\item Superscript [l] denotes the $l^{th}$ layer of the neural network.
 \end{itemize}
@@
 \begin{itemize}
-\item $m$: number of examples in the dataset.
-\item $n_{x}$: input size.
-\item $n_{y}$: output size (or number of classes).
-\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer. 
-\item $L$: number of layers in the network.
+\item $m$: number of examples in the dataset
+\item $n_{x}$: input size
+\item $n_{y}$: output size (or number of classes)
+\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer
+\item $L$: number of layers in the network
 \end{itemize}
@@
 \item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
-\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the network.
+\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the whole network.
 \item $g^{[l]}(x)$ is the $l^{th}$ activation function.  
@@
 \item $a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) $ is the general activation formula at $l^{th}$ layer.
-\item $J(x, W, b, y) or J(\hat{y}, y)$ denote the cost function.
+\item $J(x, W, b, y)$ and $J(\hat{y}, y)$ denote the cost function.
 \end{itemize}
@@
 Perceptrons are the main components of neural networks.
-They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts"\cite{13}.
+They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}.
 Today, they are called "artificial neurons" or simply "neurons".\\
-A perceptron $j$ is a function $f$ of the input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non linear activation function $g$:
+A perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
 \begin{equation}
@@
 
-Schematically, a perceptron can be represented as shown in figure \ref{perceptron_model}.
-Each input is multiplied with its corresponding weight and the sum of this result is then passed through a non linear function, called an "activation function".
-This activation function acts like a threshold that decides the proportion of the result that has to be given further.
+Schematically speaking, a perceptron can be represented as on figure \ref{perceptron_model}.
+Each input is multiplied with its corresponding weight.
+The sum of this result then goes through a non-linear function, called "activation function".
+This activation function acts like a threshold that determines the proportion of the result that goes further in the network.
 There exist multiple activation functions (see \ref{activation_functions}).
-It is extremely important to use a non linear functions instead of a linear function.
-Indeed, in a neural network the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
-Consequently, if only linear functions are used, linear outputs are given as inputs to other linear functions.
-Since the composition of two linear functions is itself a linear function, there would be no sense to assemble perceptrons to create neural networks of multiple layers.
+It is extremely important to use non-linear functions instead of a linear functions.
+In fact, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
+Consequently, if linear functions only are used throughout the network, linear outputs are given as inputs to other linear functions.
+Since the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers would not make sense anymore.
 
@@
 \label{activation_functions}
-Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result to send further.
-Activation functions have to be non-linear in order to approximate extremely complex functions.
+Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer.
+Activation functions must be non-linear in order to approximate extremely complex functions.
 In fact, neural networks are considered as universal approximators.
-Article \cite{17} claims that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense."According to \cite{18} the most commonly used activation functions are:\\\\
+Hornik et al.
+claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}.
+According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
  \noindent \textbf{Sigmoid function}\\
@@
     \end{tikzpicture}
-    \caption{The ReLU function and its derivative(not defined when x=0)}
+    \caption{The ReLU function and its derivative (undefined when $x=0$)}
   \end{center}
@@
 \label{multilayer_perceptron}
-A multilayer perceptron is an artifical neural network type. Article \cite{23} defines the multilayer perceptrons as "feedforward networks with one or more layers of units between the
-input and output layers. The output units represent a hyperplane in the space
+A multilayer perceptron is a type of artifical neural networks. Du et al. \cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
+input and output layers" where "the output units represent a hyperplane in the space
 of the input patterns."\\
-A multilayer perceptrons is composed of $L$ layers each of them having $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
+A multilayer perceptron is composed of $L$ layers, each of them composed of $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
 \begin{itemize}
 \item The input layer: it is the neural network entry point for the data.
-Generally, the data is provided to the network in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
-The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is their dimension.
+Generally speaking, the data is provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
+The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is the dimension of each sample.
 Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector.
@@
 Finally, $y^{(i)}$ is the output label for the $i^{th}$ example.
-For instance, under the assumption that the number of samples is 100 and that the batch size is 32, 4 batches in total will be given as input to the network.
+For instance, suppose the number of samples is 100 and the batch size 32.
+In this situation, the network will be feeded with 4 batches of sizes [32, 32, 32, 4] respectively.
+
 \item The hidden layer(s): hidden layers stand for all layers that are between the input layer and the output layer.
-For each of them are associated weights and biases (W, b), respectively denoted by $W^{[l]} \in \R $ and $b^{[l]}$, where $W_{ij}^{l}$ correspond to the weights associated with the connection between perceptron $j$ in layer $l$ and perceptron $i$ in layer $l+1$.
-By analogy, $b_{i}^{[l]}$ is the bias associated with perceptron $i$ in layer $l+1$.
-The weights and the biases of the network are the parameters that have to be optimized in order to have the best mapping between the inputs and the output of the network (see section \ref{training_a_neural_network} ).
+Each of them has its own weights and biases (W, b), denoted by $W^{[l]} \in \R $ and $b^{[l]}$ respectively, where $W_{ij}^{l}$ corresponds to the weights associated with the connection between perceptron $j$ in layer $l$ and perceptron $i$ in layer $l+1$.
+By analogy, $b_{i}^{[l]}$ is the bias associated with perceptron $i$ in layer $l$.
+Weights and biases are the parameters to optimize in order to obtain the best mapping between the input and the output of the network (see section \ref{training_a_neural_network} ).
 Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see section \ref{weight_initialization}).
+
 \item The output layer: it is the last layer of the neural network.
@@
 The prediction of a neural network is given by $\hat{y} \in R^{n_{y}}$ with $n_{y}$ beeing the number of different labels.
-In a classification task, whose goal is to assign to each input a specific class, the $\hat{y}$ could be the probability of belonging to each class.
-In that case, the \textit{softmax} activation function would be used.
+In a classification task, whose goal is to assign to each input a specific class, the $\hat{y}$ is usually the probability that the input belongs to each class.
+In that case, the \textit{softmax} activation function would be used at the end.
 \end{itemize}
-The advantage of organizing the weights, biases and inputs in matrices is due to the ability of modern CPU/GPU to quickly perform linear algebra computations.
-This way of structuring the network component is called "vectorization" and avoids at most using loops in the code, which could considerably slow down the computations.
-The figure \ref{multilayer_perceptrons_figure} globally illustrates the concept of a multilayer perceptrons.
+The advantage of organizing the weights, biases and inputs in matrices is due to the ability of modern CPUs/GPUs to quickly perform linear algebra computations.
+This way of structuring the network component is called "vectorization" which avoids using loops in the code, which would considerably slow down the computations.
+Figure \ref{multilayer_perceptrons_figure} illustrates the concept of multilayer perceptrons.
 In this example, the total number of layers $L$ is equal to 3, the input size $n_{x}$ is equal to 4 and the number of units in in each layer is $n_{h}^{1}= 4$, $n_{h}^{2}=2$, $n_{h}^{3}=1$.
-The network has the weights  $W^{(1)} \in \R^{2x3}$, $W^{(2)} \in \R^{1x2}$  and the biases  $b^{[1]}=2$, $b^{[2]}=1$.
+The network contains the weights  $W^{(1)} \in \R^{2x3}$, $W^{(2)} \in \R^{1x2}$ and the biases  $b^{[1]}=2$, $b^{[2]}=1$.
 
@@
 \label{training_a_neural_network}
-Training a neural network is composed of multiple steps.
-The first one is the "forward propagation step".
-It consists in giving to the untrained neural network examples that need to be classified (or segmented, depending on the task) and to transmit them through all layers of the network.
-Passing all batches as input to a network is called an "epoch".
-At the end of each batch the loss is computed using a "loss function".
+Training a neural network can be broken down into multiple steps.
+The first one is the "forward propagation" step.
+It consists in giving examples that need to be classified (or segmented, depending on the task) to the untrained neural network and to spread intermediate results through all layers of the network.
+%Passing all batches as input to a network is called an "epoch".
+After seeing every single batch, the loss is computed using a "loss function".
 The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth.
-Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minima of the loss function.
+Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minimum of the loss function.
+As soon as the neural network has seen every single batch, the end of an "epoch" is reached.
 This process is repeated for a defined number of epochs.
@@
 \subsection{Forward propagation}
-The forward propagation is used to transmit the input through the entire neural network. Mathematically, the forward propagation for a given layer $l$ is given by two equations:
+The forward propagation is used to transmit the input through the entire neural network.
+Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations.
+The first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$:
 \begin{equation}
@@
 \end{equation}
-This first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$.
 The second equation describes the effect of the activation function:
@@
 \end{equation}
-Since the output of the activation function is then given as input to all neurons of the next layer, the forward propagation step can be generally defined as:
+Since the output of the activation function is then given as input to all the neurons of the next layer, the whole forward propagation step can be defined as:
 \begin{equation}
@@
 \end{equation}
-The figure \ref{forward_propagation} illustrates the computation of the forward propagation for the $l^{th}$ layer.
-The matrix of weight $W_{jk}^{l}$ correspond to the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$.
-The latter is multiplied by the output of previous layer $a_{j}^{[l-1]}$, the biased $b^{[l]}$ is added and the entire result is given as input to the activation function.
+Figure \ref{forward_propagation} illustrates the computation of the forward propagation pass for the $l^{th}$ layer.
+The weight natrix $W_{jk}^{l}$ represents the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$.
+This matrix is multiplied by the output of the previous layer $a_{j}^{[l-1]}$ before adding the bias $b^{[l]}$.
+The result is given as input to the activation function.
 
@@
 \subsubsection{Loss computation}
-"The loss function evaluates the error performed by the neural network when it tries
-to estimate the data to be predicted"\cite{18}.
+As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
 It is therefore useful to measure the penalty for a single input.
-On the contrary, when the goal is to have a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used.
-The latter is given by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.
-There exist multiple cost functions.\\
-For a regression problem, a commonly used loss function is the mean square error:
+On the contrary, when the goal is to get a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used.
+The latter is represented by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.
+
+There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
 \begin{equation}
@@
 For classification problems, the cross entropy function is regularly used.
-We distinguish the binary classification where the number of classes $n_{y}$ = 2 and the multiclass classification where $n_{y}$ > 2.
-In binary classification, the cross entropy is:
+We distinguish the binary classification where the number of classes $n_{y}$ = 2 from the multiclass classification where $n_{y}$ > 2.
+In the case of binary classification, the cross entropy is:
 \begin{equation}
@@
 \end{equation}
-In multiclass classification, the categorical crossentropy is:
+In the case of multiclass classification, the categorical crossentropy is:
 \begin{equation}
@@
 
-Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weight in order to approach the minimum of the function.
+Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function.
 Therefore, deep learning problems can be considered as optimization problems.
 Solutions to these problems can be found using the gradient descent algorithm during backpropagation.
+
 \subsection{Backpropagation}
 Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
-Globally, "the intuition behind the backpropagation algorithm is as follows.
+Generally speaking, "the intuition behind the backpropagation algorithm is as follows.
 Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network.
@@
 How about hidden units?
-For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input"\cite{24}.\\
-In other words, after each forward pass through the entire network, backpropagation performs a backward pass with the goal of minimizing the cost function by adjusting model's parameters.
+For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.
+
+In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting parameters of the model.
 The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network.
-The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ in point $x$ is given by:
+The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ at point $x$ is given by:
 \begin{equation}
@@
 The gradient shows how much all parameters that constitute x need to change to minimize the function.
-In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$ The computation of all these gradients relies on the "chain rule".
-For weights parameters, the chain rule is:
+In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$.
+The computation of all these gradients relies on the "chain rule".
+Inthe case of weights, the chain rule is:
 \begin{equation}
@@
 \end{equation}
-Once the gradients of each parameters of the neural network are computed, they have to be updated. The weights update is described by the following equation:
+Once the gradients of each parameter are computed, these parameters are updated. The weights update is described by the following equation:
 \begin{equation}
@@
 \end{equation}
-The hyperparameter $\alpha$ of these two equations is named "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
+%The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
+The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be tuned manually.
 
@@
 \end{itemize}
-From these four categories, multiple metrics with their own specificities can be established \cite{25}:
+From these four categories, multiple metrics with their own specificities can be computed \cite{25}:
 \begin{itemize}
-\item Accuracy: it is defined as the ratio of the correctly labeled subjects to the whole pool of subjects.
+\item Accuracy: ratio of the correctly labeled subjects to the whole pool of subjects.
 \begin{equation}
@@
 
-\item Precision: it is the ratio of the correctly positive labeled subjects by the model to all positive labeled subjects.
+\item Precision: ratio of the correctly positive labeled subjects by the model to all positive labeled subjects.
 \begin{equation}
@@
 For instance drug tests can not indicate false positives but they have to cover all true negatives.
-\item AUC-ROC:
-"The ROC curve is plotted with recall against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
+\item ROC curve and AUC:
+As explained by Sarang Narkhede, "The ROC curve is plotted with recall against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
 AUC - ROC curve is a performance measurement for classification problem at various thresholds settings.
@@
 It tells how much model is capable of distinguishing between classes.
-Higher the AUC, better the model is at predicting 0s as 0s and 1s as 1s\cite{26}."
+Higher the AUC, better the model is at predicting 0s as 0s and 1s as 1s" \cite{26}.
 
@@
 \begin{itemize}
-\item The training set: this is the dataset that is used to train the neural network (the weights and biases).
+\item The training set: this is the part of the dataset that is used to train the neural network (the weights and biases).
 \item The validation set: this the dataset that is used to evaluate a trained model.
-Usually, the evaluation on the validation set is made every $x$ (a fixed number) epochs.
+Usually, the evaluation on the validation set is performed every $N$ epochs, where $N$ is a fixed number.
 The validation set needs to come from the same distribution as the training set but should contain exclusively unseen data.
-This last point is crucial since the validation set shows how well the neural network can generalize on unknown data.
-The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" (overfitting describes the behaviour of a model that too closely fits to the training set and don't generalize well).
-Indeed, if the validation loss continually increases for a certain number of epochs, going on with the training will increase the overfitting.
-This technique is called "early stopping".
-\item The test set: this last dataset is used to establish the final evaluation of the model and also contains unseen data. 
+This last point is crucial since the validation set shows how well the neural network generalizes on unknown data.
+The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" which is the behaviour of a model that fits to the training set too closely and don't generalize well.
+In fact, if the validation loss continuously increases for a certain number of epochs, going on with the training will increase the overfit.
+This fact of interrupting the training earlier is called "early stopping".
+\item The test set: this last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
 \end{itemize}
@@
 Then the remaining training samples are also split into training and validation using the ratio 80/20.
-On the contrary, if there are few data available, K-folds cross-validation is a good practice.
-This technique consists in splitting the entire dataset into k-folds.
+On the contrary, if there are few data available, k-fold cross-validation is a good practice.
+This technique consists in splitting the entire dataset into k folds.
 One fold is picked as test set and the others are considered as training sets.
@@
 Then, another test set is picked and the same process is repeated until all possible tests set are picked.
-At the end of the process, all results of all test sets are averaged, which provides a good estimation of the model performances.
+At the end of the process, the results of all test sets are averaged, which provides a good estimation of the model's performance.
 This technique is summarized on figure \ref{cross_validation}.
@@
 Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
-The initialization of the neural network weights is crucial since it will determine how quickly the network converges.
+The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum.
 The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
-If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all"\cite{27}.\\
+If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.\\
 The simplest and least efficient technique to initialize neural network weights is to randomly generate them.
@@
 \end{equation}
-where $n_{h}^{[l]}$ is the number of incoming network connection to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
-For activation functions that are not symmetric about zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative.
-It consists in multiplying the random initialized matrix weight by:
+where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
+For activation functions that are not symmetric around zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative.
+It consists in multiplying the randomly initialized  weight matrix by:
 \begin{equation}
@@
 \subsection{Hyperparameters tuning}
-Hyperparameters denotes parameters that cannot be directly learned from the data.
+Hyperparameters denote parameters that cannot be directly learned from the data.
 That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
-So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes at most the cost function keeping a good level of generalization).\\
+So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps a acceptable level of generalization).\\
 Regarding the learning rate $\alpha$, its value has to be neither too large or too small.
-A too large learning rate is recognizable analyzing the training loss curve: if "the loss is exploding or fluctuating" or if "it has stopped improving and it is wandering around a suboptimal local optima"\cite{28} the learning rate is too high and should be decreased.
-On the contrary, if "the learning is very slow and the loss is decreasing consistently" or if "the model is overfitting"\cite{28}, it is a clear sign that it should be increased.
+A too large learning rate is recognizable by analyzing the training loss curve: if "the loss is exploding or fluctuating" or if "it has stopped improving and it is wandering around a suboptimal local optima" \cite{28}, the learning rate is too high and should be decreased.
+On the contrary, if "the learning is very slow and the loss is decreasing consistently" or if "the model is overfitting" \cite{28}, it is a clear sign that it should be increased.
 The learning rate can take a wide range of values.
-Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored"\cite{28}.
-There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it at each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate schedulers).\\
+Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored" \cite{28}.
+There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).\\
 Batch size is another important hyperparameter to tune.
-Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples"\cite{29}.
-Furthermore, "networks train faster because the weights update is done after each propagation"\cite{29}.
-Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be"\cite{29}.
+Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples" \cite{29}.
+Furthermore, "networks train faster because the weights update is done after each propagation" \cite{29}.
+Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be" \cite{29}.
 Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.\\
 Finally, the number of epochs during which the network is trained has to be carefully chosen.
-In fact, from a certain point in the training, neural network don't learn anymore useful features in the data and start overfitting.
-This point corresponds to the moment where, the validation loss does not decrease anymore and starts to continually increase.
-It is exactly at this moment that the training should be stopped.
-To achieve this goal, it is possible either to directly choose the right number of epochs or to use "early stopping", which stops the training as soon as the validation loss does not decrease for $x$ epochs.
+In fact, from a certain point in the training, neural networks don't learn anymore useful features in the data and start overfitting.
+This point corresponds to the moment where, the validation loss does not decrease anymore and starts to increase continuously.
+It is exactly at this moment that the training should stop.
+To achieve this goal, it is possible either to directly choose the right number of epochs or to use "early stopping", which stops the training as soon as the validation loss does not decrease for $N$ epochs.
 
@@
 
-"Transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task.\cite{30}".
+"Transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task.
+\cite{30}".
 The model dedicated to the second task uses all or parts of the first model (i.e.
@@
 \end{itemize}
-Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated"\cite{30}.
+Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated" \cite{30}.

### 2019-12-18 Common/chapters/deep_learning.tex
Julien and Johan: finish deep learning chapter

@@
 
+Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
+They are particular in that they contain layers which perform a mathematical operation named convolution on the input data.
+CNNs are mostly used in image and video analysis.
+
+To perform a convolution, numerical input data and a "filter".
+A filter can be seen as a an $fxf$ numerical patch that moves across the entire input.
+It first moves horizontally until reaching the right-most border of the image.
+Then, it goes down a cell and starts from the border on the left-hand side.
+This process is repeated until the filter reaches the bottom-right corner of the image, which marks the end of the convolution.
+At each step, the dot product between the filter and the part of the input covered by the filter is computed.
+Figure \ref{fig:convolution} showcases a simple convolution which aims at finding vertical lines in a black and white image.
+For example, the output of the first step of the convolution (in red) is computed by evaluating the dot product between the blue filter and the red part of the input image:
+\begin{equation}
+\begin{gathered}
+\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}0
+\\ \color{black}+ \\ 
+\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
+\\ \color{black}+ \\ 
+\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
+\\ \color{black}= -2
+\end{gathered}
+\end{equation}
+
+\noindent The output of the entire convolution shows negative values in the outside parts and large positive values in the center.
+This means that the $3x3$ filter detected a vertical line in the center of the input image.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution.png}
+\caption{Convolutions - Basic convolution in a CNN}
+\label{fig:convolution}
+\end{figure}
+
+Different parameters can change the way a convolution behaves.
+First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom.
+In this case, the "stride" is equal to 1.
+Other applications could rely on a bigger stride.
+Furthermore, the previous example reduced the output size of $3x3$ in proportion to the initial input size $5x5$.
+To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s.
+Three ways of padding images are commonly used as shown on figure \ref{fig:convolution_padding}:
+\begin{itemize}
+	\item \textbf{Valid:} the input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.
+
+	\item \textbf{Same:} the input image is padded in a way that makes the output size the same as the input size.
+
+	\item \textbf{Full:} the input image is padded so that, in the first step of the convolution, only the bottom-right cell of the filter overlays the pixel value of the image, the rest overlaying padding cells.
+This makes the output size larger than the input size.
+\end{itemize}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution_padding.png}
+\caption{Convolutions - Different padding methods}
+\label{fig:convolution_padding}
+\end{figure}
+
 \section{Transfer learning}
 
-"Transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task.
-\cite{30}".
-The model dedicated to the second task uses all or parts of the first model (i.e.
+According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
+The model dedicated to the second task uses some or all parts of the first model (i.e.
 keeps the same weights and architecture or a part of them) and is then retrained on data available for this task.
-The first model can either be implemented from scratch if enough data is available or can simply be downloaded from institutions that release large and trained models for similar tasks.\\
-It mainly provides three benefits \cite{30}:
+The first model can either be implemented from scratch if enough data is available or it can simply be downloaded from institutions that release large pretrained models for similar tasks.\\
+There are three major benefits \cite{30}:
 \begin{itemize}
 \item Higher start: the initial skill (before refining the model) on the source model is higher than it otherwise would be.
-\item Higher slope: the rate of improvement of skill during training of the source model is steeper than it otherwise would be.
+\item Higher slope: the rate of skill improvement during training of the source model is steeper than it otherwise would be.
 \item Higher asymptote: the converged skill of the trained model is better than it otherwise would be.

### 2019-12-21 Common/chapters/literature_review.tex
Julien and Johan: literature review

@@
 
-Literature review.
+Song et al.
+\cite{07} proposed a DCNN method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients which were split into a training, validation and test set.
+Their data processing approach kept T2W, DWI and ADC grayscale images only.
+After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65x65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels.
+Thanks to this method, the same lesion is visible in the same area over the three channels.
+This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter.
+Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e.
+by subtracting the mean before dividing by the standard deviation.
+The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20 to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
+Most of these processing techniques were reproductible, apart from the manual lesion contouring and labelling performed by a radiologist.
+Their model is a modified version of the well-known VGG16 model, including the addition of $1x1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function.
+The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set.
+The best results were obtained by using DWI images with the highest b-value only, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
+However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.
+
+Saifeng et al.
+\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achvieving the second best performance with an AUC of $0.84$.
+The AUC on the validation set reached $0.92$.
+Their data processing approach stacked different combination of the available sequences as the three channels: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W.
+The data augmentation process differs in that the images are rotated in 3D, each lesion being sliced at 7 different orientations.
+These 2-dimensional slices were then augmented using rotation, shearing and translation of 1px, resulting in 207144 training samples.
+Both validation and testing test were also augmented in the same manner.
+This whole process allows to include 3-dimensional information in 2-dimensional images.
+The method used ensemble learning which combined different models the reach the best performance possible.
+
+Mehrtash et al.
+\cite{01} used a different approach.
+First of all, the input was feeded to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans.
+Then, each of these feature extractors' outputs are merged into a common decision maker.
+Furthermore, 3-dimensional convolutions instead were performed.
+In fact, 3-dimensional patches centered on the lesion were cropped.
+Augmentation including translation and flipping was used in order to balance the dataset.
+Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist.
+Finally, this model achieved an AUC of $0.80$ on the PROSTATEx challenge.
+To make predictions, five different models were used, averaging the predictions of the four best models.

### 2020-01-29 Common/chapters/deep_learning.tex
Johan and Julien: Bibliography generated by Zotero

@@
 As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
-From the first attempts to understand the human brain until today, huge progres was made to be establish the basic components of modern neural networks.
+From the first attempts to understand the human brain until today, huge progress was made to be establish the basic components of modern neural networks.
 One could ask why deep learning took off recently if the theory was around for a long time.\\
@@
 The first one is "supervised learning".
-It includes learning algorithms whose training samples are associtted to their labels in order to find the optimal mapping between the input and the output.
+It includes learning algorithms whose training samples are associated to their labels in order to find the optimal mapping between the input and the output.
 The second one is "unsupervised learning".

### 2020-01-30 Common/chapters/data_processing.tex
Johan and Julien: Table of contents continued and new chapters added

@@
 \label{ch:data_processing}
-
-\section{DICOM file format}
-
-\subsection{Origin}
-
-The acronym DICOM stands for Digital Imaging and Communications in Medicine.
-Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks.
-For that reason, the ACR (American College of Radiology) and the NEMA (National Electrical Manufacturers Association) created a committee to build a standard.
-After two iterations with other names, DICOM was created in 1993.
-It standardized the representation of medical images and their transmission since it provided a network protocol built on top of TCP/IP.
-
-\subsection{Data format}
-
-DICOM files can be viewed as containers of attributes, also called tags.
-The values of the pixels themselves are stored under the "Pixel Data" tag.
-Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.
-
-Other useful information such as the patient name and ID is directly stored within the DICOM files.
-This approach aims at linking each image to a specific person and event in order not to mix them up.
-Each DICOM file can be seen as part of a bigger dataset.
-
-\subsection{Processing images}
-
-When manipulating DICOM files, multiple details must be taken into account.
-
-\subsubsection{Order}
-First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
-However, this order doesn’t match the real order of the images.
-In fact, the correct order is given by the "Instance Number" tags contained in the various files.
-Therefore, converted images must be sorted by instance number.
-
-\subsubsection{Data manipulation}
-\label{sec:dicom_data_manipulation}
-CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
-DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used.
-Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations.
-Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.
-
-The procedure depends on the tags "Window Width" and "Window Center" (one always come with the other).
-These are used to represent a range of values corresponding to the pixel values in the data.
-For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.
-
-If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
-\begin{itemize}
-	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
-	\item Samples per pixel: the number of channels for each pixel
-
-\end{itemize}
-
-\noindent Examples: 
-\begin{itemize}
-\item 1 bit, 1 sample: black and white
-\item 8 bits, 1 sample: grayscale
-\item 8 bits, 3 samples: RGB
-\item 16 bits, 1 sample: grayscale
-
-\end{itemize}
-
-\noindent If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one.
-To do this, two steps are required:
-
-\begin{enumerate}
-	\item \textbf{Apply the Hounsfield correction}\newline
-	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air.
-Provided that the rescale slope and the rescale intercept are included in the DICOM header, the correction is applied thanks to the following formula:
-\begin{equation}
-	HU = m * P + b
-\end{equation}
-	where~$m$ is the rescale slope,~$P$ the pixel value, ~$b$ the rescale intercept.
-
-	\item \textbf{Apply a linear transformation}\newline
-	The result of the first operation then goes through a linear transformation based on the following conditions:
-
-\begin{equation}\label{eq:dicom1}
-\textrm{if } (P \leq c - 0.5 - \frac{w-1}{2}) \textrm{, then }y = y_{min}
-\end{equation}
-
-\begin{equation}\label{eq:dicom2}
-\textrm{else if } (P > c - 0.5 + \frac{w-1}{2}) \textrm{, then }y = y_{max}
-\end{equation}
-
-\begin{equation}\label{eq:dicom3}
-\textrm{else } y = (\frac{P - (c - 0.5)}{w-1}) + 0.5) * (y_{max} - y_{min}) + y_{min}
-\end{equation}
-
-where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255).
-Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.
-
-\end{enumerate}
-
-\section{NIfTI file format}
-
-\subsubsection{Origin}
-The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
-The main problem of the latter was that it was lacking information about orientation in space.
-Therefore, the interpretation of stored data could be problematic and inconsistent.
-For instance, there was a real confusion to determine the left and right sides of brain images.
-Hence, the NIfTI file format was defined to overcome this major issue.
-
-\subsubsection{Data format}
-Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
-This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
-The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.
-
-\subsubsection{Overview of the header structure}
-With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
-“Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
-Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}
-
-\section{RAW and MHD file formats}
-
-Some datasets use a combination of RAW and MHD files.
-The latter contain metainformation about their corresponding RAW file(s) which contain the data.
-In most cases, each MHD file points to a unique RAW file whose name is the same as the MHD file name.
-A single RAW file can be used to represent three-dimensional data, i.e.
-the combination of multiple two-dimensional images.
-Libraries such as \mbox{SimpleITK} in Python allow to manipulate RAW images in an easy way.
-
-\section{Conversion}
-
-Medical data is often represented over 16 bits.
-However, exporting images to PNG require 8-bit images.
-To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
-
-\begin{algorithm}
-    \caption{16 to 8 bits conversion}
-    \label{alg:16_to_8_bits_conversion}
-    \begin{algorithmic}[1] % The number tells where the line numbering should start
-        \Procedure{16\_to\_8\_bits\_conversion}{$pixel\_array$}
-        		\State{$Pixel_{min} \gets$ minimal pixel value in $pixel\_array$}
-        		\State{$Pixel_{max} \gets$ maximal pixel value in $pixel\_array$}
-        		\For{$pixel\_value$ in $pixel\_array$}
-        			\State $pixel\_value \gets \frac{(pixel\_value - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
-        		\EndFor
-        		\State Modify object type to 8 bits
-        		\State Export $pixel\_array$ to PNG
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
 

### 2020-01-30 Common/chapters/deep_learning.tex
Johan and Julien: Table of contents continued and new chapters added

@@
 
-\subsubsection{Loss computation}
+\subsection{Loss computation}
 As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.

### 2020-01-30 Common/chapters/medical.tex
Johan and Julien: Table of contents continued and new chapters added

@@
 
-\section{Medical imaging}
+\section{Medical imaging file formats}
+\subsection{Types of medical imaging}
 \label{sec:medical_imaging}
@@
 \end{figure}
+
+\subsection{DICOM file format}
+
+\subsubsection{Origin}
+
+The acronym DICOM stands for Digital Imaging and Communications in Medicine.
+Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks.
+For that reason, the ACR (American College of Radiology) and the NEMA (National Electrical Manufacturers Association) created a committee to build a standard.
+After two iterations with other names, DICOM was created in 1993.
+It standardized the representation of medical images and their transmission since it provided a network protocol built on top of TCP/IP.
+
+\subsubsection{Data format}
+
+DICOM files can be viewed as containers of attributes, also called tags.
+The values of the pixels themselves are stored under the "Pixel Data" tag.
+Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.
+
+Other useful information such as the patient name and ID is directly stored within the DICOM files.
+This approach aims at linking each image to a specific person and event in order not to mix them up.
+Each DICOM file can be seen as part of a bigger dataset.
+
+\subsubsection{Processing images}
+
+When manipulating DICOM files, multiple details must be taken into account.
+
+\subsubsection{Order}
+First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
+However, this order doesn’t match the real order of the images.
+In fact, the correct order is given by the "Instance Number" tags contained in the various files.
+Therefore, converted images must be sorted by instance number.
+
+\subsubsection{Data manipulation}
+\label{sec:dicom_data_manipulation}
+CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
+DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used.
+Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations.
+Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.
+
+The procedure depends on the tags "Window Width" and "Window Center" (one always come with the other).
+These are used to represent a range of values corresponding to the pixel values in the data.
+For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.
+
+If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
+\begin{itemize}
+	\item Bits allocated: the number of bits used to represent a single pixel (value: 1 or a multiple of 8)
+	\item Samples per pixel: the number of channels for each pixel
+
+\end{itemize}
+
+\noindent Examples: 
+\begin{itemize}
+\item 1 bit, 1 sample: black and white
+\item 8 bits, 1 sample: grayscale
+\item 8 bits, 3 samples: RGB
+\item 16 bits, 1 sample: grayscale
+
+\end{itemize}
+
+\noindent If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one.
+To do this, two steps are required:
+
+\begin{enumerate}
+	\item \textbf{Apply the Hounsfield correction}\newline
+	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air.
+Provided that the rescale slope and the rescale intercept are included in the DICOM header, the correction is applied thanks to the following formula:
+\begin{equation}
+	HU = m * P + b
+\end{equation}
+	where~$m$ is the rescale slope,~$P$ the pixel value, ~$b$ the rescale intercept.
+
+	\item \textbf{Apply a linear transformation}\newline
+	The result of the first operation then goes through a linear transformation based on the following conditions:
+
+\begin{equation}\label{eq:dicom1}
+\textrm{if } (P \leq c - 0.5 - \frac{w-1}{2}) \textrm{, then }y = y_{min}
+\end{equation}
+
+\begin{equation}\label{eq:dicom2}
+\textrm{else if } (P > c - 0.5 + \frac{w-1}{2}) \textrm{, then }y = y_{max}
+\end{equation}
+
+\begin{equation}\label{eq:dicom3}
+\textrm{else } y = (\frac{P - (c - 0.5)}{w-1}) + 0.5) * (y_{max} - y_{min}) + y_{min}
+\end{equation}
+
+where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255).
+Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.
+
+\end{enumerate}
+
+\subsection{NIfTI file format}
+
+\subsubsection{Origin}
+The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
+The main problem of the latter was that it was lacking information about orientation in space.
+Therefore, the interpretation of stored data could be problematic and inconsistent.
+For instance, there was a real confusion to determine the left and right sides of brain images.
+Hence, the NIfTI file format was defined to overcome this major issue.
+
+\subsubsection{Data format}
+Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
+This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
+The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate”  algorithm.
+
+\subsubsection{Overview of the header structure}
+With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+“Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
+Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}
+
+\subsection{RAW and MHD file formats}
+
+Some datasets use a combination of RAW and MHD files.
+The latter contain metainformation about their corresponding RAW file(s) which contain the data.
+In most cases, each MHD file points to a unique RAW file whose name is the same as the MHD file name.
+A single RAW file can be used to represent three-dimensional data, i.e.
+the combination of multiple two-dimensional images.
+Libraries such as \mbox{SimpleITK} in Python allow to manipulate RAW images in an easy way.
+
+\section{Conversion}
+
+Medical data is often represented over 16 bits.
+However, exporting images to PNG require 8-bit images.
+To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
+
+\begin{algorithm}
+    \caption{16 to 8 bits conversion}
+    \label{alg:16_to_8_bits_conversion}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{16\_to\_8\_bits\_conversion}{$pixel\_array$}
+        		\State{$Pixel_{min} \gets$ minimal pixel value in $pixel\_array$}
+        		\State{$Pixel_{max} \gets$ maximal pixel value in $pixel\_array$}
+        		\For{$pixel\_value$ in $pixel\_array$}
+        			\State $pixel\_value \gets \frac{(pixel\_value - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
+        		\EndFor
+        		\State Modify object type to 8 bits
+        		\State Export $pixel\_array$ to PNG
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}

### 2020-01-30 Common/chapters/paper_reproduction.tex
Johan and Julien: Table of contents continued and new chapters added

% !TEX root = ../main.tex

\chapter{Paper reproduction}
\label{ch:paper_reproduction}

What we do: improve cancer detection/classification on medical imaging
How we do it: transfer learning

\section{Goal}

\section{Process overview}

Find data -> Process data -> Verify data processing (visualization, comparing NumPy arrays and PNGs, etc.) -> Train

\section{PROSTATEx: Data processing}
\subsection{Dataset description}
\subsection{From DICOM to NumPy arrays}
\subsection{From NumPy arrays to augmented stacked images}
\subsection{From NumPy arrays to augmented non-stacked images}
\subsection{Data processing verification}
\subsubsection{Cropping verification using red dots}
\subsubsection{Alignment visualization}

\section{Training the neural network}
\subsection{Architecture}
\subsection{Criterion to save the best model}
\subsection{Script options}
\subsection{Experimental setup}
\subsection{Training verification}
\subsubsection{Gradient flow visualization}

\section{Results}

\section{Discussion}

### 2020-01-30 Common/chapters/transfer_learning.tex
Johan and Julien: Table of contents continued and new chapters added

% !TEX root = ../main.tex

\chapter{Improving performance using transfer learning}
\label{ch:transfer_learning}

What we do: improve cancer detection/classification on medical imaging
How we do it: transfer learning

\section{Goal}

\section{Process overview}

\section{Data processing}
\subsection{PROSTATEx}
\subsubsection{From DICOM to augmented DWI NumPy arrays}
Keep only the DWI because white lesion, crop the lesion etc

\subsection{LungCTChallenge}
\subsubsection{Dataset description}
\subsubsection{From DICOM to augmented NumPy arrays}

\subsection{Kaggle Brain}
\subsubsection{Dataset description}
\subsubsection{Ground truth creation}
\subsubsection{From PNG to NumPy arrays}
Manually create the csv file from scratch, crop the lesion etc

\subsection{Verification}
\subsubsection{Visual checking}

\section{Transfer learning implementation}
\subsection{Architecture generalities}
\subsection{Script options}
\subsection{Visualization of the usefulness of the features learned on each dataset}
\subsection{Experimental setup}
\section{Results}

\section{Discussion}

### 2020-01-30 Common/chapters/paper_reproduction.tex
Johan and Julien: paper reproduction chapter started

@@
 \label{ch:paper_reproduction}
+This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI"\cite{07} presented in chapter \ref{ch:literature_review}.
+It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a new transfer learning method as a way to improve the classification using other body parts and to overcome the lack of available data(see \ref{ch:transfer_learning}).\\
 
-What we do: improve cancer detection/classification on medical imaging
-How we do it: transfer learning
-
-\section{Goal}
+Furthermore, since the article does not provide the results their model got on the PROSTATEx challenge, the reproduction of the experiment can act like a benchmark to evaluate the generalization of the model to unseen data.
 
 \section{Process overview}
+Schematically, the entire experiment process can be represented as shown in figure \ref{fig:paper_reproduction_process}.
+The PROSTATEx dataset is composed of samples whose clinical significance is provided (labeled data) and samples without this information (unlabeled data).
+The latter are used for the challenge, which consists in predicting the probability of patients lesions to be malignant (class 1) or benign (class 0).
+In both cases, the data is processed before been used (see \ref{prostatex_data_processing} for details).
+After the processing, the labeled data are split into training, validation and test sets using respectively 80\%, 10\% and 10\% of all available data for each set.
+The training set is then given in batches as input to the neural network, which update its weight consequently.
+At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}).
+If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.
+When the training phase reaches the defined number of epochs, the model is then tested on the test set.
+This same model is also used to take part into the PROSTATEx challenge in order to output predictions into a .csv file for each patient and findings.
 
-Find data -> Process data -> Verify data processing (visualization, comparing NumPy arrays and PNGs, etc.) -> Train
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_process.png}
+\caption{Paper reproduction experiment}
+\label{fig:paper_reproduction_process}
+\end{figure}
 
 \section{PROSTATEx: Data processing}
+\label{prostatex_data_processing}
 \subsection{Dataset description}
+\label{prostatex_dataset_description}
 \subsection{From DICOM to NumPy arrays}
@@
 \subsection{Criterion to save the best model}
+\subsection{Tensorboard}
+\label{paper_tensorboard}
 \subsection{Script options}

### 2020-01-31 Common/chapters/deep_learning.tex
Johan and Julien: paper reproduction chapter continued

@@
 \label{ch:deep_learning}
-
+This chapter provides the theoretical foundation in deep learning for the understanding of the rest of the work.
+It starts with the historical story of the deep learning before describing what is a neural network.
+Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation, is explained.
+This part is followed with another devoted to a special type of neural networks, the "convolutional neural networks".
+They are used in many computer vision problems due to their great performance for these tasks.
+Finally, the concept of "transfer learning" is discussed since the entire chapter \ref{ch:transfer_learning} relies on it.
 \section{Introduction to deep learning}
@@
 Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc.
-The presence of deep learning techniques in these use cases clearly attests that this technology is a breakthrough.
+The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology.
 Furthermore, since this field has recently gained interest (see section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

### 2020-01-31 Common/chapters/paper_reproduction.tex
Johan and Julien: paper reproduction chapter continued

@@
 \label{ch:paper_reproduction}
-This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI"\cite{07} presented in chapter \ref{ch:literature_review}.
-It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a new transfer learning method as a way to improve the classification using other body parts and to overcome the lack of available data(see \ref{ch:transfer_learning}).\\
-
-Furthermore, since the article does not provide the results their model got on the PROSTATEx challenge, the reproduction of the experiment can act like a benchmark to evaluate the generalization of the model to unseen data.
+This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data (see chapter \ref{ch:transfer_learning}).\\
+Song et al.
+\cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is one of the biggest available dataset for prostate cancer classification.
+The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1).
+This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis (CAD) system for prostate cancer detection.
+Moreover, it also provides results about the algorithm performance, which can be used as a benchmark to compare with the results of the reproduction of the experiment.
+However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set, the reproduction of the experiment will fill that gap.\\
+This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole.
+Then, each part or theoretical notion of the experiment is deeply described.
+The structure of the dataset is detailed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented.
+After the processing of the data, the training phase is deepened and all hyperparameters, options and implementation choices are given to ensure the reproductibility of the experiment.
+This part is followed by the presentation of the raw results, which is itself followed by their analysis in the "discussion" section.
 
@@
 When the training phase reaches the defined number of epochs, the model is then tested on the test set.
-This same model is also used to take part into the PROSTATEx challenge in order to output predictions into a .csv file for each patient and findings.
+This same model is also used to take part in the PROSTATEx challenge to output predictions into a .csv file for each patient and findings.
 

### 2020-01-31 Common/chapters/paper_reproduction.tex
Johan and Julien: paper reproduction -> architecture done

@@
 \subsection{Architecture}
+The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
+This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset of 14 millions images belonging to 1000 classes.
+The figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch Framework.\\
+It is first composed by three convolution-dropout-max-pooling blocks followed by three fully connected-dropout blocks.
+Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit activation function.
+The same principle is used for the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
+The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax  function for classification.\\
+In comparison to the original VGG, this model keeps the small filter size of 3x3 or 1x1,  also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
+It differs from the traditional VGG in that it makes use of a smaller number of layer since the task is simpler than the original one, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/model_paper_manual.png}
+\caption{Model architecture with the corresponding PyTorch code}
+\label{fig:paper_model}
+\end{figure}
 \subsection{Criterion to save the best model}

### 2020-01-31 Common/chapters/paper_reproduction.tex
Johan and Julien: Tensorboard section first draft

@@
 \subsection{Tensorboard}
+During the experiment the following metrics are registered: loss, accuracy, precision, recall, F1-score, specificity and AUC.
+These metrics are computed separately on the training and on the validation sets at the end of each epoch.
+They are then stored across all epochs and plotted on the same figure thanks to Tensorboard at the end of the process.
+Subsequently, this is also the case for the metrics measured on the test set.\\
+"Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+\begin{itemize}
+\item Tracking and visualizing metrics.
+\item Visualizing the model graph (ops and layers).
+\item Viewing histograms of weights, biases, or other tensors as they change over time.
+\item Projecting embeddings to a lower dimensional space.
+\item Displaying images, text, and audio data"\cite{39}.
+\end{itemize}
+In this experiment, Tensorboard is mainly used to plot the Matplotlib figures of the model performance.
+In addition to this, written reports regarding which model was the best and the results it reached are also added on the dashboard.
 \label{paper_tensorboard}

### 2020-02-01 Common/chapters/paper_reproduction.tex
Johan and Julien: section script options done

@@
 \subsection{Script options}
+Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
+In order to achieve this goal, the Python module "argparse" is extremely useful.
+The latter automatically generates help/usage messages and displays errors when the argument given by the user are invalid.
+The table shows all options accepted by the script.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|llll|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
+\textbf{--trainingSet1}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--validationSet1}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
+\textbf{--nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
+\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
+\textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
+\textbf{--cuda-device}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
+\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
+\textbf{--optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
+\end{tabular}%
+}
+\end{table}
+
 \subsection{Experimental setup}

### 2020-02-01 Common/chapters/paper_reproduction.tex
Johan and Julien: experimental setup done

@@
 \subsection{Experimental setup}
+The best performing model was implemented with PyTorch using the ADAM optimizer to update its weights, a learning rate of 1e-8 and the cross-entropy as loss function.
+The data was organized in batches of 128 samples and the experiment ran for 1300 epochs.
+The best model was selected with respect to the area under curve.
+Regarding the weights initialization, the best model from the roulette that maximized the F1-score was chosen as initial model.\\
+The model was trained on a Nvidia GeForce GTX Titan X graphic card during approximately 15 hours.
+
 \subsection{Training verification}

### 2020-02-01 Common/chapters/paper_reproduction.tex
Johan and Julien: gradient flow section first draft

@@
 This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
-It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data (see chapter \ref{ch:transfer_learning}).\\
+It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data in the medical field (see chapter \ref{ch:transfer_learning}).\\
 Song et al.
@@
 }
+\caption{Complete list of script options}
+\label{fig:paper_reproduction_options}
 \end{table}
@@
 \subsection{Training verification}
+In deep learning, two main problems can occur during training.
+The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero.
+Consequently, the weight update is also extremely small, which makes the neural network hard or impossible to train.
+The second problem, the "exploding gradient" is the opposite.
+Concretely, in this case, large gradients are computed, which results in huge weight updates that can even reach "NaN" values.
+This makes the model unstable and unable to learn from training data.
+\\
+As a result, it is important to analyse the gradient propagation across the network to ensure not been face to such problems.
+To achieve this goal, a visualization of all gradients of the network is implemented.
+This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
 \subsubsection{Gradient flow visualization}
+The visualization displays the gradient flow through the entire neural network.
+It shows the name of each layer on the x-axis and its average gradient value on the y-axis (recall: for each layer the gradient is a matrix).
+Furthermore, it also gives information about the max value of the gradient and indicates in black if one of them is equal to zero.\\
+The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6.
+From the output layer (called "last\_layer0.weight" on the graph) to the first layer the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
 
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/gradient_flow.png}
+\caption{Gradient flow at epoch 0, batch 6 of the experiment}
+\label{fig:gradient_flow}
+\end{figure}
 \section{Results}

### 2020-02-01 Common/chapters/paper_reproduction.tex
Julien and Johan: data processing for "paper reproduction" and "transfer learning" chapters

@@
 The latter are used for the challenge, which consists in predicting the probability of patients lesions to be malignant (class 1) or benign (class 0).
-In both cases, the data is processed before been used (see \ref{prostatex_data_processing} for details).
+In both cases, the data is processed before been used (see \ref{sec:prostatex_data_processing} for details).
 After the processing, the labeled data are split into training, validation and test sets using respectively 80\%, 10\% and 10\% of all available data for each set.
@@
 \section{PROSTATEx: Data processing}
-\label{prostatex_data_processing}
+\label{sec:prostatex_data_processing}
 \subsection{Dataset description}
-\label{prostatex_dataset_description}
+\label{sec:prostatex_dataset_description}
+The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients.
+Images were taken under the sagittal, transverse and coronal planes.
+
+The PROSTATEx dataset comes with two CSV files for the training set.
+The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance.
+Multiple findings can be associated with the same patient (ProxID).
+The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding.
+Important labels are "ProxID" (patient ID), "fid" (finding ID $\in [1,\infty]$), "ClinSig" (clinical significance, TRUE or FALSE), "DCMSerNumber" (digit before the dash in the folder name containing DICOM files), "ijk" (position of the lesion: slice number $k$ at coordinates $(i,j)$, $i,j,k \in [0,\infty]$) and "VoxelSpacing" (3-dimensional vector representing the correspondance between a pixel and the space it occupies in the real world).
+Both CSV files are complementary to each other.
+Regarding the challenge patients, two analog CSV files are provided.
+The only difference is the absence of clinical significance.
+Finally, images are in the DICOM file format (see section \ref{sec:DICOM}).
+
+\subsection{Methodology}
+The methodology described in this section is the closest replication possible of the authors' data processing methodology.
+Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
+Furthermore, the authors didn't mention any participation in the official PROSTATEx challenge.
+In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented.
+Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$).
+They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets.
+Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60x$) in order to create balanced training and validation sets.
+
 \subsection{From DICOM to NumPy arrays}
+\label{sec:DICOMtoNumPy}
+Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
+Moreover, only images showing the prostate under the transverse plane were used.
+The reason for this is that the tumors are much more visible under this perspective.
+The first step consisted in converting DICOM files.
+Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays.
+The right slices were found thanks to the two CSV files.
+Important information such as the patient ID, the sequence, the lesion location and the voxel spacing were used as file output names, which allows to use these files independently for the next steps (i.e.
+without relying on the CSV files).
+
+\begin{algorithm}
+    \caption{PROSTATEx preprocessing}
+    \label{alg:PROSTATEx_preprocessing}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{main}{$dataset\_folder, findings\_CSV, slices\_CSV, output\_folder$}
+        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
+        		\State $findings \gets read\_CSV(findings\_CSV)$ \Comment{ProstateX-Findings-Train.csv}
+        		\State $slices \gets read\_CSV(slices\_CSV)$ \Comment{ProstateX-Images-Train.csv}
+            \State $meta \gets merge(findings, slices)$\Comment{Both CSV files are complementary to each other.}\\
+            \For{$row$ in $meta$}
+            		\State $patient\_id \gets row["ProxID"]$
+            		\State $finding\_id \gets row["fid"]$
+            		\State $mri\_type\_number \gets row["DCMSerNumber"]$
+            		\State $clinical\_significance \gets row["ClinSig"]$
+            		\State $img\_i, img\_j, img\_k \gets row["ijk"]$
+            		\State $slice\_number \gets img\_k + 1$ \Comment{CSV indexing in $[0,\infty]$, DICOM in $[1,\infty]$}\\
+                \For{$visit\ in\ patient\_id$'s folder}
+                		\For{$mri\_type$ in $visit$}
+                			\If {$mri\_type$ starts with "$mri\_type\_number$-"}
+                				\For{$dicom\_file$ in $mri\_type$}
+                					\If {$slice\_number == dicom\_file.InstanceNumber$}
+                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
+                						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
+                					\EndIf
+                				\EndFor
+                			\EndIf
+            			\EndFor
+            		\EndFor
+            \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
 \subsection{From NumPy arrays to augmented stacked images}
+\label{sec:numpyToAugmentedStacked}
+This step can be split into two substeps: aligning and stacking images before augmenting them.
+Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution.
+In fact, T2-weighted images have a much higher resolution than the DWI and ADC images (in this dataset at least).
+Without this operation, a single pixel on a DWI image would have covered a lot more tissue than a pixel of the corresponding T2-image.
+To perform the alignment, the lesion was localized on the three sequences than to the CSV file.
+Then, the goal was to crop a large patch which contained the same amount of tissue on the three images.
+This patch had to be centered on the lesion to ease the augmentation process.
+So, a fixed patch size was defined for the T2-image since its resolution was the highest.
+Then, using the voxel spacing information, the patch size required to cover the same amount of tissue was computed for the two other sequences (DWI, ADC).
+Finally, the three images were cropped, resized to the same resolution and stacked.
+Stacking images consists in putting each single image (grayscale) into a single array, as three different channels.
+At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel.
+This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible on the three images.
+
+An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e.
+
+\begin{equation}
+\label{eq:normalization}
+	Pixel_{{i,j}_{normalized}} = \frac{Pixel_{i,j} - \mu}{\sigma}
+\end{equation}
+
+\noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient.
+According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately help the final diagnosis.
+In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array.
+Then, the mean and standard devation of the array were computed.
+These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient.
+Same for this patient's DWI and ADC images.
+The same process was applied to all patients separately.
+
+Finally, the dataset was augmented in the same way as the authors.
+Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel).
+These techniques alone allowed to create a sufficient amount of data.
+Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not made use of.
+Images were exported as NumPy arrays.
+
 \subsection{From NumPy arrays to augmented non-stacked images}
+Another way of processing and using the data was experimented with.
+Instead of stacking the images, each image was cropped and augmented separately.
+At the end, each image was given as separate input to the neural network.
+
 \subsection{Data processing verification}
+\label{sec:dataProcessingVerification}
+
 \subsubsection{Cropping verification using red dots}
-\subsubsection{Alignment visualization}
+In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
+First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file.
+This full image was exported.
+Then, images were augmented in the exact same manner as described in section \ref{sec:numpyToAugmentedStacked}.
+The expected result satisfies the following properties:
+\begin{itemize}
+	\item The red dot is localized in the center of the cropped and augmented image.
+	\item The three cropped and augmented images within a stacked image contain the exact same tissue. 
+	\item The three cropped and augmented images within a stacked image are not rotated/shifted/flipped differently. 
+	\item A lesion is visible to the naked eye. 
+\end{itemize}
+
+\noindent Figure \ref{fig:reddot} shows an example generated during this test.
+While beeing exactly the same for the 3 images, the red dot looks bigger on the images whose resolution was smaller since it was drawn before resampling the images.
+These images come from the same exam performed on the same patient on the same day.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/test_red_dot.png}
+\caption{Red dot test - Patient 0082, Finding ID 1\\From left to right: T2, DWI, ADC.}
+\label{fig:reddot}
+\end{figure}
+
+\subsubsection{Alignment verification}
+In addition to the previous test, an additional one aiming at checking the alignment of the images was created.
+This test ensures that the three images show the exact same content at the same spot over the three channels.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/alignment.png}
+\caption{Alignment - Patient 0242 - Finding ID 1}
+\label{fig:alignment}
+\end{figure}
 

### 2020-02-01 Common/chapters/transfer_learning.tex
Johan and Julien: paper reproduction chapter started
Julien and Johan: data processing for "paper reproduction" and "transfer learning" chapters

@@
 \subsection{PROSTATEx}
-\subsubsection{From DICOM to augmented DWI NumPy arrays}
-Keep only the DWI because white lesion, crop the lesion etc
+As the other datasets used in this work don't contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
+DWI images were selected because lesions are white on this sequence, as well as on CT scans.
+Hence, choosing DWI images ensured a consistent representation of tumors from one dataset to the other, which helps the cancer detection.
+
+The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from the alignment and the stacking.
+A T2-weighted image was still used to align, crop and resize the corresponding DWI image but was not exported at the end, resulting in a single channel image (NumPy array).
+
+Ideally, transfer learning should rely on multi-sequence MRIs only.
+In fact, the extra information provided by different sequences increases the performance a lot.
+However, publicly available datasets either don't contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
 
@@
 \subsubsection{Dataset description}
+Lung CT Challenge is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
+Each patient can have 1 or multiple findings.
+Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
+For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.
+
+Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images.
+In order to facilitate the handling of information in the code, two CSV files were manually created: \mbox{\textit{TestSet.csv}} and \textit{CalibrationSet.csv}.
+Contrary to PROSTATEx, more than two labels were used in this dataset.
+Both "malignant" and "Primary lung cancer" were considered as positive, whereas "benign" and "Benign nodule" were treated as negative.
+A third label called "Suspicious malignant nodule" appeared two times.
+Since the diagnosis was not clearly defined for those images, they were not included in the training data to avoid any noise.
+
 \subsubsection{From DICOM to augmented NumPy arrays}
+As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
+
+\begin{algorithm}
+    \caption{Lung CT Challenge preprocessing}
+    \label{alg:LungCTChallenge_preprocessing}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{main}{$dataset\_folder, train\_CSV, test\_CSV, output\_folder$}
+        		\State Create output directories: $"output\_folder/True", "output\_folder/False"$\\
+        		\State $csv\_training \gets read\_CSV(train\_CSV)$ \Comment{CalibrationSet.csv}
+        		\State $csv\_test \gets read\_CSV(test\_CSV)$ \Comment{TestSet.csv}
+            \State $csv\_concatenated \gets concat(csv\_training, csv\_test)$\Comment{Both CSV files contain similar information about different patients.}\\
+            \For{$row$ in $csv\_concatenated$}
+            		\State $patient\_id \gets row["Scan\ Number"]$
+            		\State $slice\_number \gets row["Nodule\ Center\ Image"]$ \Comment{Value in $[1,\infty]$}
+            		\State $finding\_id \gets row["Nodule\ Number"]$
+            		\State $clinical\_significance \gets row["Diagnosis"]$\\
+            		\If{$clinical\_significance$ == $"malignant"$ or $"Primary\ lung\ cancer"$}
+            			\State{$clinical\_significance \gets True$}
+            		\ElsIf{$clinical\_significance$ == $"benign"$ or $"Benign\ nodule"$}
+            			\State{$clinical\_significance \gets False$}
+            		\ElsIf {$clinical\_significance$ == $"Suspicious\ malignant\ nodule"$}
+            			\State{Continue}
+            		\EndIf\\
+                \For{$visit\ in\ patient\_id$'s folder}
+                		\For{$mri\_type$ in $visit$}
+                			\For{$dicom\_file$ in $mri\_type$}
+                				\If {$slice\_number == dicom\_file.InstanceNumber$}
+                					\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
+                					\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
+                				\EndIf
+                			\EndFor
+            			\EndFor
+            		\EndFor
+            \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+The augmentation process is close to the one applied to PROSTATEx.
+An advantage Lung CT Challenge has it that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless.
+Nevertheless, lung tumors are particular.
+What differentiates them from the others is their look.
+In fact, lung tumors usually look like spider webs with a central nodule and branches coming out of it, whereas brain and prostate tumors tend to have well-defined boundaries.
+Therefore, a larger patch size was chosen in order to keep the extra information provided by the branches in the peripheral tissue.
+Furthermore, a perfect cropping method would only crop the interesting part of the tissue and a margin around it.
+However, tumor size varies a lot.
+It can either be extremely small (a dot) or large (massive nodule with long branches).
+Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors.
+Naturally, there were few outliers whose size was a lot larger than the chosen patch size.
+If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case.
+For this reason, these large tumors were treated as outliers.
+Lastly, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives).
+Therefore, each class was augmented the same amount of times (240x).
+
+Images were normalized with respect to equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
 
@@
 \subsubsection{Dataset description}
+This dataset is publicly available on Kaggle.
+All images come from single-sequence MRIs.
+It is composed of two folders: one containing images with cancerous tumors and another one containing images either without tumor or with a non-cancerous tumor.
+Images whose quality was extremely low (resolution, picture of a screen, artefacts, etc.) were manually removed.
+In other cases, the image quality was good but disturbing objects such as left/right markers, arrows pointing the tumor, etc.
+were visible.
+These objects were manually removed using the object removal tool in Photoshop.
+File names and file formats were heterogeneous.
+
 \subsubsection{Ground truth creation}
-\subsubsection{From PNG to NumPy arrays}
-Manually create the csv file from scratch, crop the lesion etc
+\label{sec:braingroundtruth}
+The particularity of this dataset is that no information apart from the clinical significance is provided.
+Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied on this one.
+
+The first step was to standardize the file names and formats.
+The chosen convention is "$[yes|no]{index}.jpg$".
+The vast majority of the files were in the JPG file format, so the few PNG images were converted.
+Then, a CSV file containing the following columns was created: "PatientID", "Nodule Center Position", "fid", "Diagnosis".
+The "PatientID" is the filename without extension that was set during the first step (for example "$yes99$" or "$no3$").
+The "Nodule Center Position" is the pixel coordinate of the lesion.
+Fortunately, brain tumors are clearly visible on brain MRIs, which eased the process.
+To find the coordinates, every single image was opened at its original size in an image visualization software.
+A rectangle selection tool was used to draw a rectangle from the top left corner of the image (because the top left pixel corresponds to the pixel at the coordinates $(0,0)$ in NumPy arrays) to the center of the lesion.
+The pixel width and height of this rectangle gave the "i,j" coordinates of the lesion.
+These coordinates were reported manually.
+Then, the "fid" corresponds to the Finding ID.
+This parameter would be more pertinent if the same "PatientID" could have multiple findings, i.e.
+if the same patient could have multiple tumors.
+Since this information is made use of during the image processing of the other datasets, it was also added here.
+In this particular case, the "fid" was always set to "1".
+Finally, the "Diagnosis" was defined as "False" or "True" according to the folder which the image was in.
+
+\subsubsection{From JPG to NumPy arrays}
+Images had to be converted to single-channel NumPy arrays (grayscale) so that the values could be normalized further on.
+First, a mere conversion involving no content modification apart from going from three to one channel was performed.
+Second, images were normalized according to equation \ref{eq:normalization}.
+The mean and standard deviation were computed over the whole dataset.
+Both classes were augmented by a common factor (200x) since both class contain a relatively close number number of instances (133 positives and 96 negatives).
+Similarly to Lung CT Challenge, tumor sizes vary substantially.
+The same methodology was applied regarding the patch size, i.e.
+choosing a size which fits the majority of large tumors perfectly.
+Even though the images all show the head under the same perspective (same height/width ratio, visible skull, black margin around the skull), images were probably collected from different sources, which explains the heterogeneity of resolutions.
+As a consequence, the patch size could not be a fixed number of pixels for all images because the cropped area would be the entire head on one image and a small portion of the tumor on another.
+The trick consisted in finding the horizontal and vertical portions of the image occupied by the tumor.
+In fact, cropping an area centered on the lesion whose height was $\frac{2}{7}$ of the image and whose width was $\frac{1}{4}$ gave good results.
 
@@
 \subsubsection{Visual checking}
+Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see section \ref{sec:braingroundtruth}).
+In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset.
+They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside.
+Like PROSTATEx (see section \ref{sec:dataProcessingVerification}), control images with red dots were generated for both datasets.
 

### 2020-02-02 Common/chapters/paper_reproduction.tex
Johan and Julien: results and discussion beginning

@@
 \section{Results}
+Figures \ref{fig:paper_reprodution_results_part_1}, \ref{fig:paper_reprodution_results_part_2} and \ref{fig:paper_reprodution_results_part_3} show the different metrics registered on the training and validation sets during the training phase.
+The best model was chosen at epoch ...
+with an accuracy of ...
+and an AUC of ...
+(the condition for a model to be chosen as best is that the current AUC should be bigger than the auc of the previous best model and the same principle apply to the accuracy).\\
+This best model was also used to take part in the ProstateX challenge. The res
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, height=\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part1.png}
+\caption{Accuracy, AUC and loss on the training and validation set}
+\label{fig:paper_reprodution_results_part_1}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, height=\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part2.png}
+\caption{F1-score, precision and recall on the training and validation set}
+\label{fig:paper_reprodution_results_part_2}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part3.png}
+\caption{Specificity on the training and validation set}
+\label{fig:paper_reprodution_results_part_3}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge.png}
+\caption{Score achieved in the PROSTATEx challenge}
+\label{fig:paper_reprodution_results_challenge}
+\end{figure}
 
 \section{Discussion}
+Thanks to all curves in the "results", many information regarding the model performance can be inferred.
+Indeed, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
+First of all, when observing the loss plot, what is noticeable is that the training and validation loss have the same behaviour until the epoch 600 approximately.
+This is a clear sign of good fit of the model on the data.
+All features learned by the model on the training data are also applicable on the validation set.
+This situation is the optimal one.
+After this phase, the model slightly start to overfit.
+In fact, the training loss continue to linearly decrease whereas the validation loss decreases less quickly until reaching  \\
+
+As the goal of the experiment was to reproduce the results of the paper,

### 2020-02-02 Common/chapters/paper_reproduction.tex
Julien and Johan: typos

@@
 \label{ch:paper_reproduction}
-This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
 It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data in the medical field (see chapter \ref{ch:transfer_learning}).\\
@@
 This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole.
-Then, each part or theoretical notion of the experiment is deeply described.
-The structure of the dataset is detailed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented.
-After the processing of the data, the training phase is deepened and all hyperparameters, options and implementation choices are given to ensure the reproductibility of the experiment.
+Then, each part or theoretical notion of the experiment is described in detail.
+The structure of the dataset is exposed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented.
+After the processing of the data, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
 This part is followed by the presentation of the raw results, which is itself followed by their analysis in the "discussion" section.
@@
 \section{Process overview}
-Schematically, the entire experiment process can be represented as shown in figure \ref{fig:paper_reproduction_process}.
+Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}.
 The PROSTATEx dataset is composed of samples whose clinical significance is provided (labeled data) and samples without this information (unlabeled data).
@@
 If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.
-When the training phase reaches the defined number of epochs, the model is then tested on the test set.
-This same model is also used to take part in the PROSTATEx challenge to output predictions into a .csv file for each patient and findings.
+When the training phase reaches the defined number of epochs, the model is tested on the test set.
+The same model is also used to generated predictions for each patient and finding of the PROSTATEx challenge dataset in order to take part in it.
 
@@
 
-\subsubsection{Alignment verification}
+\subsubsection{Alignment}
 In addition to the previous test, an additional one aiming at checking the alignment of the images was created.
@@
 This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset of 14 millions images belonging to 1000 classes.
-The figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch Framework.\\
-It is first composed by three convolution-dropout-max-pooling blocks followed by three fully connected-dropout blocks.
-Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit activation function.
-The same principle is used for the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
+Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
+It is first composed by three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks.
+Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function.
+The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
 The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax  function for classification.\\
 In comparison to the original VGG, this model keeps the small filter size of 3x3 or 1x1, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
-It differs from the traditional VGG in that it makes use of a smaller number of layer since the task is simpler than the original one, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
+It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one.
+Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
 \begin{figure}[!h]
@@
 \subsection{Tensorboard}
-During the experiment the following metrics are registered: loss, accuracy, precision, recall, F1-score, specificity and AUC.
+During the experiment the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC.
 These metrics are computed separately on the training and on the validation sets at the end of each epoch.
@@
 In this experiment, Tensorboard is mainly used to plot the Matplotlib figures of the model performance.
-In addition to this, written reports regarding which model was the best and the results it reached are also added on the dashboard.
+In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
 \label{paper_tensorboard}
@@
 The latter automatically generates help/usage messages and displays errors when the argument given by the user are invalid.
-The table shows all options accepted by the script.
+Table \ref{fig:paper_reproduction_options} shows all options accepted by the script.
 
@@
 \subsection{Experimental setup}
-The best performing model was implemented with PyTorch using the ADAM optimizer to update its weights, a learning rate of 1e-8 and the cross-entropy as loss function.
+The best performing model was implemented with PyTorch using the Adam optimizer to update its weights, a learning rate of 1e-8 and the cross-entropy as loss function.
 The data was organized in batches of 128 samples and the experiment ran for 1300 epochs.
 The best model was selected with respect to the area under curve.
-Regarding the weights initialization, the best model from the roulette that maximized the F1-score was chosen as initial model.\\
-The model was trained on a Nvidia GeForce GTX Titan X graphic card during approximately 15 hours.
+Regarding the weight initialization, the best model from the roulette that maximized the F1-score was chosen as initial model.\\
+The model was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 15 hours.
 
@@
 \\
-As a result, it is important to analyse the gradient propagation across the network to ensure not been face to such problems.
-To achieve this goal, a visualization of all gradients of the network is implemented.
+As a result, it is important to analyze the gradient propagation across the network to ensure that it is not facing such problems.
+To achieve this goal, a visualization tool to display the gradients of the entire network was implemented.
 This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
@@
 The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6.
-From the output layer (called "last\_layer0.weight" on the graph) to the first layer the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
+From the output layer (called "last\_layer0.weight" on the graph) to the first layer, the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
 
@@
 and an AUC of ...
-(the condition for a model to be chosen as best is that the current AUC should be bigger than the auc of the previous best model and the same principle apply to the accuracy).\\
-This best model was also used to take part in the ProstateX challenge. The res
+(the condition for a model to be chosen as best is that the current AUC should be higher than the AUC of the previous best model, same for the accuracy).\\
+This best model was also used to take part in the ProstateX challenge. The res XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
 
@@
 \section{Discussion}
-Thanks to all curves in the "results", many information regarding the model performance can be inferred.
-Indeed, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
-First of all, when observing the loss plot, what is noticeable is that the training and validation loss have the same behaviour until the epoch 600 approximately.
+Thanks to all curves in the "results", a lot of information regarding the model performance can be inferred.
+In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
+First of all, when observing the loss plot, what is noticeable is that the training and validation losses follow the same behaviour until the epoch 600 approximately.
 This is a clear sign of good fit of the model on the data.
-All features learned by the model on the training data are also applicable on the validation set.
+All features learned by the model on the training data are also applicable to the validation set.
 This situation is the optimal one.
-After this phase, the model slightly start to overfit.
-In fact, the training loss continue to linearly decrease whereas the validation loss decreases less quickly until reaching  \\
+After this phase, the model slightly started to overfit.
+In fact, the training loss continues to linearly decrease whereas the validation loss decreases less quickly until reaching  XXXXXXXXXXXXXXXXXXXXXXXX\\
 

### 2020-02-02 Common/chapters/transfer_learning.tex
Julien and Johan: continue transfer learning chapter

@@
 \section{Goal}
+Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.
+This technique makes sense when the quantity of data available is not sufficient and when the data of both datasets share common features.
+Both conditions are fulfilled in the case of cancer detection.
+In fact, medical data are highly confidential and cannot be shared easily.
+Provided that medical institutions agree to share it, it has to undergo an anonymization process which follows very strict rules.
+Then, to be usable as part of a deep learning classication task, data must be organized and described precisely.
+In other words, additional information such as the clinical significance, the position of the lesion, etc.
+must be collected and put together.
+This whole process requires considerable effort, which is one of the reasons why medical data is hard to find.
+In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
 
 \section{Process overview}
+The approach consists in five different steps.
+First of all, similar datasets are collected and processed in the same way.
+These datasets need to share common features.
+For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.).
+On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
+
+Second, the convolutional neural network is built and split into two parts.
+It is still a single network but the first layers are treated as the feature extractor.
+As the name suggests, the role of these layers is to extract visual features characterizing what a cancerous tumor looks like.
+On the other hand, the last layers are considered as decision maker.
+They will gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
+
+Third, the model is trained on the target dataset and the model whose performance is the best is saved.
+The target dataset is the dataset which one wants to improve the performance of the model on.
+This first training establishes a baseline of performance and of features to learn further on thanks to the other datasets.
+
+Fourth, the core of transfer learning process starts. For each non-target dataset, two different trainings take place:
+
+\begin{enumerate}
+	\item The model saved at the end of the previous training is loaded.
+Its last layers (the decision maker) are completely reset, i.e.
+the weights are reset to another initialization, as if no training occured for these layers.
+Afterwards, the first layers (the decision maker) are frozen and the model is trained.
+Since the first layers have been frozen, their weights will remain the same.
+Only the weights of the last layers will change.
+This part of the transfer learning allows the model to make use of the previously learned features while adapting to the new body part by training the decision maker.
+	\item Once the training with frozen layers is over, the best model is loaded, the frozen layers are unfrozen, and a normal training is performed on the same dataset.
+Thanks to this, the model adapts the previously learned features and tries to learn new features which could be useful for the classification of the target dataset.
+\end{enumerate}
+
+Fifth, the previous step is also applied to the target dataset.
+These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again.
+Hopefully, the new weight initialization resulting from the transfer learning will allow to reach better metric values than before the transfer learning.
+Usually, the best model
 
@@
 \subsection{Architecture generalities}
+
 \subsection{Script options}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|llll|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
+\textbf{--training-set1}     & Path to training set 1                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--validation-set1}   & Path to validation set 1                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--test-set1}     & Path to test set 1                                                                                                                                                                                                                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
+\textbf{--training-set2}     & Path to training set 2                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{--validation-set2}   & Path to validation set 2 
+& True                                                                        & String        \\ \hline
+\textbf{--training-set3}     & Path to training set 3                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--validation-set3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
+\textbf{--nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Int           \\ \hline
+\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
+\textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
+\textbf{--cuda-device}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
+\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
+\textbf{--optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{--cv} & \begin{tabular}[c]{@{}l@{}}Whether to use cross-validation or not\\{[}'True', 'False'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}True\end{tabular}             
+& String        \\ \hline
+\end{tabular}%
+}
+\caption{Complete list of script options}
+\label{fig:transfer_learning_options}
+\end{table}
+
 \subsection{Visualization of the usefulness of the features learned on each dataset}

### 2020-02-03 Common/chapters/deep_learning.tex
Julien and Johan: typos and figures

--- 2020-01-30 Common/chapters/deep_learning.tex
@@
 \label{ch:deep_learning}
+This chapter provides the theoretical foundation in deep learning which is required to understand of the rest of the work.
+It starts with an historical timeline of deep learning before describing what a neural network is.
+Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained.
+This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks".
+They are used in many computer vision applications due to their great performance for these tasks.
+Finally, the concept of "transfer learning" is discussed since the entire chapter \ref{ch:transfer_learning} relies on it.
 
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/history.png}
-\caption{Milestones - Wang et al. \cite{14}}
+\caption{Deep learning milestones - Wang et al. \cite{14}}
 \label{history}
@@
 First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom.
-In this case, the "stride" is equal to 1.
+In this case, the so-called "stride" is equal to 1.
 Other applications could rely on a bigger stride.
-Furthermore, the previous example reduced the output size of $3x3$ in proportion to the initial input size $5x5$.
+Furthermore, the previous example reduced the output size of $3x3$ in proportion to the initial input size of $5x5$.
 To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s.

### 2020-02-03 Common/chapters/medical.tex
Julien and Johan: visualization tools, chapter introduction and minors

@@
 \label{ch:medical}
+This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly.
+First of all, some notions about cancer are described in order to understand what it is and which effects it has on the human body.
+Second, medical data has its own file formats.
+To make use of them in a deep learning project, medical files must be processed in a certain way, depending on each format.
+In fact, formats represent 2D, 3D or even 4D data.
+Some of them require specific normalization in order to get the right rendering.
+Finally, some visualization tools were developped to display raw medical files easily.
 
 \section{Cancer}
-
 \subsection{Basics}
@@
 These pieces of information are used to classify cancer between four (I to IV) or sometimes five (0 to IV) different stages, reflecting the progression and the seriousness of the illness \cite{22}.
-The earlier they are detected, the higher the chances of recovering are.
+The earlier they are detected, the higher the chances of recovery are.
+This aspect makes cancer detection critical since every misjudgment can threaten someone's life.
 
@@
 \end{algorithm}
+
+\section{Visualization tools}
+Processing data manually increases the probability of making mistakes.
+For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developped.
+Their primary goal is to compare our visual representation of an image to the one obtained in professional pieces of software.
+These tools are convenient to visualize a dataset easily.
+In some cases (especially RAW/MHD images), no free software capable of reading the files was available, which made the corresponding tool useful.
+
+\subsection{DICOM}
+Visualizing DICOM files is pretty straightforward since each file represent a single two-dimensional image.
+However, the way this standard works require a lot of pixel transformation and normalization to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
+Our tools allows to display a single DICOM file as well as a sequence of files if the function is feeded with a directory.
+Users can then scroll through the Z-axis, displaying the next or previous slice.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_lung_dcm.png}
+\caption{Visualization of a folder of DICOM files}
+\label{fig:visualize_lung_dcm}
+\end{figure}
+
+\subsection{NIfTI}
+As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
+When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
+Furthermore, the three-dimensionality implies that the volume is viewable under three different perspectives.
+For example, a three-dimensional brain volume can display it from the top of the head to the bottom, from one ear to the other and from the back of the head to the person's face.
+Therefore, the user can choose a specific axis to navigate through.
+If no axis is chosen, all three perspectives are shown one after the other.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_brain_nii.png}
+\caption{Visualization of a four-dimensional NIfTI file}
+\label{fig:visualize_brain_nii}
+\end{figure}
+
+\subsection{RAW}
+Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e.
+a succession of 2D slices.
+Scrolling with the mouse goes through slices over a specific axis.
+Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part under three different perspectives.
+This tool was particularly useful since no free software capable of handling these files was available.
+Photography-related programs such as Photoshop can open RAW pictures but not these three-dimensional ones.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth, keepaspectratio=true]{./figures/visualize_liver_raw.png}
+\caption{Visualization of a three-dimensional RAW file}
+\label{fig:visualize_liver_raw}
+\end{figure}

### 2020-02-03 Common/chapters/paper_reproduction.tex
Julien and Johan: model roulette

@@
 \label{paper_tensorboard}
+
 \subsection{Script options}
@@
 In order to achieve this goal, the Python module "argparse" is extremely useful.
-The latter automatically generates help/usage messages and displays errors when the argument given by the user are invalid.
+The latter automatically generates help/usage messages and displays errors when the arguments type in by the user are invalid.
 Table \ref{fig:paper_reproduction_options} shows all options accepted by the script.
@@
 
+\subsection{Model roulette}
+As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
+Experience showed us that the exact same hyperparameters with two different initializations could give opposite results.
+Therefore, in order not to train the model vainly, a script which generates a given number of initializations was created.
+Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set.
+The model whose score was the highest on a given metric was saved and used as base model for the experiments.
+
 \subsection{Experimental setup}
@@
 This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
+
 \subsubsection{Gradient flow visualization}

### 2020-02-03 Common/chapters/literature_review.tex
Julien and Johan: continue literature review

@@
 \label{ch:literature_review}
+
+Computer-aided cancer detection is a trending topic in deep learning.
+Nowadays, articles tend to focus more on segmentation than classification, but classification still got a lot of attention over the last decade.
+
+\section{Prostate/PROSTATEx}
+PROSTATEx: section \ref{sec:PROSTATEx}.
 
@@
 Saifeng et al.
-\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achvieving the second best performance with an AUC of $0.84$.
-The AUC on the validation set reached $0.92$.
-Their data processing approach stacked different combination of the available sequences as the three channels: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W.
+\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
+The AUC on their validation set reached $0.92$.
+Regarding data processing, their approach stacked different combinations of the available sequences as the three channels instead of defining a single combination like Song et al.: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W.
 The data augmentation process differs in that the images are rotated in 3D, each lesion being sliced at 7 different orientations.
@@
 First of all, the input was feeded to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans.
-Then, each of these feature extractors' outputs are merged into a common decision maker.
-Furthermore, 3-dimensional convolutions instead were performed.
+Then, each of these feature extractors' outputs were merged into a common decision maker.
+Furthermore, 3-dimensional convolutions instead 2-dimensional ones were performed.
 In fact, 3-dimensional patches centered on the lesion were cropped.
 Augmentation including translation and flipping was used in order to balance the dataset.
-Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist.
-Finally, this model achieved an AUC of $0.80$ on the PROSTATEx challenge.
+Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist compared to the previous papers.
+These tricks allowed their model achieved an AUC of $0.80$ on the PROSTATEx challenge.
 To make predictions, five different models were used, averaging the predictions of the four best models.
+
+\section{Lung/Lung CT Challenge}
+SPIE-AAPM Lung CT Challenge is lung cancer classification dataset. It is described in detail in section \ref{sec:lungCTChallenge}.
+
+Cengil et al.
+\cite{02} built a fairly simple convolutional neural network to classify images of the Lung CT Challenge dataset.
+The model takes 4-dimensional data as input (depth, height, width and channels) and performs 3D convolutions on it.
+The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
+Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set, composed of 30 findings.
+
+\section{Brain/Kaggle Brain}
+Kaggle brain dataset: section \ref{sec:kaggleBrain}.
+
+%Hanwat et al.
+\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
+Their processing method included skull masking, i.e.
+the removal of every non-brain tissue from the image.
+According to the authors, this technique improves performance quite a lot.
+---> Check their methods (graphs look special)
+
+Saxena et al.
+\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+Their processing method used a cropping technique which removed extra black margin around the skull.
+Each border of the image merges with a part of the skull.
+Since the images come from different sources, their resolution vary quite a lot.
+Therefore, the authors resized them to $224x224x3$.
+Moreover, data was augmented by rotation and vertical/horizontal shifting.
+As the cropping was performed before augmenting the images, small parts of the brain are outside of the augmented images due to rotation and shifting.
+The data was split into a training set, a validation set and a test set.
+Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance.
+The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
+The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).

### 2020-02-03 Common/chapters/transfer_learning.tex
Julien and Johan: continue literature review

@@
 \subsection{PROSTATEx}
+\label{sec:PROSTATEx}
 As the other datasets used in this work don't contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
@@
 \subsection{LungCTChallenge}
+\label{sec:lungCTChallenge}
 \subsubsection{Dataset description}
@@
 \subsection{Kaggle Brain}
+\label{sec:kaggleBrain}
 \subsubsection{Dataset description}

### 2020-02-04 Common/chapters/literature_review.tex
Johan and Julien: continue literature review

@@
 
-Computer-aided cancer detection is a trending topic in deep learning.
-Nowadays, articles tend to focus more on segmentation than classification, but classification still got a lot of attention over the last decade.
+Traditionally building a computer-aider cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.
+Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.
+False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate.
+However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"\cite{41}.\\
+Nowadays, this two steps tend to disappear thanks to deep learning algorithms, in particular the one based on convolutional neural networks, that constitute themselves CAD of one single step.
+This is the reason why CAD systems are currently a trending topic in deep learning.\\
+There mainly exists two types of CAD systems based on convolutional neural networks.
+The first one is called CADe for "computer-aided detection".
+Their primary goal is "to increase the detection rate of diseases while reducing the false negative  rate  possibly  due  to  the  observers’  mistakes  or  fatigue"\cite{41}.
+This group includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection.
+The other type is CADx for "computer-aided diagnosis" that aim to characterize the lesions.\\
+Consequently, this chapter will go through the main articles about CADe systems using convolutional neural networks for each body part used in the experiment of section \ref{ch:transfer_learning} and \ref{ch:paper_reproduction}, that are prostate, lung and brain.
 
 \section{Prostate/PROSTATEx}
+"Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
+It was predicted that there would be 1.7 million new cancer cases by 2030.
+The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}.
 PROSTATEx: section \ref{sec:PROSTATEx}.
@@
 \section{Lung/Lung CT Challenge}
+"Lung cancer is one of the most frequent and leading causes to death all over the world.
+It was reported that there were approximately 1.8x106 new cases of lung cancer globally in 2012.
+Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"\cite{41}.
 SPIE-AAPM Lung CT Challenge is lung cancer classification dataset. It is described in detail in section \ref{sec:lungCTChallenge}.

### 2020-02-05 Common/chapters/literature_review.tex
Johan and Julien: literature review first version + minors

@@
 The other type is CADx for "computer-aided diagnosis" that aim to characterize the lesions.\\
-Consequently, this chapter will go through the main articles about CADe systems using convolutional neural networks for each body part used in the experiment of section \ref{ch:transfer_learning} and \ref{ch:paper_reproduction}, that are prostate, lung and brain.
+This chapter goes through the main articles about CADe systems using convolutional neural networks for each body part used in the experiment of section \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, that are prostate, lung and brain.
+Furthermore, it focuses on articles that used the same datasets used in this work.
 
@@
 The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}.
-PROSTATEx: section \ref{sec:PROSTATEx}.
-
+Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection.
+All the articles mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients which were split into a training, validation and test set (see \ref{sec:prostatex_dataset_description} for more details).\\
 Song et al.
-\cite{07} proposed a DCNN method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients which were split into a training, validation and test set.
+\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs.
 Their data processing approach kept T2W, DWI and ADC grayscale images only.
@@
 
+Armato et al \cite{42} summarized the results obtained by all teams that took part in the PROSTATEx challenge in 2017.
+This challenge was split into two tasks.
+The first one was devoted to the diagnostic classification of prostate lesions whereas the other was about the segmentation and the determination of Gleason Grade Group.
+Thirty-two group submitted results to the first challenge from a total of 71 different methods (each group was allowed to submit results up to three methods).
+The article indicates that "most, but not all, methods outperformed random guessing (AUC=0.5)"\cite{41}.
+The best performing method obtained an AUC value of 0.87 (standard error 0.027) and the next three methods all achieved AUC values of 0.84 (with standard errors of 0.036, 0.032, and 0.032).
+Globally, all results are illustrated on figure \ref{fig:challenge_all_results}.
+The median AUC on the challenge is approximately at 0.68.
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/challenge_all_results.png}
+\caption{AUC values achieved by the 71 methods that participated in the PROSTATEx Challenge }
+\label{fig:challenge_all_results}
+\end{figure}
+
 \section{Lung/Lung CT Challenge}
 "Lung cancer is one of the most frequent and leading causes to death all over the world.
-It was reported that there were approximately 1.8x106 new cases of lung cancer globally in 2012.
+It was reported that there were approximately $1.8x10^6$ new cases of lung cancer globally in 2012.
 Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"\cite{41}.
-SPIE-AAPM Lung CT Challenge is lung cancer classification dataset. It is described in detail in section \ref{sec:lungCTChallenge}.
+All the articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is a dataset of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for more details).
 
@@
 The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
-Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set, composed of 30 findings.
+Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set, composed of 30 findings.\\
+Armato et al.
+\cite{12} described the LUNGx Challenge and their results.
+This challenge consisted in classifying the lung nodules as benign or malignant among a training set of 10 scans and a test set of 60 scans.
+Since the training set was extremely small, training the model on other datasets was allowed.
+The article describes the results of the proposed methods and compare them with the performance of six radiologists on the same task.
+The article reports that: "ten groups applied their own methods to 73 lung nodules (37 benign and 36 malignant) that were selected to achieve approximate size matching between the two cohorts.
+Area under the receiver operating characteristic curve (AUC) values for these methods ranged from 0.50 to 0.68.
+Only three methods performed statistically better than random guessing.
+The radiologists’ AUC values ranged from 0.70 to 0.85.
+Three radiologists performed statistically better than the best-performing computer method."\cite{12}.
+Figure \ref{fig:LUNGx_challenge_all_results} gives an overview of all results.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/LUNGx_challenge_all_results.png}
+\caption{AUC values for the 11 computerized methods and six radiologists in the task of classifying malignant and benign nodules}
+\label{fig:LUNGx_challenge_all_results}
+\end{figure}
 
 \section{Brain/Kaggle Brain}
-Kaggle brain dataset: section \ref{sec:kaggleBrain}.
-
-%Hanwat et al.
-\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
-Their processing method included skull masking, i.e.
-the removal of every non-brain tissue from the image.
-According to the authors, this technique improves performance quite a lot.
----> Check their methods (graphs look special)
+Brain cancer is another major type of cancer.
+"Brain and other nervous system cancer is the 10th leading cause of death for men and women.
+It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"\cite{43}.
+All the following articles are based on the Kaggle Brain dataset.
+This dataset, unlike the ones for the prostate and the lung, does not come from a certified medical authority but from the Kaggle website.
+However, since it is the only dataset available for the brain tumor classification, some publications used it (see \ref{sec:kaggleBrain} for more details).
 
@@
 The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).
+
+%Hanwat et al.
+\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
+Their processing method included skull masking, i.e.
+the removal of every non-brain tissue from the image.
+According to the authors, this technique improves performance quite a lot.
+---> Check their methods (graphs look special)
+
+Habib Mohamed Ali \cite{04} proposed a self created convolutional neural network to classify the images of the Kaggle Brain dataset.
+First, the data was augmented.
+Then, it was cropped to contain only the brain, resized and normalized in order to scale pixel values to the range [0,1].
+After the preprocessing, the data was split into a training set (70\%), a validation set (15\%) and a test set (15\%).
+Regarding the neural network structure, the model is simple.
+It is composed of only one convolutional layer with a batch norm layer and ReLu activation function, followed by two max pool layers and a dense layer.
+Thanks to this model, an accuracy of 89\% was achieved on the test set.

### 2020-02-05 Common/chapters/transfer_learning.tex
Julien and Johan: continue transfer learning

@@
 
+\subsection{Image cropping}
+Multiple ways of cropping images were tested out.
+The differences lay in how much tissue was kept apart from the lesion itself.
+In some cases, more context around the lesion allows to capture details about the connections between the lesion and blood vessels for example.
+Moreover, the abrupt borders inducing a clear change of colors can help detect tumors.
+Experienced showed that more context helps when the images are single-sequence.
+On the other hand, cropping right around the lesion or even in the lesion gave better results for multi-sequence data, i.e.
+when T2-weighted, DWI and ADC images (MRIs) were simultaneously used to perform the tumor detection and classification.
+In the latter case, the lesion usually appears distinctively on the three sequences, and this impacts the decision a lot.
+On single-sequence images, there is a clear lack of information which makes it difficult to differentiate a tumor from another tissue without much context.
+
 \subsection{Verification}
@@
 They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside.
-Like PROSTATEx (see section \ref{sec:dataProcessingVerification}), control images with red dots were generated for both datasets.
+Like PROSTATEx (see section \ref{sec:dataProcessingVerification}), control images with red dots were also generated for the other datasets.
 
@@
 
+\subsubsection{Layer freezing}
+After each dataset switch, the model is first trained with freezed layers.
+The freezed layers are those considered as feature extractor, i.e.
+the first layers of the model.
+The exact point where the model is split is arbitrary, but is usually located after the convolutions.
+To perform this freezing, multiple steps are required:
+\begin{enumerate}
+	\item Reset the last layers.
+To do so, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance.
+	\item Iterate over the batch normalization layers and freeze them.
+Step 3 is not sufficient to freeze batch normalization layers.
+This can be achieved by calling module.eval() on each module containing such layers (after the model.train() call).
+	\item Pass the parameters of these last layers only to the optimizer.
+\end{enumerate}
+
 \subsection{Script options}
+A single Python script allows to run the entire transfer learning pipeline.
+Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc.
+Table \ref{fig:transfer_learning_options} gives a exhaustive list of the script options.
+The particularity lies in the learning rate, number of epochs, batch size and dropout options.
+In fact, the pipeline consists in seven different phases of training: dataset 1 (full model), dataset 2 (frozen model), dataset 2 (full), dataset 3 (frozen), dataset 3 (full), dataset 4 (frozen) and dataset 4 (full).
+In such a case, each phase will probably use different hyperparameters.
+Therefore, these command line options need to specify seven different parameters.
+For example, a valid value for the "--lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".
 
@@
 \textbf{--validation-set3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
-\textbf{--nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Int           \\ \hline
-\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
+\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Comma-separated ints          \\ \hline
+\textbf{--nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Comma-separated ints            \\ \hline
+\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}               & Comma-separated floats         \\ \hline
 \textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
@@
 \textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
-\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
+\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}                & Comma-separated floats          \\ \hline
 \textbf{--optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
@@
 
+\subsection{Independent training}
+To get reference values and evaluate the impact of transfer learning on the model performance, a set of trainings on each dataset was performed independently.
+These experiments involved multiple hyperparameter combinations.
+The best results were used as benchmark values.
+
+\subsubsection{PROSTATEx}
+
+\subsubsection{Kaggle brain}
+
+\subsubsection{Lung CT Challenge}
+
+\subsection{Transfer learning}
+
 \section{Discussion}

### 2020-02-06 Common/chapters/literature_review.tex
Julien and Johan: typos, remake screenshots and minors

--- 2020-02-03 Common/chapters/literature_review.tex
@@
 
-Computer-aided cancer detection is a trending topic in deep learning.
-Nowadays, articles tend to focus more on segmentation than classification, but classification still got a lot of attention over the last decade.
+Traditionally, building a computer-aider cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.
+Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.
+False-positive reduction was commonly based on traditional  machine  learning  methods  to  reduce  the  false  positive  rate.
+However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
+Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks.
+In fact, the latter can be considered as single-step CADs.
+For this reason, CAD systems are currently a trending topic in deep learning.\\
+There mainly exists two types of CAD systems based on convolutional neural networks.
+The first one is called CADe for "computer-aided detection".
+Its primary goal is "to increase the detection rate of diseases while reducing the false negative  rate  possibly  due  to  the  observers’  mistakes  or  fatigue"~\cite{41}.
+This group includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection.
+The other type is CADx for "computer-aided diagnosis" which aims at extracting the characteristics lesions.\\
+This chapter goes through the main articles about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+prostate, lung and brain.
+Furthermore, it exclusively focuses on articles that used the same datasets as wthe ones used in this work.
 
 \section{Prostate/PROSTATEx}
-PROSTATEx: section \ref{sec:PROSTATEx}.
+As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
+It was predicted that there would be 1.7 million new cancer cases by 2030.
+The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"~\cite{41}.
+Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection.
+All the articles mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see section \ref{sec:prostatex_dataset_description}).
 
-Song et al.
-\cite{07} proposed a DCNN method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients which were split into a training, validation and test set.
+Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs.
 Their data processing approach kept T2W, DWI and ADC grayscale images only.
@@
 by subtracting the mean before dividing by the standard deviation.
+The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$).
 The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20 to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
-Most of these processing techniques were reproductible, apart from the manual lesion contouring and labelling performed by a radiologist.
-Their model is a modified version of the well-known VGG16 model, including the addition of $1x1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function.
+Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist.
+Their model is a modified version of the well-known VGG16 model, including the addition of $1x1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}).
 The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set.
@@
 
-Saifeng et al.
-\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
+Saifeng et al.~\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
 The AUC on their validation set reached $0.92$.
@@
 
-Mehrtash et al.
-\cite{01} used a different approach.
+Mehrtash et al.~\cite{01} used a different approach.
 First of all, the input was feeded to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans.
@@
 
+Armato et al~\cite{42} summarized the results obtained by all teams that took part in the PROSTATEx challenge in 2017.
+This challenge was split into two separate tasks.
+The first one was devoted to the diagnosis of prostate lesions (classification) whereas the other was about the segmentation and the determination of Gleason Grade Group.
+Thirty-two groups submitted results to the first challenge out of a total of 71 different methods (each group was allowed to submit up to three methods for evaluation).
+The article indicates that "most, but not all, methods outperformed random guessing (AUC=0.5)"~\cite{41}.
+The best performing method obtained an AUC value of 0.87 (standard error 0.027) and the next three methods all achieved AUC values of 0.84 (with standard errors of 0.036, 0.032, and 0.032).
+The complete results are illustrated on figure \ref{fig:challenge_all_results}.
+The median AUC on the challenge is approximately $0.68$.
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/challenge_all_results.png}
+\caption{AUC values achieved by the 71 methods that participated in the PROSTATEx Challenge~\cite{42}}
+\label{fig:challenge_all_results}
+\end{figure}
+
 \section{Lung/Lung CT Challenge}
-SPIE-AAPM Lung CT Challenge is lung cancer classification dataset. It is described in detail in section \ref{sec:lungCTChallenge}.
+Gao et al.
+claimed that "lung cancer is one of the most frequent and leading causes to death all over the world.
+It was reported that there were approximately $1.8x10^6$ new cases of lung cancer globally in 2012.
+Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}.
+The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).
 
-Cengil et al.
-\cite{02} built a fairly simple convolutional neural network to classify images of the Lung CT Challenge dataset.
+Cengil et al.~\cite{02} built a fairly simple convolutional neural network to classify the images of the Lung CT Challenge dataset.
 The model takes 4-dimensional data as input (depth, height, width and channels) and performs 3D convolutions on it.
 The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
-Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set, composed of 30 findings.
+Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.
+
+Armato et al.~\cite{12} described the LUNGx Challenge and the overall results.
+This challenge consisted in classifying the lung nodules as benign or malignant among a training set of 10 scans and a test set of 60 scans.
+Since the training set was extremely small, training the model on other datasets was allowed.
+The article describes the results of the proposed methods and compares them with the performance of six qualified radiologists on the same task.
+The article reports that "ten groups applied their own methods to 73 lung nodules (37 benign and 36 malignant) that were selected to achieve approximate size matching between the two cohorts.
+Area under the receiver operating characteristic curve (AUC) values for these methods ranged from 0.50 to 0.68.
+Only three methods performed statistically better than random guessing.
+The radiologists’ AUC values ranged from 0.70 to 0.85.
+Three radiologists performed statistically better than the best-performing computer method."~\cite{12}.
+Figure \ref{fig:LUNGx_challenge_all_results} gives an overview of the results.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/LUNGx_challenge_all_results.png}
+\caption{AUC values for the 11 computerized methods and six radiologists in the task of classifying malignant and benign nodules~\cite{12}}
+\label{fig:LUNGx_challenge_all_results.png}
+\end{figure}
 
 \section{Brain/Kaggle Brain}
-Kaggle brain dataset: section \ref{sec:kaggleBrain}.
+Brain cancer is another major type of cancer.
+According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women.
+It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}.
+The following articles are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website.
+However, since it is the only dataset available for brain tumor classification, some publications used it.
+Further details about this dataset can be found in section \ref{sec:kaggleBrain}.
 
-%Hanwat et al.
-\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
+Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+Their processing method used a cropping technique which removed extra black margin around the skull.
+Each border of the image merged with a part of the skull.
+Since the images come from different sources, their resolution vary quite a lot.
+Therefore, the authors resized them to $224x224x3$.
+Moreover, data was augmented by rotation, vertical shifting and horizontal shifting.
+As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting.
+The data was split into a training set, a validation set and a test set.
+Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance.
+The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
+The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).
+
+%Hanwat et al.~\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
 Their processing method included skull masking, i.e.
@@
 
-Saxena et al.
-\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
-Their processing method used a cropping technique which removed extra black margin around the skull.
-Each border of the image merges with a part of the skull.
-Since the images come from different sources, their resolution vary quite a lot.
-Therefore, the authors resized them to $224x224x3$.
-Moreover, data was augmented by rotation and vertical/horizontal shifting.
-As the cropping was performed before augmenting the images, small parts of the brain are outside of the augmented images due to rotation and shifting.
-The data was split into a training set, a validation set and a test set.
-Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance.
-The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
-The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).
+Habib Mohamed Ali~\cite{04} proposed his own convolutional neural network to classify the images of the Kaggle Brain dataset.
+First, the data was augmented.
+Then, it was cropped thanks to OpenCV so that the resulting images only contained the brain itself.
+Afterwards, they were resized and normalized in order to scale pixel values to the range $[0,1]$.
+Once the processing part was over, the data was split into a training set ($70\%$), a validation set ($15\%$) and a test set ($15\%$).
+Regarding the neural network structure, the model is simple.
+It is composed of only one convolutional layer with a batch normalization layer and ReLU activation function, followed by two max-pooling layers and a dense layer.
+This model achieved an accuracy of $89\%$ on the test set.

### 2020-02-06 Common/chapters/introduction.tex
Julien and Johan: start intro

% !TEX root = ../main.tex

\chapter{Introduction}
\label{ch:introduction}

According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}. An early detection and treatment increases the chances of recovery. In this context, Computer-Aided Diagnosis (CADx) systems can play a massive role by preventing health professionals from missing positive diagnoses. Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.

Deep learning applications require a lot of data to perform well. While certain fields profit from massive amounts of publicly available data, the medical field is quite the opposite. First of all, medical information is protected by the doctor-patient confidentiality and cannot be shared freely. As a consequence, data first has to be collected, organized and anonymized. Then, additional information regarding the clinical significance of the samples, the location of the lesions, etc. must be provided so that deep learning models can make use of the data. This whole process is time consuming and medical institutions don't always see the benefits which could ensue from publishing datasets of good quality. As a consequence, the lack of data is one of the toughest challenge related to this field. In order to overcome it, techniques such as transfer learning exist. The latter aims at using independent but similar datasets in order to increase the performance of a model on a target dataset. In other words, models are trained on the former in order to learn relevant features which improve the results on the latter. In the case of cancer detection and classification, only few datasets are available for each body part. Therefore, the idea behind this work is to make use of datasets of different body parts to classify one specific type of cancer.

This work is divided into various parts. First, the related literature was reviewed in order to gather techniques and architectures which gave decent results.

Second, deep learning is presented from the ground up under the historical and technical points of view. This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.

The next chapter deals with medical knowledge related to cancer. Characteristics of the disease are presented, before focusing on the various file formats which medical imaging data are exported into. The last part is critical since a lot of data processing was performed in order to be able to make use of these files.

Then, the architecture presented in a prostate cancer classification paper was reproduced. This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.

Finally, the last part is dedicated to transfer learning and its application to cancer classification.

% The most lethal kind is the lung cancer ($1.76$ million deaths in 2018).

\section{Motivation}
xx

\section{Contributions}
xx

### 2020-02-07 Common/chapters/introduction.tex
Johan and Julien: Contribution section first draft

@@
 \section{Contributions}
-xx
+This work proposes various methods to detect cancer on MRIs, with a focus on prostate cancer. Concretely, the major contributions of this work are:
+\begin{itemize}
+\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle Brain dataset.
+This includes the conversion of DICOM/.png files to NumPy arrays, their registration (alignment for stacking and same resolution to have the same amount of tissues on each ) and their split into multiple sets to be given as input to machine learning algorithms.
+\item Visualization scripts for DICOM, NIFTI and RAW medical file formats.
+A system to navigate through 4D data (width, height, depth and time) using the directional arrows is implemented: left/right arrows to vary the time axis and up/down to navigate through the difference slices (i.e the depth) of all images of a patient.
+\item A state of the art deep learning model for prostate cancer detection.
+All steps are clearly described, the corresponding scripts are available on Github and the unadulterated results are presented.
+This allows a complete reproducibility of the experiment.
+\item Processing script for the PROSTATEx challenge.
+This goes from the processing of the test images of the challenge to the generation of the .csv file containing their output predictions, which is either "benign lesion" or "malignant tumor".
+\item A transfer learning pipeline that makes use of different body parts to increase the classification performance of one chosen body part.
+This technique is a concrete solution to overcome the lack of available data for each body part.
+Furthermore, it increases the generalization ability of the neural network.
+\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN value in images etc.
+All these scripts can be easily be imported in other projects.
+
+\end{itemize}

### 2020-02-08 Common/chapters/introduction.tex
Julien and Johan: minors

--- 2020-02-06 Common/chapters/introduction.tex
@@
 \section{Contributions}
-xx
+This work proposes various methods to detect cancer on MRIs, with a focus on prostate cancer. Concretely, the major contributions of this work are:
+\begin{itemize}
+
+\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle Brain dataset.
+This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.
+
+\item Visualization scripts for DICOM, NIfTI and RAW medical file formats.
+A system to navigate through 4D data (width, height, depth and time) using the directional arrows is implemented: left/right arrows to vary the time axis and up/down to navigate through the difference slices (i.e the depth) of all images of a patient.
+3D data is also supported with the same functionalities (apart from navigating through the fourth dimension (time)).
+
+\item A state of the art deep learning model for prostate cancer detection.
+All steps are clearly described, the corresponding scripts are available on GitHub and the raw results are presented.
+This makes the experiment completely reproductible.
+
+\item Processing scripts for the PROSTATEx challenge.
+This goes from the preprocessing of the challenge test images to the generation of the CSV file containing the predictions of a given model, which are probabilities $[0,1]$ of the lesion being malignant ("benign lesion" if $< 0.5$, "malignant tumor" otherwise).
+
+\item A transfer learning pipeline that makes use of different body parts to increase the classification performance on one chosen body part.
+This technique is a concrete solution to overcome the lack of available data for each body part.
+Furthermore, it increases the generalization ability of the neural network.
+
+\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN value in images, etc. All these scripts can easily be imported in other projects.
+\end{itemize}

### 2020-02-10 Common/chapters/transfer_learning.tex
Julien and Johan: TL: "global performance"

@@
 \section{Process overview}
+\label{sec:process_overview}
 The approach consists in five different steps.
@@
 
-\subsection{Visualization of the usefulness of the features learned on each dataset}
+\subsection{Visualization of the impact of the various datasets on the target task}
+On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
+Hence, to see how much each step was actually helping the process, the decision maker resulting from the first step of the transfer learning, i.e.
+the training on the first/target dataset, was saved.
+As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
+
+Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
+This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
+
+At the end of the transfer learning pipeline, a graph was generated for each metric.
+These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
+Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
+As expected, the curve is flat during the frozen parts. Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics don't change at all.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/global_performance_example.png}
+\caption{Please replace this figure with a real example as soon as the results are available!}
+\label{fig:global_performance_example}
+\end{figure}
+
 \subsection{Experimental setup}
+
 \section{Results}

### 2020-02-13 Common/chapters/transfer_learning.tex
Julien and Johan: TL parallelization

@@
 This can be achieved by calling module.eval() on each module containing such layers (after the model.train() call).
-	\item Pass the parameters of these last layers only to the optimizer.
+	\item Only pass the parameters of these last layers to the optimizer.
 \end{enumerate}
+
+\subsubsection{Parallelization}
+Each experiment before transfer learning was performed on a single GPU.
+While being a lot faster than training on a CPU, the computation time was still relatively high.
+In order to get results from the transfer learning in a reasonable amount of time, parallelization was required.
+In fact, the longest transfer learning experiment consisted of $(5$ folds $* 4$ learning rates $* 1200$ epochs $+ 1200$ epochs$) * 7$ parts $= 176400$ epochs.
+On a single GPU, this experiment would have taken around two weeks to complete.
+
+Thanks to parallelization, the cross-validation combinations could be run simultaneously on different GPUs (5 of them since 5-fold cross-validation was used).
+To achieve this while using Python and GPUs, we relied on a library called Ray.
+The function responsible for training the model needs to be defined as a Ray remote function thanks to a decorator (@$ray.remote$ above the function declaration).
+This new function does not return the actual return values, but object IDs pointing to placeholders for the future results, which allows to continue the execution and run other parallel functions.
+Outside of the cross-validation loop, a call to $ray.get()$ on the object IDs allows to wait for the simultaneous executions to complete and to collect the different results.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_parallel_run.png}
+\caption{Transfer learning - Parallelization}
+\label{fig:tl_parallel_run}
+\end{figure}
 

### 2020-02-15 Common/chapters/transfer_learning.tex
Julien and Johan: transfer learning overview (figure)

@@
 \label{sec:process_overview}
-The approach consists in five different steps.
+The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
 First of all, similar datasets are collected and processed in the same way.
@@
 Hopefully, the new weight initialization resulting from the transfer learning will allow to reach better metric values than before the transfer learning.
-Usually, the best model
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.2\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
+}
+\caption{Transfer learning - Overview}
+\label{fig:tl_overview}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_parallel_run.png}
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.3\textwidth, keepaspectratio=true]{./figures/tl_parallel_run.png}
+}
 \caption{Transfer learning - Parallelization}

### 2020-02-16 Common/chapters/paper_reproduction.tex
Johan and Julien: literature review first version + minors
Julien and Johan: paper reproduction - results figure

@@
 \end{figure}
-\subsection{Criterion to save the best model}
 \subsection{Tensorboard}
@@
 \centering
-\includegraphics[width=1\textwidth, height=\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part1.png}
-\caption{Accuracy, AUC and loss on the training and validation set}
-\label{fig:paper_reprodution_results_part_1}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=1\textwidth, height=\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part2.png}
-\caption{F1-score, precision and recall on the training and validation set}
-\label{fig:paper_reprodution_results_part_2}
-\end{figure}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_part3.png}
-\caption{Specificity on the training and validation set}
-\label{fig:paper_reprodution_results_part_3}
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results.png}
+\caption{Metrics on the training and validation sets}
+\label{fig:paper_reprodution_results}
 \end{figure}

### 2020-02-17 Common/chapters/medical.tex
Julien and Johan: typos, remake screenshots and minors
Julien and Johan: tools to convert medical files to PNG

@@
 \subsection{DICOM file format}
+\label{sec:DICOM}
 
@@
 
-\section{Conversion}
-
-Medical data is often represented over 16 bits.
-However, exporting images to PNG require 8-bit images.
-To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
-
-\begin{algorithm}
-    \caption{16 to 8 bits conversion}
-    \label{alg:16_to_8_bits_conversion}
-    \begin{algorithmic}[1] % The number tells where the line numbering should start
-        \Procedure{16\_to\_8\_bits\_conversion}{$pixel\_array$}
-        		\State{$Pixel_{min} \gets$ minimal pixel value in $pixel\_array$}
-        		\State{$Pixel_{max} \gets$ maximal pixel value in $pixel\_array$}
-        		\For{$pixel\_value$ in $pixel\_array$}
-        			\State $pixel\_value \gets \frac{(pixel\_value - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
-        		\EndFor
-        		\State Modify object type to 8 bits
-        		\State Export $pixel\_array$ to PNG
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
 \section{Visualization tools}
@@
 \end{figure}
+
+\section{Conversion to PNG}
+
+In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
+The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG since it makes use of the others.
+Given a root directory, it scans the latter and its subdirectories to find all DICOM, NIfTI and RAW/MHD files.
+Then, each file is converted to PNG and saved into the same directory as the one it came from.
+This library allows to keep an easily accessible visual representation of medical files without having to rely on a specific piece of software to display them.
+Finally, it also played the role of debugging tool since the conversion techniques are the same as the ones used further in this work.
+
+\subsection{8-bit conversion}
+
+Medical data is often represented over 16 bits.
+However, exporting images to PNG require 8-bit images.
+To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
+
+\begin{algorithm}
+    \caption{16 to 8 bits conversion}
+    \label{alg:16_to_8_bits_conversion}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{16\_to\_8\_bits\_conversion}{$pixel\_array$}
+        		\State{$Pixel_{min} \gets$ minimal pixel value in $pixel\_array$}
+        		\State{$Pixel_{max} \gets$ maximal pixel value in $pixel\_array$}
+        		\For{$pixel\_value$ in $pixel\_array$}
+        			\State $pixel\_value \gets \frac{(pixel\_value - Pixel_{min}) * 255.0}{Pixel_{max} - Pixel_{min}}$
+        		\EndFor
+        		\State Modify object type to 8 bits
+        		\State Export $pixel\_array$ to PNG
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}

### 2020-02-17 Common/chapters/transfer_learning.tex
Julien and Johan: TL chapter - introduction

@@
 \label{ch:transfer_learning}
-
-What we do: improve cancer detection/classification on medical imaging
-How we do it: transfer learning
+%What we do: improve cancer detection/classification on medical imaging
+%How we do it: transfer learning
+This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data. To achieve this, transfer learning is used.\\
+First, a general definition of transfer learning is given.
+Its utility in this specific application is discussed as it is useful but not in any case.
+Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied.
+Then, a section focuses on the processing of new datasets that are used in the final experiment.
+Finally, the latter and its results are discussed.
 

### 2020-02-18 Common/chapters/paper_reproduction.tex
Paper reproduction section refactored: beginning

@@
 The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1).
-This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis (CAD) system for prostate cancer detection.
-Moreover, it also provides results about the algorithm performance, which can be used as a benchmark to compare with the results of the reproduction of the experiment.
-However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set, the reproduction of the experiment will fill that gap.\\
+The dataset contains a total of 204 patients for the training (labels are given) and 208 patients for the challenge (labels are unknown).
+This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis system for prostate cancer detection.
+Moreover, it also provides results (AUC) about the algorithm performance on the test set they built (which is not the same as the challenge test set since they split the 204 patients into training,validation/test sets), which can be used as a benchmark to compare with the results of the reproduction of the experiment.
+However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels), the reproduction of the experiment will fill that gap.\\
 This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole.
@@
 After the processing of the data, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
-This part is followed by the presentation of the raw results, which is itself followed by their analysis in the "discussion" section.
+This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "discussion" section.
+Finally, the last section is devoted to the SPIE-AAPM-NCI Prostate MR Classification Challenge and presents the results that the model achieved on the challenge test set using the whole training set for the training.
+This last part is not present in the original paper and represents contribution of this work to it.
 
@@
 Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}.
-The PROSTATEx dataset is composed of samples whose clinical significance is provided (labeled data) and samples without this information (unlabeled data).
-The latter are used for the challenge, which consists in predicting the probability of patients lesions to be malignant (class 1) or benign (class 0).
-In both cases, the data is processed before been used (see \ref{sec:prostatex_data_processing} for details).
-After the processing, the labeled data are split into training, validation and test sets using respectively 80\%, 10\% and 10\% of all available data for each set.
-The training set is then given in batches as input to the neural network, which update its weight consequently.
+The first part of the experiment is about the paper reproduction, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
+The former makes use of the PROSTATEx training set.
+This dataset is only composed of samples whose clinical significance is provided (labeled data).
+The first step of the reproduction is the data processing.
+In fact, from the DICOM files to the numpy arrays that are given to the neural network, many processing steps occur such as the lesion cropping, the normalization the MRI images stacking, the data augmentation, etc (see \ref{sec:prostatex_data_processing} for details).
+After the processing, the labeled data are split into a training, a validation and a test sets using respectively 80\%, 10\% and 10\% of all available data for each set.
+The training set is then given in batches as input to the neural network, which updates its weight consequently.
 At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}).
 If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.
-When the training phase reaches the defined number of epochs, the model is tested on the test set.
-The same model is also used to generated predictions for each patient and finding of the PROSTATEx challenge dataset in order to take part in it.
+Furthermore, if the validation AUC is decreasing for a defined number of epoch, the learning rate is reduced to prevent overfitting, which avoids to progressively decrease the performance on the validation set.
+This technique contributes to maintain a good generalization level for the model and optimized the training with respect to the AUC.
+When the training phase reaches the defined number of epochs, the model is tested on the test set with their method called "enhanced prediction" and the AUC is computed.
+This method consists in predicting an output for each augmented image and then computing the average of all predictions for each lesion of each patient.
+\\
+The second part of the experiment relies on the results of the first one.
+Indeed, the same data processing pipeline is followed and the hyperparameters used to train the neural network are the same but the entire dataset is used as a training set instead of splitting the data into a training/validation/test set.
+In this way, the entire dataset is given to the neural network, which maximizes the performance of the model on the challenge test set by increasing its generalization ability.
+During the training, the model is saved at the end of each epoch.
+At the end of it, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment.
+These models are evaluated on the challenge test set according the AUC by sending their predictions for each lesion of each patient.
+The model with the best result on the challenge is the most efficient one.
 
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_process.png}
-\caption{Paper reproduction experiment}
+\caption{Overview of the whole experiment process}
 \label{fig:paper_reproduction_process}
@@
 
-\section{PROSTATEx: Data processing}
+\section{Reproducing the paper experiment}
+
+\subsection{PROSTATEx: Data processing}
 \label{sec:prostatex_data_processing}
-\subsection{Dataset description}
+This section describes all processing operations performed on the raw data until it passes through the neural network.
+All steps mentioned below are also valid for the second part of the experiment about the challenge (section \ref{sec:classification_challenge}).
+
+\subsubsection{Dataset description}
 \label{sec:prostatex_dataset_description}
@@
 Both CSV files are complementary to each other.
-Regarding the challenge patients, two analog CSV files are provided.
+Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}.
 The only difference is the absence of clinical significance.
@@
 
-\subsection{Methodology}
+\subsubsection{Methodology}
 The methodology described in this section is the closest replication possible of the authors' data processing methodology.
@@
 
-\subsection{From DICOM to NumPy arrays}
+\subsubsection{From DICOM to NumPy arrays}
 \label{sec:DICOMtoNumPy}
@@
 
-\subsection{From NumPy arrays to augmented stacked images}
+\subsubsection{From NumPy arrays to augmented stacked images}
 \label{sec:numpyToAugmentedStacked}
@@
 
-\subsection{From NumPy arrays to augmented non-stacked images}
+\subsubsection{From NumPy arrays to augmented non-stacked images}
 Another way of processing and using the data was experimented with.
@@
 
-\section{Training the neural network}
-\subsection{Architecture}
+\subsection{Training the neural network}
+Training the neural network is the most important step of the experiment after the data processing.
+This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it is trained.
+
+\subsubsection{Architecture}
 The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
@@
 \end{figure}
-\subsection{Tensorboard}
-During the experiment the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC.
-These metrics are computed separately on the training and on the validation sets at the end of each epoch.
-They are then stored across all epochs and plotted on the same figure thanks to Tensorboard at the end of the process.
-Subsequently, this is also the case for the metrics measured on the test set.\\
-"Tensorboard provides the visualization and tooling needed for machine learning experimentation:
-\begin{itemize}
-\item Tracking and visualizing metrics.
-\item Visualizing the model graph (ops and layers).
-\item Viewing histograms of weights, biases, or other tensors as they change over time.
-\item Projecting embeddings to a lower dimensional space.
-\item Displaying images, text, and audio data"\cite{39}.
-\end{itemize}
-In this experiment, Tensorboard is mainly used to plot the Matplotlib figures of the model performance.
-In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
-\label{paper_tensorboard}
 
-\subsection{Script options}
+\subsubsection{Script options}
 Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
-In order to achieve this goal, the Python module "argparse" is extremely useful.
+This requirement was solved thanks to the Python module "argparse".
 The latter automatically generates help/usage messages and displays errors when the arguments type in by the user are invalid.
-Table \ref{fig:paper_reproduction_options} shows all options accepted by the script.
+Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective type.
 
@@
 
-\subsection{Model roulette}
+\subsubsection{Tensorboard}
+The Tensorboard official website \cite{39} claims that: "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+\begin{itemize}
+\item Tracking and visualizing metrics.
+\item Visualizing the model graph (ops and layers).
+\item Viewing histograms of weights, biases, or other tensors as they change over time.
+\item Projecting embeddings to a lower dimensional space.
+\item Displaying images, text, and audio data"\cite{39}.
+\end{itemize}
+In this experiment,Tensorboard is mainly used to plot the Matplotlib figures of the model performance.
+In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC.
+These metrics are computed separately on the training and on the validation sets at the end of each epoch.
+They are then stored across all epochs and plotted on the same figure at the end of the process thanks to Tensorboard.
+In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
+\label{paper_tensorboard}
+
+\subsubsection{Model roulette}
 As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
@@
 The model whose score was the highest on a given metric was saved and used as base model for the experiments.
+It can be given as initialization model to the neural network using the option \textit{--modeltoload <path>}.
 
-\subsection{Experimental setup}
-The best performing model was implemented with PyTorch using the Adam optimizer to update its weights, a learning rate of 1e-8 and the cross-entropy as loss function.
-The data was organized in batches of 128 samples and the experiment ran for 1300 epochs.
-The best model was selected with respect to the area under curve.
-Regarding the weight initialization, the best model from the roulette that maximized the F1-score was chosen as initial model.\\
-The model was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 15 hours.
+\subsubsection{Experimental setup}
+Keeping the exact same hyperparameters as the original paper lead the reproduction experiment to disappointing results.
+Consequently, new hyperparameters had to be found.
+A grid search approach driven by the AUC obtained on the challenge test set was chosen.
+In practical terms, this consists in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on the challenge test set.
+At the end of this process, it appears that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of 1e-7, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC sequentially decreases for two epochs) and a dropout rate of 0.2.
+This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\
 
@@
 \end{figure}
-\section{Results}
+
+\subsection{Results}
 Figures \ref{fig:paper_reprodution_results_part_1}, \ref{fig:paper_reprodution_results_part_2} and \ref{fig:paper_reprodution_results_part_3} show the different metrics registered on the training and validation sets during the training phase.
@@
 
-\section{Discussion}
+\subsection{Discussion}
 Thanks to all curves in the "results", a lot of information regarding the model performance can be inferred.
@@
 As the goal of the experiment was to reproduce the results of the paper,
+
+\section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
+\label{sec:classification_challenge}
+\subsection{Training the neural network with the whole dataset}
+\subsection{Results on the challenge test set}
+\subsection{Discussion}

### 2020-02-18 Common/chapters/paper_reproduction.tex
Johan and Julien: all results for paper reproduction added, remain to describe and explain

@@
 Consequently, new hyperparameters had to be found.
-A grid search approach driven by the AUC obtained on the challenge test set was chosen.
-In practical terms, this consists in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on the challenge test set.
+A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen.
+In practical terms, this consists in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
 At the end of this process, it appears that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of 1e-7, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC sequentially decreases for two epochs) and a dropout rate of 0.2.
@@
 The second problem, the "exploding gradient" is the opposite.
-Concretely, in this case, large gradients are computed, which results in huge weight updates that can even reach "NaN" values.
+In this case, large gradients are computed, which leads to huge weight updates that can even reach "NaN" values.
 This makes the model unstable and unable to learn from training data.
 \\
-As a result, it is important to analyze the gradient propagation across the network to ensure that it is not facing such problems.
+As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems.
 To achieve this goal, a visualization tool to display the gradients of the entire network was implemented.
@@
 From the output layer (called "last\_layer0.weight" on the graph) to the first layer, the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
+The training process goes on correctly.
 
@@
 \subsection{Results}
-Figures \ref{fig:paper_reprodution_results_part_1}, \ref{fig:paper_reprodution_results_part_2} and \ref{fig:paper_reprodution_results_part_3} show the different metrics registered on the training and validation sets during the training phase.
-The best model was chosen at epoch ...
-with an accuracy of ...
-and an AUC of ...
-(the condition for a model to be chosen as best is that the current AUC should be higher than the AUC of the previous best model, same for the accuracy).\\
+Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
+The best model was chosen at epoch 21 with an accuracy of 0.7525 and an AUC of 0.765 (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model) on the validation set.\\
+Figure \ref{fig:paper_reprodution_auc_on_our_test_set}
 This best model was also used to take part in the ProstateX challenge. The res XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
@@
 \centering
-\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge.png}
-\caption{Score achieved in the PROSTATEx challenge}
-\label{fig:paper_reprodution_results_challenge}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/auc_our_test_set.png}
+\caption{AUC of the best model on our test set}
+\label{fig:paper_reprodution_auc_on_our_test_set}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
+\caption{Score achieved in the PROSTATEx challenge using only our custom training set (80\% of all available data) for the training}
+\label{fig:paper_reprodution_results_challenge_1}
 \end{figure}
@@
 \subsection{Results on the challenge test set}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge_full_dataset.png}
+\caption{Metrics on the training using all data available as training set}
+\label{fig:challenge_all_results}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge2.png}
+\caption{Score achieved in the PROSTATEx challenge using all available data as training set}
+\label{fig:paper_reprodution_results_challenge2}
+\end{figure}
+
 \subsection{Discussion}

### 2020-02-19 Common/chapters/paper_reproduction.tex
Johan and Julien: section reproducing the paper experiment done

@@
 Finally, the last section is devoted to the SPIE-AAPM-NCI Prostate MR Classification Challenge and presents the results that the model achieved on the challenge test set using the whole training set for the training.
-This last part is not present in the original paper and represents contribution of this work to it.
+This last part is not present in the original paper and represents the contribution of this work to it.
 
@@
 Furthermore, if the validation AUC is decreasing for a defined number of epoch, the learning rate is reduced to prevent overfitting, which avoids to progressively decrease the performance on the validation set.
-This technique contributes to maintain a good generalization level for the model and optimized the training with respect to the AUC.
+This technique contributes to maintain a good generalization level for the model and optimizes the training with respect to the AUC.
 When the training phase reaches the defined number of epochs, the model is tested on the test set with their method called "enhanced prediction" and the AUC is computed.
@@
 \subsection{Results}
+\label{ref:sec:results}
 Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
 The best model was chosen at epoch 21 with an accuracy of 0.7525 and an AUC of 0.765 (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model) on the validation set.\\
-Figure \ref{fig:paper_reprodution_auc_on_our_test_set}
-This best model was also used to take part in the ProstateX challenge. The res XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+This best model was evaluated on the test set built for the experiment and achieved a score of 0.75 in AUC.\\
+Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71.
+This result will be used as a benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
 
@@
 \subsection{Discussion}
-Thanks to all curves in the "results", a lot of information regarding the model performance can be inferred.
+Thanks to all curves of section \ref{ref:sec:results}, a lot of information regarding the model performance can be inferred.
 In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
-First of all, when observing the loss plot, what is noticeable is that the training and validation losses follow the same behaviour until the epoch 600 approximately.
+First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behaviour during the 50 epochs.
 This is a clear sign of good fit of the model on the data.
@@
 This situation is the optimal one.
-After this phase, the model slightly started to overfit.
-In fact, the training loss continues to linearly decrease whereas the validation loss decreases less quickly until reaching  XXXXXXXXXXXXXXXXXXXXXXXX\\
-
-As the goal of the experiment was to reproduce the results of the paper,
+Regarding the slope of the curve, a big change occurs at epoch 17.
+This moment corresponds to the one where the learning rate is reduced by a factor 100 thanks to the \textit{reducelronplateau}.
+In fact at this moment, the validation AUC was at its maximum (the training is optimized according to this metric).
+If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have go up immediately, which would have lead to overfitting.
+Here thanks to the learning rate decay, this scenario is avoided and the validation and training loss stay stable from this point.
+The same behaviour is clearly noticeable for all other metrics.\\
+Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well.
+This is particularly noticeable on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones.
+The validation accuracy indicates that the model makes the right class decision in approximately 75\% of the time.
+Furthermore, since the model achieves a validation AUC of 0.76, it shows that it has a good distinction capacity.
+In fact, the highest is the AUC, the highest is the ability for the model to predict benign lesion as benign and malignant lesion as malignant.
+Samuel G.
+et al.
+\cite{42} claim that a less-experienced radiologist can reach an AUC of 0.81 and an expert an AUC of 0.91.
+Consequently, this model is almost at the level of a radiologist, which is very promising.
+It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results could be reached with more data.
+The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the idea that the model has a good distinguishable capacity with a score of approximately 0.76 on the validation set.
+Another extremely important aspect of a computer-aider diagnosis is to avoid at most the number of false negative.
+In fact, diagnostically, it is far better to classify a benign lesion as malignant than a malignant lesion as benign.
+This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion.
+The model has a recall about 0.74 on the validation set.
+The two last metrics (precision and recall) are considered together in the f1-score that is equal to 0.76.
+Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately 0.76.
+Globally, this model is balanced, no metric is exploding at the expense of another and this stability constitutes a serious strength regarding its generalization ability.\\
+As the goal of the experiment was to reproduce the experiment of the paper, the comparison between the two models is necessary.
+The reproducted model achieved an AUC of 0.75 on our test set (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}) using their "enhanced prediction" technique whereas the authors achieved an AUC in the interval [0.876-0.994] with 95\% of confidence on their test set.
+Consequently, the model presented in this work has a slightly smaller AUC [0.126-0.244] than the one achieved in the paper.
+Multiple factors can explain this difference.
+First, the cropping of the lesion was done by a radiologist in their work while it was automatically done using the information in the csv files for this work.
+Then, the constitution of the test set can make the result vary a lot.
+In fact, the authors didn't provide any information about how they build their test set.
+It is therefore possible that the test set was chosen to maximize the performance of the model.
+In the current work, the test set is built randomly.
+Consequently, the comparison of the performance is difficult since the test set is not the same in the two works.
+The only way to seriously compare the two models would be to compare their performance on the official SPIE-AAPM-NCI Prostate MR classification challenge test set.
+The model of this section (trained with 80\% of all available data) reached an AUC of 0.71.
+Unfortunately, no information about the performance of the model of the paper is given.
+Knowing that the best result at the time of the challenge was an AUC of 0.87 on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval [0.876-0.994] with 95\% of confidence...
+Finally, another source of difference is the number of participants in the experiment and the level of experience they have.
+Indeed, the paper has seven authors including one bachelors's student and six PhD students with one of them working for Siemens Heathcare and one for hImagingTek Ltd.
+These big companies are specialized in the medical imaging and have departments dedicated to machine learning.
+At the opposite, this work was produced independently by two MSc student without previous experience in the medical imaging field.
 
@@
 \label{sec:classification_challenge}
+This section aims at maximizing the result on the SPIE-AAPM-NCI Prostat MR classification Challenge test set by using the entire available data as training set, instead of using 80\% as it was done in the previous section.
+First, the changes made to reach this goal are detailed, then the raw results are presented and finally the latter are interpreted.
 \subsection{Training the neural network with the whole dataset}

### 2020-02-19 Common/chapters/literature_review.tex
Johan and Julien: minor

@@
 
-Traditionally, building a computer-aider cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.
+Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.
 Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.
@@
 prostate, lung and brain.
-Furthermore, it exclusively focuses on articles that used the same datasets as wthe ones used in this work.
+Furthermore, it exclusively focuses on articles that used the same datasets as the ones used in this work.
 

### 2020-02-19 Common/chapters/paper_reproduction.tex
Julien and Johan: typos and minors

--- 2020-02-18 Common/chapters/paper_reproduction.tex
@@
 This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
-It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data in the medical field (see chapter \ref{ch:transfer_learning}).\\
+It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
 Song et al.
 \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is one of the biggest available dataset for prostate cancer classification.
+This dataset is one of the biggest dataset available for prostate cancer classification.
 The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1).
-The dataset contains a total of 204 patients for the training (labels are given) and 208 patients for the challenge (labels are unknown).
-This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis system for prostate cancer detection.
-Moreover, it also provides results (AUC) about the algorithm performance on the test set they built (which is not the same as the challenge test set since they split the 204 patients into training,validation/test sets), which can be used as a benchmark to compare with the results of the reproduction of the experiment.
-However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels), the reproduction of the experiment will fill that gap.\\
-This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole.
-Then, each part or theoretical notion of the experiment is described in detail.
-The structure of the dataset is exposed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented.
-After the processing of the data, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
-This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "discussion" section.
-Finally, the last section is devoted to the SPIE-AAPM-NCI Prostate MR Classification Challenge and presents the results that the model achieved on the challenge test set using the whole training set for the training.
-This last part is not present in the original paper and represents contribution of this work to it.
+The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown).
+This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection.
+Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set).
+In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set.
+These results can be used as a benchmark to compare with the results of the reproduction of the experiment.
+However, the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels).
+Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.
+
+This chapter is built in a top-down fashion.
+First, an overview of the entire process is established in order to understand the purpose of the experiment as a whole.
+Then, each theoretical notion is described in detail.
+This includes the structure of the dataset, the steps involved in its processing, and the techniques used as a verification of the proper functioning of the algorithm.Then, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
+This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "Discussion" section.
+Finally, the last section is devoted to the so-called SPIE-AAPM-NCI Prostate MR Classification Challenge.
+It presents the results that the model achieved on the challenge test set using the whole training set for the training.
+The latter section is not part of the original paper.
+%and represents the contribution of this work to it.
 
@@
 Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}.
-The first part of the experiment is about the paper reproduction, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
+The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
 The former makes use of the PROSTATEx training set.
@@
 The first step of the reproduction is the data processing.
-In fact, from the DICOM files to the numpy arrays that are given to the neural network, many processing steps occur such as the lesion cropping, the normalization the MRI images stacking, the data augmentation, etc (see \ref{sec:prostatex_data_processing} for details).
-After the processing, the labeled data are split into a training, a validation and a test sets using respectively 80\%, 10\% and 10\% of all available data for each set.
-The training set is then given in batches as input to the neural network, which updates its weight consequently.
-At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}).
-If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.
-Furthermore, if the validation AUC is decreasing for a defined number of epoch, the learning rate is reduced to prevent overfitting, which avoids to progressively decrease the performance on the validation set.
-This technique contributes to maintain a good generalization level for the model and optimized the training with respect to the AUC.
-When the training phase reaches the defined number of epochs, the model is tested on the test set with their method called "enhanced prediction" and the AUC is computed.
-This method consists in predicting an output for each augmented image and then computing the average of all predictions for each lesion of each patient.
-\\
+In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network.
+Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc.
+(see Section \ref{sec:prostatex_data_processing} for further details).\\
+Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset.
+The training set samples are then fed to the neural network in batches.
+%, which updates its weight consequently. 
+At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}).
+The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes.
+Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved.
+Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance.
+This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.
+As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC.
+The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.
+
 The second part of the experiment relies on the results of the first one.
-Indeed, the same data processing pipeline is followed and the hyperparameters used to train the neural network are the same but the entire dataset is used as a training set instead of splitting the data into a training/validation/test set.
-In this way, the entire dataset is given to the neural network, which maximizes the performance of the model on the challenge test set by increasing its generalization ability.
-During the training, the model is saved at the end of each epoch.
-At the end of it, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment.
-These models are evaluated on the challenge test set according the AUC by sending their predictions for each lesion of each patient.
-The model with the best result on the challenge is the most efficient one.
+Indeed, the same data processing pipeline is used and the hyperparameters used to train the neural network are the same.
+The difference lies in the fact that the entire dataset (i.e.
+the entire PROSTATEx training set) is used as training set instead of splitting the data into a training/validation/test set.
+In this way, the model sees more training examples, which maximizes the performance of the model on the challenge test set by increasing its generalization ability.
+During the training phase, the model is saved at the end of each epoch.
+Once the training is over, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment.
+These models are then evaluated on the challenge test set by submitting their predictions for each lesion to the Grand Challenge organization.
+The model with the best result on the challenge is considered as the most efficient one.
 
@@
 Training the neural network is the most important step of the experiment after the data processing.
-This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it is trained.
+This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it was trained.
 
@@
 The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
-This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset of 14 millions images belonging to 1000 classes.
+This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes.
 Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
-It is first composed by three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks.
+It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks.
 Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function.
@@
 The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.\\
-In comparison to the original VGG, this model keeps the small filter size of 3x3 or 1x1, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
+In comparison to the original VGG, this model keeps the small filter size of 3x3, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
 It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one.
 Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
+Also, it uses 1x1 convolutions.
+
 \begin{figure}[!h]
@@
 Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
-This requirement was solved thanks to the Python module "argparse".
-The latter automatically generates help/usage messages and displays errors when the arguments type in by the user are invalid.
-Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective type.
+This requirement was solved thanks to the Python module called "argparse".
+The latter automatically generates help/usage messages and displays errors when the arguments typed in by the user are invalid.
+Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective types.
 
@@
 \subsubsection{Tensorboard}
-The Tensorboard official website \cite{39} claims that: "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
 \begin{itemize}
@@
 \end{itemize}
-In this experiment,Tensorboard is mainly used to plot the Matplotlib figures of the model performance.
+In this experiment, Tensorboard is mainly used to plot Matplotlib figures of the model performance.
 In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC.
 These metrics are computed separately on the training and on the validation sets at the end of each epoch.
-They are then stored across all epochs and plotted on the same figure at the end of the process thanks to Tensorboard.
+They are then stored and plotted on the same figure at the end of the process thanks to Tensorboard.
 In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
@@
 The model whose score was the highest on a given metric was saved and used as base model for the experiments.
-It can be given as initialization model to the neural network using the option \textit{--modeltoload <path>}.
+It can then be used as initialization model using the script option \textit{--modeltoload <path>}.
 
 \subsubsection{Experimental setup}
-Keeping the exact same hyperparameters as the original paper lead the reproduction experiment to disappointing results.
+Keeping the exact same hyperparameters as the original paper lead the experiment to disappointing results.
 Consequently, new hyperparameters had to be found.
 A grid search approach driven by the AUC obtained on the test set (the one built for the experiment, not the one of the challenge) was chosen.
-In practical terms, this consists in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
-At the end of this process, it appears that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of 1e-7, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC sequentially decreases for two epochs) and a dropout rate of 0.2.
+In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
+At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
 This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/gradient_flow.png}
-\caption{Gradient flow at epoch 0, batch 6 of the experiment}
+\caption{Gradient flow at epoch 0 of the experiment, batch 6}
 \label{fig:gradient_flow}
@@
 \subsection{Results}
+\label{ref:sec:results}
 Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
-The best model was chosen at epoch 21 with an accuracy of 0.7525 and an AUC of 0.765 (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model) on the validation set.\\
-Figure \ref{fig:paper_reprodution_auc_on_our_test_set}
-This best model was also used to take part in the ProstateX challenge. The res XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
+The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
+This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$.\\
+Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71.
+This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
 
@@
 \subsection{Discussion}
-Thanks to all curves in the "results", a lot of information regarding the model performance can be inferred.
+Thanks to the curves of Section \ref{ref:sec:results}, a lot of information regarding the model performance can be inferred.
 In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
-First of all, when observing the loss plot, what is noticeable is that the training and validation losses follow the same behaviour until the epoch 600 approximately.
+First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behavior during the 50 epochs.
 This is a clear sign of good fit of the model on the data.
-All features learned by the model on the training data are also applicable to the validation set.
+All the features learned by the model on the training data are also applicable to the validation set.
 This situation is the optimal one.
-After this phase, the model slightly started to overfit.
-In fact, the training loss continues to linearly decrease whereas the validation loss decreases less quickly until reaching  XXXXXXXXXXXXXXXXXXXXXXXX\\
+Regarding the slope of the curve, a big change occurs at epoch 17.
+This moment corresponds to the one where the learning rate is reduced by a factor of $100$ thanks to the \textit{reducelronplateau} function.
+In fact, at this moment, the validation AUC (the training is optimized according to this metric) was at its optimum with this learning rate.
+If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have lead to overfitting.
+Here, thanks to the learning rate decay, this scenario is avoided and the validation and training losses stay stable from this point onward.
+The same behaviour is clearly noticeable for all other metrics.\\
+Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well.
+This is particularly visible on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones.
+The validation accuracy indicates that the model made the right class decision in approximately $75$\% of the time.
+Furthermore, since the model achieved a validation AUC of $0.76$, it shows that it had a good distinction capacity.
+In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant.
+Samuel G.
+et al.
+\cite{42} claim that a less-experienced radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$.
+Consequently, this model is almost at the level of a radiologist, which is very promising.
+It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results can be reached with more data.\\
+The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the hypothesis that the model has a good distinction capacity with a score of approximately $0.76$ on the validation set.\\
+Another extremely important aspect of a computer-aider diagnosis system is to avoid the number of false negatives as much as possible.
+In fact, the consequences of classifying a malignant lesion as benign are far worse than the opposite.
+This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion.
+The model has a recall of around $0.74$ on the validation set.\\
+The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$. \\
+Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.\\
+Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.
 
-As the goal of the experiment was to reproduce the results of the paper,
+As the goal of the experiment was to reproduce the experiment of the paper, a comparison between the two models is necessary.
+The reproducted model achieved an AUC of $0.75$ on our test set (see Figure \ref{fig:paper_reprodution_auc_on_our_test_set}) using the authors' "enhanced prediction" technique, whereas the authors achieved an AUC $\in [0.876-0.994]$ with 95\% of confidence on their test set.
+Consequently, the model presented in this work has a slightly smaller AUC of $[0.126-0.244]$ than the one achieved in the paper.
+Multiple factors can explain this difference.
+First, in their work, the cropping of lesions was performed by a qualified radiologist, whereas, in our case, it was automatically done using the information in the CSV files.
+Then, the way the test set is built can make the results vary a lot.
+In fact, the authors didn't provide any information about how they built their test set.
+It is therefore possible that the test set was chosen to maximize the performance of the model.
+In the current work, the test set is built randomly.
+Consequently, the comparison of the performance is difficult since the test set is not the same in the two works.
+The only way to seriously compare the two models would be to compare their performance on the official SPIE-AAPM-NCI Prostate MR classification challenge test set.
+The model of this section (trained with 80\% of all the data available) reached an AUC of $0.71$.
+Unfortunately, no information about the performance of the model of the paper is given.
+Knowing that the best result at the time of the challenge was an AUC of 0.87 on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval $[0.876-0.994]$ with 95\% of confidence...
+Finally, another source of difference is the number of participants in the experiment and the level of experience they have.
+Indeed, seven authors contributed to the paper, including one bachelors' student and six PhD with one of them working for Siemens Heathcare and one for hImagingTek Ltd.
+These big companies are specialized in the medical imaging and have departments dedicated to machine learning.
+On the opposite, this work was produced independently by two MSc students without previous experience in the medical imaging field.
 
@@
 \label{sec:classification_challenge}
+This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% as it was done in the previous section.
+First, the changes made to reach this goal are detailed, then the raw results are presented and finally the latter are interpreted.
 \subsection{Training the neural network with the whole dataset}

### 2020-02-20 Common/chapters/deep_learning.tex
Julien and Johan: typo
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

@@
 \label{ch:deep_learning}
-This chapter provides the theoretical foundation in deep learning which is required to understand of the rest of the work.
-It starts with an historical timeline of deep learning before describing what a neural network is.
+This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work.
+It starts with a historical timeline of deep learning before describing what a neural network is.
 Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained.
 This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks".
-They are used in many computer vision applications due to their great performance for these tasks.
-Finally, the concept of "transfer learning" is discussed since the entire chapter \ref{ch:transfer_learning} relies on it.
+They are used in many computer vision applications due to their great performance on these tasks.
+Finally, the concept of "transfer learning" is discussed as the entire chapter \ref{ch:transfer_learning} relies on it.
 
@@
 There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc.
-What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
+What mainly differentiates deep learning from these algorithms is the concept of neural networks (see Section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
 Neural networks are inspired from the biological neural networks of the brain.
 These systems try to learn how to solve a problem based on the data they receive as input.
-Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc.
+Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics etc.
 The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology.
-Furthermore, since this field has recently gained interest (see section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.
+Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.
 
@@
 As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
-From the first attempts to understand the human brain until today, huge progress was made to be establish the basic components of modern neural networks.
+From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks.
 One could ask why deep learning took off recently if the theory was around for a long time.\\
@@
 One other main reason concerns the lack of data.
-Since deep learning algorithms "learn" from data, learning is impossible if good-quality data are not available.
+Since deep learning algorithms "learn" from data, learning is impossible if good-quality data is not available.
 The era of Big Data enhanced deep learning possibilities.
@@
 
-From a descriptive point of view, neural networks can simply be seen as a non-linear applications that associate an input to an output with respect to certain parameters.
+From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters.
 The input can be an image, a sound or any input that can be converted into numerical features.
@@
 In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
-From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates to an input $x$ an output $y$ with respect to parameters $\theta$.
+From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
 \begin{equation}
@@
 \subsection{Supervised and unsupervised learning}
-In machine learning, two different kinds of tasks exist.
+Machine learning algorithms belongs can be classified in two classes.
 The first one is "supervised learning".
-It includes learning algorithms whose training samples are associated to their labels in order to find the optimal mapping between the input and the output.
+It includes learning algorithms whose training samples are associated with their labels in order to find the optimal mapping between the input and the output.
 The second one is "unsupervised learning".
@@
 \begin{itemize}
-\item Superscript (i) denotes the $i^{th}$ training example.
-\item Superscript [l] denotes the $l^{th}$ layer of the neural network.
+\item Superscript ($i)$ denotes the $i^{th}$ training example.
+\item Superscript $[l]$ denotes the $l^{th}$ layer of the neural network.
 \end{itemize}
@@
 \begin{itemize}
-\item $X\in \R$ is the input matrix matrix of a neural network.
+\item $X\in \R$ is the input matrix of a neural network.
 \item $x^{(i)} \in \R^{n_{x}}$ is the $i^{th}$ example (sample) represented as a column vector.
@@
 \item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
-\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted $a^{[L]}$ where $L$ is the number of layers in the whole network.
+\item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted as $a^{[L]}$ where $L$ is the number of layers in the whole network.
 \item $g^{[l]}(x)$ is the $l^{th}$ activation function.  
@@
 Today, they are called "artificial neurons" or simply "neurons".\\
-A perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
+The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
 \begin{equation}
 \label{perceptron_equation}
-a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{})
+a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{j})
 \end{equation}
@@
 In fact, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
-Consequently, if linear functions only are used throughout the network, linear outputs are given as inputs to other linear functions.
+Consequently, if only linear functions are used throughout the network, linear outputs are given as inputs to other linear functions.
 Since the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers would not make sense anymore.
@@
 \begin{itemize}
-\item The input layer: it is the neural network entry point for the data.
+\item The input layer: It is the neural network entry point for the data.
 Generally speaking, the data is provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
@@
 For instance, suppose the number of samples is 100 and the batch size 32.
-In this situation, the network will be feeded with 4 batches of sizes [32, 32, 32, 4] respectively.
+In this situation, the network will be fed with 4 batches of sizes [32, 32, 32, 4] respectively.
 
-\item The hidden layer(s): hidden layers stand for all layers that are between the input layer and the output layer.
+\item The hidden layer(s): Hidden layers stand for all layers that are between the input layer and the output layer.
 Each of them has its own weights and biases (W, b), denoted by $W^{[l]} \in \R $ and $b^{[l]}$ respectively, where $W_{ij}^{l}$ corresponds to the weights associated with the connection between perceptron $j$ in layer $l$ and perceptron $i$ in layer $l+1$.
 By analogy, $b_{i}^{[l]}$ is the bias associated with perceptron $i$ in layer $l$.
-Weights and biases are the parameters to optimize in order to obtain the best mapping between the input and the output of the network (see section \ref{training_a_neural_network} ).
-Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see section \ref{weight_initialization}).
+Weights and biases are the parameters to optimize in order to obtain the best mapping between the input and the output of the network (see Section \ref{training_a_neural_network} ).
+Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see Section \ref{weight_initialization}).
 
-\item The output layer: it is the last layer of the neural network.
+\item The output layer: It is the last layer of the neural network.
 Its role is essential since it produces the prediction of the network for a given input.
-The prediction of a neural network is given by $\hat{y} \in R^{n_{y}}$ with $n_{y}$ beeing the number of different labels.
+The prediction of a neural network is given by $\hat{y} \in R^{n_{y}}$ with $n_{y}$ representing the number of different labels.
 In a classification task, whose goal is to assign to each input a specific class, the $\hat{y}$ is usually the probability that the input belongs to each class.
@@
 The computation of all these gradients relies on the "chain rule".
-Inthe case of weights, the chain rule is:
+In the case of weights, the chain rule is:
 \begin{equation}
@@
 \begin{itemize}
-\item The training set: this is the part of the dataset that is used to train the neural network (the weights and biases).
-\item The validation set: this the dataset that is used to evaluate a trained model.
+\item The training set: This is the part of the dataset that is used to train the neural network (the weights and biases).
+\item The validation set: This dataset is used to evaluate a trained model.
 Usually, the evaluation on the validation set is performed every $N$ epochs, where $N$ is a fixed number.
@@
 This fact of interrupting the training earlier is called "early stopping".
-\item The test set: this last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
+\item The test set: This last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
 \end{itemize}
@@
 That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
-So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps a acceptable level of generalization).\\
-Regarding the learning rate $\alpha$, its value has to be neither too large or too small.
-A too large learning rate is recognizable by analyzing the training loss curve: if "the loss is exploding or fluctuating" or if "it has stopped improving and it is wandering around a suboptimal local optima" \cite{28}, the learning rate is too high and should be decreased.
-On the contrary, if "the learning is very slow and the loss is decreasing consistently" or if "the model is overfitting" \cite{28}, it is a clear sign that it should be increased.
+So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).\\
+Regarding the learning rate $\alpha$, its value has to be neither too large nor too small.
+A too large learning rate is recognizable by analyzing the training loss curve: if the loss is exploding or oscillating or if there is no improvement and it is stuck around a suboptimal local optima, the learning rate is too high and should be decreased.
+On the contrary, if the learning is very slow and the loss does not decrease or if the model is overfitting, it is a clear sign that it should be increased.
 The learning rate can take a wide range of values.
@@
 It is exactly at this moment that the training should stop.
-To achieve this goal, it is possible either to directly choose the right number of epochs or to use "early stopping", which stops the training as soon as the validation loss does not decrease for $N$ epochs.
+To achieve this goal, it is possible either to directly choose the right number of epochs or to use the so-called "early stopping" method, which stops the training as soon as the validation loss does not decrease for $N$ epochs.
 
@@
 
-To perform a convolution, numerical input data and a "filter".
-A filter can be seen as a an $fxf$ numerical patch that moves across the entire input.
+To perform a convolution, numerical input data and a "filter" are required.
+A filter can be seen as a $f$x$f$ numerical patch that moves across the entire input.
 It first moves horizontally until reaching the right-most border of the image.
@@
 Other applications could rely on a bigger stride.
-Furthermore, the previous example reduced the output size of $3x3$ in proportion to the initial input size of $5x5$.
+Furthermore, the previous example reduced the output size of $3$x$3$ in proportion to the initial input size of $5$x$5$.
 To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s.
@@
 \begin{itemize}
-	\item \textbf{Valid:} the input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.
+	\item \textbf{Valid:} The input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.
 
-	\item \textbf{Same:} the input image is padded in a way that makes the output size the same as the input size.
+	\item \textbf{Same:} The input image is padded in a way that makes the output size the same as the input size.
 
-	\item \textbf{Full:} the input image is padded so that, in the first step of the convolution, only the bottom-right cell of the filter overlays the pixel value of the image, the rest overlaying padding cells.
+	\item \textbf{Full:} The input image is padded so that, in the first step of the convolution, only the bottom-right cell of the filter overlays the pixel value of the image, the rest overlaying padding cells.
 This makes the output size larger than the input size.

### 2020-02-20 Common/chapters/literature_review.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

@@
 
-Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted in two different parts: "Lesion  detection  and  false-positive  reduction.
+Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted of two different parts: "Lesion  detection  and  false-positive  reduction.
 Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.
@@
 The other type is CADx for "computer-aided diagnosis" which aims at extracting the characteristics lesions.\\
-This chapter goes through the main articles about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+This chapter presents the main research works about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
 prostate, lung and brain.
-Furthermore, it exclusively focuses on articles that used the same datasets as the ones used in this work.
+Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.
 
@@
 It was predicted that there would be 1.7 million new cancer cases by 2030.
-The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"~\cite{41}.
+The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}.
 Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection.
-All the articles mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see section \ref{sec:prostatex_dataset_description}).
+All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).
 
@@
 Their data processing approach kept T2W, DWI and ADC grayscale images only.
-After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65x65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels.
+After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels.
 Thanks to this method, the same lesion is visible in the same area over the three channels.
@@
 The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$).
-The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20 to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
+The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
 Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist.
-Their model is a modified version of the well-known VGG16 model, including the addition of $1x1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}).
+Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}).
 The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set.
-The best results were obtained by using DWI images with the highest b-value only, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
+The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
 However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.
 
-Saifeng et al.~\cite{31} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
+Liu et al.~\cite{46} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
 The AUC on their validation set reached $0.92$.
@@
 This whole process allows to include 3-dimensional information in 2-dimensional images.
-The method used ensemble learning which combined different models the reach the best performance possible.
+The method used ensemble learning which combined different models to reach the best performance possible.
 
 Mehrtash et al.~\cite{01} used a different approach.
-First of all, the input was feeded to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans.
+First of all, the input was fed to three separated parts of the model, each one responsible for a specific sequence among ADC, maximum b-value DWI and Ktrans.
 Then, each of these feature extractors' outputs were merged into a common decision maker.
-Furthermore, 3-dimensional convolutions instead 2-dimensional ones were performed.
+Furthermore, 3-dimensional convolutions instead of 2-dimensional ones were performed.
 In fact, 3-dimensional patches centered on the lesion were cropped.
@@
 Apart from these differences, other minor differences such as normalizing the images within the range $[0,1]$ exist compared to the previous papers.
-These tricks allowed their model achieved an AUC of $0.80$ on the PROSTATEx challenge.
+These tricks allowed their model to achieve an AUC of $0.80$ on the PROSTATEx challenge.
 To make predictions, five different models were used, averaging the predictions of the four best models.
@@
 Gao et al.
-claimed that "lung cancer is one of the most frequent and leading causes to death all over the world.
-It was reported that there were approximately $1.8x10^6$ new cases of lung cancer globally in 2012.
+claimed that "lung cancer is one of the most frequent and leading causes of death all over the world.
+It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012.
 Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}.
@@
 \caption{AUC values for the 11 computerized methods and six radiologists in the task of classifying malignant and benign nodules~\cite{12}}
-\label{fig:LUNGx_challenge_all_results.png}
+\label{fig:LUNGx_challenge_all_results}
 \end{figure}
@@
 It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}.
-The following articles are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
-This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website.
+The following research papers are based on the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}.
 However, since it is the only dataset available for brain tumor classification, some publications used it.
@@
 Since the images come from different sources, their resolution vary quite a lot.
-Therefore, the authors resized them to $224x224x3$.
+Therefore, the authors resized them to $224$x$224$x$3$.
 Moreover, data was augmented by rotation, vertical shifting and horizontal shifting.

### 2020-02-20 Common/chapters/medical.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

@@
 Some of them require specific normalization in order to get the right rendering.
-Finally, some visualization tools were developped to display raw medical files easily.
+Finally, some visualization tools were developed to display raw medical files easily.
 
@@
 	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells don't wait for the body's approval to grow and proliferate, contrary to normal cells.
-They become the only responsible for their multiplication.
+They become responsible for their own multiplication.
 	\item \textbf{"Evading growth suppressors"}\\
@@
 
-MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the outputed images as a volume.
+MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the generated images as a volume.
 Different settings, usually called sequences, make the look of the output vary, as shown on figure \ref{fig:PROSTATEx-t2-adc-dwi}.
@@
 
-If they are missing, a simple conversion is sufficient. The parameters to use to convert the data are given by two tags: 
+If they are missing, a simple conversion is sufficient. The parameters used to convert the data are given by two tags: 
 \begin{itemize}
@@
 “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
-Details about the different fields contained in the header can be found here: \url{https://brainder.org/2012/09/23/the-nifti-file-format/}
+Details about the different fields contained in the header can be found here\footnote{\url{https://brainder.org/2012/09/23/the-nifti-file-format/}}.
 
@@
 Processing data manually increases the probability of making mistakes.
-For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developped.
+For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developed.
 Their primary goal is to compare our visual representation of an image to the one obtained in professional pieces of software.
@@
 \subsection{DICOM}
-Visualizing DICOM files is pretty straightforward since each file represent a single two-dimensional image.
-However, the way this standard works require a lot of pixel transformation and normalization to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
-Our tools allows to display a single DICOM file as well as a sequence of files if the function is feeded with a directory.
+Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
+However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
+Our tools allows to display a single DICOM file as well as a sequence of files if the function is fed with a directory.
 Users can then scroll through the Z-axis, displaying the next or previous slice.
@@
 When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
-Furthermore, the three-dimensionality implies that the volume is viewable under three different perspectives.
+Furthermore, the three-dimensionality implies that the volume is viewable from three different perspectives.
 For example, a three-dimensional brain volume can display it from the top of the head to the bottom, from one ear to the other and from the back of the head to the person's face.
@@
 Scrolling with the mouse goes through slices over a specific axis.
-Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part under three different perspectives.
+Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part from three different perspectives.
 This tool was particularly useful since no free software capable of handling these files was available.

### 2020-02-20 Common/chapters/paper_reproduction.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

@@
 \label{ch:paper_reproduction}
-This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
 It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
@@
 These results can be used as a benchmark to compare with the results of the reproduction of the experiment.
-However, the article gives no information about the results their model got on the official PROSTATEx challenge test set (208 patients without labels).
+However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels).
 Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.
@@
 \section{Process overview}
-Schematically speaking, the entire experiment process can be represented as shown on figure \ref{fig:paper_reproduction_process}.
+Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
 The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
@@
 The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes.
-Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved.
+Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved.
 Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance.
@@
 This section describes all processing operations performed on the raw data until it passes through the neural network.
-All steps mentioned below are also valid for the second part of the experiment about the challenge (section \ref{sec:classification_challenge}).
+All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
 
@@
 The only difference is the absence of clinical significance.
-Finally, images are in the DICOM file format (see section \ref{sec:DICOM}).
+Finally, images are in the DICOM file format (see Section \ref{sec:DICOM}).
 
@@
 They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets.
-Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60x$) in order to create balanced training and validation sets.
+Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.
 
@@
                 					\If {$slice\_number == dicom\_file.InstanceNumber$}
-                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see section \ref{sec:dicom_data_manipulation}}
+                						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see Section \ref{sec:dicom_data_manipulation}}
                 						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
@@
 At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel.
-This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible on the three images.
+This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.
 
@@
 \noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient.
-According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately help the final diagnosis.
+According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately helps the final diagnosis.
 In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array.
@@
 These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient.
-Same for this patient's DWI and ADC images.
+Same for the patient's DWI and ADC images.
 The same process was applied to all patients separately.
@@
 These techniques alone allowed to create a sufficient amount of data.
-Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not made use of.
+Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.
 Images were exported as NumPy arrays.
@@
 \subsubsection{From NumPy arrays to augmented non-stacked images}
-Another way of processing and using the data was experimented with.
+Another way of processing and using the data was experimented.
 Instead of stacking the images, each image was cropped and augmented separately.
@@
 At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
+Note that the model was initialized with the best one among 200 models coming from a roulette maximizing the f1-score.
 This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\
@@
 To achieve this goal, a visualization tool to display the gradients of the entire network was implemented.
+This implementation is based on the code of RoshanRane \cite{47}.
 This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
@@
 The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
-This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$.\\
-Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71.
+This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}).\\
+Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reprodution_results_challenge_1}).
 This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
@@
 \label{sec:classification_challenge}
-This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% as it was done in the previous section.
-First, the changes made to reach this goal are detailed, then the raw results are presented and finally the latter are interpreted.
+This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
+First, the changes needed to reach this new goal are detailed, then the raw results are presented and finally the latter are interpreted.
 \subsection{Training the neural network with the whole dataset}
+The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
+It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state of the art deep learning model for prostate lesions classification.\\
+As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data.
+In a general way, the more data are available the better will be the model.
+Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from previous section.
+Apart from this new split, the data processing and the hyperparameters keep the same.
+The only difference in the training is about the learning rate decay.
+Indeed, since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting.
+However, this does not cause any problem because, thanks to the results of the previous section, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
+Similarly, it is possible to conclude that in this case the best model should also be the one around epoch 17.
+In order to ensure this, the models at epoch 15, 17, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set.
+For each of these models, a csv file containing the predictions for each lesion of each patient was automatically generated and sent to the challenge.
 \subsection{Results on the challenge test set}
+Figure \ref{fig:challenge_all_results_experiment} presents the metrics achieved by the model trained with all available data as training set at each epoch.\\
+Figure \ref{fig:results_for_all_models_challenge} shows the performance of the model saved at different epoch on the official challenge test set.
+The best AUC achieved on the challenge is 0.76.
+Figure \ref{fig:paper_reprodution_results_challenge2} illustrates the participation to
 
@@
 \caption{Metrics on the training using all data available as training set}
-\label{fig:challenge_all_results}
+\label{fig:challenge_all_results_experiment}
 \end{figure}
+
+\begin{table}[]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\textbf{Epoch} & \textbf{15} & \textbf{17} & \textbf{19} & \textbf{20} & \textbf{21} & \textbf{22} & \textbf{23} & \textbf{30} \\ \hline
+AUC            & 0.75        & 0.76        & 0.76        & 0.76        & 0.76        & 0.76        & 0.75        & 0.75        \\ \hline
+\end{tabular}%
+}
+\caption{AUC of the model saved at different epochs on the official SPIE-AAPM-NCI Prostate MR classification test set }
+\label{fig:results_for_all_models_challenge}
+\end{table}
 

### 2020-02-20 Common/chapters/transfer_learning.tex
Johan and Julien: Corrections according to Akhansha's feedback. All done except:  - Logo p.1 - currently one of the trendiest topics p.9 - activating invasion and metastasis p.35

@@
 \label{ch:transfer_learning}
-%What we do: improve cancer detection/classification on medical imaging
-%How we do it: transfer learning
 This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data. To achieve this, transfer learning is used.\\
@@
 Provided that medical institutions agree to share it, it has to undergo an anonymization process which follows very strict rules.
-Then, to be usable as part of a deep learning classication task, data must be organized and described precisely.
+Then, to be usable as part of a deep learning classification task, data must be organized and described precisely.
 In other words, additional information such as the clinical significance, the position of the lesion, etc.
@@
 For example, DWI images (MRI sequence) and CT scans share the same color scheme and look alike (the same kind of tissue is visible, etc.).
-On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
+%On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
 

### 2020-02-20 Common/chapters/paper_reproduction.tex
Johan and Julien: Paper reproduction chapter finished :D

@@
 In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
-At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate decay on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
-Note that the model was initialized with the best one among 200 models coming from a roulette maximizing the f1-score.
+At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
+Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the f1-score.
 This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\
@@
 If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have lead to overfitting.
-Here, thanks to the learning rate decay, this scenario is avoided and the validation and training losses stay stable from this point onward.
+Here, thanks to the learning rate scheduler, this scenario is avoided and the validation and training losses stay stable from this point onward.
 The same behaviour is clearly noticeable for all other metrics.\\
@@
 Apart from this new split, the data processing and the hyperparameters keep the same.
-The only difference in the training is about the learning rate decay.
+The only difference in the training is about the learning rate scheduler.
 Indeed, since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting.
-However, this does not cause any problem because, thanks to the results of the previous section, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
-Similarly, it is possible to conclude that in this case the best model should also be the one around epoch 17.
+However, this does not cause any problem because, thanks to the results of the previous section.
+In fact, is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
+Hence, it is possible to conclude that in this case the best model should also be the one around epoch 17 (or a few epochs later).
 In order to ensure this, the models at epoch 15, 17, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set.
@@
 \subsection{Results on the challenge test set}
-Figure \ref{fig:challenge_all_results_experiment} presents the metrics achieved by the model trained with all available data as training set at each epoch.\\
-Figure \ref{fig:results_for_all_models_challenge} shows the performance of the model saved at different epoch on the official challenge test set.
-The best AUC achieved on the challenge is 0.76.
-Figure \ref{fig:paper_reprodution_results_challenge2} illustrates the participation to
+Figure \ref{fig:results_for_all_models_challenge} shows all metrics achieved by the model on the official challenge test set during the training.
+Figure \ref{fig:challenge_all_results_experiment} presents the metrics achieved by the model trained with all available data at different epochs around the epoch 20.
+The range of AUC values is contained in the interval [0.75-0.76].\\
+Finally, figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, that achieved an AUC of $0.76$.
 
@@
 \includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge2.png}
-\caption{Score achieved in the PROSTATEx challenge using all available data as training set}
+\caption{Score achieved in the PROSTATEx challenge by the model picked at epoch 20, trained on all available data as training set}
 \label{fig:paper_reprodution_results_challenge2}
@@
 \subsection{Discussion}
+Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behaviour of the model during the training.
+All metrics reach their point of convergence, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.
+In comparison to the the first part of the experiment, it is clearly visible that all metrics have an higher best value.
+This is due to the absence of learning rate scheduler during the training.
+As a result, from a certain epoch, all features learned by the model are not transposable to unseen data and the latter overfits the data.
+To determine this point, multiple models around the epoch 20 were evaluated on the challenge test set (see Figure \ref{fig:results_for_all_models_challenge}).
+It is noticeable that the model before epoch 15 did not reach its best performance and after epoch 23 was overfitting.
+All models between these two epochs achieve an AUC of $0.76$ on the challenge.\\
+In comparison to the model of the previous section trained on 80\% of all available data that achieved an AUC of $0.71$ on the challenge, this new one improves the performance by 0.05.
+This is a clear indication that the more data is available the better is the model.\\
+Finally, regarding the overall leaderboard of all challenge participants at the time of the challenge, the model is well ranked.
+Indeed, with respect to the 71 submissions to the PROSTATEx challenge (see Figure \ref{fig:challenge_all_results}), the AUC of $0.76$ achieved in this work would place the model in 15th position approximately.

### 2020-02-21 Common/chapters/paper_reproduction.tex
Julien and Johan: add processing details (patient split, MRI sequences, class balancing, etc.) +  + add new challenge results + correct typos

@@
 \label{sec:prostatex_dataset_description}
-The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients.
+The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website.
+This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients.
 Images were taken under the sagittal, transverse and coronal planes.
+These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences".
+Each sequence allows to see human tissues in a different way on the resulting images.
+The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans.
 
@@
 Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
-Furthermore, the authors didn't mention any participation in the official PROSTATEx challenge.
+Furthermore, the authors did not mention any participation in the official PROSTATEx challenge.
 In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented.
@@
 Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
-Moreover, only images showing the prostate under the transverse plane were used.
-The reason for this is that the tumors are much more visible under this perspective.
+Furthermore, DWI images regroup various parameters which create different subcategories.
+One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images.
+The higher the b-value, the stronger the diffusion effects" \cite{49}.
+The authors tested three different b-value configurations.
+This led to the conclusion that using DWI images with the highest b-value increased the performance of the model.
+For that matter, DWI images with the highest b-value only were used.
+\\
+Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
 The first step consisted in converting DICOM files.
@@
 \label{sec:numpyToAugmentedStacked}
-This step can be split into two substeps: aligning and stacking images before augmenting them.
+This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
+
 Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution.
 In fact, T2-weighted images have a much higher resolution than the DWI and ADC images (in this dataset at least).
-Without this operation, a single pixel on a DWI image would have covered a lot more tissue than a pixel of the corresponding T2-image.
+Without this operation, a single pixel on a DWI image would have covered a lot more tissue than a pixel of the corresponding T2 image.
 To perform the alignment, the lesion was localized on the three sequences than to the CSV file.
@@
 This patch had to be centered on the lesion to ease the augmentation process.
-So, a fixed patch size was defined for the T2-image since its resolution was the highest.
+So, a fixed patch size was defined for the T2 image since its resolution was the highest.
 Then, using the voxel spacing information, the patch size required to cover the same amount of tissue was computed for the two other sequences (DWI, ADC).
@@
 
-Finally, the dataset was augmented in the same way as the authors.
+Then, the dataset was augmented in the same way as the authors.
 Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel).
-These techniques alone allowed to create a sufficient amount of data.
-Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.
-Images were exported as NumPy arrays.
+These techniques alone allowed to create a large enough amount of data.
+Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.\\
+On the other hand, authors did not mention how they managed class imbalance.
+It seems like the manual lesion contouring performed by the qualified radiologist led to a natural class balance.
+However, the unmodified dataset has a ratio of 3 negative images for 1 positive.
+To solve this problem, multiple approaches were implemented.
+The first one made use of undersampling which consists in taking N elements of each class, where N is the size of the class whose number of elements is the smallest.
+As the size of the dataset is relatively small, taking a subset of it was not a good idea and led to a clear lack of data.
+The second solution solved the problem by using different augmentation factors.
+In fact, one of the script options sets the augmentation factor for the class whose number of elements is the largest.
+Then, according to this factor, the algorithm computes the augmentation factor for the other class in order to balance the final datasets.
+For example, if the dataset follows the above-mentioned 3 negatives for 1 positive ratio, and if the script option augmentation factor is set to 50, each image belonging to the negative class is going to be augmented 50 times and each image belonging to the positive class 150 times.
+At the end, the resulting dataset is balanced and makes use of all images.
+
+Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays.
+The percentages represent a percentage of patients.
+In fact, each subset contains respectively 80\%, 10\% and 10\% of the 204 patients, i.e.
+$\lfloor 0.8 * 204 \rfloor = 163$ training patients, $\lfloor 0.1 * 204 \rfloor = 20$ test patients and $204 - 163 - 20 = 21$ validation patients.
+This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset.
+In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example.
+This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.
 
 \subsubsection{From NumPy arrays to augmented non-stacked images}
-Another way of processing and using the data was experimented.
-Instead of stacking the images, each image was cropped and augmented separately.
-At the end, each image was given as separate input to the neural network.
+Another way of processing and using the data was tested.
+Instead of stacking the images of various sequences coming from the same exam, each image was treated independently.
+In other words, each of them was resized as explained previously, then cropped, augmented and exported separately.
+At the end, each image was fed as separate input to the neural network.
+
+This way of proceeding gave awful results.
+The main reason is that images from different sequences show the same tissue in an almost opposite way.
+The most striking difference to the human eye is the colors.
+Tumors are usually white on DWI images and black on ADC images.
+Moreover, they are not always as visible on the images of the various sequences.
+Therefore, finding similar features on all sequences was extremely difficult.
+Stacking the three sequences solves this problem.
+In this case, the first 4225 input neurons (4225 input neurons for a $65$x$65$px image) are dedicated to T2 images, the next 4225 neurons to DWI images and the last 4225 neurons to ADC images.
 
@@
 \subsubsection{Experimental setup}
-Keeping the exact same hyperparameters as the original paper lead the experiment to disappointing results.
+Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
 Consequently, new hyperparameters had to be found.
@@
 Unfortunately, no information about the performance of the model of the paper is given.
-Knowing that the best result at the time of the challenge was an AUC of 0.87 on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval $[0.876-0.994]$ with 95\% of confidence...
+Knowing that the best result at the time of the challenge was an AUC of $0.87$ on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval $[0.876-0.994]$ with 95\% of confidence...
 Finally, another source of difference is the number of participants in the experiment and the level of experience they have.
@@
 This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
-First, the changes needed to reach this new goal are detailed, then the raw results are presented and finally the latter are interpreted.
+First, the changes needed to reach this new goal are detailed, then the raw results are presented and the latter are finally interpreted.
+
 \subsection{Training the neural network with the whole dataset}
 The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
-It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state of the art deep learning model for prostate lesions classification.\\
+It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.\\
 As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data.
-In a general way, the more data are available the better will be the model.
+In a general way, the more available data there is, the better the model performs.
 Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from previous section.
-Apart from this new split, the data processing and the hyperparameters keep the same.
+Apart from this new split, the data processing and the hyperparameters stay the same.
 The only difference in the training is about the learning rate scheduler.
-Indeed, since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting.
-However, this does not cause any problem because, thanks to the results of the previous section.
-In fact, is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
-Hence, it is possible to conclude that in this case the best model should also be the one around epoch 17 (or a few epochs later).
-In order to ensure this, the models at epoch 15, 17, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set.
-For each of these models, a csv file containing the predictions for each lesion of each patient was automatically generated and sent to the challenge.
+Since the validation set does not exist anymore, there is no possibility to use the \textit{reducelronplateau} function to prevent overfitting.
+However, this does not cause any problem thanks to the results of the previous section.
+In fact, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
+Hence, it is possible to conclude that, in this case, the best model should also be the one around epoch 17 (or a few epochs later).
+In order to ensure this, the models at epochs 15, 16, 17, 18, 19, 20, 21, 22, 23 and 30 were evaluated on the official challenge test set.
+For each of these models, a CSV file containing the predictions for each lesion of each patient was automatically generated and submitted to the challenge.
+
 \subsection{Results on the challenge test set}
-Figure \ref{fig:results_for_all_models_challenge} shows all metrics achieved by the model on the official challenge test set during the training.
-Figure \ref{fig:challenge_all_results_experiment} presents the metrics achieved by the model trained with all available data at different epochs around the epoch 20.
-The range of AUC values is contained in the interval [0.75-0.76].\\
-Finally, figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, that achieved an AUC of $0.76$.
+Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
+The range of values is contained in the interval $[0.75-0.76]$.
+Each model was tested in order to see if a slight peak of $0.01$ could be found from one epoch to the other.\\
+Figure \ref{fig:challenge_all_results_experiment} presents the metric values the model reached while training on the whole dataset.
+The models whose results are reported in Table \ref{fig:results_for_all_models_challenge} were picked from this training.\\
+Finally, Figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, which achieved an AUC of $0.76$.
 
@@
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
 \hline
 \rowcolor[HTML]{DAE8FC} 
-\textbf{Epoch} & \textbf{15} & \textbf{17} & \textbf{19} & \textbf{20} & \textbf{21} & \textbf{22} & \textbf{23} & \textbf{30} \\ \hline
-AUC            & 0.75        & 0.76        & 0.76        & 0.76        & 0.76        & 0.76        & 0.75        & 0.75        \\ \hline
+\textbf{Epoch} & \textbf{15} & \textbf{16} & \textbf{17} & \textbf{18} & \textbf{19} & \textbf{20} & \textbf{21} & \textbf{22} & \textbf{23} & \textbf{30} \\ \hline
+AUC            & 0.75        & 0.75        & 0.76        & 0.76        & 0.76        & 0.76        & 0.76        & 0.76        & 0.75        & 0.75        \\ \hline
 \end{tabular}%
@@
 \subsection{Discussion}
-Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behaviour of the model during the training.
-All metrics reach their point of convergence, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.
-In comparison to the the first part of the experiment, it is clearly visible that all metrics have an higher best value.
+Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behavior of the model during the training.
+All metrics reach their point of convergence at some point, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.
+In comparison to the the first part of the experiment, it is clearly visible that the metrics have an higher best value.
 This is due to the absence of learning rate scheduler during the training.
-As a result, from a certain epoch, all features learned by the model are not transposable to unseen data and the latter overfits the data.
+As a result, from a certain epoch onwards, features learned by the model are not transposable to unseen data anymore and the latter overfits the data.
 To determine this point, multiple models around the epoch 20 were evaluated on the challenge test set (see Figure \ref{fig:results_for_all_models_challenge}).
-It is noticeable that the model before epoch 15 did not reach its best performance and after epoch 23 was overfitting.
-All models between these two epochs achieve an AUC of $0.76$ on the challenge.\\
-In comparison to the model of the previous section trained on 80\% of all available data that achieved an AUC of $0.71$ on the challenge, this new one improves the performance by 0.05.
-This is a clear indication that the more data is available the better is the model.\\
-Finally, regarding the overall leaderboard of all challenge participants at the time of the challenge, the model is well ranked.
-Indeed, with respect to the 71 submissions to the PROSTATEx challenge (see Figure \ref{fig:challenge_all_results}), the AUC of $0.76$ achieved in this work would place the model in 15th position approximately.
+It is noticeable that the model did not reach its best performance before epoch 15 and that it started overfitting after epoch 23.
+Every model between these two epochs achieved an AUC of $0.76$ on the challenge, which seems to be the optimum value with this data and this configuration.\\
+In comparison to the model of the previous section trained on 80\% of the available data, this new one improves the performance by $0.05$ (AUC of $0.76$ vs $0.71$).
+This is a clear indication that the more data there is, the better the model performs.\\
+Finally, regarding the overall leaderboard of all challenge participants at the time of the challenge, the model would have been well ranked.
+Indeed, according to the 71 submissions to the PROSTATEx challenge (see Figure \ref{fig:challenge_all_results}), the AUC of $0.76$ achieved in this work would place the model at the 15$^{th}$ position overall.

### 2020-02-21 Common/chapters/transfer_learning.tex
Julien and Johan: add processing details (patient split, MRI sequences, class balancing, etc.) +  + add new challenge results + correct typos

@@
 \label{sec:PROSTATEx}
-As the other datasets used in this work don't contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
+As the other datasets used in this work do not contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
 DWI images were selected because lesions are white on this sequence, as well as on CT scans.
@@
 In fact, the extra information provided by different sequences increases the performance a lot.
-However, publicly available datasets either don't contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
+However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
 
@@
 Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
-As expected, the curve is flat during the frozen parts. Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics don't change at all.
+As expected, the curve is flat during the frozen parts.
+Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all.
 

### 2020-02-22 Common/chapters/paper_reproduction.tex
Julien and Johan: script options update

@@
 \textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
-\textbf{--trainingSet1}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--validationSet1}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{--trainingset}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{--validationset}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
 \textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
 \textbf{--nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
-\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
+\textbf{--lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
 \textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
-\textbf{--cuda-device}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
-\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
+\textbf{--cudadevice}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ''\end{tabular}                 & String        \\ \hline
 \textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
-\textbf{--optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+
+\textbf{--inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}True\end{tabular} & Int        \\ \hline
+
+\textbf{--optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
 \textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline

### 2020-02-23 Common/chapters/introduction.tex
Julien and Johan: add processing details (patient split, MRI sequences, class balancing, etc.) +  + add new challenge results + correct typos
Johan and Julien: Motivation first draft

@@
 \section{Motivation}
-xx
+As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device.
+The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information".
+In these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians.
+In fact, they provide multiple advantages.\\
+First of all, used as an extra diagnostic, they decrease the probability to miss positive diagnoses.
+Then, they can also speed up the diagnosis by proposing regions of interest to clinicians, which in time, will reduce the screening costs.
+Similarly, lower screening costs mean a greater accessibility of screening tests, which results in an earlier cancer detection.
+Another benefit of using CAD systems, specially the one based on Deep Learning, is the number of cases that these latter have "seen".
+Indeed, in order to create efficient Deep Learning models, big amounts of data are needed.
+Consequently, efficient CAD systems have seen a lot of cancer examples whereas beginner clinicians have probably seen less a smallest number of them.\\
+However, all these advantages are still nowadays partially theoretical and the research is still going on to improve existent CAD systems.
+Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent.
+%Consequently, this thesis presents how to build a modern Deep Learning model for prostate lesions classification and proposes a model based on transfer learning to overcome the lack of available data in the medical field.
 
 \section{Contributions}
-This work proposes various methods to detect cancer on MRIs, with a focus on prostate cancer. Concretely, the major contributions of this work are:
+This work proposes a state of the art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field.
+Concretely, the major contributions of this work are:
 \begin{itemize}
@@
 This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.
+Also, the class imbalance problem is taked into account.
+The processing scripts automatically augment each class in order to balance the number of elements.
+In addition to this, an alternative relying on undersampling was implemented as a PyTorch sampler.
 

### 2020-02-23 Common/chapters/literature_review.tex
Johan and Julien: Motivation first draft

@@
 
-Traditionally, building a computer-aided cancer diagnosis or detection (CAD) system consisted of two different parts: "Lesion  detection  and  false-positive  reduction.
+Traditionally, building a computer-aided cancer diagnosis or detection system consisted of two different parts: "Lesion  detection  and  false-positive  reduction.
 Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.

### 2020-02-24 Common/chapters/introduction.tex
Julien and Johan: motivation

--- 2020-02-08 Common/chapters/introduction.tex
@@
 \section{Motivation}
-xx
+Cancer is a serious concern in today's society.
+An early detection is key to maximize the chances of recovery.
+As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device.
+The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information".
+Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians.
+In fact, they provide multiple advantages.
+
+First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses.
+Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.\\
+Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection.
+As stated before, detecting cancers at an early stage maximizes the probability of curing them.\\
+Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on.
+In fact, in order to built efficient deep learning models, large amounts of data are required.
+Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.
+
+However, these advantages remain partially theoretical nowadays.
+Research is still ongoing, aiming at improving existent CAD systems.
+Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent.
+Therefore, this situation can and must be improved.
+
+%Consequently, this thesis presents how to build a modern Deep Learning model for prostate lesions classification and proposes a model based on transfer learning to overcome the lack of available data in the medical field.
 
 \section{Contributions}
-This work proposes various methods to detect cancer on MRIs, with a focus on prostate cancer. Concretely, the major contributions of this work are:
+This work proposes a state of the art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field.
+Concretely, the major contributions of this work are:
 \begin{itemize}
@@
 
-\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN value in images, etc. All these scripts can easily be imported in other projects.
+\item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN values in images, etc. All these scripts can easily be used in other projects.
 \end{itemize}

### 2020-02-24 Common/chapters/conclusions.tex
Julien and Johan: conclusion first draft

% !TEX root = ../main.tex

\chapter{Conclusion}
\label{ch:conclusions}

\section{Conclusion}
This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
The starting point was the reproduction of Song et al.'s experiment~\cite{07}. Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working. This resulted in a solid baseline that was exploited in order to take part in the "SPIE-AAPM-NCI Prostate MR Classification Challenge”, also called PROSTATEx challenge. Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge.\\
Then, the work focused on overcoming one of the main issues in deep learning: the lack of data. To achieve this, transfer learning, as well as more common techniques such as data augmentation, were applied. The particularity of our transfer learning implementation is the intermediate frozen steps. During the latter, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does. Experiments showed that: XXXXXX RESULTS XXXXXX.

Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline. All these elements can serve as a baseline or as additions to future projects.

xxxxx

\section{Future Work}
% The continuation of this work includes...

This work mainly used a single model architecture based on a VGG-16 network. As MRIs are three-dimensional, using three-dimensional convolutions could improve the results. This would also require new processing techniques which would generates volumes instead of two-dimensional images. However, well-labeled classification datasets can be counted on the fingers of one hand. As deep learning models need large amounts of training data, the continuation of the same work is more or less compromised. On the other hand, tumor segmentation has become popular over the last years. Most cancer-related deep learning challenges aim at segmenting tumors in three-dimensional data, which makes such datasets easy to find. Therefore, the experience and methods acquired in this work could be transposed to segmentation tasks, which could even combine segmentation and classification.

xxxxxx

### 2020-02-24 Common/chapters/conclusions.tex
Johan and Julien: New ideas added in conclusion and future works

@@
 The starting point was the reproduction of Song et al.'s experiment~\cite{07}.
-Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working.
+Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working well.
 This resulted in a solid baseline that was exploited in order to take part in the "SPIE-AAPM-NCI Prostate MR Classification Challenge”, also called PROSTATEx challenge.
-Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge.\\
+Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge.
+This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge, which confirms the robustness of the latter.
+\\
 Then, the work focused on overcoming one of the main issues in deep learning: the lack of data.
@@
 Therefore, the experience and methods acquired in this work could be transposed to segmentation tasks, which could even combine segmentation and classification.
-
+Another possibility to deal with the small amount of publicly available datasets is to develop partnerships with hospitals in order to have access to their data.
+Obviously, this raises questions about data confidentiality and ethics regarding how the latter is used.
+New platforms ensuring their confidentiality, their public sharing and their quality have to be developed with the goal of encouraging hospitals to share them.
+This would allow to build even more better models and in an easier way.
 xxxxxx

### 2020-02-28 Common/structure/info.tex
Julien and Johan: title draft and Master

% !TEX root = ../main.tex

%----------------------------------------------------------------------------------------
% THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\cosupervisor{Co-supervisor Name} % If you have a co-supervisor, include it here. This is used in the title page, print it elsewhere with \supname
\degree{Master} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Author Name} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{Bd de Pérolles 90} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Computer Science} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{keywords, list, here} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.unifr.ch}{University of Fribourg}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{https://www3.unifr.ch/inf/fr/}{Department of Informatics}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{\href{https://www3.unifr.ch/inf/en/exascale-infolab.html}{eXascale Infolab}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{https://www3.unifr.ch/scimed/fr/}{Faculty of Science and Medicine}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

### 2020-02-28 Common/structure/title.tex
Julien and Johan: title draft and Master

% !TEX root = ../main.tex

%----------------------------------------------------------------------------------------
% TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

%\includegraphics[width=15cm]{logos/xi_logos.pdf}
%\vspace*{.06\textheight}
%\\

\begin{figure}
  \centering
    \includegraphics[width=.17\textwidth]{logos/unifr_logos.pdf}
  \hfill
    \includegraphics[width=.4\textwidth]{logos/xi_logos.pdf}
  \vspace{30mm}
\end{figure}

{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
\textsc{\Large Master's Thesis}\\[0.5cm] % Thesis type
\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
\href{mailto://david.bucher@unifr.ch}{\authorname} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
\href{https://exascale.info/phil}{\supname} % Supervisor name - remove the \href bracket to remove the link
\\\vspace*{1ex}\emph{Co-Supervisor:} \\ % Remove these two lines if no co-supervisor is involved
\href{https://exascale.info/members/}{\cosupname} % Co-supervisor name - remove the \href bracket to remove the link
\end{flushright}
\end{minipage}\\[1cm]

January 01, 1970 % date of the official defense
\vspace*{.06\textheight}

%\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\textit{ in the}\\[0.3cm] % University requirement text
\groupname\\\deptname\\ % Research group name and department name
\vfill

\footnotesize{ Boulevard de Pérolles 90 ~~$\bullet$~~ 1700~Fribourg ~~$\bullet$~~ Switzerland
            \\
            phone +41~(26)~300~84~65 ~~$\bullet$~~ \textsf{diuf-secr@unifr.ch} ~~$\bullet$~~ \textsf{www3.unifr.ch/inf}
            }

\end{center}
\end{titlepage}

### 2020-02-29 Common/chapters/introduction.tex
Julien and Johan: conclusion first draft
Johan and Julien: new goal section + result first draft and discussion structure

@@
 In the case of cancer detection and classification, only few datasets are available for each body part.
-Therefore, the idea behind this work is to make use of datasets of different body parts to classify one specific type of cancer.
+Therefore, the idea behind this work is to make use of datasets of different body parts to improve the classification of prostate lesions.
 
@@
 
-Then, the architecture presented in a prostate cancer classification paper was reproduced.
+Then, the architecture presented in a prostate cancer classification research paper was reproduced.
 This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.
@@
 
-\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle Brain dataset.
+\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
 This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.

### 2020-02-29 Common/chapters/transfer_learning.tex
Johan and Julien: new goal section + result first draft and discussion structure

@@
 \section{Goal}
-Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.
-This technique makes sense when the quantity of data available is not sufficient and when the data of both datasets share common features.
-Both conditions are fulfilled in the case of cancer detection.
+Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.\\
+This technique makes sense when the quantity of data available is not sufficient.
+This conditions is fulfilled in the case of cancer detection.
 In fact, medical data are highly confidential and cannot be shared easily.
@@
 must be collected and put together.
-This whole process requires considerable effort, which is one of the reasons why medical data is hard to find.
-In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
+This whole process requires considerable effort, which is one of the reasons why medical data is hard to find.\\
+%In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
+However, in order for the transfer learning to be useful, common low level features have to be present among the datasets on which the model is trained.
+Consequently, the goal of this experiment is to use transfer learning on three datasets of three different body parts, that are prostate, brain and lung, in order to check if they share common features that could be useful to improve the performance of the model on a specific dataset.
+In this experiment, the model trained to classify prostate lesions is then trained to classify brain tumors and finally lung lesions.
+At the end of the process, the model is once again trained on the prostate dataset and the performances of the model at the beginning and at the end of the experiment are compared.
 
@@
 the weights are reset to another initialization, as if no training occured for these layers.
-Afterwards, the first layers (the decision maker) are frozen and the model is trained.
+Afterwards, the first layers (the features extractor) are frozen and the model is trained.
 Since the first layers have been frozen, their weights will remain the same.
@@
 
-\subsection{Independent training}
-To get reference values and evaluate the impact of transfer learning on the model performance, a set of trainings on each dataset was performed independently.
+%\subsection{Independent training}
+%To get reference values and evaluate the impact of transfer learning on the model performance, a set of trainings on each dataset was performed independently.
 These experiments involved multiple hyperparameter combinations.
 The best results were used as benchmark values.
-
-\subsubsection{PROSTATEx}
-
-\subsubsection{Kaggle brain}
-
-\subsubsection{Lung CT Challenge}
+%
+%\subsubsection{PROSTATEx}
+%
+%\subsubsection{Kaggle brain}
+%
+%\subsubsection{Lung CT Challenge}
 
 \subsection{Transfer learning}
+All training graphs and the metrics they achieved on the validation sets are shown on figures XXXX.\\
+Table \ref{table:best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment.
 
+\begin{table}[]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|l|l|c|l|c|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\multicolumn{1}{|c|}{\cellcolor[HTML]{DAE8FC}\textbf{Dataset name}} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Best model saved at nth epoch} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Best model on the validation set}} & \textbf{Best model on the set set}                                                                                                                                      \\ \hline
+\textbf{PROSTATEx}                                                  & DS1/Full                                                   & 497                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6340\\ AUC: 0.7359\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.6842\\ Precision: 0.3333\\ Recall: 0.5\\ F1-score: 0.4\\ Specificity: 0.7333\\ AUC: 0.6833\end{tabular}}     \\ \hline
+\textbf{Kaggle Brain}                                               & DS2/Frozen                                                 & 594                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7799\\ AUC: 0.8379\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{Kaggle Brain}                                               & DS2/Full                                                   & 264                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.8233\\ AUC: 0.8566\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{LungCTChallenge}                                            & DS3/Frozen                                                 & 999                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6617\\ AUC: 0.7019\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{LungCTChallenge}                                            & DS3/Full                                                   & 345                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6924\\ AUC: 0.7197\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{PROSTATEx}                                                  & DS4/Frozen                                                 & 999                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7803\\ AUC: 0.7743\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{PROSTATEx}                                                  & DS4/Full                                                   & 50                                     & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7307\\ AUC: 0.7592\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.7368\\ Precision: 0.4286\\ Recall: 0.75\\ F1-score: 0.5454\\ Specificity: 0.7333\\ AUC: 0.7667\end{tabular}} \\ \hline
+\end{tabular}%
+}
+\caption{Best model performance at each step of the transfer learning pipeline}
+\label{table:best_model_performance}
+\end{table}
 \section{Discussion}
+1) All steps of the transfer learning are correctly tuned\\
+	--> Description of the shapes of the curves at each step \\
+
+Consequences:\\	
+2) Features of brain and lung images are useful to classify prostate lesions\\
+	--> Explain that the frozen curve immediately start to increase, and then get stable. Furthermore they allow us to reach good performance metrics, describe them.\\
+	--> Explain the global performance graph in details\\\\
+3) Transfer learning improves the performance of the target model\\
+ --> Comparison between the results on the test set before and after the TL\\
+
+ 4) Summarizing the advantage of usins this pipeline \\
+ --> Increase the generalization ability of the model and its performance\\
+ --> Overcome the lack of data for one specific body part

### 2020-03-01 Common/chapters/paper_reproduction.tex
Julien and Johan: transfer learning results, figures and minors

@@
 \subsection{Results}
-\label{ref:sec:results}
+\label{sec:paper_reproduction_results}
 Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
@@
 \subsection{Discussion}
-Thanks to the curves of Section \ref{ref:sec:results}, a lot of information regarding the model performance can be inferred.
+Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
 In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\

### 2020-03-01 Common/chapters/transfer_learning.tex
Julien and Johan: transfer learning results, figures and minors

@@
 This technique makes sense when the quantity of data available is not sufficient.
-This conditions is fulfilled in the case of cancer detection.
-In fact, medical data are highly confidential and cannot be shared easily.
+This condition is fulfilled in the case of cancer detection.
+In fact, medical data is highly confidential and cannot be shared easily.
 Provided that medical institutions agree to share it, it has to undergo an anonymization process which follows very strict rules.
@@
 must be collected and put together.
-This whole process requires considerable effort, which is one of the reasons why medical data is hard to find.\\
+This whole process requires considerable effort, which is one of the reasons why usable medical data is hard to find.\\
 %In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
 However, in order for the transfer learning to be useful, common low level features have to be present among the datasets on which the model is trained.
-Consequently, the goal of this experiment is to use transfer learning on three datasets of three different body parts, that are prostate, brain and lung, in order to check if they share common features that could be useful to improve the performance of the model on a specific dataset.
+Consequently, the goal of this experiment is to use transfer learning on three datasets coming from three different body parts that are prostate, brain and lung, in order to check if they share common features that could be useful to improve the performance of the model on a specific dataset.
 In this experiment, the model trained to classify prostate lesions is then trained to classify brain tumors and finally lung lesions.
@@
 Fifth, the previous step is also applied to the target dataset.
+The only difference lies in the frozen part.
+Instead of dropping the old last layers and initialize its weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model.
 These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again.
@@
 \subsection{Visualization of the impact of the various datasets on the target task}
+\label{sec:global_performance}
 On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
@@
 
-\subsection{Transfer learning}
-All training graphs and the metrics they achieved on the validation sets are shown on figures XXXX.\\
-Table \ref{table:best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment.
+%\subsection{Transfer learning}
+Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
+Each graph displays the metric values achieved on the training and the validation set on the Y-axis, whereas the X-axis represents the number of epochs.
+Table \ref{table:tl_best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment. 
+Finally, Figure \ref{fig:tl_results_test_set} compares the performance of the model resulting from the first training to the one which underwent the transfer learning experiment.
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS1_full.png}
+}
+\caption{Transfer learning - DS1 Full}
+\label{fig:tl_ds1_full}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS2_frozen.png}
+}
+\caption{Transfer learning - DS2 Frozen}
+\label{fig:tl_ds2_frozen}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS2_full.png}
+}
+\caption{Transfer learning - DS2 Full}
+\label{fig:tl_ds2_full}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS3_frozen.png}
+}
+\caption{Transfer learning - DS3 Frozen}
+\label{fig:tl_ds3_frozen}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS3_full.png}
+}
+\caption{Transfer learning - DS3 Full}
+\label{fig:tl_ds3_full}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS4_frozen.png}
+}
+\caption{Transfer learning - DS4 Frozen}
+\label{fig:tl_ds4_frozen}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS4_full.png}
+}
+\caption{Transfer learning - DS4 Full}
+\label{fig:tl_ds4_full}
+\end{figure}
 
@@
 \caption{Best model performance at each step of the transfer learning pipeline}
-\label{table:best_model_performance}
+\label{table:tl_best_model_performance}
 \end{table}
+
 \section{Discussion}
+
+\subsection{Training}
+This transfer learning experiment was performed manually.
+This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.
+
+First of all, the best weight initialization among 200 random initializations was loaded and trained on the PROSTATEx dataset.
+With a learning rate of $1*10^{-8}$ and a dropout of $0.4$, the model learned genuine features in a smooth way (accuracy of $0.6340$ and AUC of $0.7359$ on the validation set).
+Obviously, the values are lower than the ones reached in Section \ref{sec:paper_reproduction_results} (accuracy of $0.7525$ and AUC of $0.765$) since the current experiment only uses DWI images.
+Hence, this model could only learn from one third of the information the first model had at disposal.
+Furthermore, the training and validation curves remained close during the whole experiment.
+This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.
+
+Second, the first frozen training took place on the brain dataset.
+During the first epochs, the metrics remained low.
+For example, the accuracy was below $0.5$, which means that a model classifying the input randomly would have performed better.
+This was expected since the last layers were reset to new random initialization.
+In these early stages, the model was relying on a feature extractor capable of classifying prostate cancer and a new decision maker which did not have the time to adapt to the new task.
+Thanks to a learning rate of $1*10^{-7}$ and a dropout probability of $0.3$, the metrics started increasing early, reaching impressive values (accuracy of $0.7799$ and AUC of $0.8379$).
+Since the feature extractor is frozen, this behavior indicates that the features allowing to detect prostate cancer transfer well to brain cancer.
+Once more, the training and validation curves have the same shape for every metric, which implies a healthy learning process.
+% Mention no reducelr ?
+
+Then, the full model was trained on the same dataset.
+The model resulting from the previous set was composed of decision maker trained on this dataset and a feature extractor resulting from the training on the prostate dataset (whose features seem to suit this dataset well).
+Therefore, the first metric values are already high.
+The rest of the training aimed at learning new useful features from the brain images, which could then be used for lung and prostate cancer classification.
+As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).
+
+The model was then frozen again and trained on the third dataset composed of lung images.
+Here again, the metrics peaked early.
+This is partly due to the larger learning rate of $1*10^{-5}$ and to the common features that prostate, brain and lung cancers share.
+The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.
+
+The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.
+
+Finally, the model was trained on the first dataset again.
+At the beginning of the frozen part, the last layers were not initialized randomly.
+In fact, the last layers coming from the model resulting from the training on the first dataset were used as last layers.
+Since these last layers were already trained on the same dataset, the improvement occuring during this training stemed from new features learned from the brain and lung datasets.
+Even with a large learning rate of $1*10^{-5}$, the curves went up slowly and regularly throughout the whole training.
+Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.
+
+Regarding the full training, the curves contain quite some noise with a already small learning rate of $1*10^{-9}$.
+An even smaller learning rate of $1*10^{-10}$ was tested and it turned out to be too small since the training and validation curves were completely flat from the beginning to the end.
+Here, the validation curves show an improvement during the first 50 epochs approximately, before going down while the training metrics keeps increasing.
+At this point, the model reached its optimum.
+Moreover, the fact that the validation metrics are much higher than the training ones with a dropout probability of 0.0 indicates that the model generalizes well thanks to transfer learning.
+
+\subsection{Global performance}
+The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task (see Section \ref{sec:global_performance}).
+In this case, they show how much the features learned from each dataset help classify prostate lesions.
+As a reminder, a temporary model was created and tested at the end of each epoch.
+The latter was composed of the current feature extractor attached to the decision maker of the best model resulting from the first training on the prostate dataset.
+
+Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
+As expected, the first training on PROSTATEx generated the same curve as on \ref{fig:tl_ds1_full}).
+The latter looks steeper because of the difference in the X-axis scale.\\
+Then, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line.
+Reason for that is that the feature extractor and the decision maker do not change here.
+In fact, the former is frozen and the latter is always the same.
+As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
+\\
+Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.\\
+During the full training on the lung dataset, the AUC starts at a high value after the first epoch.
+The combination of the prostate, brain and lung features seem to help the model generalize.
+This hypothesis is confirmed by the values reached during last training on the prostate dataset.
+In fact, the AUC value is higher than the one reached during the first training.
+Interestingly, the curve starts high, improves for 50 epochs and starts decreasing.
+This implies that the model reached a high level of generalization during transfer learning.
+These 50 epochs were enough to keep the low level features learned thanks to the other datasets while adapting the model to the target task again.
+
+%To draw these graphs, a temporary model was created and tested at the end of each epoch.
+The latter is composed of the current feature extractor attached to the decision maker of the best model resulting from the first training on the prostate dataset.
+Doing so ensures that the model is able to classify prostate cancer (thanks to the trained decision maker) but making use of the features learned during transfer learning exclusively.\\
+%The frozen steps are flat lines since the feature extractor and the decision maker do not change here. In fact, the former is frozen and the latter is always the same.
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_auc.png}
+}
+\caption{Transfer learning - Global Performance - Validation AUC}
+\label{fig:tl_global_validation_auc.png}
+\end{figure}
+
+\subsection{Testing}
+To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
+Figure \ref{fig:tl_results_test_set} summarizes the various metrics.
+To obtain these numbers, both models were tested on our test set which has never been seen by the model before.
+The transfer learning allowed to increase every single metric.
+First, the AUC increased by approximately 10\%, going from $0.68$ to $0.77$.
+The recall went from $0.5$ to $0.75$.
+Both metrics are key metrics in a medical context.
+In fact, classifying a positive case as negative is far worse than the opposite.
+Hence, maximizing the AUC (which is an indicator of how well the model is able to differentiate each class) and the recall (which shows how many positive cases have been correctly detected by the model).
+Apart from these metrics, the accuracy gained around 10\% ($0.68$ to $0.74$).
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_results_test_set.png}
+}
+\caption{Transfer learning - Performance comparison between the best model obtained during DS1 Full and DS4 Full}
+\label{fig:tl_results_test_set}
+\end{figure}
+
+\subsection{Conclusion}
+This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
+Moreover, it also increased the generalization ability of the model by learning features coming from different body parts.
+This makes the model more robust and more likely to classify new unseen data well.
+These improvements are a huge step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.
+
+----------------------------------------------------------\\
 1) All steps of the transfer learning are correctly tuned\\
@@
 2) Features of brain and lung images are useful to classify prostate lesions\\
-	--> Explain that the frozen curve immediately start to increase, and then get stable. Furthermore they allow us to reach good performance metrics, describe them.\\
-	--> Explain the global performance graph in details\\\\
+	--> Explain that the frozen curve immediately starts to increase, and then gets stable. Furthermore they allow us to reach good performance metrics, describe them.\\
+	--> Explain the global performance graph in detail\\\\
 3) Transfer learning improves the performance of the target model\\

### 2020-03-02 Common/structure/info.tex
Julien and Johan: title draft + remake figure

@@
 
-\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach Using Multiple Body Parts} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname

### 2020-03-03 Common/chapters/deep_learning.tex
Johan and Julien: First part of final corrections

@@
 \begin{itemize}
-\item Accuracy: ratio of the correctly labeled subjects to the whole pool of subjects.
+\item Accuracy: "Ratio of the correctly labeled subjects to the whole pool of subjects".
 \begin{equation}
@@
 
-\item Precision: ratio of the correctly positive labeled subjects by the model to all positive labeled subjects.
+\item Precision: "Ratio of the correctly positive labeled subjects by the model to all positive labeled subjects".
 \begin{equation}
@@
 
-\item Recall (sensitivity):
+\item Recall (sensitivity): "Ratio of the correctly positive labeled subjects to all subjects whose class is actually positive".
+
 \begin{equation}
@@
 
-\item F1-score:
+\item F1-score: Harmonic mean(average) of the precision and recall.
 \begin{equation}
-F1-score = \frac{2* recall * precision}{(recall + precision)}
+F1\operatorname{-}score = \frac{2* recall * precision}{(recall + precision)}
 \end{equation}
@@
 
-\item Specificity:
+\item Specificity: "Ratio of the correctly negative labeled subjects to all subjects whose class is actually negative".
 \begin{equation}
@@
 \item ROC curve and AUC:
-As explained by Sarang Narkhede, "The ROC curve is plotted with recall against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
-AUC - ROC curve is a performance measurement for classification problem at various thresholds settings.
+As explained by Sarang Narkhede ~\cite{26}, "the ROC curve is plotted with the true positive rate (=recall) against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
+AUC-ROC curve is a performance measurement for classification problem at various thresholds settings.
 ROC is a probability curve and AUC represents degree or measure of separability.
 It tells how much model is capable of distinguishing between classes.
-Higher the AUC, better the model is at predicting 0s as 0s and 1s as 1s" \cite{26}.
+Higher the AUC, better the model is at predicting 0s as 0s and 1s as 1s".
 
@@
 \begin{itemize}
-\item Higher start: the initial skill (before refining the model) on the source model is higher than it otherwise would be.
-\item Higher slope: the rate of skill improvement during training of the source model is steeper than it otherwise would be.
-\item Higher asymptote: the converged skill of the trained model is better than it otherwise would be.
+\item Higher start: The initial skill (before refining the model) on the source model is higher than it otherwise would be.
+\item Higher slope: The rate of skill improvement during training of the source model is steeper than it otherwise would be.
+\item Higher asymptote: The converged skill of the trained model is better than it otherwise would be.
 \end{itemize}

### 2020-03-03 Common/chapters/introduction.tex
Johan and Julien: First part of final corrections

@@
 An early detection and treatment increases the chances of recovery.
-In this context, Computer-Aided Diagnosis (CADx) systems can play a massive role by preventing health professionals from missing positive diagnoses.
+In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses.
 Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.
@@
 Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians.
-In fact, they provide multiple advantages.
-
+A complete CAD system is composed of two parts.
+The first one is called CADe for "computer-aided detection".
+This part includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection.
+The other part is CADx for "computer-aided diagnosis" which aims at extracting the characteristics of lesions to classify them according to their malignancy.
+CAD systems provide multiple advantages.\\
 First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses.
@@
 Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection.
-As stated before, detecting cancers at an early stage maximizes the probability of curing them.\\
+As stated before, detecting cancers at an early stage maximizes the probability of curing them.
 Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on.
 In fact, in order to built efficient deep learning models, large amounts of data are required.
-Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.
-
+Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.\\
 However, these advantages remain partially theoretical nowadays.
@@
 \section{Contributions}
-This work proposes a state of the art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field.
+This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field.
 Concretely, the major contributions of this work are:
 \begin{itemize}
+\item A state-of-the-art deep learning model for prostate cancer detection.
+All steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented.
+This makes the experiment completely reproductible.
+
+\item A transfer learning pipeline that makes use of different body parts to increase the classification performance on one chosen body part.
+This technique is a concrete solution to overcome the lack of available data for each body part.
+Furthermore, it increases the generalization ability of the neural network.
 
 \item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
-This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.
-Also, the class imbalance problem is taked into account.
+This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images into multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.
+Also, the class imbalance problem is taken into account.
 The processing scripts automatically augment each class in order to balance the number of elements.
@@
 
-\item A state of the art deep learning model for prostate cancer detection.
-All steps are clearly described, the corresponding scripts are available on GitHub and the raw results are presented.
-This makes the experiment completely reproductible.
-
 \item Processing scripts for the PROSTATEx challenge.
@@
 
-\item A transfer learning pipeline that makes use of different body parts to increase the classification performance on one chosen body part.
-This technique is a concrete solution to overcome the lack of available data for each body part.
-Furthermore, it increases the generalization ability of the neural network.
-
+\end{itemize}
+Other minor contributions:
+\begin{itemize}
 \item Verification scripts to check the gradient flow of a neural network, the cropping of images, the presence of NaN values in images, etc. All these scripts can easily be used in other projects.

### 2020-03-03 Common/chapters/literature_review.tex
Johan and Julien: First part of final corrections

@@
 \label{ch:literature_review}
-
-Traditionally, building a computer-aided cancer diagnosis or detection system consisted of two different parts: "Lesion  detection  and  false-positive  reduction.
+%Traditionally, building a CAD system consisted of two different steps: "Lesion  detection (CADe) and  false-positive  reduction (CADx).
 Lesion  detection  was  primarily  based  on  algorithms  specific  to  the  detection task, resulting in many candidate lesions.
@@
 However,  even  with  these  complicated and sophisticated programs, the general performance of conventional CAD systems was not good, thus hampering their widespread usage in clinical practice"~\cite{41}.\\
-Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks.
-In fact, the latter can be considered as single-step CADs.
+%Nowadays, these two steps tend to disappear thanks to deep learning algorithms, in particular the ones based on convolutional neural networks.
+In fact, the latter can be considered as single-step CAD system.
 For this reason, CAD systems are currently a trending topic in deep learning.\\
-There mainly exists two types of CAD systems based on convolutional neural networks.
-The first one is called CADe for "computer-aided detection".
-Its primary goal is "to increase the detection rate of diseases while reducing the false negative  rate  possibly  due  to  the  observers’  mistakes  or  fatigue"~\cite{41}.
-This group includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection.
-The other type is CADx for "computer-aided diagnosis" which aims at extracting the characteristics lesions.\\
-This chapter presents the main research works about CADe systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+Since this work focuses on lesions classification, this chapter presents the main research works about CADx systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
 prostate, lung and brain.

### 2020-03-03 Common/chapters/medical.tex
Johan and Julien: First part of final corrections

@@
 
-\section{Medical imaging file formats}
-\subsection{Types of medical imaging}
+\section{Types of medical imaging}
 \label{sec:medical_imaging}
@@
 %, PET (Positron Emission Tomography) scans
-and mammographies.
+and mammograms.
 
@@
 
-\subsection{DICOM file format}
+\section{DICOM file format}
 \label{sec:DICOM}
 
-\subsubsection{Origin}
+\subsection{Origin}
 
@@
 
-\subsubsection{Data format}
+\subsection{Data format}
 
@@
 
-\subsubsection{Processing images}
+\subsection{Processing images}
 
@@
 
-\subsection{NIfTI file format}
+\section{NIfTI file format}
 
-\subsubsection{Origin}
+\subsection{Origin}
 The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
@@
 
-\subsubsection{Data format}
+\subsection{Data format}
 Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
@@
 
-\subsubsection{Overview of the header structure}
+\subsection{Overview of the header structure}
 With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
@@
 
-\subsection{RAW and MHD file formats}
+\section{RAW and MHD file formats}
 

### 2020-03-03 Common/chapters/paper_reproduction.tex
Johan and Julien: First part of final corrections

@@
 
-\chapter{Paper reproduction}
+\chapter{Research paper experiment reproduction}
 \label{ch:paper_reproduction}
@@
 \label{sec:prostatex_dataset_description}
-The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website.
+The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}.
 This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients.
@@
 Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set.
-The model whose score was the highest on a given metric was saved and used as base model for the experiments.
+The model whose score was the highest on a given metric (the AUC in this experiment) was saved and used as base model for the experiments.
 It can then be used as initialization model using the script option \textit{--modeltoload <path>}.
@@
 In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
-At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-^{7}$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
-Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the f1-score.
+At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-7$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
+Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the AUC.
 This model was implemented with PyTorch using the Adam optimizer to update its weights and was trained on an Nvidia GeForce GTX Titan X graphics card during approximately 3 hours.\\
@@
 Finally, another source of difference is the number of participants in the experiment and the level of experience they have.
-Indeed, seven authors contributed to the paper, including one bachelors' student and six PhD with one of them working for Siemens Heathcare and one for hImagingTek Ltd.
+Indeed, seven authors contributed to the paper, including one bachelors' student and six PhD with one of them working for Siemens Healthcare and one for hImagingTek Ltd.
 These big companies are specialized in the medical imaging and have departments dedicated to machine learning.

### 2020-03-03 Common/chapters/transfer_learning.tex
Johan and Julien: First part of final corrections

@@
 \subsubsection{Dataset description}
-Lung CT Challenge is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
+Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
 Each patient can have 1 or multiple findings.
@@
 \subsubsection{Dataset description}
-This dataset is publicly available on Kaggle.
+This dataset is publicly available on Kaggle~\cite{51}.
 All images come from single-sequence MRIs.
@@
 File names and file formats were heterogeneous.
+At the end of this process, the dataset was composed of 96 benign lesions and 133 malignant ones.
 

### 2020-03-03 Common/structure/info.tex
Johan and Julien: First part of final corrections

--- 2020-02-28 Common/structure/info.tex
@@
 
-\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Prostate Cancer Classification: A Freeze/Unfreeze Transfer Learning Approach Using Multiple Body Parts} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname

### 2020-03-03 Common/chapters/abstract.tex
Johan and Julien: Abstract first draft

% !TEX root = ../main.tex

Automating the detection of cancer participates to an an early detection and treatment, which increases the chances of recovery. Recent algorithms in artificial intelligence relying on deep learning have shown promising results in this field. Indeed, the usefulness of convolutional neural networks (CNNs) for segmentation or classification tasks is no longer to be proven. However, the performance of these models is often limited by the amount of available data to train the algorithm.\\
This thesis first presents a state-of-the-art convolutional neural network for prostate lesion classification. All steps from the data processing to the smallest details regarding the neural network training are explained, ensuring a complete reproducibility of the experiment. This model was evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset and achieved an AUC of 0.76. This result constitutes a solid baseline and confirms the correct functioning of our implementation.\\
On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) was built. This method shows that integrating information from diverse datasets improves automated prostate cancer diagnosis. Indeed, it appears that lesions of these different types of cancer share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier. This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancer can take advantage of it.

### 2020-03-03 Common/chapters/transfer_learning.tex
Johan and Julien: experimental setup section added in TL

@@
 \subsection{Experimental setup}
+The whole transfer learning is composed of multiple steps.
+In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
+Figure \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each time of the experiment.
+Note that, as it was the case in the research paper reproduction, the first model was also initialized picking a the best model among 200 initializations.
+The model was implemented with PyTorch using the Adam optimizer to update its weight and the complete succession of the whole process took approximately 50 hours on an Tesla v100 XSM2 graphic card.
+
+\begin{table}[]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|l|l|c|l|c|l|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\textbf{Dataset}         & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Learning rate} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Batch size}} & \textbf{Dropout} & \textbf{Number of epochs} \\ \hline
+\textbf{PROSTATEx}       & DS1/Full                                                   & 1e-8                   & 128                                                              & 0.4              & 1000                      \\ \hline
+\textbf{Kaggle Brain}    & DS2/Frozen                                                 & 1e-7                   & 128                                                              & 0.3              & 1000                      \\ \hline
+\textbf{Kaggle Brain}    & DS2/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 1000                      \\ \hline
+\textbf{LungCTChallenge} & DS3/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 1000                      \\ \hline
+\textbf{LungCTChallenge} & DS3/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 1000                      \\ \hline
+\textbf{PROSTATEx}       & DS4/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 1000                      \\ \hline
+\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0                & 1000                      \\ \hline
+\end{tabular}%
+}
+\caption{Hyperparameters used at each step of the transfer learning pipeline}
+\label{table:tl_hyperparameters_at_each_step}
+\end{table}
 

### 2020-03-04 Common/structure/info.tex
Johan and Julien: Title page updated

--- 2020-03-02 Common/structure/info.tex
@@
 \supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
-\cosupervisor{Co-supervisor Name} % If you have a co-supervisor, include it here. This is used in the title page, print it elsewhere with \supname
+\cosupervisor{\href{https://exascale.info/members/giuseppe-cuccu/}{Dr.
+Giuseppe Cuccu},\\ \href{https://exascale.info/members/akansha/}{Akansha Bhardwaj}} % If you have a co-supervisor, include it here.
+This is used in the title page, print it elsewhere with \supname
 \degree{Master} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{Author Name} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
+\author{\href{mailto:julienclement22@gmail.com}{Julien Clément}, \href{mailto:joh.job13@gmail.com}{Johan Jobin}}% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
 \addresses{Bd de Pérolles 90} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
@@
 \subject{Computer Science} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
-\keywords{keywords, list, here} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
+\keywords{Prostate cancer classification, Convolutional Neural Networks (CNN), Transfer learning, PROSTATEx, Machine learning, Artificial intelligence} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
 \university{\href{http://www.unifr.ch}{University of Fribourg}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname

### 2020-03-04 Common/structure/title.tex
Johan and Julien: Title page updated

@@
 \begin{flushleft} \large
-\emph{Author:}\\
-\href{mailto://david.bucher@unifr.ch}{\authorname} % Author name - remove the \href bracket to remove the link
+\emph{Authors:}\\
+{\authorname} % Author name - remove the \href bracket to remove the link
 \end{flushleft}
@@
 \href{https://exascale.info/phil}{\supname} % Supervisor name - remove the \href bracket to remove the link
-\\\vspace*{1ex}\emph{Co-Supervisor:} \\ % Remove these two lines if no co-supervisor is involved
-\href{https://exascale.info/members/}{\cosupname} % Co-supervisor name - remove the \href bracket to remove the link
+\\\vspace*{1ex}\emph{Co-Supervisors:} \\ % Remove these two lines if no co-supervisor is involved
+{\cosupname} % Co-supervisor name - remove the \href bracket to remove the link
 \end{flushright}
@@
 
-January 01, 1970 % date of the official defense
+March XX, 2020 % date of the official defense
 \vspace*{.06\textheight}

### 2020-03-04 Common/chapters/transfer_learning.tex
Julien and Johan: more info about the image cropping (larger patches, resizing to 65x65)

@@
 \label{sec:PROSTATEx}
+Ideally, transfer learning should rely on multi-sequence MRIs only.
+In fact, the extra information provided by different sequences increases the performance a lot.
+However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
+
 As the other datasets used in this work do not contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
@@
 
-The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from the alignment and the stacking.
-A T2-weighted image was still used to align, crop and resize the corresponding DWI image but was not exported at the end, resulting in a single channel image (NumPy array).
+The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from a few details.
+First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of stacked images.
+As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.
 
-Ideally, transfer learning should rely on multi-sequence MRIs only.
-In fact, the extra information provided by different sequences increases the performance a lot.
-However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
+Second, the way of cropping images was slighty modified.
+Previously, a 65x65px patch centered on the lesion was cropped, which sometimes resulted in patches cropped inside the lesion (i.e.
+patches which did not contain the entire lesion).
+As the two missing channels reduce the amount of information contained in each image, cropping inside lesions complicates the work of the neural network.
+Therefore, a larger patch of 130x130px was cropped, before being resized to 65x65px using a bicubic interpolation.
+These dimensions were found to be the best compromise across the whole dataset since they allowed to crop the vast majority of lesions with a small tissue margin around them.
+Consequently, the borders of the lesions are also visible on the images, which should help the detection process.
 
@@
 Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
-Each patient can have 1 or multiple findings.
+Each patient can have one or multiple findings.
 Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
@@
 The augmentation process is close to the one applied to PROSTATEx.
-An advantage Lung CT Challenge has it that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless.
+An advantage Lung CT Challenge has is that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless.
 Nevertheless, lung tumors are particular.
@@
 For this reason, these large tumors were treated as outliers.
-Lastly, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives).
-Therefore, each class was augmented the same amount of times (240x).
+The chosen dimensions are 1/6 of the image width and height (i.e.
+85x85px since the image size is 512x512px).
+These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.
 
-Images were normalized with respect to equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
+Also, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives). Therefore, each class was augmented the same amount of times (240x).
+
+Lastly, images were normalized with respect to Equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
 
@@
 were visible.
-These objects were manually removed using the object removal tool in Photoshop.
+These objects were manually removed using the object removal tool in Photoshop or by cropping the image if the disturbing object was far from the skull.
 File names and file formats were heterogeneous.
@@
 The particularity of this dataset is that no information apart from the clinical significance is provided.
-Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied on this one.
+Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.
 
@@
 The "PatientID" is the filename without extension that was set during the first step (for example "$yes99$" or "$no3$").
-The "Nodule Center Position" is the pixel coordinate of the lesion.
+The "Nodule Center Position" represents the pixel coordinates of the lesion.
 Fortunately, brain tumors are clearly visible on brain MRIs, which eased the process.
@@
 First, a mere conversion involving no content modification apart from going from three to one channel was performed.
-Second, images were normalized according to equation \ref{eq:normalization}.
+Second, images were normalized according to Equation \ref{eq:normalization}.
 The mean and standard deviation were computed over the whole dataset.
-Both classes were augmented by a common factor (200x) since both class contain a relatively close number number of instances (133 positives and 96 negatives).
+Both classes were augmented by a common factor (200x) since both classes contain a relatively close number of instances (133 positives and 96 negatives).
 Similarly to Lung CT Challenge, tumor sizes vary substantially.
@@
 In fact, cropping an area centered on the lesion whose height was $\frac{2}{7}$ of the image and whose width was $\frac{1}{4}$ gave good results.
+Finally, images were also resized to 65x65px using a bicubic interpolation before being saved as NumPy arrays.
 
@@
 Multiple ways of cropping images were tested out.
-The differences lay in how much tissue was kept apart from the lesion itself.
+The differences lay in how much tissue was kept around the lesion.
 In some cases, more context around the lesion allows to capture details about the connections between the lesion and blood vessels for example.
@@
 \subsubsection{Visual checking}
-Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see section \ref{sec:braingroundtruth}).
+Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
 In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset.

### 2020-03-04 Common/chapters/acknowledgements.tex
Johan and Julien: Acknowledgements

\textbf{Common acknowledgements:}\\
First of all, we would like to thank Giuseppe Cuccu, for his valuable advices, the exciting discussions we had with him and all the time he took to supervise the project. Even when the results were not there and we were struggling to find the cause, he always encouraged us and pushed us to give our best. We are extremely grateful to him. We also would like to greatly thank Akansha Bhardwaj for her supervision throughout the project and the amazing work she did reviewing the thesis. Last but not least, we would like to express sincere thanks to Philippe Cudré-Mauroux for having supervised the project. Having the opportunity to work with such a team and having access to amazingly powerful machines was literally awesome.\\

\noindent \textbf{Julien Clément's acknowledgements:}\\
XXX\\

\noindent \textbf{Johan Jobin's acknowledgements:}\\
On a personal note, I would like to acknowledge with gratitude the support of my family, specially my parents (Jean-Bernard and Marie Jobin), my brother (Lucas Jobin), my uncle (Xavier Marchand) and my girlfriend (Julie Timmermans). They have allowed me to be in the best conditions to carry out this work and have expressed great interest in it, which gave me even more motivation. A special thanks goes to my close friends that have always encouraged me. Finally, I would like to thank Julien Clément for the great work we did together on this thesis but also for all the years we have spent at the university. The realization of this work would not have been possible without him.

### 2020-03-04 Common/chapters/introduction.tex
Johan and Julien: datasets names uniformisation

@@
 \section{Contributions}
-This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning to overcome the lack of publicly available data in the medical field.
+This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
 Concretely, the major contributions of this work are:
@@
 
-\item Processing scripts for the PROSTATEx dataset, the Lung CT Challenge dataset and the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+\item Processing scripts for the SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, the Lung CT Challenge dataset and the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
 This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images into multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.

### 2020-03-04 Common/chapters/transfer_learning.tex
Julien and Johan: new figure and minors

@@
 
-\subsubsection{Parallelization}
-Each experiment before transfer learning was performed on a single GPU.
-While being a lot faster than training on a CPU, the computation time was still relatively high.
-In order to get results from the transfer learning in a reasonable amount of time, parallelization was required.
-In fact, the longest transfer learning experiment consisted of $(5$ folds $* 4$ learning rates $* 1200$ epochs $+ 1200$ epochs$) * 7$ parts $= 176400$ epochs.
-On a single GPU, this experiment would have taken around two weeks to complete.
-
-Thanks to parallelization, the cross-validation combinations could be run simultaneously on different GPUs (5 of them since 5-fold cross-validation was used).
-To achieve this while using Python and GPUs, we relied on a library called Ray.
-The function responsible for training the model needs to be defined as a Ray remote function thanks to a decorator (@$ray.remote$ above the function declaration).
-This new function does not return the actual return values, but object IDs pointing to placeholders for the future results, which allows to continue the execution and run other parallel functions.
-Outside of the cross-validation loop, a call to $ray.get()$ on the object IDs allows to wait for the simultaneous executions to complete and to collect the different results.
+\noindent Figure \ref{fig:tl_model_split} shows the chosen model split for this experience.
+The first fully connected layer is part of the feature extractor because it can be considered as a feature gatherer and assembler.
+The next ones are part of the decision maker.
 
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.3\textwidth, keepaspectratio=true]{./figures/tl_parallel_run.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_model_split.png}
 }
-\caption{Transfer learning - Parallelization}
-\label{fig:tl_parallel_run}
+\caption{Transfer learning - Model split}
+\label{fig:tl_model_split}
 \end{figure}
+
+%\subsubsection{Parallelization}
+%Each experiment before transfer learning was performed on a single GPU.
+While being a lot faster than training on a CPU, the computation time was still relatively high.
+In order to get results from the transfer learning in a reasonable amount of time, parallelization was required.
+In fact, the longest transfer learning experiment consisted of $(5$ folds $* 4$ learning rates $* 1200$ epochs $+ 1200$ epochs$) * 7$ parts $= 176400$ epochs.
+On a single GPU, this experiment would have taken around two weeks to complete.
+%
+%Thanks to parallelization, the cross-validation combinations could be run simultaneously on different GPUs (5 of them since 5-fold cross-validation was used).
+To achieve this while using Python and GPUs, we relied on a library called Ray.
+The function responsible for training the model needs to be defined as a Ray remote function thanks to a decorator (@$ray.remote$ above the function declaration).
+This new function does not return the actual return values, but object IDs pointing to placeholders for the future results, which allows to continue the execution and run other parallel functions.
+Outside of the cross-validation loop, a call to $ray.get()$ on the object IDs allows to wait for the simultaneous executions to complete and to collect the different results.
+%
+%\begin{figure}[!h]
+%\centering
+%\noindent
+%\makebox[\textwidth]{
+%\includegraphics[width=1.3\textwidth, keepaspectratio=true]{./figures/tl_parallel_run.png}
+%}
+%\caption{Transfer learning - Parallelization}
+%\label{fig:tl_parallel_run}
+%\end{figure}
 
@@
 \subsection{Experimental setup}
-The whole transfer learning is composed of multiple steps.
+The whole transfer learning consists in multiple steps.
 In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
-Figure \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each time of the experiment.
-Note that, as it was the case in the research paper reproduction, the first model was also initialized picking a the best model among 200 initializations.
-The model was implemented with PyTorch using the Adam optimizer to update its weight and the complete succession of the whole process took approximately 50 hours on an Tesla v100 XSM2 graphic card.
+Table \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each step of the experiment.
+Note that, as it was the case in the paper experiment reproduction, the first model was also initialized by loading the best model among 200 initializations.
+The model was implemented with PyTorch using the Adam optimizer to update its weights.
+The complete succession of the process took approximately 60 hours on an Nvidia GeForce RTX 2080 Ti.
+%Tesla v100 XSM2 graphics card.
 

### 2020-03-04 Common/chapters/abstract.tex
Julien and Johan: work repartition + minors

@@
 This model was evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset and achieved an AUC of 0.76.
-This result constitutes a solid baseline and confirms the correct functioning of our implementation.\\
+This result constitutes a solid baseline and confirms the correct functioning of the implementation.\\
 On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) was built.
@@
 Indeed, it appears that lesions of these different types of cancer share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier.
-This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancer can take advantage of it.
+This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancers can take advantage of it.

### 2020-03-04 Common/chapters/acknowledgements.tex
Julien and Johan: work repartition + minors

@@
 \textbf{Common acknowledgements:}\\
-First of all, we would like to thank Giuseppe Cuccu, for his valuable advices, the exciting discussions we had with him and all the time he took to supervise the project.
-Even when the results were not there and we were struggling to find the cause, he always encouraged us and pushed us to give our best.
-We are extremely grateful to him.
-We also would like to greatly thank Akansha Bhardwaj for her supervision throughout the project and the amazing work she did reviewing the thesis.
-Last but not least, we would like to express sincere thanks to Philippe Cudré-Mauroux for having supervised the project.
-Having the opportunity to work with such a team and having access to amazingly powerful machines was literally awesome.\\
+First of all, we would like to thank Dr.
+Giuseppe Cuccu for his valuable advices, the exciting discussions we had with him and the precious time he took to supervise the project.
+Even we were struggling to find the cause of bad results for weeks, he always encouraged us and pushed us to give our best.
+We are extremely grateful to him for his support.
+We would also like to greatly thank Akansha Bhardwaj for her supervision throughout the project and the amazing work she did reviewing the thesis.
+Last but not least, we would like to express sincere thanks to Prof.
+Dr.
+Philippe Cudré-Mauroux for supervising the project.
+Having the opportunity to work with such a wonderful team and having access to amazingly powerful machines was incredible.\\
 

### 2020-03-04 Common/chapters/introduction.tex
Julien and Johan: work repartition + minors

@@
 \end{itemize}
+
+\section{Work repartition}
+This thesis was jointly written by two authors.
+The name(s) of the contributor(s) in the margins of the following chapters only concern the written thesis.
+Both of them contributed equally to the work leading to the writing of the latter.
+The work was either done together or in parallel, depending on the tasks.
+In any case, parts written by one of the authors were reread by the other.
+Finally, the order in which the names are cited does not represent the quantity of work performed and was chosen arbitrarily.

### 2020-03-04 Common/chapters/conclusions.tex
Johan and Julien: all names for each section added

@@
 \section{Conclusion}
-This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
 The starting point was the reproduction of Song et al.'s experiment~\cite{07}.
@@
 \section{Future Work}
-% The continuation of this work includes...
-
-This work mainly used a single model architecture based on a VGG-16 network.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This work mainly used a single model architecture based on a VGG-16 network.
 As MRIs are three-dimensional, using three-dimensional convolutions could improve the results.

### 2020-03-04 Common/chapters/deep_learning.tex
Johan and Julien: all names for each section added

@@
 \label{ch:deep_learning}
-This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work.
 It starts with a historical timeline of deep learning before describing what a neural network is.
@@
 \section{Introduction to deep learning}
-Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence.
 Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them.
@@
 \label{historical_background}
-As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
 From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks.
@@
 \label{what_is_a_neural_network}
-
-From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters.
 The input can be an image, a sound or any input that can be converted into numerical features.
@@
 \subsection{Supervised and unsupervised learning}
-Machine learning algorithms belongs can be classified in two classes.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Machine learning algorithms belongs can be classified in two classes.
 The first one is "supervised learning".
@@
 \subsection{Notation}
-In order to keep the mathematical description of neural network consistent, this work will use Andrew Y. Ng's notation \cite{16}, who was a pioneer in deep learning.\\
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In order to keep the mathematical description of neural network consistent, this work will use Andrew Y.
+Ng's notation \cite{16}, who was a pioneer in deep learning.\\
 
@@
 \label{perceptron}
-Perceptrons are the main components of neural networks.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Perceptrons are the main components of neural networks.
 They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}.
@@
 \label{activation_functions}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
 Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer.
@@
 \label{multilayer_perceptron}
-A multilayer perceptron is a type of artifical neural networks. Du et al. \cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}A multilayer perceptron is a type of artifical neural networks.
+Du et al.
+\cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
 input and output layers" where "the output units represent a hyperplane in the space
@@
 \label{training_a_neural_network}
-Training a neural network can be broken down into multiple steps.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Training a neural network can be broken down into multiple steps.
 The first one is the "forward propagation" step.
@@
 \subsection{Forward propagation}
-The forward propagation is used to transmit the input through the entire neural network.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The forward propagation is used to transmit the input through the entire neural network.
 Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations.
@@
 \subsection{Loss computation}
-As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
 It is therefore useful to measure the penalty for a single input.
@@
 \subsection{Backpropagation}
-Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
 Generally speaking, "the intuition behind the backpropagation algorithm is as follows.
@@
 \subsection{Metrics}
-In classification tasks, four separate output labels can occur:
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In classification tasks, four separate output labels can occur:
 \begin{itemize}
@@
 \subsection{Data}
-In deep learning, data is essential.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In deep learning, data is essential.
 As seen previously, neural networks learn features from it.
@@
 \label{weight_initialization}
-Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
 The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum.
@@
 \subsection{Hyperparameters tuning}
-Hyperparameters denote parameters that cannot be directly learned from the data.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Hyperparameters denote parameters that cannot be directly learned from the data.
 That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
@@
 
-Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
 They are particular in that they contain layers which perform a mathematical operation named convolution on the input data.
@@
 \section{Transfer learning}
-
-According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
 The model dedicated to the second task uses some or all parts of the first model (i.e.

### 2020-03-04 Common/chapters/introduction.tex
Johan and Julien: all names for each section added

@@
 
-According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
 An early detection and treatment increases the chances of recovery.
@@
 \section{Motivation}
-Cancer is a serious concern in today's society.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Cancer is a serious concern in today's society.
 An early detection is key to maximize the chances of recovery.
@@
 \section{Contributions}
-This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
 Concretely, the major contributions of this work are:
@@
 \section{Work repartition}
+
 This thesis was jointly written by two authors.

### 2020-03-04 Common/chapters/literature_review.tex
Johan and Julien: all names for each section added

@@
 For this reason, CAD systems are currently a trending topic in deep learning.\\
-Since this work focuses on lesions classification, this chapter presents the main research works about CADx systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesions classification, this chapter presents the main research works about CADx systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
 prostate, lung and brain.
@@
 \section{Prostate/PROSTATEx}
-As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
 It was predicted that there would be 1.7 million new cancer cases by 2030.
@@
 \section{Lung/Lung CT Challenge}
-Gao et al.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Gao et al.
 claimed that "lung cancer is one of the most frequent and leading causes of death all over the world.
@@
 \section{Brain/Kaggle Brain}
-Brain cancer is another major type of cancer.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Brain cancer is another major type of cancer.
 According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women.

### 2020-03-04 Common/chapters/medical.tex
Johan and Julien: all names for each section added

@@
 \label{ch:medical}
-This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly.
 First of all, some notions about cancer are described in order to understand what it is and which effects it has on the human body.
@@
 \subsection{Basics}
-An accumulation of cells forming a mass is called a tumor.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}An accumulation of cells forming a mass is called a tumor.
 These tumors are detectable thanks to medical imaging (see section \ref{sec:medical_imaging}) and other symptoms.
@@
 \\Apoptosis is the programmed death of cells, which is part of the continuous regeneration of every cell within a body. Cancerous cells survive this programmed death. 
-
 \end{itemize}
@@
 \subsection{Seriousness}
-
-Most cancers can be staged thanks to the TNM system.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Most cancers can be staged thanks to the TNM system.
 The T corresponds to the tumor size and its location; the N corresponds to whether or not the tumor has spread to draining lymph nodes; the M corresponds to the presence or absence of metastases in other parts of the body \cite{21}.
@@
 
-%\subsection{Chronological evolution}
-
 \section{Types of medical imaging}
 \label{sec:medical_imaging}
-
-Multiple types of medical imaging exist. The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Multiple types of medical imaging exist.
+The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans
 %, PET (Positron Emission Tomography) scans
@@
 \label{sec:DICOM}
-
 \subsection{Origin}
-
-The acronym DICOM stands for Digital Imaging and Communications in Medicine.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The acronym DICOM stands for Digital Imaging and Communications in Medicine.
 Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks.
@@
 \subsection{Data format}
-
-DICOM files can be viewed as containers of attributes, also called tags.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}DICOM files can be viewed as containers of attributes, also called tags.
 The values of the pixels themselves are stored under the "Pixel Data" tag.
@@
 \subsection{Processing images}
-
-When manipulating DICOM files, multiple details must be taken into account.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}When manipulating DICOM files, multiple details must be taken into account.
 
 \subsubsection{Order}
-First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
 However, this order doesn’t match the real order of the images.
@@
 \label{sec:dicom_data_manipulation}
-CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
 DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used.
@@
 Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.
-
 \end{enumerate}
@@
 \section{NIfTI file format}
-
 \subsection{Origin}
-The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
 The main problem of the latter was that it was lacking information about orientation in space.
@@
 \subsection{Data format}
-Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
 This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
@@
 \subsection{Overview of the header structure}
-With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
 “Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
@@
 \section{RAW and MHD file formats}
-
-Some datasets use a combination of RAW and MHD files.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Some datasets use a combination of RAW and MHD files.
 The latter contain metainformation about their corresponding RAW file(s) which contain the data.
@@
 \section{Visualization tools}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
 Processing data manually increases the probability of making mistakes.
@@
 \subsection{DICOM}
-Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
 However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
@@
 \subsection{NIfTI}
-As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
 When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
@@
 \subsection{RAW}
-Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e.
 a succession of 2D slices.
@@
 \section{Conversion to PNG}
-
-In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
 The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG since it makes use of the others.
@@
 \subsection{8-bit conversion}
-
-Medical data is often represented over 16 bits.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Medical data is often represented over 16 bits.
 However, exporting images to PNG require 8-bit images.

### 2020-03-04 Common/chapters/paper_reproduction.tex
Johan and Julien: all names for each section added

@@
 \label{ch:paper_reproduction}
-This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
 It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
@@
 \section{Process overview}
-Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
 The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
@@
 \label{sec:prostatex_data_processing}
-This section describes all processing operations performed on the raw data until it passes through the neural network.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This section describes all processing operations performed on the raw data until it passes through the neural network.
 All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
@@
 \label{sec:prostatex_dataset_description}
-The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}.
 This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients.
@@
 \subsubsection{Methodology}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
 The methodology described in this section is the closest replication possible of the authors' data processing methodology.
@@
 \label{sec:DICOMtoNumPy}
-Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
 Furthermore, DWI images regroup various parameters which create different subcategories.
@@
 \label{sec:numpyToAugmentedStacked}
-This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
 
@@
 \subsubsection{From NumPy arrays to augmented non-stacked images}
-Another way of processing and using the data was tested.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Another way of processing and using the data was tested.
 Instead of stacking the images of various sequences coming from the same exam, each image was treated independently.
@@
 \label{sec:dataProcessingVerification}
-
 \subsubsection{Cropping verification using red dots}
-In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
 First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file.
@@
 \subsection{Training the neural network}
-Training the neural network is the most important step of the experiment after the data processing.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Training the neural network is the most important step of the experiment after the data processing.
 This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it was trained.
@@
 \subsubsection{Architecture}
-The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
 This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes.
@@
 \subsubsection{Script options}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
 Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
@@
 \subsubsection{Tensorboard}
-Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
 \begin{itemize}
@@
 \subsubsection{Model roulette}
-As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
 Experience showed us that the exact same hyperparameters with two different initializations could give opposite results.
@@
 \subsubsection{Experimental setup}
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
 Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
@@
 \subsection{Training verification}
-In deep learning, two main problems can occur during training.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In deep learning, two main problems can occur during training.
 The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero.
@@
 \label{sec:paper_reproduction_results}
-Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
 The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
@@
 \subsection{Discussion}
-Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
 In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
@@
 \label{sec:classification_challenge}
-This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
 First, the changes needed to reach this new goal are detailed, then the raw results are presented and the latter are finally interpreted.
@@
 \subsection{Training the neural network with the whole dataset}
-The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
 It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.\\
@@
 \subsection{Results on the challenge test set}
-Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
 The range of values is contained in the interval $[0.75-0.76]$.
@@
 \subsection{Discussion}
-Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behavior of the model during the training.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behavior of the model during the training.
 All metrics reach their point of convergence at some point, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.

### 2020-03-04 Common/chapters/transfer_learning.tex
Johan and Julien: all names for each section added

@@
 \label{ch:transfer_learning}
-This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data. To achieve this, transfer learning is used.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data.
+To achieve this, transfer learning is used.\\
 First, a general definition of transfer learning is given.
@@
 \section{Goal}
-Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.\\
 This technique makes sense when the quantity of data available is not sufficient.
@@
 \label{sec:process_overview}
-The approach consists in five different steps. Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The approach consists in five different steps.
+Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
 First of all, similar datasets are collected and processed in the same way.
@@
 \label{sec:PROSTATEx}
-Ideally, transfer learning should rely on multi-sequence MRIs only.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Ideally, transfer learning should rely on multi-sequence MRIs only.
 In fact, the extra information provided by different sequences increases the performance a lot.
@@
 \subsubsection{Dataset description}
-Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
 Each patient can have one or multiple findings.
@@
 \subsubsection{From DICOM to augmented NumPy arrays}
-As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
 
@@
 \subsubsection{Dataset description}
-This dataset is publicly available on Kaggle~\cite{51}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This dataset is publicly available on Kaggle~\cite{51}.
 All images come from single-sequence MRIs.
@@
 \label{sec:braingroundtruth}
-The particularity of this dataset is that no information apart from the clinical significance is provided.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The particularity of this dataset is that no information apart from the clinical significance is provided.
 Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.
@@
 \subsubsection{From JPG to NumPy arrays}
-Images had to be converted to single-channel NumPy arrays (grayscale) so that the values could be normalized further on.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Images had to be converted to single-channel NumPy arrays (grayscale) so that the values could be normalized further on.
 First, a mere conversion involving no content modification apart from going from three to one channel was performed.
@@
 \subsection{Image cropping}
-Multiple ways of cropping images were tested out.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Multiple ways of cropping images were tested out.
 The differences lay in how much tissue was kept around the lesion.
@@
 \subsubsection{Visual checking}
-Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
 In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset.
@@
 \subsection{Architecture generalities}
-
 \subsubsection{Layer freezing}
-After each dataset switch, the model is first trained with freezed layers.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}After each dataset switch, the model is first trained with freezed layers.
 The freezed layers are those considered as feature extractor, i.e.
@@
 \subsection{Script options}
-A single Python script allows to run the entire transfer learning pipeline.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}A single Python script allows to run the entire transfer learning pipeline.
 Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc.
@@
 \label{sec:global_performance}
-On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
 As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
@@
 \subsection{Experimental setup}
-The whole transfer learning consists in multiple steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The whole transfer learning consists in multiple steps.
 In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
@@
 %\subsection{Transfer learning}
-Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
 Each graph displays the metric values achieved on the training and the validation set on the Y-axis, whereas the X-axis represents the number of epochs.
@@
 \subsection{Training}
-This transfer learning experiment was performed manually.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This transfer learning experiment was performed manually.
 This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.
@@
 \subsection{Global performance}
-The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task (see Section \ref{sec:global_performance}).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task (see Section \ref{sec:global_performance}).
 In this case, they show how much the features learned from each dataset help classify prostate lesions.
@@
 \subsection{Testing}
-To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
 Figure \ref{fig:tl_results_test_set} summarizes the various metrics.
@@
 \subsection{Conclusion}
-This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
 Moreover, it also increased the generalization ability of the model by learning features coming from different body parts.
@@
 
-----------------------------------------------------------\\
-1) All steps of the transfer learning are correctly tuned\\
-	--> Description of the shapes of the curves at each step \\
-
-Consequences:\\	
-2) Features of brain and lung images are useful to classify prostate lesions\\
-	--> Explain that the frozen curve immediately starts to increase, and then gets stable. Furthermore they allow us to reach good performance metrics, describe them.\\
-	--> Explain the global performance graph in detail\\\\
-3) Transfer learning improves the performance of the target model\\
- --> Comparison between the results on the test set before and after the TL\\
-
- 4) Summarizing the advantage of usins this pipeline \\
- --> Increase the generalization ability of the model and its performance\\
- --> Overcome the lack of data for one specific body part
+%----------------------------------------------------------\\
+%1) All steps of the transfer learning are correctly tuned\\
+%	--> Description of the shapes of the curves at each step \\
+%
+%Consequences:\\	
+%2) Features of brain and lung images are useful to classify prostate lesions\\
+%	--> Explain that the frozen curve immediately starts to increase, and then gets stable. Furthermore they allow us to reach good performance metrics, describe them.\\
+%	--> Explain the global performance graph in detail\\\\
+%3) Transfer learning improves the performance of the target model\\
+% --> Comparison between the results on the test set before and after the TL\\
+% 
+% 4) Summarizing the advantage of usins this pipeline \\
+% --> Increase the generalization ability of the model and its performance\\
+% --> Overcome the lack of data for one specific body part
+%

### 2020-03-05 Common/chapters/medical.tex
Johan and Julien: Validation process corrected + nifti reference added

@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
-“Some fields were reused, some were preserved, but ignored, and some were entirely overwritten”.
-Details about the different fields contained in the header can be found here\footnote{\url{https://brainder.org/2012/09/23/the-nifti-file-format/}}.
+M.
+Winkler confirms this fact by claiming "some fields were reused, some were preserved, but ignored, and some were entirely overwritten"~\cite{52}.
+Details about the different fields contained in the header can be found in the reference of the previous citation.
 

### 2020-03-05 Common/chapters/paper_reproduction.tex
Johan and Julien: Validation process corrected + nifti reference added

@@
 \label{sec:numpyToAugmentedStacked}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
 
@@
 The percentages represent a percentage of patients.
-In fact, each subset contains respectively 80\%, 10\% and 10\% of the 204 patients, i.e.
-$\lfloor 0.8 * 204 \rfloor = 163$ training patients, $\lfloor 0.1 * 204 \rfloor = 20$ test patients and $204 - 163 - 20 = 21$ validation patients.
+Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set.
+In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class.
+At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients.
+Class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients.
+Then, during the stacking process, some patients are not used since the three sequences are not available for them.
+These patients are simply discarded.
+At the end, the "false" class is composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class is composed of 55 training patients, 8 validation patients and 4 test patients.
 This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset.
@@
 This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
+\caption{Splitting process}
+\label{fig:paper_reproduction_split}
+\end{figure}
 

### 2020-03-05 Common/chapters/literature_review.tex
Julien and Johan: patient split figure + global performance + minors

@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesions classification, this chapter presents the main research works about CADx systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
 prostate, lung and brain.

### 2020-03-05 Common/chapters/paper_reproduction.tex
Julien and Johan: patient split figure + global performance + minors

@@
 \subsubsection{Methodology}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
-The methodology described in this section is the closest replication possible of the authors' data processing methodology.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The methodology described in this section is the closest replication possible of the authors' data processing methodology.
 Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
@@
 They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets.
-Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.
+Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class (which was augmented $60$x) in order to create balanced training and validation sets.
 
@@
 At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients.
-Class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients.
-Then, during the stacking process, some patients are not used since the three sequences are not available for them.
-These patients are simply discarded.
-At the end, the "false" class is composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class is composed of 55 training patients, 8 validation patients and 4 test patients.
+On the other hand, class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients.
+Then, during the stacking process, some patients were not used since the three sequences were not available for them.
+These patients were simply discarded.
+At the end, the "false" class was composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class was composed of 55 training patients, 8 validation patients and 4 test patients.
 This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset.
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
-\caption{Splitting process}
+\caption{Patient splitting process}
 \label{fig:paper_reproduction_split}

### 2020-03-05 Common/chapters/transfer_learning.tex
Julien and Johan: patient split figure + global performance + minors

@@
 
-\subsection{Visualization of the impact of the various datasets on the target task}
-\label{sec:global_performance}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+%\subsection{Visualization of the impact of the various datasets on the target task}
+%\label{sec:global_performance}
+%\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
 As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
@@
 In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
-
-Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
+%
+%Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
 This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
-
-At the end of the transfer learning pipeline, a graph was generated for each metric.
+%
+%At the end of the transfer learning pipeline, a graph was generated for each metric.
 These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
@@
 Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
-As expected, the curve is flat during the frozen parts.
+%As expected, the curve is flat during the frozen parts.
 Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/global_performance_example.png}
-\caption{Please replace this figure with a real example as soon as the results are available!}
-\label{fig:global_performance_example}
-\end{figure}
+%
+%\begin{figure}[!h]
+%\centering
+%\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/global_performance_example.png}
+%\caption{Please replace this figure with a real example as soon as the results are available!}
+%\label{fig:global_performance_example}
+%\end{figure}
 
@@
 
-\subsection{Global performance}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task (see Section \ref{sec:global_performance}).
+\subsection{Visualization of the impact of the various datasets on the target task}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task.
 In this case, they show how much the features learned from each dataset help classify prostate lesions.
-As a reminder, a temporary model was created and tested at the end of each epoch.
-The latter was composed of the current feature extractor attached to the decision maker of the best model resulting from the first training on the prostate dataset.
+On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
+Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e.
+the training on the first/target dataset, was saved.
+As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.\\
+Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
+This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.\\
+At the end of the transfer learning pipeline, a graph was generated for each metric.
+These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
 
 Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
-As expected, the first training on PROSTATEx generated the same curve as on \ref{fig:tl_ds1_full}).
+As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}.
 The latter looks steeper because of the difference in the X-axis scale.\\
-Then, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line.
+Moreover, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line.
 Reason for that is that the feature extractor and the decision maker do not change here.
@@
 As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
-\\
-Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.\\
+
+Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.
+
 During the full training on the lung dataset, the AUC starts at a high value after the first epoch.
@@
 These 50 epochs were enough to keep the low level features learned thanks to the other datasets while adapting the model to the target task again.
+
+%\subsection{Visualization of the impact of the various datasets on the target task}
+%\label{sec:global_performance}
+%\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
+Hence, to see how much each step was actually helping the process, the decision maker resulting from the first step of the transfer learning, i.e.
+the training on the first/target dataset, was saved.
+As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
+%
+%Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
+This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
+%
+%At the end of the transfer learning pipeline, a graph was generated for each metric.
+These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
+Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
+%As expected, the curve is flat during the frozen parts.
+Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all.
 
@@
 \caption{Transfer learning - Global Performance - Validation AUC}
-\label{fig:tl_global_validation_auc.png}
+\label{fig:tl_global_validation_auc}
 \end{figure}

### 2020-03-05 Common/structure/info.tex
Julien and Johan: keywords

@@
 \subject{Computer Science} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
-\keywords{Prostate cancer classification, Convolutional Neural Networks (CNN), Transfer learning, PROSTATEx, Machine learning, Artificial intelligence} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
+\keywords{Prostate cancer classification, Convolutional Neural Network (CNN), Transfer learning, PROSTATEx, Machine Learning (ML), Deep Learning (DL), Artificial Intelligence (AI)} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
 \university{\href{http://www.unifr.ch}{University of Fribourg}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname

### 2020-03-05 Common/chapters/conclusions.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 During the latter, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does.
-Experiments showed that: XXXXXX RESULTS XXXXXX.
-
+Experiments showed that: XXXXXX RESULTS XXXXXX.\\
 Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline.
-All these elements can serve as a baseline or as additions to future projects.
-
+All these elements can serve as a baseline or as additions to future projects.\\
 xxxxx

### 2020-03-05 Common/chapters/deep_learning.tex
Johan and Julien: Validation process corrected + nifti reference added
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 \label{perceptron}
-
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Perceptrons are the main components of neural networks.
@@
 \end{equation}
-
 Schematically speaking, a perceptron can be represented as on figure \ref{perceptron_model}.
@@
 \label{activation_functions}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
+\setlength{\marginparwidth}{2cm}\leavevmode \marginnote{\textbf{Johan}}
 Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer.
@@
 \end{equation}
-
 Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function.
@@
 How about hidden units?
-For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.
-
+For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.\\
 In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting parameters of the model.
@@
 %The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
-The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be tuned manually.
+The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to manually be tuned.
 
@@
 They are particular in that they contain layers which perform a mathematical operation named convolution on the input data.
-CNNs are mostly used in image and video analysis.
-
+CNNs are mostly used in image and video analysis.\\
 To perform a convolution, numerical input data and a "filter" are required.
@@
 \end{figure}
-
-Different parameters can change the way a convolution behaves.
+\noindent Different parameters can change the way a convolution behaves.
 First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom.

### 2020-03-05 Common/chapters/introduction.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses.
-Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.
-
+Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.\\
 Deep learning applications require a lot of data to perform well.
@@
 In the case of cancer detection and classification, only few datasets are available for each body part.
-Therefore, the idea behind this work is to make use of datasets of different body parts to improve the classification of prostate lesions.
-
-This work is divided into various parts. First, the related literature was reviewed in order to gather techniques and architectures which gave decent results.
-
+Therefore, the idea behind this work is to make use of datasets of different body parts to improve the classification of prostate lesions.\\
+This work is divided into various parts. First, the related literature was reviewed in order to gather techniques and architectures which gave decent results.\\
 Second, deep learning is presented from the ground up under the historical and technical points of view.
-This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.
-
+This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.\\
 The next chapter deals with medical knowledge related to cancer.
 Characteristics of the disease are presented, before focusing on the various file formats which medical imaging data are exported into.
-The last part is critical since a lot of data processing was performed in order to be able to make use of these files.
-
+The last part is critical since a lot of data processing was performed in order to be able to make use of these files.\\
 Then, the architecture presented in a prostate cancer classification research paper was reproduced.
-This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.
-
+This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.\\
 Finally, the last part is dedicated to transfer learning and its application to cancer classification.

### 2020-03-05 Common/chapters/literature_review.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).
-
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).\\ \\
 Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs.
@@
 The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
-However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.
-
+However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.\\ \\
 Liu et al.~\cite{46} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
@@
 This whole process allows to include 3-dimensional information in 2-dimensional images.
-The method used ensemble learning which combined different models to reach the best performance possible.
-
+The method used ensemble learning which combined different models to reach the best performance possible.\\ \\
 Mehrtash et al.~\cite{01} used a different approach.
@@
 These tricks allowed their model to achieve an AUC of $0.80$ on the PROSTATEx challenge.
-To make predictions, five different models were used, averaging the predictions of the four best models.
-
+To make predictions, five different models were used, averaging the predictions of the four best models.\\ \\
 Armato et al~\cite{42} summarized the results obtained by all teams that took part in the PROSTATEx challenge in 2017.
@@
 \section{Lung/Lung CT Challenge}
-
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Gao et al.
@@
 Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}.
-The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).
-
+The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).\\ \\
 Cengil et al.~\cite{02} built a fairly simple convolutional neural network to classify the images of the Lung CT Challenge dataset.
@@
 The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
-Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.
-
+Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.\\ \\
 Armato et al.~\cite{12} described the LUNGx Challenge and the overall results.
@@
 This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}.
-However, since it is the only dataset available for brain tumor classification, some publications used it.
-Further details about this dataset can be found in section \ref{sec:kaggleBrain}.
-
+However, as it is the only dataset available for brain tumor classification, some publications used it.
+Further details can be found in section \ref{sec:kaggleBrain}.\\ \\
 Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
@@
 The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
-The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).
-
+The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).\\ \\
 %Hanwat et al.~\cite{32} implemented a convolutional neural network to classify brain tumors of the Kaggle dataset.
@@
 ---> Check their methods (graphs look special)
-
 Habib Mohamed Ali~\cite{04} proposed his own convolutional neural network to classify the images of the Kaggle Brain dataset.

### 2020-03-05 Common/chapters/medical.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 These tumors are detectable thanks to medical imaging (see section \ref{sec:medical_imaging}) and other symptoms.
-However, not every tumor is as dangerous as the other, as it can be benign (does not contain cancerous cells) or malignant (contains cancerous cells).
-
+However, not every tumor is as dangerous as the other, as it can be benign (does not contain cancerous cells) or malignant (contains cancerous cells).\\
 The term cancer refers to different phenomena which involve mutation, abnormal multiplication and spreading of cells.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Multiple types of medical imaging exist.
-The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans
-%, PET (Positron Emission Tomography) scans
-and mammograms.
-
+The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans and mammograms.\\
 MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the generated images as a volume.
@@
 The values of the pixels themselves are stored under the "Pixel Data" tag.
-Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.
-
+Every single DICOM file usually represents a 2-dimensional image, which will form a 3-dimensional volume when put all together.\\
 Other useful information such as the patient name and ID is directly stored within the DICOM files.
@@
 Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations.
-Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.
-
+Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.\\
 The procedure depends on the tags "Window Width" and "Window Center" (one always come with the other).
 These are used to represent a range of values corresponding to the pixel values in the data.
-For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.
-
+For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.\\
 If they are missing, a simple conversion is sufficient. The parameters used to convert the data are given by two tags: 
@@
 \end{itemize}
-
+\newpage
 \noindent Examples: 
@@
 The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate” algorithm.
+M.
+Winkler affirms that this is particularly useful as "it is very common for images to have large areas of solid background, or files describing masks and regions of interest containing just a few unique values that appear repeated many time"~\cite{52}.
 
@@
 Users can then scroll through the Z-axis, displaying the next or previous slice.
+Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung at three different Z-axis coordinates.
+This allows the user to simply have a complete overview of the body part he want to analyse.
 

### 2020-03-05 Common/chapters/paper_reproduction.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
-The former makes use of the PROSTATEx training set.
+\begin{itemize}
+
+\item The former makes use of the PROSTATEx training set.
 This dataset is only composed of samples whose clinical significance is provided (labeled data).
@@
 Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance.
-This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.
+This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.\\
 As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC.
@@
 
-The second part of the experiment relies on the results of the first one.
+\item The second part of the experiment relies on the results of the first one.
 Indeed, the same data processing pipeline is used and the hyperparameters used to train the neural network are the same.
@@
 The model with the best result on the challenge is considered as the most efficient one.
+\end{itemize}
+\clearpage
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \centering
@@
 The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans.
+The PROSTATEx dataset comes with two CSV files for the training set:
+\begin{itemize}
+\item The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can be associated with the same patient (ProxID).
 
-The PROSTATEx dataset comes with two CSV files for the training set.
-The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance.
-Multiple findings can be associated with the same patient (ProxID).
-The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding.
+\item The second one, \textit{ProstateX-Images-Train.csv}, gives information about where to find the right DICOM file for each patient and each finding.
 Important labels are "ProxID" (patient ID), "fid" (finding ID $\in [1,\infty]$), "ClinSig" (clinical significance, TRUE or FALSE), "DCMSerNumber" (digit before the dash in the folder name containing DICOM files), "ijk" (position of the lesion: slice number $k$ at coordinates $(i,j)$, $i,j,k \in [0,\infty]$) and "VoxelSpacing" (3-dimensional vector representing the correspondance between a pixel and the space it occupies in the real world).
 Both CSV files are complementary to each other.
-Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}.
+\end{itemize}
+
+\noindent Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}.
 The only difference is the absence of clinical significance.
@@
 Furthermore, the authors did not mention any participation in the official PROSTATEx challenge.
+\\
 In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented.
@@
 without relying on the CSV files).
+These information are then used for the next step that consists in stacking and augmenting the images.
+Furthermore, they are also extremely useful for the visualization scripts in order to provide concrete details regarding the images.
 
@@
 
+\newpage
+
 \subsubsection{From NumPy arrays to augmented stacked images}
 \label{sec:numpyToAugmentedStacked}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
-
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.\\
 Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution.
@@
 At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel.
-This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.
-
+This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.\\
 An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e. 
-
 \begin{equation}
@@
 \end{equation}
-
 \noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient.
@@
 The same process was applied to all patients separately.
-
-Then, the dataset was augmented in the same way as the authors.
+\\ Then, the dataset was augmented in the same way as the authors.
 Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel).
@@
 For example, if the dataset follows the above-mentioned 3 negatives for 1 positive ratio, and if the script option augmentation factor is set to 50, each image belonging to the negative class is going to be augmented 50 times and each image belonging to the positive class 150 times.
-At the end, the resulting dataset is balanced and makes use of all images.
-
+At the end, the resulting dataset is balanced and makes use of all images.\\
 Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays.
@@
 \centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
+\includegraphics[width=0.95\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
 \caption{Patient splitting process}
@@
 In other words, each of them was resized as explained previously, then cropped, augmented and exported separately.
-At the end, each image was fed as separate input to the neural network.
-
+At the end, each image was fed as separate input to the neural network.\\
 This way of proceeding gave awful results.
@@
 Stacking the three sequences solves this problem.
-In this case, the first 4225 input neurons (4225 input neurons for a $65$x$65$px image) are dedicated to T2 images, the next 4225 neurons to DWI images and the last 4225 neurons to ADC images.
+%In this case, the first 4225 input neurons (4225 input neurons for a $65$x$65$px image) are dedicated to T2 images, the next 4225 neurons to DWI images and the last 4225 neurons to ADC images.
 
@@
 In addition to the previous test, an additional one aiming at checking the alignment of the images was created.
-This test ensures that the three images show the exact same content at the same spot over the three channels.
+This test ensures that the three images show the exact same content at the same spot over the three channels (see Figure \ref{fig:alignment}).
 
@@
 % If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
-\begin{table}[]
+\begin{table}[!h]
 \resizebox{\textwidth}{!}{%
@@
 \label{sec:paper_reproduction_results}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:paper_reprodution_results} shows the different metrics registered on the training and validation sets during the training phase.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:paper_reproduction_results} shows the different metrics registered on the training and validation sets during the training phase.
 The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
-This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reprodution_auc_on_our_test_set}).\\
-Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reprodution_results_challenge_1}).
+This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reproduction_auc_on_our_test_set}).\\
+Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reproduction_results_challenge_1}).
 This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
+
+\newpage
 
@@
 \centering
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results.png}
 \caption{Metrics on the training and validation sets}
-\label{fig:paper_reprodution_results}
+\label{fig:paper_reproduction_results}
 \end{figure}
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \centering
@@
 \caption{AUC of the best model on our test set}
-\label{fig:paper_reprodution_auc_on_our_test_set}
+\label{fig:paper_reproduction_auc_on_our_test_set}
 \end{figure}
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \centering
-\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
+\includegraphics[height=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
 \caption{Score achieved in the PROSTATEx challenge using only our custom training set (80\% of all available data) for the training}
-\label{fig:paper_reprodution_results_challenge_1}
+\label{fig:paper_reproduction_results_challenge_1}
 \end{figure}
@@
 Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.\\
-Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.
-
+Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.\\
 As the goal of the experiment was to reproduce the experiment of the paper, a comparison between the two models is necessary.
@@
 
+\newpage
 \section{SPIE-AAPM-NCI Prostate MR Classification Challenge}
@@
 \subsection{Results on the challenge test set}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
+\setlength{\marginparwidth}{1.5cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:challenge_all_results_experiment} presents the metric values the model reached while training on the whole dataset.
+The models whose results are reported in Table \ref{fig:results_for_all_models_challenge} were picked from this training.\\
+Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
 The range of values is contained in the interval $[0.75-0.76]$.
 Each model was tested in order to see if a slight peak of $0.01$ could be found from one epoch to the other.\\
-Figure \ref{fig:challenge_all_results_experiment} presents the metric values the model reached while training on the whole dataset.
-The models whose results are reported in Table \ref{fig:results_for_all_models_challenge} were picked from this training.\\
 Finally, Figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, which achieved an AUC of $0.76$.
+
+\newpage
 
@@
 \centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge_full_dataset.png}
+\includegraphics[width=0.95\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge_full_dataset.png}
 \caption{Metrics on the training using all data available as training set}
@@
 
-\begin{table}[]
+\begin{table}[!t]
 \resizebox{\textwidth}{!}{%
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \centering
-\includegraphics[height=0.4\textheight, keepaspectratio=true]{./figures/paper_reproduction_results_challenge2.png}
+\includegraphics[height=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge2.png}
 \caption{Score achieved in the PROSTATEx challenge by the model picked at epoch 20, trained on all available data as training set}
@@
 \end{figure}
+
+\clearpage
 

### 2020-03-05 Common/chapters/transfer_learning.tex
Johan and Julien: Thesis layout updated until transfer learning :)

@@
 Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
-First of all, similar datasets are collected and processed in the same way.
+\begin{itemize}
+\item First of all, similar datasets are collected and processed in the same way.
 These datasets need to share common features.
@@
 %On the contrary, mammographies are completely different in that they the tissue is less visible FIGURE.
-
-Second, the convolutional neural network is built and split into two parts.
+\item Second, the convolutional neural network is built and split into two parts.
 It is still a single network but the first layers are treated as the feature extractor.
@@
 They will gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
-
-Third, the model is trained on the target dataset and the model whose performance is the best is saved.
+\item Third, the model is trained on the target dataset and the model whose performance is the best is saved.
 The target dataset is the dataset which one wants to improve the performance of the model on.
 This first training establishes a baseline of performance and of features to learn further on thanks to the other datasets.
-
-Fourth, the core of transfer learning process starts. For each non-target dataset, two different trainings take place:
+\item Fourth, the core of transfer learning process starts. For each non-target dataset, two different trainings take place:
 
@@
 \end{enumerate}
-
-Fifth, the previous step is also applied to the target dataset.
+\item Fifth, the previous step is also applied to the target dataset.
 The only difference lies in the frozen part.
@@
 
+\end{itemize}
 \begin{figure}[!h]
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.2\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
 }
@@
 In fact, the extra information provided by different sequences increases the performance a lot.
-However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
-
+However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).\\
 As the other datasets used in this work do not contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
 DWI images were selected because lesions are white on this sequence, as well as on CT scans.
-Hence, choosing DWI images ensured a consistent representation of tumors from one dataset to the other, which helps the cancer detection.
-
+Hence, choosing DWI images ensured a consistent representation of tumors from one dataset to the other, which helps the cancer detection.\\
 The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from a few details.
 First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of stacked images.
-As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.
-
+As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.\\
 Second, the way of cropping images was slighty modified.
@@
 Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
-For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.
-
+For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.\\
 Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images.
@@
 \subsubsection{From DICOM to augmented NumPy arrays}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.\\
+The augmentation process is close to the one applied to PROSTATEx.
+An advantage Lung CT Challenge has is that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless.
+Nevertheless, lung tumors are particular.
+What differentiates them from the others is their look.
+In fact, lung tumors usually look like spider webs with a central nodule and branches coming out of it, whereas brain and prostate tumors tend to have well-defined boundaries.
+Therefore, a larger patch size was chosen in order to keep the extra information provided by the branches in the peripheral tissue.
+Furthermore, a perfect cropping method would only crop the interesting part of the tissue and a margin around it.
+However, tumor size varies a lot.
+It can either be extremely small (a dot) or large (massive nodule with long branches).
+Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors.
+Naturally, there were few outliers whose size was a lot larger than the chosen patch size.
+If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case.
+For this reason, these large tumors were treated as outliers.
+The chosen dimensions are 1/6 of the image width and height (i.e.
+85x85px since the image size is 512x512px).
+These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.\\
+Also, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives). Therefore, each class was augmented the same amount of times (240x).\\
+Lastly, images were normalized with respect to Equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
 
@@
 
-The augmentation process is close to the one applied to PROSTATEx.
-An advantage Lung CT Challenge has is that all images have the same resolution (512x512 pixels), making the alignment process used for PROSTATEx useless.
-Nevertheless, lung tumors are particular.
-What differentiates them from the others is their look.
-In fact, lung tumors usually look like spider webs with a central nodule and branches coming out of it, whereas brain and prostate tumors tend to have well-defined boundaries.
-Therefore, a larger patch size was chosen in order to keep the extra information provided by the branches in the peripheral tissue.
-Furthermore, a perfect cropping method would only crop the interesting part of the tissue and a margin around it.
-However, tumor size varies a lot.
-It can either be extremely small (a dot) or large (massive nodule with long branches).
-Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors.
-Naturally, there were few outliers whose size was a lot larger than the chosen patch size.
-If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case.
-For this reason, these large tumors were treated as outliers.
-The chosen dimensions are 1/6 of the image width and height (i.e.
-85x85px since the image size is 512x512px).
-These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.
-
-Also, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives). Therefore, each class was augmented the same amount of times (240x).
-
-Lastly, images were normalized with respect to Equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
-
 \subsection{Kaggle Brain}
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The particularity of this dataset is that no information apart from the clinical significance is provided.
-Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.
-
+Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.\\
 The first step was to standardize the file names and formats.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This transfer learning experiment was performed manually.
-This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.
-
+This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.\\
 First of all, the best weight initialization among 200 random initializations was loaded and trained on the PROSTATEx dataset.
@@
 Furthermore, the training and validation curves remained close during the whole experiment.
-This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.
-
+This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.\\
 Second, the first frozen training took place on the brain dataset.
@@
 Since the feature extractor is frozen, this behavior indicates that the features allowing to detect prostate cancer transfer well to brain cancer.
-Once more, the training and validation curves have the same shape for every metric, which implies a healthy learning process.
+Once more, the training and validation curves have the same shape for every metric, which implies a healthy learning process.\\
 % Mention no reducelr ?
-
 Then, the full model was trained on the same dataset.
@@
 The rest of the training aimed at learning new useful features from the brain images, which could then be used for lung and prostate cancer classification.
-As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).
-
+As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).\\
 The model was then frozen again and trained on the third dataset composed of lung images.
@@
 This is partly due to the larger learning rate of $1*10^{-5}$ and to the common features that prostate, brain and lung cancers share.
-The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.
-
-The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.
-
+The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.\\
+The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.\\
 Finally, the model was trained on the first dataset again.
@@
 Even with a large learning rate of $1*10^{-5}$, the curves went up slowly and regularly throughout the whole training.
-Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.
-
+Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.\\
 Regarding the full training, the curves contain quite some noise with a already small learning rate of $1*10^{-9}$.
@@
 These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
-If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
-
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.\\
 Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
@@
 In fact, the former is frozen and the latter is always the same.
-As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
-
-Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.
-
+As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.\\
+Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.\\
 During the full training on the lung dataset, the AUC starts at a high value after the first epoch.

### 2020-03-06 Common/chapters/paper_reproduction.tex
Johan and Julien: Hardcode thesis title in abstract (no new line) + layout finished (TL included) + undefined references removed

@@
 As the goal of the experiment was to reproduce the experiment of the paper, a comparison between the two models is necessary.
-The reproducted model achieved an AUC of $0.75$ on our test set (see Figure \ref{fig:paper_reprodution_auc_on_our_test_set}) using the authors' "enhanced prediction" technique, whereas the authors achieved an AUC $\in [0.876-0.994]$ with 95\% of confidence on their test set.
+The reproducted model achieved an AUC of $0.75$ on our test set (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}) using the authors' "enhanced prediction" technique, whereas the authors achieved an AUC $\in [0.876-0.994]$ with 95\% of confidence on their test set.
 Consequently, the model presented in this work has a slightly smaller AUC of $[0.126-0.244]$ than the one achieved in the paper.

### 2020-03-06 Common/chapters/transfer_learning.tex
Johan and Julien: Hardcode thesis title in abstract (no new line) + layout finished (TL included) + undefined references removed

@@
 Finally, the "Diagnosis" was defined as "False" or "True" according to the folder which the image was in.
+This allows to simply use the same processing techniques as it is done for the other datasets.
 
@@
 The mean and standard deviation were computed over the whole dataset.
-Both classes were augmented by a common factor (200x) since both classes contain a relatively close number of instances (133 positives and 96 negatives).
+Both classes were augmented by a common factor (200x) since both classes contain a relatively close number of instances (133 positives and 96 negatives).\\
 Similarly to Lung CT Challenge, tumor sizes vary substantially.
@@
 
-\begin{table}[]
+\begin{table}[!h]
 \resizebox{\textwidth}{!}{%
@@
 
+\newpage
+
 \begin{figure}[!h]
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS1_full.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_DS1_full.png}
 }
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
 \centering
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS2_frozen.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_DS2_frozen.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS2_full.png}
+\includegraphics[width=0.95\textwidth, keepaspectratio=true]{./figures/tl_DS2_full.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS3_frozen.png}
+\includegraphics[width=0.98\textwidth, keepaspectratio=true]{./figures/tl_DS3_frozen.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS3_full.png}
+\includegraphics[width=0.98\textwidth, keepaspectratio=true]{./figures/tl_DS3_full.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS4_frozen.png}
+\includegraphics[width=0.98\textwidth, keepaspectratio=true]{./figures/tl_DS4_frozen.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_DS4_full.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_DS4_full.png}
 }
@@
 \makebox[\textwidth]{
-\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_results_test_set.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_results_test_set.png}
 }

### 2020-03-06 Common/chapters/abstract.tex
Johan and Julien: Thesis layout updated until transfer learning :)
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

@@
 % !TEX root = ../main.tex
-
 Automating the detection of cancer participates to an an early detection and treatment, which increases the chances of recovery.
@@
 Indeed, the usefulness of convolutional neural networks (CNNs) for segmentation or classification tasks is no longer to be proven.
-However, the performance of these models is often limited by the amount of available data to train the algorithm.\\
-This thesis first presents a state-of-the-art convolutional neural network for prostate lesion classification.
+However, the performance of these models is often limited by the amount of available data to train the algorithm.
+
+This thesis first presents a state-of-the-art convolutional neural network for pro\-state lesion classification.
 All steps from the data processing to the smallest details regarding the neural network training are explained, ensuring a complete reproducibility of the experiment.
 This model was evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset and achieved an AUC of 0.76.
-This result constitutes a solid baseline and confirms the correct functioning of the implementation.\\
+This result constitutes a solid baseline and confirms the correct functioning of the implementation.
+
 On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) was built.

### 2020-03-06 Common/chapters/deep_learning.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

--- 2020-03-03 Common/chapters/deep_learning.tex
@@
 \label{ch:deep_learning}
-This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work.
 It starts with a historical timeline of deep learning before describing what a neural network is.
@@
 \section{Introduction to deep learning}
-Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence.
 Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them.
@@
 There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc.
-What mainly differentiates deep learning from these algorithms is the concept of neural networks (see Section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
+What mainly differentiates deep learning from these algorithms is the concept of neural networks (see Section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.
+
 Neural networks are inspired from the biological neural networks of the brain.
@@
 \label{historical_background}
-As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
 From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks.
-One could ask why deep learning took off recently if the theory was around for a long time.\\
+One could ask why deep learning took off recently if the theory was around for a long time.
+
 As stated by Goodfellow et al.
@@
 \label{what_is_a_neural_network}
-
-From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters.
 The input can be an image, a sound or any input that can be converted into numerical features.
 The output of a neural network depends on the problem it tries to solve.
-In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
+In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).
+
 From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
@@
 \subsection{Supervised and unsupervised learning}
-Machine learning algorithms belongs can be classified in two classes.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Machine learning algorithms belongs can be classified in two classes.
 The first one is "supervised learning".
@@
 \subsection{Notation}
-In order to keep the mathematical description of neural network consistent, this work will use Andrew Y. Ng's notation \cite{16}, who was a pioneer in deep learning.\\
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In order to keep the mathematical description of neural network consistent, this work will use Andrew Y.
+Ng's notation \cite{16}, who was a pioneer in deep learning.\\
 
@@
 \label{perceptron}
-Perceptrons are the main components of neural networks.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Perceptrons are the main components of neural networks.
 They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}.
-Today, they are called "artificial neurons" or simply "neurons".\\
+Today, they are called "artificial neurons" or simply "neurons".
+
 The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
@@
 	\begin{tikzpicture}
-
 	\node[functions] (center) {};
@@
 \label{activation_functions}
-Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer.
-Activation functions must be non-linear in order to approximate extremely complex functions.
+\setlength{\marginparwidth}{2cm}\leavevmode \marginnote{\textbf{Jobin}}
+Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function.
+The latter must be non-linear in order to approximate extremely complex functions.
 In fact, neural networks are considered as universal approximators.
@@
 \end{figure}
+
+ \newpage
 
@@
 \label{multilayer_perceptron}
-A multilayer perceptron is a type of artifical neural networks. Du et al. \cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}A multilayer perceptron is a type of artifical neural networks.
+Du et al.
+\cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
 input and output layers" where "the output units represent a hyperplane in the space
-of the input patterns."\\
-A multilayer perceptron is composed of $L$ layers, each of them composed of $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
+of the input patterns". A multilayer perceptron is composed of $L$ layers, each of them composed of $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
 \begin{itemize}
@@
 \label{training_a_neural_network}
-Training a neural network can be broken down into multiple steps.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Training a neural network can be broken down into multiple steps.
 The first one is the "forward propagation" step.
@@
 \subsection{Forward propagation}
-The forward propagation is used to transmit the input through the entire neural network.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The forward propagation is used to transmit the input through the entire neural network.
 Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations.
@@
 \subsection{Loss computation}
-As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
 It is therefore useful to measure the penalty for a single input.
@@
 The latter is represented by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.
-
-There exist multiple cost functions. For a regression problem, a commonly used loss function is the mean square error:
+There exist multiple cost functions.\\
+For a regression problem, a commonly used loss function is the mean square error:
 \begin{equation}
@@
 \subsection{Backpropagation}
-Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$.
 Generally speaking, "the intuition behind the backpropagation algorithm is as follows.
@@
 %The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
-The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be tuned manually.
+The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to manually be tuned.
 
 \subsection{Metrics}
-In classification tasks, four separate output labels can occur:
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In classification tasks, four separate output labels can occur:
 \begin{itemize}
@@
 \subsection{Data}
-In deep learning, data is essential.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In deep learning, data is essential.
 As seen previously, neural networks learn features from it.
@@
 \label{weight_initialization}
-Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation.
 The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum.
 The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
-If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.\\
+If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.
+
 The simplest and least efficient technique to initialize neural network weights is to randomly generate them.
@@
 \subsection{Hyperparameters tuning}
-Hyperparameters denote parameters that cannot be directly learned from the data.
+
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Hyperparameters denote parameters that cannot be directly learned from the data.
 That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
-So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).\\
+So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).
+
 Regarding the learning rate $\alpha$, its value has to be neither too large nor too small.
@@
 Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored" \cite{28}.
-There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).\\
+There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).
+
 Batch size is another important hyperparameter to tune.
@@
 Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be" \cite{29}.
-Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.\\
+Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.
+
 Finally, the number of epochs during which the network is trained has to be carefully chosen.
@@
 
-Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
 They are particular in that they contain layers which perform a mathematical operation named convolution on the input data.
@@
 \section{Transfer learning}
-
-According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
 The model dedicated to the second task uses some or all parts of the first model (i.e.
 keeps the same weights and architecture or a part of them) and is then retrained on data available for this task.
-The first model can either be implemented from scratch if enough data is available or it can simply be downloaded from institutions that release large pretrained models for similar tasks.\\
+The first model can either be implemented from scratch if enough data is available or it can simply be downloaded from institutions that release large pretrained models for similar tasks.
 There are three major benefits \cite{30}:

### 2020-03-06 Common/chapters/introduction.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

--- 2020-03-04 Common/chapters/introduction.tex
@@
 \label{ch:introduction}
-
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}
+According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
 An early detection and treatment increases the chances of recovery.
@@
 \section{Motivation}
-
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Cancer is a serious concern in today's society.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Cancer is a serious concern in today's society.
 An early detection is key to maximize the chances of recovery.
@@
 The other part is CADx for "computer-aided diagnosis" which aims at extracting the characteristics of lesions to classify them according to their malignancy.
-CAD systems provide multiple advantages.\\
+CAD systems provide multiple advantages.
+
 First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses.
-Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.\\
+Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.
+
 Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection.
@@
 In fact, in order to built efficient deep learning models, large amounts of data are required.
-Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.\\
+Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.
+
 However, these advantages remain partially theoretical nowadays.
@@
 \section{Contributions}
-
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
 Concretely, the major contributions of this work are:

### 2020-03-06 Common/chapters/literature_review.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

--- 2020-03-05 Common/chapters/literature_review.tex
@@
 For this reason, CAD systems are currently a trending topic in deep learning.\\
-
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
 prostate, lung and brain.
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
 It was predicted that there would be 1.7 million new cancer cases by 2030.
@@
 All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) for a total of 204 training patients and 208 challenge patients (see Section \ref{sec:prostatex_dataset_description}).
+This dataset is composed of prostate lesions multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans)(see Section \ref{sec:prostatex_dataset_description}).
 
@@
 The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
-Most of the processing techniques are reproductible, apart from the manual lesion contouring and labelling performed by a radiologist.
 Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}).
@@
 \section{Lung/Lung CT Challenge}
-
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Gao et al.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Gao et al.
 claimed that "lung cancer is one of the most frequent and leading causes of death all over the world.
@@
 Three radiologists performed statistically better than the best-performing computer method."~\cite{12}.
-Figure \ref{fig:LUNGx_challenge_all_results} gives an overview of the results.
+Figure \ref{fig:LUNGx_challenge_all_results} reports all methods used to classify the lung lesions with their associated performance.
+Furthermore, it also provides the performance of all radiologists.
 
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Brain cancer is another major type of cancer.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Brain cancer is another major type of cancer.
 According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women.
@@
 This dataset, unlike the ones for the other body parts, does not come from a certified medical authority but from the Kaggle website \cite{45}.
-However, since it is the only dataset available for brain tumor classification, some publications used it.
-Further details about this dataset can be found in section \ref{sec:kaggleBrain}.
+However, as it is the only dataset available for brain tumor classification, some publications used it.
+Further details can be found in section \ref{sec:kaggleBrain}.
 
@@
 ---> Check their methods (graphs look special)
-
 Habib Mohamed Ali~\cite{04} proposed his own convolutional neural network to classify the images of the Kaggle Brain dataset.
 First, the data was augmented.
-Then, it was cropped thanks to OpenCV so that the resulting images only contained the brain itself.
+Then, it was cropped so that the resulting images only contained the brain itself.
 Afterwards, they were resized and normalized in order to scale pixel values to the range $[0,1]$.

### 2020-03-06 Common/chapters/medical.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

--- 2020-03-05 Common/chapters/medical.tex
@@
 \label{ch:medical}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This chapter gives an overview of the basic medical knowledge that is required to apprehend the following chapters smoothly.
 First of all, some notions about cancer are described in order to understand what it is and which effects it has on the human body.
@@
 \subsection{Basics}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}An accumulation of cells forming a mass is called a tumor.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}An accumulation of cells forming a mass is called a tumor.
 These tumors are detectable thanks to medical imaging (see section \ref{sec:medical_imaging}) and other symptoms.
@@
 \subsection{Seriousness}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Most cancers can be staged thanks to the TNM system.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Most cancers can be staged thanks to the TNM system.
 The T corresponds to the tumor size and its location; the N corresponds to whether or not the tumor has spread to draining lymph nodes; the M corresponds to the presence or absence of metastases in other parts of the body \cite{21}.
@@
 \label{sec:medical_imaging}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Multiple types of medical imaging exist.
-The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans
-%, PET (Positron Emission Tomography) scans
-and mammograms.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Multiple types of medical imaging exist.
+The most commonly used to detect cancer are Magnetic Resonance Imaging (MRI), CT (Computed Tomography) scans and mammograms.
 
@@
 \subsection{Origin}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The acronym DICOM stands for Digital Imaging and Communications in Medicine.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The acronym DICOM stands for Digital Imaging and Communications in Medicine.
 Before the 1980’s, images resulting from CT scans and MRIs were only decodable by machine manufacturers, while the medical community needed to export and share them for other tasks.
@@
 \subsection{Data format}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}DICOM files can be viewed as containers of attributes, also called tags.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}DICOM files can be viewed as containers of attributes, also called tags.
 The values of the pixels themselves are stored under the "Pixel Data" tag.
@@
 \subsection{Processing images}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}When manipulating DICOM files, multiple details must be taken into account.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}When manipulating DICOM files, multiple details must be taken into account.
 
 \subsubsection{Order}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
 However, this order doesn’t match the real order of the images.
@@
 \label{sec:dicom_data_manipulation}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}CT and MRI machines, as well as monitors, differ from one manufacturer to the other and even from one model to the other.
 DICOM takes this problematic into account by providing specific tags that allow to display the exact same representation of the data, no matter the hardware used.
@@
 These are used to represent a range of values corresponding to the pixel values in the data.
-For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.
-
+For instance, a window center of 0 and a window width of 200 imply pixel values between -100 and 100.\\
 If they are missing, a simple conversion is sufficient. The parameters used to convert the data are given by two tags: 
@@
 \subsection{Origin}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
 The main problem of the latter was that it was lacking information about orientation in space.
@@
 \subsection{Data format}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
 This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
 The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate” algorithm.
+M.
+Winkler affirms that this is particularly useful as "it is very common for images to have large areas of solid background, or files describing masks and regions of interest containing just a few unique values that appear repeated many time"~\cite{52}.
 
 \subsection{Overview of the header structure}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
 M.
@@
 \section{RAW and MHD file formats}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Some datasets use a combination of RAW and MHD files.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Some datasets use a combination of RAW and MHD files.
 The latter contain metainformation about their corresponding RAW file(s) which contain the data.
@@
 \section{Visualization tools}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}
 Processing data manually increases the probability of making mistakes.
@@
 \subsection{DICOM}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
 However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
@@
 \subsection{NIfTI}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
 When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
@@
 \subsection{RAW}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Medical RAW files are three-dimensional, which means that a single file contains a volume, i.e.
 a succession of 2D slices.
@@
 \section{Conversion to PNG}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
 The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG since it makes use of the others.
@@
 \subsection{8-bit conversion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Medical data is often represented over 16 bits.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Medical data is often represented over 16 bits.
 However, exporting images to PNG require 8-bit images.

### 2020-03-06 Common/chapters/paper_reproduction.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

@@
 \label{ch:paper_reproduction}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
-It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).
+
 Song et al.
@@
 \section{Process overview}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
 The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
@@
 \label{sec:prostatex_data_processing}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This section describes all processing operations performed on the raw data until it passes through the neural network.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes all processing operations performed on the raw data until it passes through the neural network.
 All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
@@
 \label{sec:prostatex_dataset_description}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}.
 This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients.
@@
 \subsubsection{Methodology}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The methodology described in this section is the closest replication possible of the authors' data processing methodology.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The methodology described in this section is the closest replication possible of the authors' data processing methodology.
 Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
 Furthermore, the authors did not mention any participation in the official PROSTATEx challenge.
-\\
+
 In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented.
@@
 \label{sec:DICOMtoNumPy}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside.
 Furthermore, DWI images regroup various parameters which create different subcategories.
@@
 For that matter, DWI images with the highest b-value only were used.
-\\
+
 Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
@@
 \label{sec:numpyToAugmentedStacked}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}This step can be split into two substeps: aligning and stacking images before augmenting them and splitting them across a training set, a validation set and a test set.
+
 Once full images were converted to NumPy arrays, the images related to a specific lesion of a specific patient needed to be aligned and resampled to the same resolution.
@@
 At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel.
-This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.\\
+This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.
+
 An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e. 
@@
 In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array.
-Then, the mean and standard devation of the array were computed.
+Then, the mean and standard deviation of the array were computed.
 These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient.
@@
 The same process was applied to all patients separately.
-\\ Then, the dataset was augmented in the same way as the authors.
+Then, the dataset was augmented in the same way as the authors.
 Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel).
 These techniques alone allowed to create a large enough amount of data.
-Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.\\
+Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.
+
 On the other hand, authors did not mention how they managed class imbalance.
@@
 For example, if the dataset follows the above-mentioned 3 negatives for 1 positive ratio, and if the script option augmentation factor is set to 50, each image belonging to the negative class is going to be augmented 50 times and each image belonging to the positive class 150 times.
-At the end, the resulting dataset is balanced and makes use of all images.\\
+At the end, the resulting dataset is balanced and makes use of all images.
+
 Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays.
@@
 \subsubsection{From NumPy arrays to augmented non-stacked images}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Another way of processing and using the data was tested.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Another way of processing and using the data was tested.
 Instead of stacking the images of various sequences coming from the same exam, each image was treated independently.
 In other words, each of them was resized as explained previously, then cropped, augmented and exported separately.
-At the end, each image was fed as separate input to the neural network.\\
+At the end, each image was fed as separate input to the neural network.
+
 This way of proceeding gave awful results.
@@
 \subsubsection{Cropping verification using red dots}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
 First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file.
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Training the neural network is the most important step of the experiment after the data processing.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Training the neural network is the most important step of the experiment after the data processing.
 This section first describes the architecture of the model, then presents the script used for the training and finally provides details about the experimental setup such as the training procedure, the hyperparameters of the model and the configuration of the machine on which it was trained.
@@
 \subsubsection{Architecture}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
 This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes.
-Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.\\
+Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.
+
 It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks.
@@
 The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
-The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.\\
+The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.
+
 In comparison to the original VGG, this model keeps the small filter size of 3x3, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
@@
 \subsubsection{Script options}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}
 Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
@@
 \subsubsection{Tensorboard}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
 \begin{itemize}
@@
 \subsubsection{Model roulette}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}.
 Experience showed us that the exact same hyperparameters with two different initializations could give opposite results.
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}
 Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
@@
 \subsection{Training verification}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In deep learning, two main problems can occur during training.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In deep learning, two main problems can occur during training.
 The first one is called the "vanishing gradient" problem and refers to the fact that it is possible for the loss function to compute extremely small gradients, near zero.
@@
 This makes the model unstable and unable to learn from training data.
-\\
+
 As a result, it is important to analyse the gradient propagation across the network to ensure that it is not facing such problems.
@@
 It shows the name of each layer on the x-axis and its average gradient value on the y-axis (recall: for each layer the gradient is a matrix).
-Furthermore, it also gives information about the max value of the gradient and indicates in black if one of them is equal to zero.\\
+Furthermore, it also gives information about the max value of the gradient and indicates in black if one of them is equal to zero.
+
 The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6.
@@
 \label{sec:paper_reproduction_results}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:paper_reproduction_results} shows the different metrics registered on the training and validation sets during the training phase.
-The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).\\
-This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reproduction_auc_on_our_test_set}).\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Figure \ref{fig:paper_reproduction_results} shows the different metrics registered on the training and validation sets during the training phase.
+The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).
+
+This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reproduction_auc_on_our_test_set}).
+
 Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reproduction_results_challenge_1}).
@@
 \subsection{Discussion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
-In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
+In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.
+
 First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behavior during the 50 epochs.
@@
 Here, thanks to the learning rate scheduler, this scenario is avoided and the validation and training losses stay stable from this point onward.
-The same behaviour is clearly noticeable for all other metrics.\\
+The same behaviour is clearly noticeable for all other metrics.
+
 Regarding the overall performance of the best model chosen at epoch 21, the latter generalizes well.
@@
 Consequently, this model is almost at the level of a radiologist, which is very promising.
-It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results can be reached with more data.\\
-The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the hypothesis that the model has a good distinction capacity with a score of approximately $0.76$ on the validation set.\\
+It is also important to keep in mind that the model was trained using only 164 patients (80\% of the 204 training patients), which suggests that even better results can be reached with more data.
+
+The precision plot, which is used to measure the proportion of malignant lesions correctly classified among all lesions classified as malignant lesions, confirms the hypothesis that the model has a good distinction capacity with a score of approximately $0.76$ on the validation set.
+
 Another extremely important aspect of a computer-aider diagnosis system is to avoid the number of false negatives as much as possible.
@@
 This concept is measured by the recall, which quantifies the proportion of malignant lesions correctly classified among all malignant lesion.
-The model has a recall of around $0.74$ on the validation set.\\
-The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$. \\
-Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.\\
-Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.\\
+The model has a recall of around $0.74$ on the validation set.
+
+The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$.
+
+Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.
+
+Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.
+
 As the goal of the experiment was to reproduce the experiment of the paper, a comparison between the two models is necessary.
@@
 Multiple factors can explain this difference.
+
 First, in their work, the cropping of lesions was performed by a qualified radiologist, whereas, in our case, it was automatically done using the information in the CSV files.
@@
 \label{sec:classification_challenge}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This section aims at maximizing the result on the SPIE-AAPM-NCI Prostate MR classification Challenge test set by using the entire dataset as training set, instead of using 80\% of it, as it was done in the previous section.
 First, the changes needed to reach this new goal are detailed, then the raw results are presented and the latter are finally interpreted.
@@
 \subsection{Training the neural network with the whole dataset}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
-It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
+It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.
+
 As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data.
@@
 \subsection{Results on the challenge test set}
-\setlength{\marginparwidth}{1.5cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:challenge_all_results_experiment} presents the metric values the model reached while training on the whole dataset.
-The models whose results are reported in Table \ref{fig:results_for_all_models_challenge} were picked from this training.\\
+\setlength{\marginparwidth}{1.5cm}\leavevmode \marginnote{\textbf{Jobin}}Figure \ref{fig:challenge_all_results_experiment} presents the metric values the model reached while training on the whole dataset.
+The models whose results are reported in Table \ref{fig:results_for_all_models_challenge} were picked from this training.
+
 Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
 The range of values is contained in the interval $[0.75-0.76]$.
-Each model was tested in order to see if a slight peak of $0.01$ could be found from one epoch to the other.\\
+Each model was tested in order to see if a slight peak of $0.01$ could be found from one epoch to the other.
+
 Finally, Figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, which achieved an AUC of $0.76$.
@@
 \subsection{Discussion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behavior of the model during the training.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Figure \ref{fig:challenge_all_results_experiment} illustrates once again the correct behavior of the model during the training.
 All metrics reach their point of convergence at some point, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.
@@
 It is noticeable that the model did not reach its best performance before epoch 15 and that it started overfitting after epoch 23.
-Every model between these two epochs achieved an AUC of $0.76$ on the challenge, which seems to be the optimum value with this data and this configuration.\\
+Every model between these two epochs achieved an AUC of $0.76$ on the challenge, which seems to be the optimum value with this data and this configuration.
+
 In comparison to the model of the previous section trained on 80\% of the available data, this new one improves the performance by $0.05$ (AUC of $0.76$ vs $0.71$).
-This is a clear indication that the more data there is, the better the model performs.\\
+This is a clear indication that the more data there is, the better the model performs.
+
 Finally, regarding the overall leaderboard of all challenge participants at the time of the challenge, the model would have been well ranked.

### 2020-03-06 Common/chapters/transfer_learning.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

@@
 \label{ch:transfer_learning}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data.
 To achieve this, transfer learning is used.\\

### 2020-03-06 Common/structure/info.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning

@@
 \degree{Master} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{\href{mailto:julienclement22@gmail.com}{Julien Clément}, \href{mailto:joh.job13@gmail.com}{Johan Jobin}}% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
+\author{\href{mailto:julienclement22@gmail.com}{Julien Clément}\\ \href{mailto:joh.job13@gmail.com}{Johan Jobin}}% Your name, this is used in the title page and abstract, print it elsewhere with \authorname
 \addresses{Bd de Pérolles 90} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

### 2020-03-06 Common/chapters/conclusions.tex
Johan and Julien: new indents  + surnames finished

--- 2020-02-24 Common/chapters/conclusions.tex
@@
 \section{Conclusion}
-This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This work presented the process leading to the development of a deep learning system to classify potentially cancerous lesions, as well as strategies to overcome field-related issues such as the lack of data.
+
 The starting point was the reproduction of Song et al.'s experiment~\cite{07}.
@@
 This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge, which confirms the robustness of the latter.
-\\
+
 Then, the work focused on overcoming one of the main issues in deep learning: the lack of data.
@@
 \section{Future Work}
-% The continuation of this work includes...
-
-This work mainly used a single model architecture based on a VGG-16 network.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}This work mainly used a single model architecture based on a VGG-16 network.
 As MRIs are three-dimensional, using three-dimensional convolutions could improve the results.

### 2020-03-06 Common/chapters/transfer_learning.tex
Johan and Julien: new indents  + surnames finished

@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}This chapter proposes a technique to improve cancer detection and classification despite the lack of publicly available medical data.
-To achieve this, transfer learning is used.\\
+To achieve this, transfer learning is used.
+
 First, a general definition of transfer learning is given.
@@
 \section{Goal}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.
+
 This technique makes sense when the quantity of data available is not sufficient.
@@
 must be collected and put together.
-This whole process requires considerable effort, which is one of the reasons why usable medical data is hard to find.\\
+This whole process requires considerable effort, which is one of the reasons why usable medical data is hard to find.
+
 %In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
@@
 \label{sec:process_overview}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The approach consists in five different steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The approach consists in five different steps.
 Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
@@
 \label{sec:PROSTATEx}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Ideally, transfer learning should rely on multi-sequence MRIs only.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Ideally, transfer learning should rely on multi-sequence MRIs only.
 In fact, the extra information provided by different sequences increases the performance a lot.
-However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).\\
+However, publicly available datasets either do not contain the clinical significance of the samples or were purely made for segmentation purposes (where labels are also images).
+
 As the other datasets used in this work do not contain multi-sequence exams (single-sequence MRIs and CT scans), it was required to extract single channel images from the PROSTATEx dataset.
 DWI images were selected because lesions are white on this sequence, as well as on CT scans.
-Hence, choosing DWI images ensured a consistent representation of tumors from one dataset to the other, which helps the cancer detection.\\
+Hence, choosing DWI images ensured a consistent representation of tumors from one dataset to the other, which helps the cancer detection.
+
 The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from a few details.
 First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of stacked images.
-As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.\\
+As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.
+
 Second, the way of cropping images was slighty modified.
@@
 \subsubsection{Dataset description}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Lung CT Challenge ~\cite{12, 36, 38} is composed of two different subdatasets of CT scans: one is called "calibration set" (10 patients) and the other "test set" (60 patients).
 Each patient can have one or multiple findings.
 Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
-For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.\\
+For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.
+
 Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images.
@@
 \subsubsection{From DICOM to augmented NumPy arrays}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
+
 The augmentation process is close to the one applied to PROSTATEx.
@@
 85x85px since the image size is 512x512px).
-These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.\\
-Also, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives). Therefore, each class was augmented the same amount of times (240x).\\
-Lastly, images were normalized with respect to Equation \ref{eq:normalization}. The mean and standard deviation were computed over the whole dataset.
+These patches were then resized to 65x65px thanks to a bicubic interpolation, before being exported as NumPy arrays.
+
+Also, and contrary to PROSTATEx, the dataset was more or less balanced (39 positives and 42 negatives).
+Therefore, each class was augmented the same amount of times (240x).
+Lastly, images were normalized with respect to Equation \ref{eq:normalization}.
+The mean and standard deviation were computed over the whole dataset.
 
@@
 \subsubsection{Dataset description}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This dataset is publicly available on Kaggle~\cite{51}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This dataset is publicly available on Kaggle~\cite{51}.
 All images come from single-sequence MRIs.
@@
 \label{sec:braingroundtruth}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The particularity of this dataset is that no information apart from the clinical significance is provided.
-Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The particularity of this dataset is that no information apart from the clinical significance is provided.
+Therefore, it was necessary to create a ground truth CSV file from scratch so that the same processing methods as the other datasets could be applied to this one.
+
 The first step was to standardize the file names and formats.
@@
 \subsubsection{From JPG to NumPy arrays}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Images had to be converted to single-channel NumPy arrays (grayscale) so that the values could be normalized further on.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Images had to be converted to single-channel NumPy arrays (grayscale) so that the values could be normalized further on.
 First, a mere conversion involving no content modification apart from going from three to one channel was performed.
@@
 The mean and standard deviation were computed over the whole dataset.
-Both classes were augmented by a common factor (200x) since both classes contain a relatively close number of instances (133 positives and 96 negatives).\\
+Both classes were augmented by a common factor (200x) since both classes contain a relatively close number of instances (133 positives and 96 negatives).
+
 Similarly to Lung CT Challenge, tumor sizes vary substantially.
@@
 \subsection{Image cropping}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Multiple ways of cropping images were tested out.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Multiple ways of cropping images were tested out.
 The differences lay in how much tissue was kept around the lesion.
@@
 \subsubsection{Visual checking}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
 In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset.
@@
 \subsubsection{Layer freezing}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}After each dataset switch, the model is first trained with freezed layers.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers.
 The freezed layers are those considered as feature extractor, i.e.
@@
 \subsection{Script options}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}A single Python script allows to run the entire transfer learning pipeline.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}A single Python script allows to run the entire transfer learning pipeline.
 Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc.
@@
 \subsection{Experimental setup}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The whole transfer learning consists in multiple steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The whole transfer learning consists in multiple steps.
 In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
@@
 %\subsection{Transfer learning}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan \& Julien}}Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
 Each graph displays the metric values achieved on the training and the validation set on the Y-axis, whereas the X-axis represents the number of epochs.
@@
 \subsection{Training}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This transfer learning experiment was performed manually.
-This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning experiment was performed manually.
+This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.
+
 First of all, the best weight initialization among 200 random initializations was loaded and trained on the PROSTATEx dataset.
@@
 Furthermore, the training and validation curves remained close during the whole experiment.
-This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.\\
+This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.
+
 Second, the first frozen training took place on the brain dataset.
@@
 Since the feature extractor is frozen, this behavior indicates that the features allowing to detect prostate cancer transfer well to brain cancer.
-Once more, the training and validation curves have the same shape for every metric, which implies a healthy learning process.\\
+Once more, the training and validation curves have the same shape for every metric, which implies a healthy learning process.
+
 % Mention no reducelr ?
@@
 The rest of the training aimed at learning new useful features from the brain images, which could then be used for lung and prostate cancer classification.
-As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).\\
+As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).
+
 The model was then frozen again and trained on the third dataset composed of lung images.
@@
 This is partly due to the larger learning rate of $1*10^{-5}$ and to the common features that prostate, brain and lung cancers share.
-The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.\\
-The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.\\
+The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.
+
+The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.
+
 Finally, the model was trained on the first dataset again.
@@
 Even with a large learning rate of $1*10^{-5}$, the curves went up slowly and regularly throughout the whole training.
-Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.\\
+Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.
+
 Regarding the full training, the curves contain quite some noise with a already small learning rate of $1*10^{-9}$.
@@
 \subsection{Visualization of the impact of the various datasets on the target task}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task.
 In this case, they show how much the features learned from each dataset help classify prostate lesions.
@@
 As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
-In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.\\
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
+
 Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
-This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.\\
+This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
+
 At the end of the transfer learning pipeline, a graph was generated for each metric.
 These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
-If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.\\
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
+
 Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
 As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}.
-The latter looks steeper because of the difference in the X-axis scale.\\
+The latter looks steeper because of the difference in the X-axis scale.
+
 Moreover, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line.
@@
 In fact, the former is frozen and the latter is always the same.
-As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.\\
-Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.\\
+As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
+
+Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.
+
 During the full training on the lung dataset, the AUC starts at a high value after the first epoch.
@@
 \subsection{Testing}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
 Figure \ref{fig:tl_results_test_set} summarizes the various metrics.
@@
 \subsection{Conclusion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
 Moreover, it also increased the generalization ability of the model by learning features coming from different body parts.
@@
 These improvements are a huge step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.
-
-%----------------------------------------------------------\\
-%1) All steps of the transfer learning are correctly tuned\\
-%	--> Description of the shapes of the curves at each step \\
-%
-%Consequences:\\	
-%2) Features of brain and lung images are useful to classify prostate lesions\\
-%	--> Explain that the frozen curve immediately starts to increase, and then gets stable. Furthermore they allow us to reach good performance metrics, describe them.\\
-%	--> Explain the global performance graph in detail\\\\
-%3) Transfer learning improves the performance of the target model\\
-% --> Comparison between the results on the test set before and after the TL\\
-% 
-% 4) Summarizing the advantage of usins this pipeline \\
-% --> Increase the generalization ability of the model and its performance\\
-% --> Overcome the lack of data for one specific body part
-%

### 2020-03-06 Common/chapters/transfer_learning.tex
Julien and Johan: new metrics with 2000 epochs + introduce global performance before "discussion"

@@
 
+\subsection{Visualization of the impact of the various datasets on the target task}
+\label{sec:global_performance}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task.
+On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
+As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
+Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e.
+the training on the first/target dataset, was saved.
+As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
+
+Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
+This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
+
+At the end of the transfer learning pipeline, a graph was generated for each metric.
+These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
+If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
+In this case, the graphs show how much the features learned from each dataset help classify prostate lesions.
+
 %\subsection{Visualization of the impact of the various datasets on the target task}
@@
 \textbf{Dataset}         & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Learning rate} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Batch size}} & \textbf{Dropout} & \textbf{Number of epochs} \\ \hline
-\textbf{PROSTATEx}       & DS1/Full                                                   & 1e-8                   & 128                                                              & 0.4              & 1000                      \\ \hline
-\textbf{Kaggle Brain}    & DS2/Frozen                                                 & 1e-7                   & 128                                                              & 0.3              & 1000                      \\ \hline
-\textbf{Kaggle Brain}    & DS2/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 1000                      \\ \hline
-\textbf{LungCTChallenge} & DS3/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 1000                      \\ \hline
-\textbf{LungCTChallenge} & DS3/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 1000                      \\ \hline
-\textbf{PROSTATEx}       & DS4/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 1000                      \\ \hline
-\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0                & 1000                      \\ \hline
+\textbf{PROSTATEx}       & DS1/Full                                                   & 1e-8                   & 128                                                              & 0.4              & 2000                      \\ \hline
+\textbf{Kaggle Brain}    & DS2/Frozen                                                 & 1e-7                   & 128                                                              & 0.3              & 2000                      \\ \hline
+\textbf{Kaggle Brain}    & DS2/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 2000                      \\ \hline
+\textbf{Lung CT Challenge} & DS3/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 2000                      \\ \hline
+\textbf{Lung CT Challenge} & DS3/Full                                                   & 1e-8                   & 128                                                              & 0.3              & 2000                      \\ \hline
+\textbf{PROSTATEx}       & DS4/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 2000                      \\ \hline
+\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0                & 2000                      \\ \hline
 \end{tabular}%
@@
 \multicolumn{1}{|c|}{\cellcolor[HTML]{DAE8FC}\textbf{Dataset name}} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Best model saved at nth epoch} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Best model on the validation set}} & \textbf{Best model on the set set}                                                                                                                                      \\ \hline
-\textbf{PROSTATEx}                                                  & DS1/Full                                                   & 497                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6340\\ AUC: 0.7359\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.6842\\ Precision: 0.3333\\ Recall: 0.5\\ F1-score: 0.4\\ Specificity: 0.7333\\ AUC: 0.6833\end{tabular}}     \\ \hline
-\textbf{Kaggle Brain}                                               & DS2/Frozen                                                 & 594                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7799\\ AUC: 0.8379\end{tabular}                 & -                                                                                                                                                                       \\ \hline
-\textbf{Kaggle Brain}                                               & DS2/Full                                                   & 264                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.8233\\ AUC: 0.8566\end{tabular}                 & -                                                                                                                                                                       \\ \hline
-\textbf{LungCTChallenge}                                            & DS3/Frozen                                                 & 999                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6617\\ AUC: 0.7019\end{tabular}                 & -                                                                                                                                                                       \\ \hline
-\textbf{LungCTChallenge}                                            & DS3/Full                                                   & 345                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6924\\ AUC: 0.7197\end{tabular}                 & -                                                                                                                                                                       \\ \hline
-\textbf{PROSTATEx}                                                  & DS4/Frozen                                                 & 999                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7803\\ AUC: 0.7743\end{tabular}                 & -                                                                                                                                                                       \\ \hline
-\textbf{PROSTATEx}                                                  & DS4/Full                                                   & 50                                     & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7307\\ AUC: 0.7592\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.7368\\ Precision: 0.4286\\ Recall: 0.75\\ F1-score: 0.5454\\ Specificity: 0.7333\\ AUC: 0.7667\end{tabular}} \\ \hline
+\textbf{PROSTATEx}                                                  & DS1/Full                                                   & 527                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6375\\ AUC: 0.7334\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.6842\\ Precision: 0.3333\\ Recall: 0.5000\\ F1-score: 0.4000\\ Specificity: 0.7333\\ AUC: 0.6833\end{tabular}}     \\ \hline
+\textbf{Kaggle Brain}                                               & DS2/Frozen                                                 & 1999                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7498\\ AUC: 0.8483\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{Kaggle Brain}                                               & DS2/Full                                                   & 166                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.8172\\ AUC: 0.8596\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{LungCTChallenge}                                            & DS3/Frozen                                                 & 1852                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6734\\ AUC: 0.7755\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{LungCTChallenge}                                            & DS3/Full                                                   & 167                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7367\\ AUC: 0.7691\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{PROSTATEx}                                                  & DS4/Frozen                                                 & 1914                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7561\\ AUC: 0.7777\end{tabular}                 & -                                                                                                                                                                       \\ \hline
+\textbf{PROSTATEx}                                                  & DS4/Full                                                   & 391                                     & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.7662\\ AUC: 0.7749\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.6842\\ Precision: 0.3333\\ Recall: 0.5000\\ F1-score: 0.4000\\ Specificity: 0.7333\\ AUC: 0.8000\end{tabular}} \\ \hline
 \end{tabular}%
@@
 \subsection{Visualization of the impact of the various datasets on the target task}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task.
-In this case, they show how much the features learned from each dataset help classify prostate lesions.
-On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
-As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
-Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e.
-the training on the first/target dataset, was saved.
-As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
-In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
-
-Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
-This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
-
-At the end of the transfer learning pipeline, a graph was generated for each metric.
-These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
-If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.
-
-Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
 As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}.

### 2020-03-07 Common/chapters/conclusions.tex
Julien and Johan: new results and discussion with 2000 epochs (TL, global, ...) + minors

@@
 Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge.
-This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge, which confirms the robustness of the latter.
+This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge~\cite{42}, which confirms the robustness of the latter.
 
@@
 To achieve this, transfer learning, as well as more common techniques such as data augmentation, were applied.
-The particularity of our transfer learning implementation is the intermediate frozen steps.
-During the latter, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does.
-Experiments showed that: XXXXXX RESULTS XXXXXX.
+Our transfer learning implementation alternated between frozen and unfrozen steps and made use of brain and lung datasets to increase the model performance on a prostate dataset.
+During frozen steps, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does.
+Experiments showed that showed that our method allowed to increase the AUC on our test set by approximately 18\%, from $0.68$ before transfer learning to $0.80$ after transfer learning.
+Since a less-experienced radiologist can aim for an AUC of $0.81$~\cite{42}, this model could serve as a second diagnosis.
+These results could even be improved with another model, another way of processing data, etc.
 
 Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline.
-All these elements can serve as a baseline or as additions to future projects.
-
-xxxxx
+All these elements can be used as a baseline for future workds or as additions to existing projects.
 
@@
 Therefore, the experience and methods acquired in this work could be transposed to segmentation tasks, which could even combine segmentation and classification.
+
 Another possibility to deal with the small amount of publicly available datasets is to develop partnerships with hospitals in order to have access to their data.
@@
 New platforms ensuring their confidentiality, their public sharing and their quality have to be developed with the goal of encouraging hospitals to share them.
-This would allow to build even more better models and in an easier way.
-xxxxxx
+This would allow to build even better models in an easier way.

### 2020-03-07 Common/chapters/transfer_learning.tex
Julien and Johan: new results and discussion with 2000 epochs (TL, global, ...) + minors

@@
 \section{Goal}
+\label{sec:tl_goal}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.
@@
 
+In this experiment, the goal is to increase the model performance on the test set, particularly the AUC.
+In fact, the latter represents a degree of separability, i.e.
+how well the model distinguishes between classes.
+In cancer classification, the diagnosis must be clear.
+For instance, Model A and Model B can classify a sample X as benign (prediction: class 0) but with different probabilities of 0.10 and 0.45 respectively, where the latter are the probabilities of belonging to class 1 (malignant).
+While Model A is sure of choice, Model B is not trustworthy and may misdiagnose samples.
+
 \section{Process overview}
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The approach consists in five different steps.
-Figure \ref{fig:tl_overview} gives a visual representation of the following explanations.\\
-\begin{itemize}
+Figure \ref{fig:tl_overview} gives a visual representation of the following explanations:\\
+\begin{enumerate}
 \item First of all, similar datasets are collected and processed in the same way.
@@
 
-\end{itemize}
+\end{enumerate}
 \begin{figure}[!h]
@@
 \section{Data processing}
+\label{sec:tl_data_processing}
 \subsection{PROSTATEx}
@@
 \section{Transfer learning implementation}
-\subsection{Architecture generalities}
-\subsubsection{Layer freezing}
+\subsection{Layer freezing}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers.
@@
 
+\subsection{Conditions to choose the best model}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}At each step of the transfer learning, the best model was saved in order to be used during the next transfer learning step.
+After each epoch, the current model was compared to the previous best model.
+If it was better, the current model became the best model.
+A model was considered better if its AUC was higher than the best model's AUC, and if its accuracy was higher than the best model's accuracy.
+As the goal was to optimize the AUC, the first condition makes immediate sense.
+The second condition was put in place in order to prevent a sudden spike in AUC to set an unreacheable value.
+This scenario did not occur during this experiment but during previous ones with random initialization.
+The AUC at early epochs could be really high, saving an untrained model the best model.
+
 %\subsubsection{Parallelization}
@@
 \subsection{Script options}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}A single Python script allows to run the entire transfer learning pipeline.
+\label{sec:tl_script_options}
+
+\subsubsection{Automated transfer learning}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The first Python script runs the entire transfer learning pipeline by itself.
 Multiple command line arguments are required in order to load the right datasets and the right model, to define the hyperparameters, etc.
-Table \ref{fig:transfer_learning_options} gives a exhaustive list of the script options.
+Table \ref{table:transfer_learning_options} gives an exhaustive list of the script options.
 The particularity lies in the learning rate, number of epochs, batch size and dropout options.
@@
 Therefore, these command line options need to specify seven different parameters.
-For example, a valid value for the "--lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".
+For example, a valid value for the "-{}-lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".
+
+\subsubsection{Manual transfer learning}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The second Python script allows to perform each transfer learning step by hand.
+Thanks to this, a good combination of hyperparameters can be searched for without having to run the whole pipeline.
+The command line arguments slighty differ from the automated version.
+Table \ref{table:manual_transfer_learning_options} lists them.
+First of all, a single training set and a single validation set have to be specificied.
+In addition to this, a second dataset is required: the reference/target one, which does not change throughout the process.
+This dataset is useful for the global performance graphs (see Section \ref{sec:global_performance}).
+Another parameter used for the global performance graphs is the '-{}-last{\_}layer'.
+Its role is limited to this, unless '-{}-{}attach{\_}DM' is set to True.
+In this case, the decision maker loaded as '-{}-last{\_}layer' is attached to the current model, provided the current step is a frozen one.
+This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well)	.
+Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.
+
+Furthermore, arguments which were comma-separated values become single values, since this file runs a single transfer learning step instead of seven.
+Also, an output name for Tensorboard needs to be specified.
+In the automated version, this output name is automatically handled by the code itself.
+The latter allows to organize the various graphs generated by Tensorboard within the Tensorboard interface.
+If it is used in a correct way, Tensorboard can then be run on the root directory containing all graphs (which may be in subdirectories).
 
@@
 \textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
-\textbf{--training-set1}     & Path to training set 1                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--validation-set1}   & Path to validation set 1                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--test-set1}     & Path to test set 1                                                                                                                                                                                                                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
-\textbf{--training-set2}     & Path to training set 2                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
-\textbf{--validation-set2}   & Path to validation set 2 
+\textbf{-{}-training-set1}     & Path to training set 1                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validation-set1}   & Path to validation set 1                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-test-set1}     & Path to test set 1                                                                                                                                                                                                                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
+\textbf{-{}-training-set2}     & Path to training set 2                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{-{}-validation-set2}   & Path to validation set 2 
 & True                                                                        & String        \\ \hline
-\textbf{--training-set3}     & Path to training set 3                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--validation-set3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Comma-separated ints          \\ \hline
-\textbf{--nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Comma-separated ints            \\ \hline
-\textbf{--lr}               & Learning rate used in the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}               & Comma-separated floats         \\ \hline
-\textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
-\textbf{--cuda-device}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
-\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
-\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}                & Comma-separated floats          \\ \hline
-\textbf{--optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
-\textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
-\textbf{--cv} & \begin{tabular}[c]{@{}l@{}}Whether to use cross-validation or not\\{[}'True', 'False'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}True\end{tabular}             
+\textbf{-{}-training-set3}     & Path to training set 3                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validation-set3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Comma-separated ints          \\ \hline
+\textbf{-{}-nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Comma-separated ints            \\ \hline
+\textbf{-{}-lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}               & Comma-separated floats         \\ \hline
+\textbf{-{}-lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
+\textbf{-{}-cuda-device}      & \begin{tabular}[c]{@{}l@{}}GPU name to run the experiment\\For example: 'cuda:0', 'cuda:1', ...\end{tabular}                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{-{}-modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
+\textbf{-{}-dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}                & Comma-separated floats          \\ \hline
+\textbf{-{}-optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{-{}-outputdirectory}  & Root of the output directory. Used to save the models and Tensorboard's files.
+& \begin{tabular}[c]{@{}l@{}}False\\Default: './runs'\end{tabular}
+& String        \\ \hline
+\textbf{-{}-cv} & \begin{tabular}[c]{@{}l@{}}Whether to use cross-validation or not\\{[}'True', 'False'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}True\end{tabular}             
 & String        \\ \hline
@@
 }
-\caption{Complete list of script options}
-\label{fig:transfer_learning_options}
+\caption{Automated transfer learning -- List of script options}
+\label{table:transfer_learning_options}
+\end{table}
+
+%%%%%%% Manual transfer learning
+\begin{table}[!h]
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|llll|}
+\hline
+\rowcolor[HTML]{DAE8FC} 
+\textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
+\textbf{-{}-trainingset}     & Path to training set                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validationset}   & Path to validation set                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-reference{\_}training}     & Path to target training set (for the global performance graphs)                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
+\textbf{-{}-reference{\_}validation}   & Path to target validation set (for the global performance graphs)
+& True                                                                        & String        \\ \hline
+\textbf{-{}-batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int          \\ \hline
+\textbf{-{}-nbepochs}         & Number of epochs the training phase lasts                                                                                                                                                                  & True                                                                        & Int           \\ \hline
+\textbf{-{}-lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}               & Float         \\ \hline
+\textbf{-{}-lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
+\textbf{-{}-cudadevice}      & \begin{tabular}[c]{@{}l@{}}GPU name to run the experiment\\For example: 'cuda:0', 'cuda:1', ...\end{tabular}                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{-{}-modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
+\textbf{-{}-dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}                & Float          \\ \hline
+\textbf{-{}-inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & True
+& String        \\ \hline
+\textbf{-{}-optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{-{}-freeze}     & \begin{tabular}[c]{@{}l@{}}True to freeze the feature extractor, False otherwise.\\{[}'True', 'False'{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'False'\end{tabular} & String        \\
+\hline
+\textbf{-{}-last{\_}layer}     & If DS1: Skip.
+Otherwise: path to last{\_}layer.pckl file of previous step                                                                                                                                                                                                  & \begin{tabular}[c]{@{}l@{}}False\\Default: 'no{\_}path'                                                                       \end{tabular} & String        \\ \hline
+\textbf{-{}-attach{\_}DM}     & \begin{tabular}[c]{@{}l@{}}True to use the "-{}-last{\_}layer" as Decision Maker for the model.
+Use it for DS4 Frozen.\\{[}'True', 'False'{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'False'\end{tabular} & String        \\ \hline
+\textbf{-{}-tensorboard{\_}output{\_}name}     & \begin{tabular}[c]{@{}l@{}}Step of the transfer learning pipeline.
+Used by Tensorboard.\\{[}'DS1/Full', 'DS2/Frozen', 'DS2/Full', 'DS3/Frozen', 'DS3/Full', 'DS4/Frozen', 'DS4/Full'{]}\end{tabular}                                                                                                            & True
+& String        \\ \hline
+\textbf{-{}-outputdirectory}  & Root of the output directory. Used to save the models and Tensorboard's files. 
+& \begin{tabular}[c]{@{}l@{}}False\\Default: './runs'\end{tabular}                                                                        
+& String        \\ \hline
+\end{tabular}%
+}
+\caption{Manual transfer learning -- List of script options}
+\label{table:manual_transfer_learning_options}
 \end{table}
@@
 Each graph displays the metric values achieved on the training and the validation set on the Y-axis, whereas the X-axis represents the number of epochs.
-Table \ref{table:tl_best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment. 
-Finally, Figure \ref{fig:tl_results_test_set} compares the performance of the model resulting from the first training to the one which underwent the transfer learning experiment.
 
-\newpage
-
-\begin{figure}[!h]
+\begin{figure}[H]
 \centering
@@
 
-\begin{figure}[!t]
+\begin{figure}[H]
 \centering
@@
 \end{figure}
+
+\noindent Table \ref{table:tl_best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment. 
+Finally, Figure \ref{fig:tl_results_test_set} compares the performance of the model resulting from the first training to the one which underwent the transfer learning experiment.
 
@@
 \subsection{Training}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning experiment was performed manually.
-This means that each step was run sequentially, each time tuning the hyperparameters (learning rate and dropout probability) to reach the best performance possible.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning experiment was performed in a manual way.
+In other words, each step was run sequentially using the "Manual transfer learning" script (see Section \ref{sec:tl_script_options}).
+The learning rate and dropout probability were tuned by hand to reach the best performance possible.
 
-First of all, the best weight initialization among 200 random initializations was loaded and trained on the PROSTATEx dataset.
-With a learning rate of $1*10^{-8}$ and a dropout of $0.4$, the model learned genuine features in a smooth way (accuracy of $0.6340$ and AUC of $0.7359$ on the validation set).
-Obviously, the values are lower than the ones reached in Section \ref{sec:paper_reproduction_results} (accuracy of $0.7525$ and AUC of $0.765$) since the current experiment only uses DWI images.
+\subsubsection{DS1}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}First of all, the best initialization among 200 random initializations was loaded and trained on the PROSTATEx dataset.
+The best weight initialization is the one which achieved the best AUC on the PROSTATEx validation set.
+With a learning rate of $1*10^{-8}$ and a dropout of $0.4$, the model learned genuine features in a smooth way (accuracy of $0.6375$ and AUC of $0.7334$ on the validation set).
+Obviously, the values are lower than the ones reached in Section \ref{sec:paper_reproduction_results} (accuracy of $0.7525$ and AUC of $0.765$) since the current experiment only uses DWI images instead of stacked images (T2, DWI and ADC).
 Hence, this model could only learn from one third of the information the first model had at disposal.
 Furthermore, the training and validation curves remained close during the whole experiment.
-This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best case scenario.
+This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best-case scenario.
 
-Second, the first frozen training took place on the brain dataset.
+\subsubsection{DS2}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The first frozen training took place on the brain dataset with a frozen version of the best model resulting from the previous step.
 During the first epochs, the metrics remained low.
-For example, the accuracy was below $0.5$, which means that a model classifying the input randomly would have performed better.
-This was expected since the last layers were reset to new random initialization.
-In these early stages, the model was relying on a feature extractor capable of classifying prostate cancer and a new decision maker which did not have the time to adapt to the new task.
-Thanks to a learning rate of $1*10^{-7}$ and a dropout probability of $0.3$, the metrics started increasing early, reaching impressive values (accuracy of $0.7799$ and AUC of $0.8379$).
+For example, the accuracy was below $0.575$ and the AUC below $0.5$, which means that a model classifying the input randomly could have performed better.
+This was expected since the last layers were reset to a new random initialization.
+In these early stages, the model was relying on a feature extractor capable of classifying prostate cancer.
+The new decision maker did not have the time to adapt to the new task and was classifying samples randomly.
+Thanks to a learning rate of $1*10^{-7}$ and a dropout probability of $0.3$, the metrics started increasing relatively fast, reaching impressive values (accuracy of $0.7498$ and AUC of $0.8483$).
 Since the feature extractor is frozen, this behavior indicates that the features allowing to detect prostate cancer transfer well to brain cancer.
@@
 % Mention no reducelr ?
-Then, the full model was trained on the same dataset.
-The model resulting from the previous set was composed of decision maker trained on this dataset and a feature extractor resulting from the training on the prostate dataset (whose features seem to suit this dataset well).
+Second, the frozen layers were unfrozen and the full model was trained on the same dataset.
+The model resulting from the previous step is composed of a feature extractor resulting from the training on the prostate dataset (whose features seem to suit this dataset well) and a decision maker trained on this dataset.
 Therefore, the first metric values are already high.
 The rest of the training aimed at learning new useful features from the brain images, which could then be used for lung and prostate cancer classification.
-As expected, this step increased the metrics a bit (accuracy of $0.8233$ and AUC of $0.8566$).
+As expected, this step increased the metrics a bit (accuracy of $0.8172$ and AUC of $0.8596$).
 
-The model was then frozen again and trained on the third dataset composed of lung images.
-Here again, the metrics peaked early.
+\subsubsection{DS3}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The model was then frozen again and trained on the third dataset composed of lung images.
+Again, the metrics peaked early.
 This is partly due to the larger learning rate of $1*10^{-5}$ and to the common features that prostate, brain and lung cancers share.
@@
 
-The full training on the lung dataset went smoothly and allowed to increase the accuracy by $0.03$ and the AUC by $0.2$ before reaching stability.
+The full training on the lung dataset allowed to increase the accuracy by $0.06$ and the AUC by $0.1$ before reaching stability.
+Even with a small learning rate of $1*10^{-8}$, the optimal model was saved pretty early (at epoch 167 out of 2000).
 
-Finally, the model was trained on the first dataset again.
-At the beginning of the frozen part, the last layers were not initialized randomly.
+\subsubsection{DS4}
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Finally, the model was trained on the first dataset again.
+At the beginning of this frozen part, the last layers were not reset randomly.
 In fact, the last layers coming from the model resulting from the training on the first dataset were used as last layers.
-Since these last layers were already trained on the same dataset, the improvement occuring during this training stemed from new features learned from the brain and lung datasets.
-Even with a large learning rate of $1*10^{-5}$, the curves went up slowly and regularly throughout the whole training.
-Therefore, the features acquired thanks to transfer learning are really helpful for the prostate cancer classification.
+Since these last layers had already been trained on the same dataset, the improvement occuring during this training stemed from new features learned from the brain and lung datasets.
+With a large learning rate of $1*10^{-5}$, the curves went up quickly at the beginning, then slowly and regularly until the completion of the training.
+The initial spike and the progressive increase following it show that the low-level features acquired thanks to transfer learning are helpful for prostate cancer classification.
 
-Regarding the full training, the curves contain quite some noise with a already small learning rate of $1*10^{-9}$.
-An even smaller learning rate of $1*10^{-10}$ was tested and it turned out to be too small since the training and validation curves were completely flat from the beginning to the end.
-Here, the validation curves show an improvement during the first 50 epochs approximately, before going down while the training metrics keeps increasing.
+Regarding the full training, the curves contain quite some noise with an already small learning rate of $1*10^{-9}$.
+An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small since the training and validation curves were completely flat from the beginning to the end.
+The noise is also very visible on specific graphs such as the precision and specificity since the Y-axis scale is small.
+During this last part of the transfer learning, the validation metrics improved during the first 50 epochs approximately, before gently going down as the training metrics kept increasing.
 At this point, the model reached its optimum.
-Moreover, the fact that the validation metrics are much higher than the training ones with a dropout probability of 0.0 indicates that the model generalizes well thanks to transfer learning.
+Moreover, the fact that the validation metrics are much higher than the training ones with a dropout probability of 0.0 indicates that the model generalizes well.
 
@@
 As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
-
-Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.
+Furthermore, the full training curve shows that brain cancer features are actually helpful.
+If they were not, the metric would not have increased as fast and would not have reached such values.
 
@@
 \makebox[\textwidth]{
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_auc.png}
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_auc.png}
 }
-\caption{Transfer learning - Global Performance - Validation AUC}
+\caption{Transfer learning -- Global Performance -- Validation AUC}
 \label{fig:tl_global_validation_auc}
@@
 
-\subsection{Testing}
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_accuracy.png}
+}
+\caption{Transfer learning -- Global Performance -- Validation accuracy}
+\label{fig:tl_global_validation_accuracy}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_f1score.png}
+}
+\caption{Transfer learning -- Global Performance -- Validation f1-score}
+\label{fig:tl_global_validation_f1score}
+\end{figure}
+
+\subsection{Performance on the test set}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}To test the effectiveness of this transfer learning pipeline, the performance of the model before and after transfer learning was compared.
 Figure \ref{fig:tl_results_test_set} summarizes the various metrics.
-To obtain these numbers, both models were tested on our test set which has never been seen by the model before.
-The transfer learning allowed to increase every single metric.
-First, the AUC increased by approximately 10\%, going from $0.68$ to $0.77$.
-The recall went from $0.5$ to $0.75$.
-Both metrics are key metrics in a medical context.
-In fact, classifying a positive case as negative is far worse than the opposite.
-Hence, maximizing the AUC (which is an indicator of how well the model is able to differentiate each class) and the recall (which shows how many positive cases have been correctly detected by the model).
-Apart from these metrics, the accuracy gained around 10\% ($0.68$ to $0.74$).
+To obtain these numbers, both models were tested on our test set composed of 19 samples which have never been seen by the model before.
+The "enhanced prediction" technique proposed by Song et al.~\cite{07} was applied.
+As a reminder, the 19 test samples were augmented 11 times following the usual augmentation process (see Section \ref{sec:tl_data_processing}).
+The model predicted a value for each of the 11 versions of a sample.
+These values were then averaged.
+
+As explained in Section \ref{sec:tl_goal}, the primary goal of this technique was to increase the AUC as much as possible, as it evaluates how well the model can distinguish between classes.
+The transfer learning allowed to increase the AUC increased by approximately 18\%, going from $0.68$ to $0.80$.
+Armato.
+et al.
+\cite{42} stated that a less-experienced human radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$.
+This performance indicates that the model was able to learn low-level features from each body part in order to generalize.
+Hence, the goal of a creating a robust enough model which can serve as a second diagnosis for radiologists is reached.
+
+Regarding the other metrics, the values stayed the same.
+The reason for that is that the AUC is computed using exact predictions (float values $\in [0.0,1.0]$ expressing the probability that a sample belongs to the positive class), whereas the other metrics are based on class predictions (either 0 or 1).
+In other words, the model made the same predictions but its predictions were a lot more distinct and trustworthy.
+The fact that this test set is relatively small can play a role in this outcome since the validation metrics were far better during the last training on PROSTATEx than during the first one (for instance, the validation accuracy went from $0.6375$ to $0.7749$ when the best model was saved, which is clearly visible on Figures \ref{fig:tl_global_validation_auc}, \ref{fig:tl_global_validation_accuracy} and \ref{fig:tl_global_validation_f1score} when comparing "DS1{\_}Full" with "DS4{\_}Full").
 
@@
 \subsection{Conclusion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the performance of the model up to 10\% depending on the metric.
-Moreover, it also increased the generalization ability of the model by learning features coming from different body parts.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the AUC of the model by 15\%.
+It increased the generalization ability of the model by learning features coming from different body parts.
 This makes the model more robust and more likely to classify new unseen data well.
-These improvements are a huge step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.
+As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), it could be used as a second diagnosis.
+These improvements are a step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.

### 2020-03-08 Common/chapters/conclusions.tex
Johan and Julien: minors

@@
 During frozen steps, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does.
-Experiments showed that showed that our method allowed to increase the AUC on our test set by approximately 18\%, from $0.68$ before transfer learning to $0.80$ after transfer learning.
+Experiments showed that our method allowed to increase the AUC on our test set by approximately 18\%, from $0.68$ before transfer learning to $0.80$ after transfer learning.
 Since a less-experienced radiologist can aim for an AUC of $0.81$~\cite{42}, this model could serve as a second diagnosis.

### 2020-03-08 Common/chapters/transfer_learning.tex
Julien and Johan: new model split figure

@@
 In this case, the decision maker loaded as '-{}-last{\_}layer' is attached to the current model, provided the current step is a frozen one.
-This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well)	.
+This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well).
 Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.
@@
 \textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
-\textbf{-{}-training-set1}     & Path to training set 1                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{-{}-validation-set1}   & Path to validation set 1                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{-{}-test-set1}     & Path to test set 1                                                                                                                                                                                                                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
-\textbf{-{}-training-set2}     & Path to training set 2                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
-\textbf{-{}-validation-set2}   & Path to validation set 2 
+\textbf{-{}-trainingset1}     & Path to training set 1                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validationset1}   & Path to validation set 1                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-testset1}     & Path to test set 1                                                                                                                                                                                                                                                                                                                                                                                                   & True                                                                        & String        \\ \hline
+\textbf{-{}-trainingset2}     & Path to training set 2                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{-{}-validationset2}   & Path to validation set 2 
 & True                                                                        & String        \\ \hline
-\textbf{-{}-training-set3}     & Path to training set 3                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{-{}-validation-set3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-trainingset3}     & Path to training set 3                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validationset3}   & Path to validation set 3                                                                                                                                                                                                                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
 \textbf{-{}-batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Comma-separated ints          \\ \hline
@@
 \textbf{-{}-lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
-\textbf{-{}-cuda-device}      & \begin{tabular}[c]{@{}l@{}}GPU name to run the experiment\\For example: 'cuda:0', 'cuda:1', ...\end{tabular}                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{-{}-cudadevice}      & \begin{tabular}[c]{@{}l@{}}GPU name to run the experiment\\For example: 'cuda:0', 'cuda:1', ...\end{tabular}                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
 \textbf{-{}-modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ""\end{tabular}                 & String        \\ \hline
 \textbf{-{}-dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}True\end{tabular}                & Comma-separated floats          \\ \hline
-\textbf{-{}-optimized-metric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{-{}-optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
 \textbf{-{}-outputdirectory}  & Root of the output directory. Used to save the models and Tensorboard's files.

### 2020-03-08 Common/chapters/paper_reproduction.tex
Julien and Johan: new results and discussion with 2000 epochs (TL, global, ...) + minors
Johan and Julien: Resize one figure

@@
 \centering
-\includegraphics[width=0.95\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
 \caption{Patient splitting process}
@@
 \centering
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/auc_our_test_set.png}
+\includegraphics[width=0.6\textwidth, keepaspectratio=true]{./figures/auc_our_test_set.png}
 \caption{AUC of the best model on our test set}
@@
 \centering
-\includegraphics[height=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
+\includegraphics[width=0.6\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
 \caption{Score achieved in the PROSTATEx challenge using only our custom training set (80\% of all available data) for the training}
@@
 In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant.
-Samuel G.
+Armato.
 et al.

### 2020-03-09 Common/chapters/abstract.tex
Julien and Johan: proofread until deep learning (included)

@@
 % !TEX root = ../main.tex
-Automating the detection of cancer participates to an an early detection and treatment, which increases the chances of recovery.
+Automating the detection of cancer participates to an early detection and treatment, which increases the chances of recovery.
 Recent algorithms in artificial intelligence relying on deep learning have shown promising results in this field.
 Indeed, the usefulness of convolutional neural networks (CNNs) for segmentation or classification tasks is no longer to be proven.
-However, the performance of these models is often limited by the amount of available data to train the algorithm.
+However, the performance of these models is often limited by the amount of data which is available to train the algorithm.
 
 This thesis first presents a state-of-the-art convolutional neural network for pro\-state lesion classification.
-All steps from the data processing to the smallest details regarding the neural network training are explained, ensuring a complete reproducibility of the experiment.
-This model was evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset and achieved an AUC of 0.76.
-This result constitutes a solid baseline and confirms the correct functioning of the implementation.
+All steps from the data processing to the smallest detail regarding the neural network training are explained, ensuring a complete reproducibility of the experiment.
+This model is then evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, achieving an AUC of 0.76.
+This result constitutes a solid baseline and confirms the proper functioning of the implementation.
 
-On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) was built.
+On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) is built.
 This method shows that integrating information from diverse datasets improves automated prostate cancer diagnosis.
-Indeed, it appears that lesions of these different types of cancer share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier.
-This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancers can take advantage of it.
+Indeed, it appears that cancerous lesions coming from various body parts share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier.
+This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancer can be taken advantage of.
+Thanks to this technique, the AUC achieved on our test set increases by 18\% (from 0.68 to 0.80).

### 2020-03-09 Common/chapters/acknowledgements.tex
Johan and Julien: Authors on a new line + authors in abstract hardcoded + indent everywhere and surname everywhere until section deep learning
Julien and Johan: proofread until deep learning (included)

@@
-\textbf{Common acknowledgements:}\\
+\textbf{Common acknowledgements}\\
 First of all, we would like to thank Dr.
-Giuseppe Cuccu for his valuable advices, the exciting discussions we had with him and the precious time he took to supervise the project.
-Even we were struggling to find the cause of bad results for weeks, he always encouraged us and pushed us to give our best.
+Giuseppe Cuccu for his valuable advices, the exciting discussions we had and the precious time he took to supervise the project.
+Even we were struggling to find the cause of disappointing results for weeks, he always encouraged us and pushed us to give our best.
 We are extremely grateful to him for his support.
@@
 Philippe Cudré-Mauroux for supervising the project.
-Having the opportunity to work with such a wonderful team and having access to amazingly powerful machines was incredible.\\
+Having the opportunity to work with such a wonderful team and having access to amazingly powerful machines was awesome.\\
 
-\noindent \textbf{Julien Clément's acknowledgements:}\\
-XXX\\
+\noindent \textbf{Julien Clément's acknowledgements}\\
+I wish to thank my close friends and relatives who supported me throughout this work and understood the time-consuming nature of it.
 
-\noindent \textbf{Johan Jobin's acknowledgements:}\\
+\noindent Also, I’m sincerely grateful for having had the opportunity to work with Johan Jobin.
+His work ethic and determination to reach new goals despite the obstacles made the completion of this work possible.
+His kind and wise words were really valuable during tougher times.
+
+\noindent Finally, a special thought goes to my father Joseph who encouraged me to strive for the best throughout his whole life.
+Having him in mind gave meaning to this cancer-related project and allowed me to keep pushing when needed.\\
+
+\noindent \textbf{Johan Jobin's acknowledgements}\\
 On a personal note, I would like to acknowledge with gratitude the support of my family, specially my parents (Jean-Bernard and Marie Jobin), my brother (Lucas Jobin), my uncle (Xavier Marchand) and my girlfriend (Julie Timmermans).

### 2020-03-09 Common/chapters/deep_learning.tex
Julien and Johan: new figures with 2000 epochs
Julien and Johan: proofread until deep learning (included)

@@
 These systems try to learn how to solve a problem based on the data they receive as input.
-Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics etc.
-The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology.
+Many concrete applications make use of neural networks: autonomous vehicles, translators, computer-aided diagnosis systems, personal assistants, art creation, robotics, etc.
+The presence of deep learning techniques in these use cases clearly confirm the enthusiasm of many for this technology.
 Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As described on Figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers.
 From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks.
@@
 
-As stated by Goodfellow et al.
-\cite{15}, the first part of the answer is computing power.
-In fact, deep learning algorithms need a lot of data to work properly, which requires powerful CPUs/GPUs that either didn't exist or were only within few people's reach.
+As stated by Goodfellow et al.~\cite{15}, the first part of the answer is "computing power".
+In fact, deep learning algorithms need a lot of data to work properly, which requires powerful CPUs/GPUs that either did not exist or were only within few people's reach.
 One other main reason concerns the lack of data.
-Since deep learning algorithms "learn" from data, learning is impossible if good-quality data is not available.
+Since deep learning algorithms "learn" from data, learning is impossible if large amounts of good-quality data are not available.
 The era of Big Data enhanced deep learning possibilities.
@@
 Finally, before the year 2012, the abilities of neural networks were still to be proven.
-This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluate their algorithms on several visual recognition tasks).
+This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluated their algorithms on several visual recognition tasks).
 In fact, the deep convolutional neural network called "AlexNet" achieved 16\% of classification error rate, whereas the previous best scores were around 25\%.
@@
 
-From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
+\noindent From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
 \begin{equation}
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Machine learning algorithms belongs can be classified in two classes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Machine learning algorithms can be belong to two classes.
 The first one is "supervised learning".
@@
 Its main goal is to infer the natural structure  present in the data.
-Since the models presented in this work belong to the "supervised learning" category, notions explained below refer to this kind of algorithms.
+As the models presented in this work belong to the "supervised learning" category, notions explained below refer to this kind of algorithms.
 
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In order to keep the mathematical description of neural network consistent, this work will use Andrew Y.
-Ng's notation \cite{16}, who was a pioneer in deep learning.\\
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}In order to keep the mathematical description of neural networks consistent, this work will use Andrew Y.
+Ng's notation~\cite{16}, who is a pioneer in deep learning.\\
 
-\noindent \textbf{General comment}
+\noindent \textbf{General comments}
 \begin{itemize}
@@
 \item $n_{y}$: output size (or number of classes)
-\item $n_{h}^{[l]}$: number of hidden units (i.e neurons) of the $l^{th}$ layer
+\item $n_{h}^{[l]}$: number of hidden units (i.e. neurons) of the $l^{th}$ layer
 \item $L$: number of layers in the network
@@
 \item $W^{[l]} \in \R$ \textsuperscript{\# of neurons in the next layer = j  x \# of neurons in the previous layer = k} is the weight matrix at layer $[l]$.
-\item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector in the $l^{th}$ layer.
+\item $b^{[l]} \in R$\textsuperscript{\# of units in next layer} is the bias vector at the $l^{th}$ layer.
 \item $\hat{y} \in R^{n_{y}}$ is the predicted output vector. It can also be denoted as $a^{[L]}$ where $L$ is the number of layers in the whole network.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Perceptrons are the main components of neural networks.
-They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}.
+They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work of Warren McCulloch and Walter Pitts"~\cite{13}.
 Today, they are called "artificial neurons" or simply "neurons".
@@
 \end{equation}
-Schematically speaking, a perceptron can be represented as on figure \ref{perceptron_model}.
+Schematically speaking, a perceptron can be represented as on Figure \ref{perceptron_model}.
 Each input is multiplied with its corresponding weight.
-The sum of this result then goes through a non-linear function, called "activation function".
+The sum of these multiplications then goes through a non-linear function, called "activation function".
 This activation function acts like a threshold that determines the proportion of the result that goes further in the network.
-There exist multiple activation functions (see \ref{activation_functions}).
-It is extremely important to use non-linear functions instead of a linear functions.
-In fact, the output of a perceptron is given as input to the others (see \ref{multilayer_perceptron}).
-Consequently, if only linear functions are used throughout the network, linear outputs are given as inputs to other linear functions.
-Since the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers would not make sense anymore.
+There exist multiple activation functions (see Section \ref{activation_functions}).
+It is extremely important to use non-linear functions instead of linear functions.
+In fact, the output of a perceptron is given as input to the others (see Section \ref{multilayer_perceptron}).
+Consequently, if linear functions only are used throughout the network, linear outputs are given as inputs to other linear functions.
+As the composition of two linear functions is itself a linear function, assembling perceptrons to create neural networks of multiple layers does not make sense in this case.
 
@@
 \label{activation_functions}
-\setlength{\marginparwidth}{2cm}\leavevmode \marginnote{\textbf{Jobin}}
-Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function.
+\setlength{\marginparwidth}{2cm}\leavevmode \marginnote{\textbf{Jobin}}Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function.
 The latter must be non-linear in order to approximate extremely complex functions.
@@
 Hornik et al.
-claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}.
-According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
+claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense"~\cite{17}.
+According to Thomas Epelbaum~\cite{18}, the most commonly used activation functions are:\\\\
  \noindent \textbf{Sigmoid function}\\
@@
  \begin{cases}
-    1 & \text{if } x\geq 0\\
+    x & \text{if } x> 0\\
     0              & \text{otherwise}
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}A multilayer perceptron is a type of artifical neural networks.
-Du et al.
-\cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
+Du et al.~\cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
 input and output layers" where "the output units represent a hyperplane in the space
@@
 \item The input layer: It is the neural network entry point for the data.
-Generally speaking, the data is provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
-The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is the dimension of each sample.
+Generally speaking, the data are provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$.
+The batch size defines the number of samples that will be fed to the network at the same time and $n_{x}$ is the dimension of each sample.
 Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector.
@@
 \item The hidden layer(s): Hidden layers stand for all layers that are between the input layer and the output layer.
-Each of them has its own weights and biases (W, b), denoted by $W^{[l]} \in \R $ and $b^{[l]}$ respectively, where $W_{ij}^{l}$ corresponds to the weights associated with the connection between perceptron $j$ in layer $l$ and perceptron $i$ in layer $l+1$.
+Each of them has its own weights and biases (W, b), denoted by $W^{[l]} \in \R $ and $b^{[l]}$ respectively, where $W_{ij}^{l}$ corresponds to the weights associated with the connection between perceptron $j$ in \mbox{layer $l$} and perceptron $i$ in layer $l+1$.
 By analogy, $b_{i}^{[l]}$ is the bias associated with perceptron $i$ in layer $l$.
-Weights and biases are the parameters to optimize in order to obtain the best mapping between the input and the output of the network (see Section \ref{training_a_neural_network} ).
-Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization"(see Section \ref{weight_initialization}).
+Weights and biases are the parameters to optimize in order to obtain the best mapping between the input and the output of the network (see Section \ref{training_a_neural_network}).
+Before training the neural network, the weights can be randomly initialized or initialized with more sophisticated methods such as "Xavier initialization" or "Kaiming initialization" (see Section \ref{weight_initialization}).
 
@@
 The prediction of a neural network is given by $\hat{y} \in R^{n_{y}}$ with $n_{y}$ representing the number of different labels.
-In a classification task, whose goal is to assign to each input a specific class, the $\hat{y}$ is usually the probability that the input belongs to each class.
-In that case, the \textit{softmax} activation function would be used at the end.
+In a classification task, whose goal is to assign a specific class to each input, $\hat{y}$ is usually the probability that the input belongs to each class.
+In this case, the \textit{softmax} activation function would be used at the end.
 \end{itemize}
 The advantage of organizing the weights, biases and inputs in matrices is due to the ability of modern CPUs/GPUs to quickly perform linear algebra computations.
-This way of structuring the network component is called "vectorization" which avoids using loops in the code, which would considerably slow down the computations.
+This way of structuring the network components is called "vectorization" which avoids using loops in the code, which would considerably slow down the computations.
 Figure \ref{multilayer_perceptrons_figure} illustrates the concept of multilayer perceptrons.
-In this example, the total number of layers $L$ is equal to 3, the input size $n_{x}$ is equal to 4 and the number of units in in each layer is $n_{h}^{1}= 4$, $n_{h}^{2}=2$, $n_{h}^{3}=1$.
-The network contains the weights  $W^{(1)} \in \R^{2x3}$, $W^{(2)} \in \R^{1x2}$ and the biases  $b^{[1]}=2$, $b^{[2]}=1$.
+In this example, the total number of layers $L$ is equal to 3, the input size $n_{x}$ is equal to 4 and the number of units of each layer is $n_{h}^{1}= 4$, $n_{h}^{2}=2$, $n_{h}^{3}=1$.
+The network contains the weights  $W^{(1)} \in \R^{2x3}$, $W^{(2)} \in \R^{1x2}$ and the biases  $b^{[1]} \in \R^2$, $b^{[2]} \in \R^1$.
 
@@
 After seeing every single batch, the loss is computed using a "loss function".
-The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth.
-Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minimum of the loss function.
+The latter is used to evaluate the predictions of the neural network in comparison to their ground truth.
+Then, the weights and biases of the network are updated during a process called "backpropagation" in order to find the global minimum of the loss function.
 As soon as the neural network has seen every single batch, the end of an "epoch" is reached.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The forward propagation is used to transmit the input through the entire neural network.
-Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations.
+Mathematically speaking, the forward propagation step for a specific \mbox{layer $l$} is represented by two equations.
 The first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$:
@@
 Figure \ref{forward_propagation} illustrates the computation of the forward propagation pass for the $l^{th}$ layer.
-The weight natrix $W_{jk}^{l}$ represents the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$.
+The weight matrix $W_{jk}^{l}$ represents the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$.
 This matrix is multiplied by the output of the previous layer $a_{j}^{[l-1]}$ before adding the bias $b^{[l]}$.
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted"~\cite{18}.
 It is therefore useful to measure the penalty for a single input.
@@
 There exist multiple cost functions.\\
-For a regression problem, a commonly used loss function is the mean square error:
+For a regression problem, a commonly used loss function is the mean squared error:
 \begin{equation}
@@
 For classification problems, the cross entropy function is regularly used.
-We distinguish the binary classification where the number of classes $n_{y}$ = 2 from the multiclass classification where $n_{y}$ > 2.
+The binary classification where the number of classes $n_{y}$ = 2 is distinguished from the multiclass classification where $n_{y}$ > 2.
 In the case of binary classification, the cross entropy is:
@@
 \end{equation}
-In the case of multiclass classification, the categorical crossentropy is:
+In the case of multiclass classification, the categorical crossentropy is given by:
 \begin{equation}
@@
 \end{equation}
-Since the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function.
+As the cost function gives an estimation of the overall error of the network, the main objective of training a neural network is to update its weights in order to approach the minimum of the function.
 Therefore, deep learning problems can be considered as optimization problems.
@@
 How about hidden units?
-For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.
+For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input"~\cite{24}.
 
-In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting parameters of the model.
+In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting the parameters of the model.
 The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network.
@@
 \begin{equation}
-\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{1}}]
+\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{m}}]
 \end{equation}
-The gradient shows how much all parameters that constitute x need to change to minimize the function.
+The gradient shows how much the parameters that constitute x need to change to minimize the function.
 In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$.
@@
 \end{equation}
-Once the gradients of each parameter are computed, these parameters are updated. The weights update is described by the following equation:
+Once the gradients of each parameter are computed, the corresponding parameters are updated. The weight update is described by the following equation:
 \begin{equation}
@@
 \end{equation}
-The biases update corresponds to:
+The bias update corresponds to:
 \begin{equation}
@@
 %The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
-The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to manually be tuned.
+The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be manually tuned.
 
@@
 \end{itemize}
-From these four categories, multiple metrics with their own specificities can be computed \cite{25}:
+From these four categories, multiple metrics with their own specificities can be computed~\cite{25}:
 \begin{itemize}
@@
 \end{equation}
-Accuracy is a great measure in the case of symmetric data (i.e the number of FN $\approx$ FP and their cost is similar).
+Accuracy is a great measure in the case of symmetric data (i.e.
+the number of FN $\approx$ FP and their cost is similar).
 When this condition is not fulfilled, accuracy can lead to bad models.
-For instance, let's define a binary classification model that always outputs class 0.
-If the data is composed of 99 samples from class 0 and 1 sample from class 1, the accuracy is equal to 99\% but the model is not smart.
+For instance, suppose that a binary classification model always outputs class 0.
+If the data is composed of 99 samples from class 0 and 1 sample from class 1, the accuracy is equal to 99\%, but the model is not smart.
 Consequently, this metric has to be used in addition to other metrics.
 
-\item Precision: "Ratio of the correctly positive labeled subjects by the model to all positive labeled subjects".
+\item Precision: "Ratio of the correctly labeled positive subjects to all positively labeled subjects"
 \begin{equation}
@@
 This metric is recommended when the confidence of the true positives predicted by the model is important.
-For instance, this happens in spam blockers where it is preferable to have a spam in mailbox rather than a regular mail in the spam box.
+For instance, this happens in the case of spam blockers where it is preferable to have a spam in the mailbox rather than a regular mail in the spam folder.
 
-\item Recall (sensitivity): "Ratio of the correctly positive labeled subjects to all subjects whose class is actually positive".
+\item Recall (sensitivity): "Ratio of the correctly labeled positive subjects to all subjects whose class is actually positive".
 
@@
 
-\item F1-score: Harmonic mean(average) of the precision and recall.
+\item F1-score: Harmonic mean (average) of the precision and recall.
 \begin{equation}
@@
 
-\item Specificity: "Ratio of the correctly negative labeled subjects to all subjects whose class is actually negative".
+\item Specificity: "Ratio of the correctly labeled negative subjects to all subjects whose class is actually negative."
 \begin{equation}
@@
 For instance drug tests can not indicate false positives but they have to cover all true negatives.
+
 \item ROC curve and AUC:
-As explained by Sarang Narkhede ~\cite{26}, "the ROC curve is plotted with the true positive rate (=recall) against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
+As explained by Sarang Narkhede~\cite{26}, "the ROC curve is plotted with the true positive rate (=recall) against the false positive rate (1-specificity) where recall is on y-axis and the false positive rate is on the x-axis.
 AUC-ROC curve is a performance measurement for classification problem at various thresholds settings.
@@
 Usually, the evaluation on the validation set is performed every $N$ epochs, where $N$ is a fixed number.
-The validation set needs to come from the same distribution as the training set but should contain exclusively unseen data.
+The validation set needs to come from the same distribution as the training set but should exclusively contain unseen data.
 This last point is crucial since the validation set shows how well the neural network generalizes on unknown data.
-The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" which is the behaviour of a model that fits to the training set too closely and don't generalize well.
+The validation set can also be used as an indicator to decide when the training should be stopped in order to prevent "overfitting" which is the behaviour of a model that fits to the training set too closely and do not generalize well.
 In fact, if the validation loss continuously increases for a certain number of epochs, going on with the training will increase the overfit.
-This fact of interrupting the training earlier is called "early stopping".
+The fact of interrupting the training earlier is called "early stopping".
 \item The test set: This last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
 \end{itemize}
-Regarding the way these three sets are split, it mostly depends on the number of samples available.
-If the latter is big, the data is split into training and testing using the ratio 80/20.
-Then the remaining training samples are also split into training and validation using the ratio 80/20.
+The way these three sets are split mostly depends on the number of samples available.
+If the latter is big, the data is split into training and test sets using the ratio 80/20.
+Then, the remaining training samples are also split into training and validation using the ratio 80/20.
 On the contrary, if there are few data available, k-fold cross-validation is a good practice.
@@
 One fold is picked as test set and the others are considered as training sets.
-The model is trained on training folds and tested on the test set.
+The model is trained on training folds and tested on the test fold.
 Then, another test set is picked and the same process is repeated until all possible tests set are picked.
 At the end of the process, the results of all test sets are averaged, which provides a good estimation of the model's performance.
-This technique is summarized on figure \ref{cross_validation}.
+This technique is summarized on Figure \ref{cross_validation}.
 
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/cross_validation.png}
-\caption{5-folds cross-validation }
+\caption{5-fold cross-validation}
 \label{cross_validation}
@@
 The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
-If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.
+If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all"~\cite{27}.
 
@@
 The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions.
-Consequently, the slope of the gradient changes slowly and the learning takes a lot of time.\\
+Consequently, the slope of the gradient changes slowly and the learning process takes a lot of time.\\
 To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
@@
 where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
-For activation functions that are not symmetric around zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative.
+For activation functions that are not symmetric around zero and do not have outputs inside the range $[-1,1]$ such as ReLU or ELU, Kaiming initialization is an alternative.
 It consists in multiplying the randomly initialized weight matrix by:
@@
 \end{equation} 
-where $n_{h}^{[l]}$ is the number of incoming connections coming into a given layer from the previous layer's output.
+where $n_{h}^{[l]}$ is the number of incoming connections coming to a given layer from the previous layer's output.
 
-\subsection{Hyperparameters tuning}
+\subsection{Hyperparameter tuning}
 
@@
 That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections.
-So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).
+So, these parameters have to be manually tuned in order to find the best configuration (i.e.
+the one that minimizes the cost function and that keeps an acceptable level of generalization).
 
 Regarding the learning rate $\alpha$, its value has to be neither too large nor too small.
-A too large learning rate is recognizable by analyzing the training loss curve: if the loss is exploding or oscillating or if there is no improvement and it is stuck around a suboptimal local optima, the learning rate is too high and should be decreased.
-On the contrary, if the learning is very slow and the loss does not decrease or if the model is overfitting, it is a clear sign that it should be increased.
+A too large learning rate is recognizable by analyzing the training loss curve: if the loss is exploding, oscillating or if there is no improvement and it is stuck around a suboptimal local optima, the learning rate is too high and should be decreased.
+On the contrary, if the learning is very slow and the loss does not decrease, or if the model is overfitting, it is a clear sign that it should be increased.
 The learning rate can take a wide range of values.
-Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored" \cite{28}.
-There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).
+Consequently, the most used technique to find the optimal learning rate is simply the "trial and error" method, which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored"~\cite{28}.
+There also exist methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).
 
 Batch size is another important hyperparameter to tune.
-Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples" \cite{29}.
-Furthermore, "networks train faster because the weights update is done after each propagation" \cite{29}.
-Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be" \cite{29}.
-Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.
+Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples"~\cite{29}.
+Furthermore, "networks train faster because the weight update is done after each propagation" \cite{29}.
+Nevertheless, "the smaller the batch, the less accurate the estimate of the gradient will be"~\cite{29}.
+Indeed, due to the high weight update frequency, the gradient fluctuates much more than if it was computed after a bigger number of samples.
 
 Finally, the number of epochs during which the network is trained has to be carefully chosen.
-In fact, from a certain point in the training, neural networks don't learn anymore useful features in the data and start overfitting.
-This point corresponds to the moment where, the validation loss does not decrease anymore and starts to increase continuously.
+In fact, from a certain point in the training onwards, neural networks do not learn anymore useful features in the data and start overfitting.
+This point corresponds to the moment where the validation loss does not decrease anymore and starts to increase continuously.
 It is exactly at this moment that the training should stop.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks.
-They are particular in that they contain layers which perform a mathematical operation named convolution on the input data.
+They are particular in that they contain layers which perform a mathematical operation named "convolution" on the input data.
 CNNs are mostly used in image and video analysis.
@@
 To perform a convolution, numerical input data and a "filter" are required.
-A filter can be seen as a $f$x$f$ numerical patch that moves across the entire input.
+A filter can be seen as an $f$x$f$ numerical patch that moves across the entire input.
 It first moves horizontally until reaching the right-most border of the image.
@@
 To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s.
-Three ways of padding images are commonly used as shown on figure \ref{fig:convolution_padding}:
+Three ways of padding images are commonly used as shown on Figure \ref{fig:convolution_padding}:
 \begin{itemize}
@@
 
-	\item \textbf{Full:} The input image is padded so that, in the first step of the convolution, only the bottom-right cell of the filter overlays the pixel value of the image, the rest overlaying padding cells.
+	\item \textbf{Full:} The input image is padded so that, in the first step of the convolution, only the bottom-right cell of the filter overlays the pixel values of the image, the rest overlaying padding cells.
 This makes the output size larger than the input size.
@@
 \section{Transfer learning}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task" \cite{30}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task"~\cite{30}.
 The model dedicated to the second task uses some or all parts of the first model (i.e.
@@
 The first model can either be implemented from scratch if enough data is available or it can simply be downloaded from institutions that release large pretrained models for similar tasks.
-There are three major benefits \cite{30}:
+There are three major benefits~\cite{30}:
 \begin{itemize}
@@
 \end{itemize}
-Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated" \cite{30}.
+Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated"~\cite{30}.

### 2020-03-09 Common/chapters/introduction.tex
Julien and Johan: proofread until deep learning (included)

@@
 \label{ch:introduction}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}
-According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
-An early detection and treatment increases the chances of recovery.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}According to the World Health Organization, cancer caused $9.6$ million deaths in 2018, making it the second leading cause of death~\cite{44}.
+An early detection and treatment increase the chances of recovery.
 In this context, Computer-Aided Diagnosis (CAD) systems can play a massive role by preventing health professionals from missing positive diagnoses.
-Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transfered the advantages of deep learning to cancer detection, diagnosis and localization tasks.
+Thanks to the attention that deep learning has gotten over the last decade, newer and better diagnosis systems have transferred the advantages of deep learning to cancer detection, diagnosis and localization tasks.
 
@@
 must be provided so that deep learning models can make use of the data.
-This whole process is time consuming and medical institutions don't always see the benefits which could ensue from publishing datasets of good quality.
-As a consequence, the lack of data is one of the toughest challenge related to this field.
-In order to overcome it, techniques such as transfer learning exist.
+This whole process is time-consuming and medical institutions do not always see the benefits which could ensue from publishing good-quality datasets.
+As a consequence, the lack of data is one of the toughest challenges related to this field.
+To deal with it, techniques like transfer learning exist.
 The latter aims at using independent but similar datasets in order to increase the performance of a model on a target dataset.
@@
 Second, deep learning is presented from the ground up under the historical and technical points of view.
-This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible without any background in the field.
+This section demystifies the topic by making an overview of the mathematical concepts and definitions related to it, which makes the understanding of the technical part possible (without any background in the field).
 
 The next chapter deals with medical knowledge related to cancer.
-Characteristics of the disease are presented, before focusing on the various file formats which medical imaging data are exported into.
+Characteristics of the disease are presented, before focusing on the most common medical file formats.
 The last part is critical since a lot of data processing was performed in order to be able to make use of these files.
 
-Then, the architecture presented in a prostate cancer classification research paper was reproduced.
-This chapter allows to set a baseline proving that the methods and implementations work, from the image preprocessing to the neural network.
+Then, the experiment presented in a prostate cancer classification research paper was reproduced.
+This chapter allows to set a baseline proving that our methods and implementations work, from the image preprocessing to the training of the neural network.
 
-Finally, the last part is dedicated to transfer learning and its application to cancer classification.
+Finally, the last part is dedicated to a transfer learning method making use of multiple body parts and its application to prostate cancer classification.
 
@@
 An early detection is key to maximize the chances of recovery.
-As explained by Nicholas Petrick~\cite{50}, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device.
-The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information".
+As explained by Nicholas Petrick, "it has long been recognized that clinicians do not always make optimal use of the data acquired by an imaging device.
+The limitations of the human eye-brain system, limitations in training and experience, and factors such as fatigue, distraction, and satisfaction of search may all contribute to suboptimal use of available information" ~\cite{50}.
 Under these circumstances, CAD systems can contribute to address some of these issues by being used as an aid for clinicians.
 A complete CAD system is composed of two parts.
-The first one is called CADe for "computer-aided detection".
+The first one is called CADe which stands for "computer-aided detection".
 This part includes medical  image  analysis  tasks  such  as  segmentation,  identification,  localization  and  detection.
-The other part is CADx for "computer-aided diagnosis" which aims at extracting the characteristics of lesions to classify them according to their malignancy.
+The other part is CADx for "computer-aided diagnosis", which aims at extracting the characteristics of lesions to classify them according to their malignancy.
 CAD systems provide multiple advantages.
 
-First of all, used as an extra diagnosis, they decrease the probability of missing positive diagnoses.
+First of all, used as an extra diagnosis, they can decrease the probability of missing positive diagnoses.
 Then, they can also speed up the diagnosis process by proposing regions of interest to clinicians, which, in time, could reduce the screening costs.
-
 Similarly, lower screening costs imply a greater accessibility to screening tests, which can result in an earlier cancer detection.
 As stated before, detecting cancers at an early stage maximizes the probability of curing them.
-Another benefit of using CAD systems, especially the ones based on deep learning, is the number of cases that the latter are trained on.
+Another benefit of using CAD systems (especially the ones based on deep learning) is the number of cases that the latter are trained on.
 In fact, in order to built efficient deep learning models, large amounts of data are required.
@@
 Research is still ongoing, aiming at improving existent CAD systems.
-Problems like poor detection rates, small amounts of available data and bad levels of generalization are still recurrent.
+Problems like poor detection rates, small amounts of available data and poor levels of generalization are still recurrent.
 Therefore, this situation can and must be improved.
@@
 \section{Contributions}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}%This work proposes a state-of-the-art deep learning model for prostate lesion classification and proposes a method based on transfer learning using different body parts to overcome the lack of publicly available data in the medical field.
 Concretely, the major contributions of this work are:
+The major contributions of this work are:
 \begin{itemize}
-\item A state-of-the-art deep learning model for prostate cancer detection.
-All steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented.
-This makes the experiment completely reproductible.
+\item A state-of-the-art deep learning model for prostate cancer classification.
+To make the experiment completely reproductible, all steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented.
 
@@
 
-\item Processing scripts for the SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, the Lung CT Challenge dataset and the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+\item Processing scripts for the SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, the SPIE-AAPM Lung CT Challenge dataset and the Kaggle Brain MRI Images for Brain Tumor Detection dataset.
 This includes the conversion of DICOM and PNG files to NumPy arrays, their registration (alignment and resizing to the same resolution for stacking purpose, ensuring the same amount of tissue on each image), their augmentation and the organization of the resulting images into multiple subsets (training, validation and test) that can be used as input to machine learning algorithms.
@@
 \item Visualization scripts for DICOM, NIfTI and RAW medical file formats.
-A system to navigate through 4D data (width, height, depth and time) using the directional arrows is implemented: left/right arrows to vary the time axis and up/down to navigate through the difference slices (i.e the depth) of all images of a patient.
-3D data is also supported with the same functionalities (apart from navigating through the fourth dimension (time)).
+A system to navigate through 4D data (width, height, depth and time) using the directional arrows is available: left/right arrows to make the time axis vary and scrolling up/down to navigate through the different slices (i.e the depth) of a patient's images.
+3D data is also supported with the same functionalities (apart from the navigation through the fourth dimension (time)).
 
-\item Processing scripts for the PROSTATEx challenge.
-This goes from the preprocessing of the challenge test images to the generation of the CSV file containing the predictions of a given model, which are probabilities $[0,1]$ of the lesion being malignant ("benign lesion" if $< 0.5$, "malignant tumor" otherwise).
+\item Scripts for the PROSTATEx challenge.
+This goes from the preprocessing of the challenge test images to the generation of the CSV file containing the predictions of a given model, which are probabilities $\in [0,1]$ of the lesion being malignant ("benign lesion" if $< 0.5$, "malignant tumor" otherwise).
 

### 2020-03-09 Common/chapters/literature_review.tex
Julien and Johan: proofread until deep learning (included)

@@
 For this reason, CAD systems are currently a trending topic in deep learning.\\
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e.
-prostate, lung and brain.
-Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}As this work focuses on lesion classification, this chapter presents the main research papers about CAD systems based on convolutional neural networks for each body part used in Chapters \ref{ch:paper_reproduction} and \ref{ch:transfer_learning} (i.e.
+prostate, lung and brain).
+Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this thesis.
 
-\section{Prostate/PROSTATEx}
+\section{Prostate -- PROSTATEx}
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by Gao et al., "Prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by Gao et al., "prostate  cancer  is  the  most  common  malignancies  among  men  and  remains  a  second  leading  cause to deaths in men globally.
 It was predicted that there would be 1.7 million new cancer cases by 2030.
-The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"\cite{41}.
+The early detection and diagnosis of prostate cancer can help to survive nine out of 10 men for the last five years"~\cite{41}.
 Therefore, researchers proposed many different models to achieve good performance in prostate cancer detection.
-All the research works mentioned below are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is composed of prostate lesions multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans)(see Section \ref{sec:prostatex_dataset_description}).
+All the below-mentioned research papers are based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is composed of multiparametric MRIs (T2W, DWI, ADC, DCE, PD, Ktrans) of prostate lesions (see Section \ref{sec:prostatex_dataset_description}).
 
@@
 by subtracting the mean before dividing by the standard deviation.
-The data was split into a training ($80\%$), validation ($10\%$) and test set ($10\%$).
-The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal flipping, vertical sliding of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
-Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer, and the use of the ELU activation function (see figure \ref{fig:paper_model}).
-The evaluation method for each patient and finding made an average of the 11 predictions resulting from the 11-time augmentation of the test set.
-The best results were obtained by using DWI images only with the highest b-value, reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
-However, this model was not tested on the official PROSTATEx challenge images, which is an interesting benchmark to evaluate how well a model generalizes.
+The data was split into a training set ($80\%$), validation set ($10\%$) and test set ($10\%$).
+The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal and vertical flipping, horizontal and vertical shifting of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
+Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer.
+They also use the ELU activation function instead of ReLU (see figure \ref{fig:paper_model}).
+The evaluation method for each patient and finding made an average of the 11 predictions resulting from the augmentation of the test set.
+The best results were obtained by using DWI images with the highest b-value (stacked with T2 and ADC images), reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$).
+However, this model was not tested on the official PROSTATEx challenge, which would have been an interesting benchmark to evaluate how well the model generalizes.
 
-Liu et al.~\cite{46} created another architecture called XMasNet which was tested on the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
+Liu et al.~\cite{46} created another architecture called XMasNet.
+They took part in the actual PROSTATEx challenge, achieving the second best performance at the time with an AUC of $0.84$.
 The AUC on their validation set reached $0.92$.
-Regarding data processing, their approach stacked different combinations of the available sequences as the three channels instead of defining a single combination like Song et al.: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W.
-The data augmentation process differs in that the images are rotated in 3D, each lesion being sliced at 7 different orientations.
-These 2-dimensional slices were then augmented using rotation, shearing and translation of 1px, resulting in 207144 training samples.
-Both validation and testing test were also augmented in the same manner.
+Contrary to Song et al., their data processing approach stacked different combinations of the various sequences as the three channels instead of defining a single combination: DWI-ADC-Ktrans, DWI-ADC-T2W, ADC-Ktrans-T2W and DWI-Ktrans-T2W.
+The data augmentation process differs in that it uses a three-dimensional process.
+First, each lesion was sliced seven times using seven different orientations.
+These seven 2-dimensional slices were then augmented using rotation, shearing and translation of 1px, resulting in 207144 training samples.
+Both validation and test sets were also augmented in the same manner.
 This whole process allows to include 3-dimensional information in 2-dimensional images.
-The method used ensemble learning which combined different models to reach the best performance possible.
+The training method relied on ensemble learning which combined different models to reach the best performance possible.
 
@@
 
-Armato et al~\cite{42} summarized the results obtained by all teams that took part in the PROSTATEx challenge in 2017.
+Armato et al.~\cite{42} summarized the results obtained by all the teams that took part in the PROSTATEx challenge in 2017.
 This challenge was split into two separate tasks.
-The first one was devoted to the diagnosis of prostate lesions (classification) whereas the other was about the segmentation and the determination of Gleason Grade Group.
-Thirty-two groups submitted results to the first challenge out of a total of 71 different methods (each group was allowed to submit up to three methods for evaluation).
+The first one was devoted to prostate lesion diagnosis (classification), whereas the other was about segmenting lesions and determining their Gleason Grade Group.
+Thirty-two groups submitted their results to the first challenge, resulting in a total of 71 different methods (each group was allowed to submit up to three methods for evaluation).
 The article indicates that "most, but not all, methods outperformed random guessing (AUC=0.5)"~\cite{41}.
-The best performing method obtained an AUC value of 0.87 (standard error 0.027) and the next three methods all achieved AUC values of 0.84 (with standard errors of 0.036, 0.032, and 0.032).
-The complete results are illustrated on figure \ref{fig:challenge_all_results}.
-The median AUC on the challenge is approximately $0.68$.
+The best performing method obtained an AUC value of 0.87 (standard error 0.027) and the next three methods all achieved an AUC of 0.84 (with standard errors of 0.036, 0.032, and 0.032).
+Figure \ref{fig:challenge_all_results} shows the different results.
+Finally, the median AUC on the challenge is $0.68$.
 \begin{figure}[!h]
@@
 Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}.
-The articles cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans devoted to the classification of lung nodules (see \ref{sec:lungCTChallenge} for further details).
+The papers cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans, and are devoted to the classification of lung nodules (see Section \ref{sec:lungCTChallenge} for further details).
 
@@
 Figure \ref{fig:LUNGx_challenge_all_results} reports all methods used to classify the lung lesions with their associated performance.
-Furthermore, it also provides the performance of all radiologists.
+It also provides the performance achieved by the radiologists.
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Brain cancer is another major type of cancer.
-According to American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women.
+According to the American Society of Clinical Oncology, "brain and other nervous system cancer is the 10th leading cause of death for men and women.
 It is estimated that 17760 adults (9910 men and 7850 women) will die from primary cancerous brain and central nervous system tumors this year"~\cite{43}.
@@
 However, as it is the only dataset available for brain tumor classification, some publications used it.
-Further details can be found in section \ref{sec:kaggleBrain}.
+Further details can be found in \mbox{Section \ref{sec:kaggleBrain}}.
 
@@
 Their processing method used a cropping technique which removed extra black margin around the skull.
-Each border of the image merged with a part of the skull.
+Each border of the image was merged with a part of the skull.
 Since the images come from different sources, their resolution vary quite a lot.
@@
 The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
-The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy of $0.55$).
+The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy \mbox{of $0.55$}).
 

### 2020-03-09 Common/chapters/deep_learning.tex
Julien and Johan: minors

@@
 The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum.
-The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
+According to James Dellinger, the idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network.
 If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all"~\cite{27}.
@@
 This means that the $3x3$ filter detected a vertical line in the center of the input image.
+Like other parameters, filters are learned while training.
 
@@
 keeps the same weights and architecture or a part of them) and is then retrained on data available for this task.
-The first model can either be implemented from scratch if enough data is available or it can simply be downloaded from institutions that release large pretrained models for similar tasks.
-There are three major benefits~\cite{30}:
+The first model can either be implemented from scratch if enough data is available or downloaded from institutions that release large pretrained models for similar tasks.
+Transfer learning provides three major benefits~\cite{30}:
 \begin{itemize}

### 2020-03-09 Common/chapters/medical.tex
Julien and Johan: medical ok

@@
 In fact, formats represent 2D, 3D or even 4D data.
-Some of them require specific normalization in order to get the right rendering.
-Finally, some visualization tools were developed to display raw medical files easily.
+Some of them require specific normalization in order to obtain the right rendering.
+Finally, some visualization tools that were developed to display raw medical files easily are presented.
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}An accumulation of cells forming a mass is called a tumor.
-These tumors are detectable thanks to medical imaging (see section \ref{sec:medical_imaging}) and other symptoms.
+These tumors are detectable thanks to medical imaging (see Section \ref{sec:medical_imaging}) and other symptoms.
 However, not every tumor is as dangerous as the other, as it can be benign (does not contain cancerous cells) or malignant (contains cancerous cells).
@@
 As stated by Hanahan et al.
-in "The Hallmarks of Cancer" \cite{19} and "The Hallmarks of Cancer: The Next Generation" \cite{20}, every malignant tumor acquires six different capabilities during its evolution:
+in "The Hallmarks of Cancer"~\cite{19} and "The Hallmarks of Cancer: The Next Generation"~\cite{20}, every malignant tumor acquires six different capabilities during its evolution:
 \begin{itemize}
-	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells don't wait for the body's approval to grow and proliferate, contrary to normal cells.
+	\item \textbf{"Sustaining proliferative signaling"}\\ Cancerous cells do not wait for the body's approval to grow and proliferate, contrary to normal cells.
 They become responsible for their own multiplication.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Most cancers can be staged thanks to the TNM system.
-The T corresponds to the tumor size and its location; the N corresponds to whether or not the tumor has spread to draining lymph nodes; the M corresponds to the presence or absence of metastases in other parts of the body \cite{21}.
-These pieces of information are used to classify cancer between four (I to IV) or sometimes five (0 to IV) different stages, reflecting the progression and the seriousness of the illness \cite{22}.
+The T corresponds to the tumor size and its location; the N corresponds to whether or not the tumor has spread to draining lymph nodes; the M corresponds to the presence or absence of metastases in other parts of the body~\cite{21}.
+These pieces of information are used to classify cancer between four (I to IV) or sometimes five (0 to IV) different stages, reflecting the progression and the seriousness of the illness~\cite{22}.
 The earlier they are detected, the higher the chances of recovery are.
@@
 MRI relies on magnetic fields to provide a three-dimensional view of body parts, which allows to see the generated images as a volume.
-Different settings, usually called sequences, make the look of the output vary, as shown on figure \ref{fig:PROSTATEx-t2-adc-dwi}.
+Different settings, usually called sequences, make the look of the output vary, as shown on Figure \ref{fig:PROSTATEx-t2-adc-dwi}.
 Unlike MRI, CT is based on X-rays instead of magnetic fields, but still provides a three-dimensional representation of a body part. Figure \ref{fig:visualize_lung_dcm} shows a lung CT scan.
@@
 \includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/PROSTATEx-t2-adc-dwi.png}
-\caption{MRI - PROSTATEx - From left to right: T2-weighted, ADC and DWI}
+\caption{MRI sequences - PROSTATEx - From left to right: T2-weighted, ADC and DWI}
 \label{fig:PROSTATEx-t2-adc-dwi}
@@
 After two iterations with other names, DICOM was created in 1993.
-It standardized the representation of medical images and their transmission since it provided a network protocol built on top of TCP/IP.
+It standardized the representation of medical images and their transmission as it provided a network protocol built on top of TCP/IP.
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}First of all, the name of the files within datasets is a 6-digit number, from 000000 to the number of images minus one.
-However, this order doesn’t match the real order of the images.
-In fact, the correct order is given by the "Instance Number" tags contained in the various files.
+However, this order does not match the real order of the images.
+In fact, the correct order is given by the "Instance Number" tag contained in the various files.
 Therefore, converted images must be sorted by instance number.
@@
 Otherwise, physicians may struggle to detect anomalies because of color and exposition-related variations.
-Therefore, before displaying or converting an image to any format (such as png), pixel data must be normalized.
+Therefore, before displaying or converting an image to any format (such as PNG or NumPy array), pixel data must be normalized.
 
@@
 \noindent If they are included in the DICOM header, a linear transformation must be done to convert the stored representation of the pixels to the correct visualizable one.
-To do this, two steps are required:
+To achieve this, two steps are required:
 
@@
 	\item \textbf{Apply the Hounsfield correction}\newline
-	Hounsfield Units (HU) are used in CT images it is a measure of radio-density, calibrated to distilled water and free air.
+	Hounsfield Units (HU) are used in CT images.
+It is a measure of radio-density, calibrated to distilled water and free air.
 Provided that the rescale slope and the rescale intercept are included in the DICOM header, the correction is applied thanks to the following formula:
@@
 \subsection{Origin}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
 The main problem of the latter was that it was lacking information about orientation in space.
@@
 \subsection{Data format}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single file “.nii” but keeps this split between the real data and the header for compatibility.
-This has the advantage to facilitate the use of the data and avoid storing the data without the meta-data.
-The NIfTI format can also be compressed/decompressed on-the-fly using the “deflate” algorithm.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Unlike the ANALYZE format that used two files to store the metadata and the actual data, the NIfTI file format stores them in one single “.nii” file, but keeps this split between the real data and the header for compatibility.
+This has the advantage to facilitate the use of the data and avoid storing the data without the metadata.
+The NIfTI format can also be compressed/decompressed on the fly using the “deflate” algorithm.
+A.
 M.
@@
 \subsection{Overview of the header structure}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}With the goal of preserving the compatibility between the ANALYZE and the NIfTI formats, both headers have the same size of 348 bytes.
+A.
 M.
-Winkler confirms this fact by claiming "some fields were reused, some were preserved, but ignored, and some were entirely overwritten"~\cite{52}.
-Details about the different fields contained in the header can be found in the reference of the previous citation.
+Winkler confirms this fact by claiming that "some fields were reused, some were preserved, but ignored, and some were entirely overwritten"~\cite{52}.
+Details about the different fields contained in the header can be found in the references of the previous citation.
 
@@
 \section{Visualization tools}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}
-Processing data manually increases the probability of making mistakes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Processing data manually increases the probability of making mistakes.
 For that matter, visualization tools relying on the same processing code as the ones used to generate training images were developed.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Visualizing DICOM files is pretty straightforward since each file represents a single two-dimensional image.
-However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
-Our tools allows to display a single DICOM file as well as a sequence of files if the function is fed with a directory.
+However, a lot of pixel transformations and normalizations have to be applied to obtain the desired result (see Section \ref{sec:dicom_data_manipulation}), which may be a source of errors.
+Our tool allows to display a single DICOM file as well as a sequence of files if the function is fed with a directory.
 Users can then scroll through the Z-axis, displaying the next or previous slice.
 Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung at three different Z-axis coordinates.
-This allows the user to simply have a complete overview of the body part he want to analyse.
+This allows the user to have a complete overview of the body part he wants to analyze.
 
@@
 \subsection{NIfTI}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allows to jump to the same slice at another timestamp.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As a single NIfTI file can contain three- or four-dimensional data (the fourth dimension being time), our visualization tool takes this aspect into consideration by allowing to scroll through them: scrolling with the mouse goes through slices belonging to a specific timestamp over a specific axis, while the left and right arrows allow to jump to the same slice at another timestamp.
 When reaching the end of a timestamp with the mouse, the first slice of the next timestamp is displayed.
@@
 Scrolling with the mouse goes through slices over a specific axis.
-Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part from three different perspectives.
+Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part under three different perspectives.
 This tool was particularly useful since no free software capable of handling these files was available.
@@
 \section{Conversion to PNG}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developped.
-The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG since it makes use of the others.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developed.
+The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG as it makes use of the others.
 Given a root directory, it scans the latter and its subdirectories to find all DICOM, NIfTI and RAW/MHD files.
@@
 This library allows to keep an easily accessible visual representation of medical files without having to rely on a specific piece of software to display them.
-Finally, it also played the role of debugging tool since the conversion techniques are the same as the ones used further in this work.
+Finally, it also played the role of a debugging tool since the conversion techniques are the same as the ones used further in this work.
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Medical data is often represented over 16 bits.
-However, exporting images to PNG require 8-bit images.
+However, exporting images to PNG requires 8-bit images.
 To transpose a 16-bit image to an 8-bit one, the procedure described by algorithm \ref{alg:16_to_8_bits_conversion} was applied.
@@
         		\State Modify object type to 8 bits
-        		\State Export $pixel\_array$ to PNG
+        		\State Export $pixel\_array$ as PNG
         \EndProcedure

### 2020-03-10 Common/chapters/paper_reproduction.tex
Julien and Johan: paper reproduction proofread

@@
 \label{ch:paper_reproduction}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in Chapter \ref{ch:literature_review}.
 It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).
 
-Song et al.
-\cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
-This dataset is one of the biggest dataset available for prostate cancer classification.
+Song et al.~\cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset.
+This dataset is one of the biggest datasets available for prostate cancer classification.
 The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1).
@@
 However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels).
-Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.
+Therefore, the reproduction of the experiment will fill that gap by taking part in the actual challenge.
 
@@
 Then, each theoretical notion is described in detail.
-This includes the structure of the dataset, the steps involved in its processing, and the techniques used as a verification of the proper functioning of the algorithm.Then, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
+This includes the structure of the dataset, the steps involved in its processing, and the techniques used as a verification of the proper functioning of the algorithm.
+Then, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment.
 This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "Discussion" section.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
-The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
-\begin{itemize}
+The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.
 
-\item The former makes use of the PROSTATEx training set.
+\subsubsection{Experiment reproduction}
+The first part of the experiment makes use of the PROSTATEx training set.
 This dataset is only composed of samples whose clinical significance is provided (labeled data).
@@
 %, which updates its weight consequently. 
-At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}).
+At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted in Tensorboard (see Section \ref{paper_tensorboard}).
 The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes.
-Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved.
-Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance.
+Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved.
+Furthermore, if the validation AUC decreases for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a progressive decrease in validation performance.
 This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.\\
 As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC.
-The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.
+The latter method consists in predicting an output for each augmented test image before averaging the predictions for each lesion of each patient.
 
-\item The second part of the experiment relies on the results of the first one.
-Indeed, the same data processing pipeline is used and the hyperparameters used to train the neural network are the same.
+\subsubsection{PROSTATEx challenge}
+The second part of the experiment relies on the results of the first one.
+Indeed, the same data processing pipeline is applied and the hyperparameters used to train the neural network are the same.
 The difference lies in the fact that the entire dataset (i.e.
-the entire PROSTATEx training set) is used as training set instead of splitting the data into a training/validation/test set.
+the entire PROSTATEx training set) is used as training set instead of splitting the data into a training, validation and test sets.
 In this way, the model sees more training examples, which maximizes the performance of the model on the challenge test set by increasing its generalization ability.
@@
 The model with the best result on the challenge is considered as the most efficient one.
-\end{itemize}
+
 \clearpage
@@
 \label{sec:prostatex_data_processing}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes all processing operations performed on the raw data until it passes through the neural network.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes the processing operations performed on the raw data until it passes through the neural network.
 All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
@@
 These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences".
-Each sequence allows to see human tissues in a different way on the resulting images.
+Each sequence allows to see human tissue in a different way on the resulting images.
 The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans.
@@
 \begin{itemize}
-\item The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can be associated with the same patient (ProxID).
+\item The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can belong to the same patient (ProxID).
 
@@
 The only difference is the absence of clinical significance.
-Finally, images are in the DICOM file format (see Section \ref{sec:DICOM}).
+Finally, images are in the DICOM file format (see \mbox{Section \ref{sec:DICOM}}).
 
@@
 One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images.
-The higher the b-value, the stronger the diffusion effects" \cite{49}.
+The higher the b-value, the stronger the diffusion effects"~\cite{49}.
 The authors tested three different b-value configurations.
 This led to the conclusion that using DWI images with the highest b-value increased the performance of the model.
-For that matter, DWI images with the highest b-value only were used.
+For that matter, DWI images with the highest b-value only were used in this work.
+Moreover, only images showing the prostate under the transverse plane were used.
+The reason for this is that the tumors are much more visible under this perspective.
 
-Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
-The first step consisted in converting DICOM files.
-Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays.
+The first step consisted in converting PROSTATEx's DICOM files to NumPy arrays.
+Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in this process.
 The right slices were found thanks to the two CSV files.
-Important information such as the patient ID, the sequence, the lesion location and the voxel spacing were used as file output names, which allows to use these files independently for the next steps (i.e.
-without relying on the CSV files).
-These information are then used for the next step that consists in stacking and augmenting the images.
-Furthermore, they are also extremely useful for the visualization scripts in order to provide concrete details regarding the images.
+Important information such as the patient ID, the sequence, the lesion location and the voxel spacing was included in the file output names, which allows to use these files independently for the next steps (i.e.
+without relying on the CSV files) that consist in stacking and augmenting the images.
+Furthermore, this information is also extremely useful for the visualization and verification scripts in order to provide concrete details regarding the displayed images.
 
@@
 \noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient.
-According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately helps the final diagnosis.
+According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately help the final diagnosis.
 In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array.
@@
 These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient.
-Same for the patient's DWI and ADC images.
-The same process was applied to all patients separately.
+The equivalent was applied to the patient's DWI and ADC images, and the same process was employed for all patients separately.
 Then, the dataset was augmented in the same way as the authors.
-Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel).
+Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and shifted horizontally (value $\in [-1, 0, 1]$ pixel).
 These techniques alone allowed to create a large enough amount of data.
-Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.
+Therefore, two other augmentation methods used by the authors (horizontal stretching by a \mbox{factor $\in [0.9,1.1]$} and vertical shifting) were not used.
 
@@
 These patients were simply discarded.
-At the end, the "false" class was composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class was composed of 55 training patients, 8 validation patients and 4 test patients.
+At the end, the "false" class contained 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class contained 55 training patients, 8 validation patients and 4 test patients.
 This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated.
-First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file.
+First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute of the CSV file.
 This full image was exported.
-Then, images were augmented in the exact same manner as described in section \ref{sec:numpyToAugmentedStacked}.
+Then, images were augmented in the exact same manner as described in Section \ref{sec:numpyToAugmentedStacked}.
 The expected result satisfies the following properties:
@@
 \noindent Figure \ref{fig:reddot} shows an example generated during this test.
-While beeing exactly the same for the 3 images, the red dot looks bigger on the images whose resolution was smaller since it was drawn before resampling the images.
+While being exactly the same size for the 3 images, the red dot looks bigger on the images whose resolution was smaller since it was drawn before resampling the images.
 These images come from the same exam performed on the same patient on the same day.
@@
 \subsubsection{Alignment}
-In addition to the previous test, an additional one aiming at checking the alignment of the images was created.
+In addition to the previous test, another one aiming at checking the alignment of the images was implemented.
 This test ensures that the three images show the exact same content at the same spot over the three channels (see Figure \ref{fig:alignment}).
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG).
-This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes.
+This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 million images belonging to 1000 classes.
 Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.
@@
 The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
-The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.
+The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced with a softmax function for classification.
 
@@
 \subsubsection{Script options}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}
-Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary.
 This requirement was solved thanks to the Python module called "argparse".
@@
 \textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
-\textbf{--trainingset}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
-\textbf{--validationset}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
-\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
-\textbf{--nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
-\textbf{--lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
-\textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
-\textbf{--cudadevice}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
-\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ''\end{tabular}                 & String        \\ \hline
-\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
+\textbf{-{}-trainingset}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
+\textbf{-{}-validationset}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{-{}-batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
+\textbf{-{}-nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
+\textbf{-{}-lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
+\textbf{-{}-lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
+\textbf{-{}-cudadevice}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
+\textbf{-{}-modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ''\end{tabular}                 & String        \\ \hline
+\textbf{-{}-dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline
 
-\textbf{--inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}True\end{tabular} & Int        \\ \hline
+\textbf{-{}-inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}True\end{tabular} & Int        \\ \hline
 
-\textbf{--optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
-\textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
+\textbf{-{}-optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
+\textbf{-{}-outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
 \end{tabular}%
@@
 \subsubsection{Tensorboard}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Tensorboard's official website~\cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
 \begin{itemize}
@@
 \item Projecting embeddings to a lower dimensional space.
-\item Displaying images, text, and audio data"\cite{39}.
+\item Displaying images, text, and audio data"~\cite{39}.
 \end{itemize}
@@
 The model whose score was the highest on a given metric (the AUC in this experiment) was saved and used as base model for the experiments.
-It can then be used as initialization model using the script option \textit{--modeltoload <path>}.
+It can then be used as initialization model using the script option \textit{-{}-modeltoload <path>}.
 
@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}
-Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
 Consequently, new hyperparameters had to be found.
@@
 In practical terms, this consisted in trying multiple combinations of hyperparameters and keeping the one that produced the best AUC on this test set.
-At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1e-7$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
+At the end of this process, it appeared that the best model was obtained when the data was organized in batches of 32 samples, with an initial learning rate of $1*10^{-7}$, a learning rate scheduler on plateau using a factor of 100 with a patience of 2 (the learning rate is divided by 100 when the validation AUC decreases for two consecutive epochs) and a dropout rate of 0.2.
 Note that the model was initialized picking the best one among 200 models generated from a roulette maximizing the AUC.
@@
 To achieve this goal, a visualization tool to display the gradients of the entire network was implemented.
-This implementation is based on the code of RoshanRane \cite{47}.
-This allows to visually notice the gradient flow at a glance and to see its evolution during all epochs (one visualization per batch is generated).
+This implementation is based on the code of RoshanRane~\cite{47}.
+This allows to visually notice the gradient flow at a glance and to see its evolution throughout all epochs (one visualization per batch is generated).
 
@@
 
-The figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6.
+Figure \ref{fig:gradient_flow} shows the gradient flow at epoch 0, batch 6.
 From the output layer (called "last\_layer0.weight" on the graph) to the first layer, the gradient is well propagated: no huge average gradients are registered, neither extremely small ones.
@@
 
-This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see figure \ref{fig:paper_reproduction_auc_on_our_test_set}).
+\noindent This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}).
 
-Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see figure \ref{fig:paper_reproduction_results_challenge_1}).
-This result will be used as benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
+Finally, the model was also tested on the official SPIE-AAPM-NCI Prostate MR Classification Challenge test set and achieved an AUC of 0.71 (see Figure \ref{fig:paper_reproduction_results_challenge_1}).
+This result will be used as a benchmark in comparison with the model trained on the whole dataset in the second part of the experiment.
 
@@
 \includegraphics[width=0.6\textwidth, keepaspectratio=true]{./figures/paper_reproduction_results_challenge1.png}
-\caption{Score achieved in the PROSTATEx challenge using only our custom training set (80\% of all available data) for the training}
+\caption{Score achieved on the PROSTATEx challenge only using our custom training set (80\% of all available data) for the training}
 \label{fig:paper_reproduction_results_challenge_1}
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
-In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"\cite{40}.
+In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"~\cite{40}.
 
@@
 In fact, at this moment, the validation AUC (the training is optimized according to this metric) was at its optimum with this learning rate.
-If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have lead to overfitting.
-Here, thanks to the learning rate scheduler, this scenario is avoided and the validation and training losses stay stable from this point onward.
+If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have led to overfitting.
+Here, thanks to the learning rate scheduler, this scenario is avoided as the validation and training losses stay stable from this point onwards.
 The same behaviour is clearly noticeable for all other metrics.
@@
 In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant.
-Armato.
-et al.
-\cite{42} claim that a less-experienced radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$.
+Armato et al.~\cite{42} claim that a less-experienced radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$.
 Consequently, this model is almost at the level of a radiologist, which is very promising.
@@
 
-Another extremely important aspect of a computer-aider diagnosis system is to avoid the number of false negatives as much as possible.
+\noindent Another extremely important aspect of a computer-aided diagnosis system is to avoid the number of false negatives as much as possible.
 In fact, the consequences of classifying a malignant lesion as benign are far worse than the opposite.
@@
 
-The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$.
+\noindent The two last metrics (precision and recall) are both considered in the f1-score that is equal to $0.76$.
 
-Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.
+\noindent Finally, the specificity indicates the proportion of correctly classified benign lesions among all benign lesions and is equal to approximately $0.76$.
 
-Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.
+\noindent Globally, this model is balanced as no metric is exploding at the expense of another. This stability constitutes a serious strength regarding its generalization ability.
 
@@
 Then, the way the test set is built can make the results vary a lot.
-In fact, the authors didn't provide any information about how they built their test set.
+In fact, the authors did not provide any information about how they built their test set.
 It is therefore possible that the test set was chosen to maximize the performance of the model.
@@
 Unfortunately, no information about the performance of the model of the paper is given.
-Knowing that the best result at the time of the challenge was an AUC of $0.87$ on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval $[0.876-0.994]$ with 95\% of confidence...
+Knowing that the best result at the time of the challenge was an AUC of $0.87$ on the challenge test set, it is highly surprising not to take part in it when claiming an AUC in the interval $[0.876-0.994]$ with 95\% of confidence.
 Finally, another source of difference is the number of participants in the experiment and the level of experience they have.
-Indeed, seven authors contributed to the paper, including one bachelors' student and six PhD with one of them working for Siemens Healthcare and one for hImagingTek Ltd.
-These big companies are specialized in the medical imaging and have departments dedicated to machine learning.
+Indeed, seven authors contributed to the paper, including one student with a Bachelor degree and six PhD with one of them working for Siemens Healthcare and one for hImagingTek Ltd.
+These big companies are specialized in medical imaging and have departments dedicated to machine learning.
 On the opposite, this work was produced independently by two MSc students without previous experience in the medical imaging field.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The first part of the experiment, in which the whole dataset was split into a training (80\%), a validation (10\%) and a test set (10\%), was extremely useful to find the right hyperparameters.
-It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesions classification.
+It also gave the opportunity to check that the model and the processing of the data worked as expected with the goal of creating a state-of-the-art deep learning model for prostate lesion classification.
 
-As a complement, this part of the experiment focuses on the challenge results in order to improve the $0.71$ in AUC achieved by the model trained with 80\% of all available data.
+As a complement, this part of the experiment focuses on the challenge results in order to improve the AUC of $0.71$ achieved by the model trained with 80\% of all available data.
 In a general way, the more available data there is, the better the model performs.
@@
 However, this does not cause any problem thanks to the results of the previous section.
-In fact, it is clearly visible that the model would have started to overfit at epoch 17 if no learning rate correction had been applied.
+In fact, it is clearly visible that the model would have started overfitting at epoch 17 if no learning rate correction had been applied.
 Hence, it is possible to conclude that, in this case, the best model should also be the one around epoch 17 (or a few epochs later).
@@
 
-Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
+\noindent Table \ref{fig:results_for_all_models_challenge} shows the AUC values achieved by the model on the official challenge test set.
 The range of values is contained in the interval $[0.75-0.76]$.
@@
 
-Finally, Figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, which achieved an AUC of $0.76$.
+\noindent Finally, Figure \ref{fig:paper_reprodution_results_challenge2} shows our participation to the challenge with the model picked at epoch 20, which achieved an AUC of $0.76$.
 
@@
 All metrics reach their point of convergence at some point, the curves do not oscillate much and the loss curve is continually decreasing, which is a good learning indicator.
-In comparison to the the first part of the experiment, it is clearly visible that the metrics have an higher best value.
+In comparison to the the first part of the experiment, it is clearly visible that the metrics have a higher best value.
 This is due to the absence of learning rate scheduler during the training.
 As a result, from a certain epoch onwards, features learned by the model are not transposable to unseen data anymore and the latter overfits the data.
-To determine this point, multiple models around the epoch 20 were evaluated on the challenge test set (see Figure \ref{fig:results_for_all_models_challenge}).
+To determine this point, multiple models around \mbox{epoch 20} were evaluated on the challenge test set (see Figure \ref{fig:results_for_all_models_challenge}).
 It is noticeable that the model did not reach its best performance before epoch 15 and that it started overfitting after epoch 23.
@@
 Finally, regarding the overall leaderboard of all challenge participants at the time of the challenge, the model would have been well ranked.
-Indeed, according to the 71 submissions to the PROSTATEx challenge (see Figure \ref{fig:challenge_all_results}), the AUC of $0.76$ achieved in this work would place the model at the 15$^{th}$ position overall.
+Indeed, according to the 71 submissions to the PROSTATEx challenge (see Figure \ref{fig:challenge_all_results}), the AUC of $0.76$ achieved in this work would have placed the model at the 15$^{th}$ position overall.

### 2020-03-10 Common/chapters/medical.tex
Julien and Johan: minors

@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The Neuroimaging Informatics Technology Initiative (NIfTI) file format is the successor of the ANALYZE file format.
-The main problem of the latter was that it was lacking information about orientation in space.
+The main problem of the latter was a lack of information about orientation in space.
 Therefore, the interpretation of stored data could be problematic and inconsistent.
@@
 Our tool allows to display a single DICOM file as well as a sequence of files if the function is fed with a directory.
-Users can then scroll through the Z-axis, displaying the next or previous slice.
-Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung at three different Z-axis coordinates.
+Users can then scroll through the z-axis, displaying the next or previous slice.
+Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung at three different z-axis coordinates.
 This allows the user to have a complete overview of the body part he wants to analyze.
@@
 Like the NIfTI file format, the user can profit from the three-dimensionality by displaying the body part under three different perspectives.
-This tool was particularly useful since no free software capable of handling these files was available.
+This tool is particularly useful as no free software capable of handling these files is available.
 Photography-related programs such as Photoshop can open RAW pictures but not these three-dimensional ones.
@@
 \section{Conversion to PNG}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In addition to the visualization tools, libraries allowing to convert medical files into PNG files were developed.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In addition to the visualization tools, scripts allowing to convert medical files into PNG files were developed.
 The main one is called "anydatasettopng" and is the one to use to convert any file format to PNG as it makes use of the others.
@@
 Then, each file is converted to PNG and saved into the same directory as the one it came from.
-This library allows to keep an easily accessible visual representation of medical files without having to rely on a specific piece of software to display them.
+This script allows to keep an easily accessible visual representation of medical files without having to rely on a specific piece of software to display them.
 Finally, it also played the role of a debugging tool since the conversion techniques are the same as the ones used further in this work.

### 2020-03-10 Common/chapters/paper_reproduction.tex
Julien and Johan: minors

@@
 The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown).
-This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection.
+This paper explains all the steps to follow with the goal of building a deep learning model for prostate cancer detection.
 Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set).
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}.
-The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.
+The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PROSTATEx challenge since the authors of the article did not take part in it.
 
@@
 The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes.
-Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved.
+Then, if the current model reaches a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved.
 Furthermore, if the validation AUC decreases for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a progressive decrease in validation performance.
@@
 At this point, the pixel $i,j$ of the three channels represents the exact same tissue, which means that the lesion is at the same position and is covered by the same number of pixels on each channel.
-This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necessarily as visible on the three images.
+This approach increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, as the latter is not necessarily as visible on the three images.
 
@@
 Moreover, they are not always as visible on the images of the various sequences.
-Therefore, finding similar features on all sequences was extremely difficult.
+Therefore, finding similar features on all sequences is extremely difficult.
 Stacking the three sequences solves this problem.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Thanks to the curves of Section \ref{sec:paper_reproduction_results}, a lot of information regarding the model performance can be inferred.
-In fact, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"~\cite{40}.
+In fact, as stated by Jason Brownlee, "reviewing learning curves of models during training can be used to diagnose problems with learning, such as an underfit or overfit model, as well as whether the training and validation datasets are suitably representative"~\cite{40}.
 
-First of all, when observing the loss plot, what is noticeable is that the training and validation losses have the same behavior during the 50 epochs.
+First of all, when observing the loss plot, it is noticeable that the training and validation losses have the same behavior during the 50 epochs.
 This is a clear sign of good fit of the model on the data.
@@
 The validation accuracy indicates that the model made the right class decision in approximately $75$\% of the time.
-Furthermore, since the model achieved a validation AUC of $0.76$, it shows that it had a good distinction capacity.
+Furthermore, as the model achieves a validation AUC of $0.76$, it shows that it has a good distinction capacity.
 In fact, the highest the AUC is, the highest is the ability of the model to predict benign lesions as benign and malignant lesions as malignant.
@@
 
-\noindent Another extremely important aspect of a computer-aided diagnosis system is to avoid the number of false negatives as much as possible.
+\noindent Another extremely important aspect in the medical field is to avoid the number of false negatives as much as possible.
 In fact, the consequences of classifying a malignant lesion as benign are far worse than the opposite.
@@
 
-First, in their work, the cropping of lesions was performed by a qualified radiologist, whereas, in our case, it was automatically done using the information in the CSV files.
+First, in their work, the lesion cropping was performed by a qualified radiologist, whereas, in our case, it was automatically done using the information in the CSV files.
 Then, the way the test set is built can make the results vary a lot.

### 2020-03-10 Common/chapters/transfer_learning.tex
Julien and Johan: minors

@@
 \label{sec:tl_goal}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of the task on a dataset B.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Transfer learning consists in using a network which was pretrained on a dataset A to improve the performance of a similar task on a dataset B.
 
@@
 
-In this experiment, the goal is to increase the model performance on the test set, particularly the AUC.
+Throughout the whole pipeline, the target metric to optimize is the AUC.
 In fact, the latter represents a degree of separability, i.e.
@@
 For instance, Model A and Model B can classify a sample X as benign (prediction: class 0) but with different probabilities of 0.10 and 0.45 respectively, where the latter are the probabilities of belonging to class 1 (malignant).
-While Model A is sure of choice, Model B is not trustworthy and may misdiagnose samples.
+While Model A is sure of its choice, Model B is not trustworthy and may misdiagnose samples.
 

### 2020-03-10 Common/chapters/paper_reproduction.tex
Julien and Johan: minors

@@
 This moment corresponds to the one where the learning rate is reduced by a factor of $100$ thanks to the \textit{reducelronplateau} function.
-In fact, at this moment, the validation AUC (the training is optimized according to this metric) was at its optimum with this learning rate.
+In fact, at this moment, the validation AUC (the training is optimized according to this metric) is at its optimum with this learning rate.
 If this technique had not been used, the training loss would have continued to go down whereas the validation loss would have gone up immediately, which would have led to overfitting.
@@
 This is particularly visible on the accuracy and AUC plots where the validation metrics are higher or equal to the training ones.
-The validation accuracy indicates that the model made the right class decision in approximately $75$\% of the time.
+The validation accuracy indicates that the model makes the right class decision in approximately $75$\% of the time.
 Furthermore, as the model achieves a validation AUC of $0.76$, it shows that it has a good distinction capacity.
@@
 As the goal of the experiment was to reproduce the experiment of the paper, a comparison between the two models is necessary.
-The reproducted model achieved an AUC of $0.75$ on our test set (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}) using the authors' "enhanced prediction" technique, whereas the authors achieved an AUC $\in [0.876-0.994]$ with 95\% of confidence on their test set.
+The reproduced model achieves an AUC of $0.75$ on our test set (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}) using the authors' "enhanced prediction" technique, whereas the authors achieved an AUC $\in [0.876-0.994]$ with 95\% of confidence on their test set.
 Consequently, the model presented in this work has a slightly smaller AUC of $[0.126-0.244]$ than the one achieved in the paper.

### 2020-03-11 Common/chapters/conclusions.tex
Julien and Johan: proofread til the end

@@
 The starting point was the reproduction of Song et al.'s experiment~\cite{07}.
-Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods were working well.
+Reaching good performance in this part despite being able to reproduce every single trick showed that our processing and training methods worked well.
 This resulted in a solid baseline that was exploited in order to take part in the "SPIE-AAPM-NCI Prostate MR Classification Challenge”, also called PROSTATEx challenge.
 Various hyperparameters and ways of processing the data were tested in order to reach an AUC of $0.76$ on this challenge.
-This score would have placed the model at the 15$^th$ position out of 71 submissions at the time of the challenge~\cite{42}, which confirms the robustness of the latter.
+This score would have placed the model at the 15$^{th}$ position out of 71 submissions at the time of the challenge~\cite{42}, which confirms the robustness of the latter.
 
@@
 Our transfer learning implementation alternated between frozen and unfrozen steps and made use of brain and lung datasets to increase the model performance on a prostate dataset.
-During frozen steps, the first part of the model (the feature extractor) does not update its weights at all, whereas the second part (the decision maker) does.
+During frozen steps, the first part of the model (the feature extractor) did not update its weights at all, whereas the second part (the decision maker) did.
 Experiments showed that our method allowed to increase the AUC on our test set by approximately 18\%, from $0.68$ before transfer learning to $0.80$ after transfer learning.
-Since a less-experienced radiologist can aim for an AUC of $0.81$~\cite{42}, this model could serve as a second diagnosis.
-These results could even be improved with another model, another way of processing data, etc.
+As a less-experienced radiologist can aim for an AUC of $0.81$~\cite{42}, some adjustments (better model architecture, other processing and augmentation techniques, more data, more time to optimize the hyperparameters, ...) could produce a trustworthy model capable of helping diagnose cancer.
 
-Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), an all-in-one training and testing file which can be adapted to new models and datasets, an end-to-end transfer learning pipeline.
-All these elements can be used as a baseline for future workds or as additions to existing projects.
+Throughout this thesis, various reusable tools were developed: visualization of medical imaging files, conversion of medical imaging files to PNG, a PyTorch sampler using undersampling, easy-to-use processing scripts for multiple datasets (PROSTATEx, Kaggle Brain, Lung CT Challenge), processing verification tools (red dot images), training verification tools (gradient flow graphs, metrics plots using Tensorboard), all-in-one training and testing files which can be adapted to new models and datasets, an end-to-end transfer learning pipeline.
+All these elements can be used as a baseline for future works or as additions to existing projects.
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}This work mainly used a single model architecture based on a VGG-16 network.
-As MRIs are three-dimensional, using three-dimensional convolutions could improve the results.
-This would also require new processing techniques which would generates volumes instead of two-dimensional images.
+Trying other architectures with other kernel sizes and a different number of layers can lead to better performance.
+Moreover, as MRIs are three-dimensional, using three-dimensional convolutions could improve the results as MRIs are three-dimensional.
+This would also require new processing techniques to generate volumes instead of two-dimensional images.
+
 However, well-labeled classification datasets can be counted on the fingers of one hand.
-As deep learning models need large amounts of training data, the continuation of the same work is more or less compromised.
+As deep learning models need large amounts of training data, the continuation of the same work is more or less compromised due to the lack of data.
 On the other hand, tumor segmentation has become popular over the last years.
-Most cancer-related deep learning challenges aim at segmenting tumors in three-dimensional data, which makes such datasets easy to find.
-Therefore, the experience and methods acquired in this work could be transposed to segmentation tasks, which could even combine segmentation and classification.
+Currently, most cancer-related deep learning challenges aim at segmenting tumors in three-dimensional data, which makes such datasets easy to find.
+Therefore, the experience acquired in this work could be transposed to segmentation tasks, which could even combine segmentation and classification.
 
@@
 New platforms ensuring their confidentiality, their public sharing and their quality have to be developed with the goal of encouraging hospitals to share them.
-This would allow to build even better models in an easier way.
+This would allow to build even better models in an easier way and make research move forward.

### 2020-03-11 Common/chapters/transfer_learning.tex
Julien and Johan: proofread til the end

@@
 First, a general definition of transfer learning is given.
-Its utility in this specific application is discussed as it is useful but not in any case.
+Its utility in this specific application is discussed as it can be useful but not in any case.
 Second, the different steps involved in the process are described in order to get an overview of the whole transfer learning pipeline that was applied.
@@
 \item Second, the convolutional neural network is built and split into two parts.
-It is still a single network but the first layers are treated as the feature extractor.
+It is still a single network but the first layers are treated as a feature extractor.
 As the name suggests, the role of these layers is to extract visual features characterizing what a cancerous tumor looks like.
-On the other hand, the last layers are considered as decision maker.
-They will gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
+On the other hand, the last layers are considered as a decision maker.
+They gather the features collected by the feature extractor and process them to make a decision on the class of the input sample.
 \item Third, the model is trained on the target dataset and the model whose performance is the best is saved.
 The target dataset is the dataset which one wants to improve the performance of the model on.
-This first training establishes a baseline of performance and of features to learn further on thanks to the other datasets.
-\item Fourth, the core of transfer learning process starts. For each non-target dataset, two different trainings take place:
+This first training establishes a baseline of performance and a baseline of features which will be improved further on.
+\item Fourth, the core of the transfer learning process starts. For each non-target dataset, two different trainings take place:
 
@@
 the weights are reset to another initialization, as if no training occured for these layers.
-Afterwards, the first layers (the features extractor) are frozen and the model is trained.
-Since the first layers have been frozen, their weights will remain the same.
-Only the weights of the last layers will change.
+Afterwards, the first layers (the feature extractor) are frozen and the model is trained.
+Since the first layers have been frozen, their weights remain the same.
+Only the weights of the last layers change during this training.
 This part of the transfer learning allows the model to make use of the previously learned features while adapting to the new body part by training the decision maker.
@@
 The only difference lies in the frozen part.
-Instead of dropping the old last layers and initialize its weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model.
+Instead of dropping the old last layers and initialize their weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model.
 These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again.
@@
 \makebox[\textwidth]{
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
+\includegraphics[width=1.1\textwidth, keepaspectratio=true]{./figures/tl_overview.png}
 }
@@
 
-The overall process is the same as the one described in section \ref{sec:numpyToAugmentedStacked}, apart from a few details.
-First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of stacked images.
-As DWI images are not of the same size, the fact of using the corresonding T2 images to resize them was the easiest way to make the new implementation go without a hitch.
+The overall process is the same as the one described in Section \ref{sec:numpyToAugmentedStacked}, apart from a few details.
+First, a T2-weighted image was still used to align and resize the corresponding DWI image, but was not exported at the end, resulting in a single-channel DWI image (NumPy array) instead of a stacked image.
+As DWI images are not of the same size, the fact of using the corresponding T2 images to resize them was the easiest way to make the new implementation go without a hitch.
 
@@
 Each patient can have one or multiple findings.
-Since labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
+As labels were provided for both sets and the amount of data is fairly low, they were merged and used as a global training set.
 For each finding, an abnormal mass is visible in the lungs, which means that the classification is going to differentiate malignant from benign tumors.
 
-Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain labels for these images.
+Regarding labelling, two Excel files, \textit{TestSet\_NoduleData\_PublicRelease\_wTruth} and \textit{CalibrationSet\_NoduleData}, contain the corresponding labels for these images.
 In order to facilitate the handling of information in the code, two CSV files were manually created: \mbox{\textit{TestSet.csv}} and \textit{CalibrationSet.csv}.
@@
 Both "malignant" and "Primary lung cancer" were considered as positive, whereas "benign" and "Benign nodule" were treated as negative.
-A third label called "Suspicious malignant nodule" appeared two times.
+A fifth label called "Suspicious malignant nodule" appeared two times.
 Since the diagnosis was not clearly defined for those images, they were not included in the training data to avoid any noise.
@@
 \subsubsection{From DICOM to augmented NumPy arrays}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays.) Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Like PROSTATEx, the processing was split into two parts: preprocessing (DICOM to NumPy arrays) and augmentation (NumPy arrays to cropped augmented NumPy arrays).
+Algorithm \ref{alg:LungCTChallenge_preprocessing} shows the various preprocessing steps.
 
@@
 Consequently, the selected patch size makes the largest tumors fit perfectly in the cropped area, inducing more tissue for smaller tumors.
-Naturally, there were few outliers whose size was a lot larger than the chosen patch size.
+Naturally, there were some outliers whose size was a lot larger than the chosen patch size.
 If the patch size was set accordingly, entire lungs were sometimes visible in other patients' case.
@@
 Images whose quality was extremely low (resolution, picture of a screen, artefacts, etc.) were manually removed.
-In other cases, the image quality was good but disturbing objects such as left/right markers, arrows pointing the tumor, etc.
-were visible.
+In other cases, the image quality was good but disturbing objects such as left/right markers or arrows pointing the tumor were visible.
 These objects were manually removed using the object removal tool in Photoshop or by cropping the image if the disturbing object was far from the skull.
-File names and file formats were heterogeneous.
-At the end of this process, the dataset was composed of 96 benign lesions and 133 malignant ones.
+At the end of this cleaning process, the dataset was composed of 96 benign lesions and 133 malignant ones.
+File names and file formats were still heterogeneous.
 
@@
 Finally, the "Diagnosis" was defined as "False" or "True" according to the folder which the image was in.
-This allows to simply use the same processing techniques as it is done for the other datasets.
+This allows to simply use the same processing techniques as for the other datasets.
 
@@
 Even though the images all show the head under the same perspective (same height/width ratio, visible skull, black margin around the skull), images were probably collected from different sources, which explains the heterogeneity of resolutions.
-As a consequence, the patch size could not be a fixed number of pixels for all images because the cropped area would be the entire head on one image and a small portion of the tumor on another.
+As a consequence, the patch size could not be a fixed number of pixels for all images because the cropped area would have been the entire head on one image and a small portion of the tumor on another.
 The trick consisted in finding the horizontal and vertical portions of the image occupied by the tumor.
@@
 Moreover, the abrupt borders inducing a clear change of colors can help detect tumors.
-Experienced showed that more context helps when the images are single-sequence.
-On the other hand, cropping right around the lesion or even in the lesion gave better results for multi-sequence data, i.e.
-when T2-weighted, DWI and ADC images (MRIs) were simultaneously used to perform the tumor detection and classification.
-In the latter case, the lesion usually appears distinctively on the three sequences, and this impacts the decision a lot.
+Experience showed that more context helps when the images are single-sequence.
+On the other hand, cropping right around the lesion or even in the lesion gives better results for multi-sequence data, i.e.
+when T2-weighted, DWI and ADC images (MRIs) are simultaneously used to perform the tumor detection and classification.
+In the latter case, the lesion usually appears distinctively on the three sequences, which impacts the decision a lot.
 On single-sequence images, there is a clear lack of information which makes it difficult to differentiate a tumor from another tissue without much context.
@@
 \subsubsection{Visual checking}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Regarding the Kaggle brain dataset, the data generated was paid close attention to since the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Regarding the Kaggle brain dataset, the data generated was paid close attention to as the ground truth CSV file was created by hand (see Section \ref{sec:braingroundtruth}).
 In fact, each generated image was taken a close look at, which allowed to spot 3 mistakes over the whole dataset.
-They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside.
-Like PROSTATEx (see section \ref{sec:dataProcessingVerification}), control images with red dots were also generated for the other datasets.
+They concerned typos in the "Nodule Center Position", which created pitch black images since the cropped area was outside of the original images.
+Like PROSTATEx (see Section \ref{sec:dataProcessingVerification}), images with red dots were also generated for the lung and brain datasets.
 
@@
 the first layers of the model.
-The exact point where the model is split is arbitrary, but is usually located after the convolutions.
+The exact point where the model is split is arbitrary.
+In our case, it is located after the convolutions.
 To perform this freezing, multiple steps are required:
@@
 	\item Reset the last layers.
-To do so, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance.
+To achieve this, an instance of the model is created by loading the state dictionary of the previous one, before replacing the module containing the last layers with a new instance (i.e.
+a new initialization).
 	\item Iterate over the batch normalization layers and freeze them.
 Step 3 is not sufficient to freeze batch normalization layers.
-This can be achieved by calling module.eval() on each module containing such layers (after the model.train() call).
+In an implementation based on PyTorch, this can be achieved by calling \textit{module.eval()} on each module containing such layers (after the \textit{model.train()} call).
 	\item Only pass the parameters of these last layers to the optimizer.
@@
 \subsection{Conditions to choose the best model}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}At each step of the transfer learning, the best model was saved in order to be used during the next transfer learning step.
-After each epoch, the current model was compared to the previous best model.
-If it was better, the current model became the best model.
-A model was considered better if its AUC was higher than the best model's AUC, and if its accuracy was higher than the best model's accuracy.
-As the goal was to optimize the AUC, the first condition makes immediate sense.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}At each step of the transfer learning, the best model is saved in order to be used during the next transfer learning step.
+After each epoch, the current model is compared to the previous best model.
+If it is better, the current model becomes the best model.
+A model is considered as better if its AUC is higher than the best model's AUC, and if its accuracy is higher than the best model's accuracy.
+As the goal is to optimize the AUC, the first condition makes immediate sense.
 The second condition was put in place in order to prevent a sudden spike in AUC to set an unreacheable value.
 This scenario did not occur during this experiment but during previous ones with random initialization.
-The AUC at early epochs could be really high, saving an untrained model the best model.
+The AUC at early epochs could be really high, saving an untrained model as best model.
 
@@
 In fact, the pipeline consists in seven different phases of training: dataset 1 (full model), dataset 2 (frozen model), dataset 2 (full), dataset 3 (frozen), dataset 3 (full), dataset 4 (frozen) and dataset 4 (full).
-In such a case, each phase will probably use different hyperparameters.
-Therefore, these command line options need to specify seven different parameters.
+In such a case, each phase uses different hyperparameters.
+Therefore, seven different parameters need to be specified for these command line options.
 For example, a valid value for the "-{}-lr" option could be "1e-8,1e-5,1e-6,1e-6,1e-5,1e-7,1e-8".
@@
 This dataset is useful for the global performance graphs (see Section \ref{sec:global_performance}).
-Another parameter used for the global performance graphs is the '-{}-last{\_}layer'.
-Its role is limited to this, unless '-{}-{}attach{\_}DM' is set to True.
-In this case, the decision maker loaded as '-{}-last{\_}layer' is attached to the current model, provided the current step is a frozen one.
+Another parameter used for the global performance graphs is the "-{}-last{\_}layer".
+Its role is limited to this, unless "-{}-{}attach{\_}DM" is set to "True".
+In this case, the decision maker loaded as "-{}-last{\_}layer" is attached to the current model, provided the current step is a frozen one.
 This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well).
 Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.
-
-Furthermore, arguments which were comma-separated values become single values, since this file runs a single transfer learning step instead of seven.
+Furthermore, arguments which were comma-separated values become single values, as this file runs a single transfer learning step instead of seven.
 Also, an output name for Tensorboard needs to be specified.
@@
 & \begin{tabular}[c]{@{}l@{}}False\\Default: './runs'\end{tabular}
-& String        \\ \hline
-\textbf{-{}-cv} & \begin{tabular}[c]{@{}l@{}}Whether to use cross-validation or not\\{[}'True', 'False'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}True\end{tabular}             
 & String        \\ \hline
@@
 On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset).
-As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
-Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e.
-the training on the first/target dataset, was saved.
+As described in Section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker.
+Hence, to see how much each step is actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e.
+the training on the first/target dataset, is saved.
 As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter.
-In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.
+In other words, if the target dataset is a prostate dataset, it will classify prostate tumors accurately.
 
-Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor.
-This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.
+Then, after each epoch of every other transfer learning step, the saved decision maker is attached to the current feature extractor.
+This temporary model is then tested on the target training and validation sets, which results in different training and validation metrics.
 
-At the end of the transfer learning pipeline, a graph was generated for each metric.
+At the end of the transfer learning pipeline, a graph is generated for each metric.
 These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task.
@@
 \subsection{Experimental setup}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The whole transfer learning consists in multiple steps.
-In order to maximize the performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}%The whole transfer learning consists in multiple steps.
+In order to maximize performance during the whole process, the hyperparameters were manually tuned at each step, following a grid search approach to find the best ones.
 Table \ref{table:tl_hyperparameters_at_each_step} presents the best hyperparameters used at each step of the experiment.
-Note that, as it was the case in the paper experiment reproduction, the first model was also initialized by loading the best model among 200 initializations.
+Note that, as it was the case in the paper experiment reproduction (see Section \ref{sec:paper_reproduction_experimental_setup}), the first model was also initialized by loading the best model among 200 initializations.
 The model was implemented with PyTorch using the Adam optimizer to update its weights.
@@
 
-\begin{table}[!h]
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{|l|l|c|l|c|l|}
+\begin{table}[H]
+\centering
+\resizebox{0.9\textwidth}{!}{%
+\begin{tabular}{|l|l|c|c|c|c|}
 \hline
@@
 \textbf{PROSTATEx}       & DS4/Frozen                                                 & 1e-5                   & 128                                                              & 0.3              & 2000                      \\ \hline
-\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0                & 2000                      \\ \hline
+\textbf{PROSTATEx}       & DS4/Full                                                   & 1e-9                   & 128                                                              & 0.0              & 2000                      \\ \hline
 \end{tabular}%
 }
-\caption{Hyperparameters used at each step of the transfer learning pipeline}
+\caption{Transfer learning -- Hyperparameters at each step}
 \label{table:tl_hyperparameters_at_each_step}
@@
 
-%\subsection{Independent training}
-%To get reference values and evaluate the impact of transfer learning on the model performance, a set of trainings on each dataset was performed independently.
-These experiments involved multiple hyperparameter combinations.
-The best results were used as benchmark values.
-%
-%\subsubsection{PROSTATEx}
-%
-%\subsubsection{Kaggle brain}
-%
-%\subsubsection{Lung CT Challenge}
-
-%\subsection{Transfer learning}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Figures \ref{fig:tl_ds1_full} to \ref{fig:tl_ds4_full} show the performance of the model at each step of the transfer learning.
-Each graph displays the metric values achieved on the training and the validation set on the Y-axis, whereas the X-axis represents the number of epochs.
+Each graph displays the metric values achieved on the training and the validation set on the y-axis, whereas the x-axis represents the number of epochs.
 
@@
 
-\noindent Table \ref{table:tl_best_model_performance} summaries the performance of the best model on the validation and test sets at each step of the experiment. 
+\noindent Table \ref{table:tl_best_model_performance} summarizes the performance of the best model on the validation and test sets at each step of the experiment. 
 Finally, Figure \ref{fig:tl_results_test_set} compares the performance of the model resulting from the first training to the one which underwent the transfer learning experiment.
@@
 \rowcolor[HTML]{DAE8FC} 
-\multicolumn{1}{|c|}{\cellcolor[HTML]{DAE8FC}\textbf{Dataset name}} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Best model saved at nth epoch} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Best model on the validation set}} & \textbf{Best model on the set set}                                                                                                                                      \\ \hline
+\multicolumn{1}{|c|}{\cellcolor[HTML]{DAE8FC}\textbf{Dataset name}} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Step}} & \textbf{Best model saved at epoch} & \multicolumn{1}{c|}{\cellcolor[HTML]{DAE8FC}\textbf{Best model on the validation set}} & \textbf{Best model on the set set}                                                                                                                                      \\ \hline
 \textbf{PROSTATEx}                                                  & DS1/Full                                                   & 527                                    & \begin{tabular}[c]{@{}l@{}}Accuracy: 0.6375\\ AUC: 0.7334\end{tabular}                 & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}Accuracy: 0.6842\\ Precision: 0.3333\\ Recall: 0.5000\\ F1-score: 0.4000\\ Specificity: 0.7333\\ AUC: 0.6833\end{tabular}}     \\ \hline
@@
 
+\begin{figure}[!h]
+\centering
+\noindent
+\makebox[\textwidth]{
+\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_results_test_set.png}
+}
+\caption{Transfer learning - Performance comparison between the best model obtained during DS1 Full and DS4 Full}
+\label{fig:tl_results_test_set}
+\end{figure}
+
 \section{Discussion}
@@
 \subsection{Training}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning experiment was performed in a manual way.
+%\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning experiment was performed in a manual way.
 In other words, each step was run sequentially using the "Manual transfer learning" script (see Section \ref{sec:tl_script_options}).
@@
 Hence, this model could only learn from one third of the information the first model had at disposal.
-Furthermore, the training and validation curves remained close during the whole experiment.
+Furthermore, the training and validation curves remain close during the whole experiment.
 This behavior indicates that the model actually learns features on the training set and generalizes well on new data (validation set), which is the best-case scenario.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The first frozen training took place on the brain dataset with a frozen version of the best model resulting from the previous step.
-During the first epochs, the metrics remained low.
-For example, the accuracy was below $0.575$ and the AUC below $0.5$, which means that a model classifying the input randomly could have performed better.
-This was expected since the last layers were reset to a new random initialization.
+During the first epochs, the metrics remain low.
+For example, the accuracy is below $0.575$ and the AUC below $0.5$, which means that a model classifying the input randomly could have performed better.
+This is to be expected as the last layers were reset to a new random initialization.
 In these early stages, the model was relying on a feature extractor capable of classifying prostate cancer.
@@
 The rest of the training aimed at learning new useful features from the brain images, which could then be used for lung and prostate cancer classification.
-As expected, this step increased the metrics a bit (accuracy of $0.8172$ and AUC of $0.8596$).
+As expected, the metrics increased a bit (accuracy of $0.8172$ and AUC of $0.8596$).
 
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The model was then frozen again and trained on the third dataset composed of lung images.
-Again, the metrics peaked early.
+Again, the metrics peak early.
 This is partly due to the larger learning rate of $1*10^{-5}$ and to the common features that prostate, brain and lung cancers share.
-The metrics went not as high as for the brain dataset but still showed good signs of the usefulness of the prostate and brain features.
+The metrics go not as high as for the brain dataset but still show good signs of the usefulness of the prostate and brain features.
 
@@
 Since these last layers had already been trained on the same dataset, the improvement occuring during this training stemed from new features learned from the brain and lung datasets.
-With a large learning rate of $1*10^{-5}$, the curves went up quickly at the beginning, then slowly and regularly until the completion of the training.
+With a large learning rate of $1*10^{-5}$, the curves go up quickly at the beginning, then slowly and regularly until the completion of the training.
 The initial spike and the progressive increase following it show that the low-level features acquired thanks to transfer learning are helpful for prostate cancer classification.
@@
 Regarding the full training, the curves contain quite some noise with an already small learning rate of $1*10^{-9}$.
-An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small since the training and validation curves were completely flat from the beginning to the end.
-The noise is also very visible on specific graphs such as the precision and specificity since the Y-axis scale is small.
+An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small as the training and validation curves were completely flat from the beginning to the end.
+The noise is also very visible on specific graphs such as the precision and specificity because the Y-axis scale is small.
 During this last part of the transfer learning, the validation metrics improved during the first 50 epochs approximately, before gently going down as the training metrics kept increasing.
@@
 As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.
-Furthermore, the full training curve shows that brain cancer features are actually helpful.
-If they were not, the metric would not have increased as fast and would not have reached such values.
+Furthermore, the full training curves show that brain cancer features are actually helpful.
+If they were not, the metrics would not have increased as fast and would not have reached such values.
 
-During the full training on the lung dataset, the AUC starts at a high value after the first epoch.
+During the full training on the lung dataset, the AUC started at a high value after the first epoch.
 The combination of the prostate, brain and lung features seem to help the model generalize.
@@
 This implies that the model reached a high level of generalization during transfer learning.
-These 50 epochs were enough to keep the low level features learned thanks to the other datasets while adapting the model to the target task again.
+These 50 epochs were enough to keep the low-level features learned thanks to the other datasets while adapting the model to the target task again.
 
@@
 
-As explained in Section \ref{sec:tl_goal}, the primary goal of this technique was to increase the AUC as much as possible, as it evaluates how well the model can distinguish between classes.
-The transfer learning allowed to increase the AUC increased by approximately 18\%, going from $0.68$ to $0.80$.
-Armato.
-et al.
+As explained in Section \ref{sec:tl_goal}, the primary goal of this technique is to increase the AUC as much as possible, as it evaluates how well the model can distinguish between classes.
+The transfer learning allowed to increase the AUC by approximately 18\%, going from $0.68$ to $0.80$.
+Armato et al.
 \cite{42} stated that a less-experienced human radiologist can reach an AUC of $0.81$ and an expert an AUC of $0.91$.
 This performance indicates that the model was able to learn low-level features from each body part in order to generalize.
-Hence, the goal of a creating a robust enough model which can serve as a second diagnosis for radiologists is reached.
+Hence, the fact that this model is close to humans is promising.
+%Hence, the goal of a creating a robust enough model which can serve as a second diagnosis for radiologists is reached.
 
@@
 In other words, the model made the same predictions but its predictions were a lot more distinct and trustworthy.
-The fact that this test set is relatively small can play a role in this outcome since the validation metrics were far better during the last training on PROSTATEx than during the first one (for instance, the validation accuracy went from $0.6375$ to $0.7749$ when the best model was saved, which is clearly visible on Figures \ref{fig:tl_global_validation_auc}, \ref{fig:tl_global_validation_accuracy} and \ref{fig:tl_global_validation_f1score} when comparing "DS1{\_}Full" with "DS4{\_}Full").
-
-\begin{figure}[!h]
-\centering
-\noindent
-\makebox[\textwidth]{
-\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_results_test_set.png}
-}
-\caption{Transfer learning - Performance comparison between the best model obtained during DS1 Full and DS4 Full}
-\label{fig:tl_results_test_set}
-\end{figure}
+The fact that this test set is relatively small can play a role in this outcome, as the validation metrics were far better during the last training on PROSTATEx than during the first one (for instance, the validation accuracy went from $0.6375$ to $0.7749$ when the best model was saved, which is clearly visible on Figures \ref{fig:tl_global_validation_auc}, \ref{fig:tl_global_validation_accuracy} and \ref{fig:tl_global_validation_f1score} when comparing "DS1{\_}Full" with "DS4{\_}Full").
 
@@
 This makes the model more robust and more likely to classify new unseen data well.
-As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), it could be used as a second diagnosis.
+As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), a few adjustments can create a robust model capable of being used as second diagnosis.
 These improvements are a step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.

### 2020-03-11 Common/chapters/transfer_learning.tex
Julien and Johan: minors

@@
 the first layers of the model.
-The exact point where the model is split is arbitrary.
-In our case, it is located after the convolutions.
+The exact point where the model is split is right after the convolutions.
 To perform this freezing, multiple steps are required:
@@
 An even smaller learning rate of $1*10^{-10}$ was tested out but it turned out to be too small as the training and validation curves were completely flat from the beginning to the end.
-The noise is also very visible on specific graphs such as the precision and specificity because the Y-axis scale is small.
+The noise is also very visible on specific graphs such as the precision and specificity because the y-axis scale is small.
 During this last part of the transfer learning, the validation metrics improved during the first 50 epochs approximately, before gently going down as the training metrics kept increasing.
@@
 As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}.
-The latter looks steeper because of the difference in the X-axis scale.
+The latter looks steeper because of the difference in the x-axis scale.
 
@@
 Furthermore, the full training curves show that brain cancer features are actually helpful.
-If they were not, the metrics would not have increased as fast and would not have reached such values.
+If they were not, the metrics would not have increased as fast and would not have reached higher values than the ones reached at DS1 Full.
 
@@
 These 50 epochs were enough to keep the low-level features learned thanks to the other datasets while adapting the model to the target task again.
+Overall, it is noticeable that the maximum AUC is higher at each step.
 
@@
 This makes the model more robust and more likely to classify new unseen data well.
-As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), a few adjustments can create a robust model capable of being used as second diagnosis.
+As this model reached an AUC ($0.80$) similar to the one a less-experienced radiologist can aim for ($0.81$), the goal of creating a model which generalizes well is reached.
 These improvements are a step forward in a field where the nonavailability of data is a real obstacle that prevents deep learning-based cancer classification systems from making a major breakthrough.

### 2020-03-13 Common/structure/info.tex
Julien and Johan: new model split figure
Julien and Johan: title

@@
 
-\thesistitle{Prostate Cancer Classification:\\A Freeze/Unfreeze Transfer Learning Approach Using Multiple Body Parts} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Prostate Cancer Classification:\\A Transfer Learning Approach\\to Integrate Information\\From Diverse Body Parts} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Prof. Dr. Philippe Cudré-Mauroux} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
 \cosupervisor{\href{https://exascale.info/members/giuseppe-cuccu/}{Dr.
-Giuseppe Cuccu},\\ \href{https://exascale.info/members/akansha/}{Akansha Bhardwaj}} % If you have a co-supervisor, include it here.
+Giuseppe Cuccu}\\ \href{https://exascale.info/members/akansha/}{Akansha Bhardwaj}} % If you have a co-supervisor, include it here.
 This is used in the title page, print it elsewhere with \supname

### 2020-03-14 Common/chapters/abstract.tex
Julien and Johan: typo
Julien and Johan: minors

@@
 % !TEX root = ../main.tex
-Automating the detection of cancer participates to an early detection and treatment, which increases the chances of recovery.
+Automating the detection of cancer contributes to an early detection and treatment, which increases the chances of recovery.
 Recent algorithms in artificial intelligence relying on deep learning have shown promising results in this field.
@@
 This thesis first presents a state-of-the-art convolutional neural network for pro\-state lesion classification.
-All steps from the data processing to the smallest detail regarding the neural network training are explained, ensuring a complete reproducibility of the experiment.
+All the steps from the data processing to the smallest detail regarding the neural network training are explained, ensuring a complete reproducibility of the experiment.
 This model is then evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, achieving an AUC of 0.76.
@@
 Indeed, it appears that cancerous lesions coming from various body parts share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier.
-This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancer can be taken advantage of.
+This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancers can be taken advantage of.
 Thanks to this technique, the AUC achieved on our test set increases by 18\% (from 0.68 to 0.80).

### 2020-03-14 Common/chapters/deep_learning.tex
Julien and Johan: minors

@@
 It starts with a historical timeline of deep learning before describing what a neural network is.
-Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained.
+Then, the notion of training a neural network, with all that is involved (like forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation), is explained.
 This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks".
 They are used in many computer vision applications due to their great performance on these tasks.
-Finally, the concept of "transfer learning" is discussed as the entire chapter \ref{ch:transfer_learning} relies on it.
+Finally, the concept of "transfer learning" is discussed as Chapter \ref{ch:transfer_learning} entirely relies on it.
 
@@
 Many concrete applications make use of neural networks: autonomous vehicles, translators, computer-aided diagnosis systems, personal assistants, art creation, robotics, etc.
-The presence of deep learning techniques in these use cases clearly confirm the enthusiasm of many for this technology.
+The presence of deep learning techniques in these use cases clearly confirms the enthusiasm of many for this technology.
 Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.
@@
           ylabel={$y$},
-          xlabel=X Axis, 
-          ylabel=Y Axis,
+          xlabel=x-axis, 
+          ylabel=y-axis,
           samples=41, 
@@
 \end{equation}
-The gradient shows how much the parameters that constitute x need to change to minimize the function.
+The gradient shows how much the parameters that constitute $x$ need to change to minimize the function.
 In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$.
@@
 The simplest and least efficient technique to initialize neural network weights is to randomly generate them.
-The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions.
+The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values that lead to values near 0 or 1 for most activation functions.
 Consequently, the slope of the gradient changes slowly and the learning process takes a lot of time.\\
@@
 In fact, from a certain point in the training onwards, neural networks do not learn anymore useful features in the data and start overfitting.
-This point corresponds to the moment where the validation loss does not decrease anymore and starts to increase continuously.
-It is exactly at this moment that the training should stop.
-To achieve this goal, it is possible either to directly choose the right number of epochs or to use the so-called "early stopping" method, which stops the training as soon as the validation loss does not decrease for $N$ epochs.
+This point corresponds to the moment where the validation loss does not decrease anymore and starts increasing continuously.
+It is at this moment that the training should stop.
+To achieve this goal, it is either possible to directly choose the right number of epochs or to use the so-called "early stopping" method which stops the training as soon as the validation loss does not decrease for $N$ epochs.
 
@@
 To perform a convolution, numerical input data and a "filter" are required.
-A filter can be seen as an $f$x$f$ numerical patch that moves across the entire input.
+A filter can be seen as an $f*f$ numerical patch that moves across the entire input.
 It first moves horizontally until reaching the right-most border of the image.
@@
 \noindent The output of the entire convolution shows negative values in the outside parts and large positive values in the center.
-This means that the $3x3$ filter detected a vertical line in the center of the input image.
+This means that the $3*3$ filter detected a vertical line in the center of the input image.
 Like other parameters, filters are learned while training.
@@
 Other applications could rely on a bigger stride.
-Furthermore, the previous example reduced the output size of $3$x$3$ in proportion to the initial input size of $5$x$5$.
+Furthermore, the previous example reduced the output size of $3*3$ in proportion to the initial input size of $5*5$.
 To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s.
@@
 The model dedicated to the second task uses some or all parts of the first model (i.e.
-keeps the same weights and architecture or a part of them) and is then retrained on data available for this task.
+keeps the same weights and architecture or a part of them) and is then retrained on the data that is available for this task.
 The first model can either be implemented from scratch if enough data is available or downloaded from institutions that release large pretrained models for similar tasks.

### 2020-03-14 Common/chapters/literature_review.tex
Julien and Johan: minors

@@
 The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal and vertical flipping, horizontal and vertical shifting of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
-Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer.
+Their model is a modified version of the well-known VGG-
+
+16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer.
 They also use the ELU activation function instead of ReLU (see figure \ref{fig:paper_model}).
@@
 
-\section{Lung/Lung CT Challenge}
+\section{Lung -- Lung CT Challenge}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Gao et al.
@@
 The model takes 4-dimensional data as input (depth, height, width and channels) and performs 3D convolutions on it.
-The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
+The model consists of an input layer, five layers of 3D convolutions (the first is associated with a ReLU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end.
 Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.
@@
 
-\section{Brain/Kaggle Brain}
+\section{Brain -- Kaggle Brain}
 
@@
 
-Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
+Saxena et al.~\cite{31} implemented three convolutional neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset.
 Their processing method used a cropping technique which removed extra black margin around the skull.
 Each border of the image was merged with a part of the skull.
-Since the images come from different sources, their resolution vary quite a lot.
+Since the images come from different sources, their resolution varies quite a lot.
 Therefore, the authors resized them to $224$x$224$x$3$.
-Moreover, data was augmented by rotation, vertical shifting and horizontal shifting.
+Moreover, data was augmented using rotation, vertical shifting and horizontal shifting.
 As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting.
@@
 The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$).
-The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy \mbox{of $0.55$}).
+The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 did not perform well (AUC of $0.55$ and accuracy \mbox{of $0.55$}).
 

### 2020-03-14 Common/chapters/medical.tex
Julien and Johan: minors

@@
 The earlier they are detected, the higher the chances of recovery are.
-This aspect makes cancer detection critical since every misjudgment can threaten someone's life.
+This aspect makes cancer detection critical as every misjudgment can threaten someone's life.
 
@@
 
-Other useful information such as the patient name and ID is directly stored within the DICOM files.
+Other useful information such as the patient's name and ID is directly stored within DICOM files.
 This approach aims at linking each image to a specific person and event in order not to mix them up.
@@
 
-where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, ~$y$ the pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255).
+where~$c$ is the window center,~$w$ window width, ~$P$ the pixel input value, \mbox{$y$ the} pixel output value, ~$y_{min}$ the minimal value of the output range (usually 0), ~$y_{max}$ the maximal value of the output range (usually 255).
 Equations \ref{eq:dicom1}, \ref{eq:dicom2} and \ref{eq:dicom3} ensure that the pixel values are correctly distributed within the output range.
@@
 M.
-Winkler affirms that this is particularly useful as "it is very common for images to have large areas of solid background, or files describing masks and regions of interest containing just a few unique values that appear repeated many time"~\cite{52}.
+Winkler affirms that this is particularly useful as "it is very common for images to have large areas of solid background, or files describing masks and regions of interest containing just a few unique values that appear repeated many times"~\cite{52}.
 
@@
 Users can then scroll through the z-axis, displaying the next or previous slice.
-Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung at three different z-axis coordinates.
+Figure \ref{fig:visualize_lung_dcm} shows the visualization of a lung CT scan at three different z-axis coordinates.
 This allows the user to have a complete overview of the body part he wants to analyze.

### 2020-03-14 Common/chapters/paper_reproduction.tex
Julien and Johan: proofread til the end
Julien and Johan: minors

@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes the processing operations performed on the raw data until it passes through the neural network.
-All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
+All the steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).
 
@@
                 						\State{$slice \gets normalize\_dicom(dicom\_file)$} \Comment{see Section \ref{sec:dicom_data_manipulation}}
-                						\State{Save $slice$ in $"output\_folder/clinical\_significance"$}
+                						\State{Save $slice$ into $"output\_folder/clinical\_significance"$}
                 					\EndIf
@@
 Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set.
-In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class.
+Each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" classes.
 At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients.
@@
 In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example.
-This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.
+This would be problematic since the validation set and the test set need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.
 
@@
 This full image was exported.
-Then, images were augmented in the exact same manner as described in Section \ref{sec:numpyToAugmentedStacked}.
+Then, images were augmented in the exact same manner as described in \mbox{Section \ref{sec:numpyToAugmentedStacked}}.
 The expected result satisfies the following properties:
@@
 This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 million images belonging to 1000 classes.
-Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.
+\mbox{Figure \ref{fig:paper_model}} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.
 
@@
 Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function.
-The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit.
+The same principle applies to the fully connected layer box (in orange) that is divided in a fully connected layer followed by the exponential linear unit.
 The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced with a softmax function for classification.
 
-In comparison to the original VGG, this model keeps the small filter size of 3x3, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
+In comparison to the original VGG, this model keeps the small filter size of $3*3$, also doubles the number of filters after each convolution-dropout-max-pooling block and has a stride of 1 for all convolutions.
 It differs from the traditional VGG in that it makes use of a smaller number of layers since the task is simpler than the original one.
 Moreover, it uses exponential linear units instead of rectified linear units as activation functions and adds dropout and batch normalization layers.
-Also, it uses 1x1 convolutions.
+Also, it uses $1*1$ convolutions.
 
@@
 \subsubsection{Experimental setup}
-
+\label{sec:paper_reproduction_experimental_setup}
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Keeping the exact same hyperparameters as the original paper led the experiment to disappointing results.
@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Figure \ref{fig:paper_reproduction_results} shows the different metrics registered on the training and validation sets during the training phase.
-The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy should be higher than the AUC and accuracy of the previous best model).
+The best model was chosen at epoch 21 with an accuracy of $0.7525$ and an AUC of $0.765$ on the validation set (recall: the condition for a model to be chosen as best is that the current AUC and accuracy must be higher than the AUC and accuracy of the previous best model).
 
-\noindent This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using their "enhanced prediction" technique (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}).
+\noindent This best model was evaluated on the test set built for the experiment and achieved an AUC of $0.75$ using the authors' "enhanced prediction" technique (see Figure \ref{fig:paper_reproduction_auc_on_our_test_set}).
 
@@
 In a general way, the more available data there is, the better the model performs.
-Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from previous section.
+Consequently, the idea behind this section is to use the entire dataset as training set, discarding the validation and test sets from the previous section.
 Apart from this new split, the data processing and the hyperparameters stay the same.

### 2020-03-14 Common/chapters/transfer_learning.tex
Julien and Johan: minors

@@
 %In this case, transfer learning aims at using various small datasets in order to achieve the same or better performance than with a single huge dataset.
-However, in order for the transfer learning to be useful, common low level features have to be present among the datasets on which the model is trained.
+However, in order for the transfer learning to be useful, common low-level features have to be present among the datasets on which the model is trained.
 Consequently, the goal of this experiment is to use transfer learning on three datasets coming from three different body parts that are prostate, brain and lung, in order to check if they share common features that could be useful to improve the performance of the model on a specific dataset.

### 2020-03-14 Common/chapters/transfer_learning.tex
Julien and Johan: minors

@@
 \subsection{Visualization of the impact of the various datasets on the target task}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx.
-As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}\mbox{Figure \ref{fig:tl_global_validation_auc}} shows the evolution of the AUC on the validation set of PROSTATEx.
+As expected, the first training on PROSTATEx generated the same curve as on \mbox{Figure \ref{fig:tl_ds1_full}}.
 The latter looks steeper because of the difference in the x-axis scale.

### 2020-03-25 Common/chapters/deep_learning.tex
Julien and Johan: typos

@@
 
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Machine learning algorithms can be belong to two classes.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Machine learning algorithms can belong to two classes.
 The first one is "supervised learning".
@@
 The model is trained on training folds and tested on the test fold.
-Then, another test set is picked and the same process is repeated until all possible tests set are picked.
+Then, another test set is picked and the same process is repeated until all possible test sets are picked.
 At the end of the process, the results of all test sets are averaged, which provides a good estimation of the model's performance.

### 2020-03-25 Common/chapters/introduction.tex
Julien and Johan: typos

@@
 Another benefit of using CAD systems (especially the ones based on deep learning) is the number of cases that the latter are trained on.
-In fact, in order to built efficient deep learning models, large amounts of data are required.
+In fact, in order to build efficient deep learning models, large amounts of data are required.
 Consequently, efficient CAD systems usually see a lot more cancer cases than beginner clinicians.
@@
 \item A state-of-the-art deep learning model for prostate cancer classification.
-To make the experiment completely reproductible, all steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented.
+To make the experiment completely reproducible, all steps are clearly described, the corresponding scripts are available on GitHub\footnote{\url{https://github.com/eXascaleInfolab/2019_Hospital-Fribourg}} and the raw results are presented.
 

### 2020-03-25 Common/chapters/paper_reproduction.tex
Julien and Johan: typos

@@
 \setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The methodology described in this section is the closest replication possible of the authors' data processing methodology.
-Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
+Most of it were reproducible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case.
 Furthermore, the authors did not mention any participation in the official PROSTATEx challenge.
@@
 
-An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e. 
+An important part of this process is the normalization. In fact, images were normalized before being stacked, based on a Z-score, i.e. 
 \begin{equation}

### 2020-03-25 Common/chapters/transfer_learning.tex
Julien and Johan: typos

@@
 This option is used at the last frozen step (target dataset), as this decision maker comes from the first training (target dataset as well).
-Doing this instead of a using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.
+Doing this instead of using a new random decision maker (as it is the case for the non-target datasets) allows to start with a skilled decision maker and optimize it according to the features learned during the transfer learning.
 Furthermore, arguments which were comma-separated values become single values, as this file runs a single transfer learning step instead of seven.

### 2020-03-26 Common/chapters/transfer_learning.tex
Julien and Johan: minor

@@
 \subsection{Conclusion}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the AUC of the model by 15\%.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This transfer learning approach using frozen and unfrozen parts allowed to increase the AUC of the model by 18\%.
 It increased the generalization ability of the model by learning features coming from different body parts.

### 2020-03-26 Common/chapters/transfer_learning.tex
Johan and Julien: minor
Julien and Johan: minor

@@
 The only difference lies in the frozen part.
-Instead of dropping the old last layers and initialize their weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model.
+Instead of dropping the old last layers and initializing their weights randomly, the decision maker of the best model found during the training on the first dataset is attached to the model.
 These final two trainings aim at keeping the interesting features from the other datasets while learning specializing the model in the target dataset again.

### 2020-03-26 Common/chapters/transfer_learning.tex
Julien and Johan: minor

@@
 \subsection{Layer freezing}
-\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with freezed layers.
-The freezed layers are those considered as feature extractor, i.e.
+\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}After each dataset switch, the model is first trained with frozen layers.
+The frozen layers are those considered as feature extractor, i.e.
 the first layers of the model.
@@
 Then, after each epoch of every other transfer learning step, the saved decision maker is attached to the current feature extractor.
-This temporary model is then tested on the target training and validation sets, which results in different training and validation metrics.
+This temporary model is then tested on the target validation set, which results in different validation metrics that summarize the whole process.
 
